### 2016-02-26 Draft.tex
Add draft.tex and .bib file

\documentclass[useAMS,usenatbib,referee,12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage[round]{natbib}
\usepackage{color}
\usepackage{caption}
\usepackage{tabularx, booktabs}
\usepackage{float}
\usepackage{adjustbox}
\usepackage[toc,page]{appendix}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{setspace}
\doublespacing

\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
\newenvironment{indpar}[1]%
     {\begin{list}{}%
             {\setlength{\leftmargin}{#1}}%
             \item[]%
     }
     {\end{list}}

\begin{document}

%\tableofcontents

\section{Abstract}

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate.  However, the assumption of constant detection rates during surveys is often not justified.  Detection rates can be influenced by organismal behaviors, including responses to observer presence and heterogeneity among population subgroups, and also by survey methods that affect observer effort.  Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  Under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals.  We instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.  Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.  Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.

{\bf Keywords:} abundance; availability; N-mixture model; survival analysis; Bayesian

\section{Introduction}\label{sec:intro}

\adam{WRITE A SPELLBINDING OPENING PARAGRAPH}

%Farnsworth's removal assumptions:\\
%(i) Closed population during survey\\
%(ii) no double-counting\\
%(iii) easy-to-detect group is entirely counted during first interval\\
%(iv) TTDD for hard-to-detect group is correct after the end of the observation period\\
%(v) if using limited radius counts, they are accurate
% Johnson2008: all birds present are available for detection

Removal sampling is a survey methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.  Assuming that the probability of capturing a not-yet-trapped individual is the same for all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time \citep{Moran1951, Zippin1958, Seber1982}.  \citet{Farnsworth2002} adapted removal sampling to avian point-count surveys, suggesting that observers record just the first time-to-detection for each bird and mentally remove it thereafter.  Analysis proceeds by subdividing the observation period into intervals of equal duration and assuming that the rate of detection is the same at all times (i.e., following a Poisson process); then, each interval can be treated as a trapping event, and the interval-specific probability of detection can be estimated as with conventional removal sampling.

However, the assumption of a constant detection rate throughout the observation period is not always justified.  Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.  The act of observation may affect animal behavior.  Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002}.  Survey methodology itself can affect detection rates.  In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time.  Observers may also become distracted as the survey progresses \citep{Johnson2008}.  And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.

In this study, we model scenarios where detection rates are not constant.  We model time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function $F_T(t)$.  We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  The instantaneous detection rate, a.k.a. hazard function $h(t)$, can be found from the TTDD and vice versa, by observing $h(t) = f_T(t) / (1-F_T(t))$.

Our model is different from most survival analysis in that the number of individuals $N$ present at a survey is unknown.  To address this data truncation, we embed the TTDD in a hierarchical framework for multinomial counts sometimes termed an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.  An N-mixture model treats abundance as a latent variable modeled according to a Poisson or other discrete distribution, and it models detection probabilities independently.  Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014}.  For our purposes, the N-mixture framework provides three clear benefits.  First, its multinomial data framework accords with the interval censored data collection that is customary in point-count surveys \citep{Ralph1995}.  Second, the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}.  Third, for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).

The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.  Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Refs}.  Such an approach is not possible when detection rates are not constant.  Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.  Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection models can only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.

\subsection{TTDDs as behavior and methodology}

A TTDD expresses the range of detection patterns possible in an analysis.  The most basic is an exponential TTDD, which imposes a constant detection rate.  If species behavior and/or observation methodology may cause a systematic non-constant detection regime, then we should choose a TTDD accordingly.  For example, we may wish to model an observer effect where: (i) the observer's arrival either suppresses or stimulates detectable cues from organisms at a site, but (ii) as organisms acclimate to the observer, cues gradually return to an undisturbed baseline of constant detection.  As illustrated by the hazard functions in Figure \ref{HazardFxns}, a gamma TTDD would be appropriate for this scenario.

Increasing hazard functions result in a probability distribution function (pdf) for $T$ that has a mode away from zero; we term these TTDDs 'peaked'.  We contrast these to decreasing hazard functions with a mode at zero, which we term 'non-peaked'.  For our purposes, the exponential TTDD is neither peaked nor non-peaked.

One issue related to non-constant detection is detection heterogeneity.  Heterogeneity across population subgroups can be modeled using a mixture of detection distributions \citep{Pledger2000}\adam{longer history here... do I even need a reference? According to NorrisPollock1996, this goes all the way back to Burnham1972}.  In particular, \citet{Farnsworth2002} introduce a partial finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.   Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  A similar marginal detection pattern can be achieved by modeling non-constant detection using a non-peaked TTDD.  Detectability models constructed to address heterogeneity may provide reasonable estimates when detection rates change for behavioral reasons and vice versa \citep{Mantyniemi2005}.

\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}

\begin{figure}[h!]\centering
\includegraphics[width=0.68\textwidth]{Hazards.pdf}
\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
\end{figure}

In this study, we construct a time-to-detection N-mixture model for estimating abundance from removal-sampled point-count surveys.  We explore model behavior through a series of simulations testing: (i) how well constant-detection models fit non-constant detection data and vice versa, (ii) the utility of a partial finite-mixtures to model detection heterogeneity involving a non-constant detection subgroup, and (iii) model performance with covariates and random effects affecting both abundance and detection.  Finally, we apply our model to Ovenbird data from the Minnesota Forest Breeding Bird Project.

\section{Model}\label{sec:model}

\subsection{Data structure}\label{sec:data}

Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$.  We present the model for times to detection censored into $i=1,\dotso,I$ intervals, but it is straight-forward to simplify the model for exact time-to-detection data.  Our data therefore take the form of counts indexed by site, time period, and interval: $n_{sri}$.  For any given survey (uniquely defined by site and time period), we write the vector of counts across all $I$ intervals as $\textbf{n}_{sr}$.

\subsection{Framework}

We construct a hierarchical model of removal-sampled point-counts patterned after an N-mixture framework \citep{Wyatt2002, Royle2004NMixture}.  We treat the hierarchical model in three pieces: (a) a multinomial data model, (b) a detection model based upon the partitioned cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.  Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.

\subsubsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the time-to-detection is independent, identically distributed for all individuals of a species at any given survey.  Given that an organism is detected, we denote the probability it is first detected during interval $i$ as $p_{sri}$ such that $\sum_{i=1}^I p_{sri} = 1$.  We write $\textbf{p}_{sr}$ to denote the vector of interval-specific detection probabilities at any given survey.  The data model is then:

\begin{align}\label{eq:datamodel}
\textbf{n}_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
\end{align}

where for each survey, $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$ is the number of individuals counted across all intervals, $N_{sr}$ is the total number of individuals present (counted and uncounted), and $p_{sr}^{(det)}$ is the probability that an individual is detected at any time during the observation period.  The number of uncounted individuals at any survey can be written: $n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)}$.

\subsubsection{Detection Model}\label{sec:detectionmodel}
We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by means of the partitioned cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$.  Here, $T$ is a random variable for the first time at which an individual is detected, $\textbf{X}_{sr}^D$ and $\textbf{Z}_{sr}^D$ are detection-related covariates and random effect factor levels to be used in a mixed effects linear regression, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.  In this study, we consider eight different forms for the TTDD.  The first four TTDDs we employ are:
\begin{enumerate}
\item $T_{srq} \overset{ind}{\sim} \text{Exponential}(\varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Gamma}(\alpha, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Weibull}(k, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Lognormal}(\mu_{sr}, \sigma_{det}^2)$
\end{enumerate}
where $q = 1,\dotso,N_{sr}$ indexes individuals present at survey $sr$, and $\varphi_{sr}$ is always a survey-specific detection rate parameter.  We treat the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ as constant across all surveys.

The other four TTDDs we use are partial finite-mixtures of a point mass at zero paired with each of the above distributions.  They can be written: $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$ where $\gamma \in (0,1)$ is a mixing parameter and $F_{T1}(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$ is one of the four above distributions.  Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.

For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{sr}$ via mixed effects regression using survey-specific covariates.  In so doing, we are equivalently modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  For the lognormal distribution, we fit a regression to $-\mu_{sr}$.  Thus, the link functions in our detection model are:
\begin{alignat}{3}
&\text{Exponential:\;} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}])\\
&\text{Gamma:} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log(\alpha)\\
&\text{Weibull:}  &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
&\text{LogNormal:} &&-\mu_{sr} &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \sigma_{det}^2/2
\end{alignat}
where $\boldsymbol{\beta}^D$ is a vector of fixed effects, and $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.  When the TTDD is a mixture distribution, we model $F_{T1}$ portion of the mixture according to the above link functions and treat the mixing parameter $\gamma$ as constant across all surveys.

For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  Conditioned on an individual being detected, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.  In summary, the detection model is:
\begin{align}
p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
\end{align}

To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq. (\ref{eq:datamodel}).  Instead of modeling the interval-specific counts $\textbf{n}_{sr}$, we model the times to detection directly.  Defining $\textbf{T}_{sr}$ as a vector of all observed detection times for a given survey, and continuing to treat all detections as independent, we write
\[
Pr\left(\textbf{T}_{sr}|n_{sr}^{(obs)}\right) = \prod\limits_{q=1}^{n_{sr}^{(obs)}} f_T(t_{srq}|t_{srq}<C_I)
\]
The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.

\subsubsection{Abundance Model}
For simplicity, we model abundance at each site during each year $N_{sr}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{sr}$, though any non-negative integer-valued distribution may be used.  We can decompose $N_{sr}$ into observed and unobserved portions.  The unobserved portion follows a binomial-Poisson hierarchy, resulting in Equation \ref{eq:binPois}.  We model expected abundance $\lambda_{sr}$ using a mixed effects linear regression with a log link.

\begin{align}
\label{eq:binPois} n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)} &\sim \mbox{Poisson}\left(\lambda_{sr}(1-p_{sr}^{(det)})\right)\\
\log (\lambda_{sr}) &= \text{Intercept}^A + \textbf{X}_{sr}^A\boldsymbol{\beta}^A + \textbf{Z}_{sr}^A\boldsymbol{\xi}^A
\end{align}
where $\textbf{X}^A$ is a matrix of covariates associated with the vector of fixed effects $\boldsymbol{\beta}^A$, and $\textbf{Z}^A$ is a matrix identifying factor levels for $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ random effects.

Since we use this hierarchical model in a Bayesian context, the complete model formulation requires priors for $\text{Intercept}^D, \boldsymbol{\beta}^D, \sigma_{D}^2, \boldsymbol{\theta}, \text{Intercept}^A, \boldsymbol{\beta}^A, \text{ and } \sigma_{A}^2$, such as we specify in Section \ref{sec:priors}.

\subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}

We analyze point-count surveys conducted in Chippewa National Forest over a six-year period from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  The MNFB design is described in \citet{Hanowski1995}, and we only summarize relevant details here.  Point-count time-to-detection surveys are conducted by trained observers at a fixed set of sampling locations once annually.  Forest stands follow a stratified random design, with three survey sites in each stand.  Survey durations are 10 minutes, with times to detection being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  We restrict survey sites in our analysis to those of the most abundant habitat type (sawtimber red pine); in so doing, we focus modeling efforts upon the detection model and skirt the complication of modeling abundance across habitat types.  We further constrain analysis to 65 sites with a site origin year between 1870-1970.  In total, we use 381 site-year surveys in our analysis with a total of 947 Ovenbirds counted, and a maximum of eight birds counted in any single survey.

For the abundance half of our model, we use four covariates plus two random effects.  The covariates are: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.  We associate random effects with each survey year and each stand.  For the detection half of our model, we use covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.  69\% of surveys in our dataset involve observers in their first year at MNFB.  Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  Finally, we associate random effects with each observer.  We center and standardize all continuous covariates prior to fitting models.

% Effects: * - means signif
% Observer variation - Farnsworth2002, LinkSauer98, LinkSauer97, Dief07 cites Sauer94 & Dief03, Simons et al03
% First-year observer - LinkSauer98
% Time of day - Farnsworth2002, Soly13*, Amundson
% Season (jdate) - Farnsworth2002, Soly13*, Amundson, Dief07
% Quadratic effects - Soly13
% Tree cover (density) - Soly13
% Year (on detection!) - Reidy11, see also Norvell03
% See Table 1 of Johnson2008, see McShea and Rappole, lit review from Warren2013, Rosenstock02

We fit eight models to the Ovenbird data, once using each of the time-to-detection distributions described in Section \ref{sec:detectionmodel}.

\subsubsection{Priors}\label{sec:priors}
Our goal in selecting priors is to be non-informative within a reasonable range of values.  Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  All priors are independent.
\begin{align*}
\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
\text{Intercept}^A &\sim N(1, 1)\\
\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
\boldsymbol{\beta}^A &\sim N(0, 1)\\
\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
\gamma &\sim \text{Beta}(1,1)
\end{align*}

\subsection{Simulation studies}

We conduct three simulation studies, the first two of which implement intercept-only models.  The objectives for the studies are to assess: (i) the importance of modeling detection heterogeneity via a partial finite-mixture component, (ii) the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to a dataset simulated from a Weibull TTDD), and (iii) model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.  Because model behavior may differ between datasets with peaked and non-peaked TTDDs, we simulate both kinds of dataset for gamma, Weibull, and lognormal models.  For all families, we simulate both mixture and non-mixture datasets.  In total, we simulate 14 different dataset types.  Counts are censored into nine-intervals as with the Ovenbird dataset.

In the first simulation study (Sim 1), for all 14 dataset types, we fit models with both mixture and non-mixture TTDD where the family is correctly specified.  In Sim 2, we consider only mixture TTDD datasets, and we fit them with mixture TTDD models from all four families.  For both Sim 1 and Sim 2, we simulate and model 16 replicates.  In Sim 3, we repeat both the first two studies, but: (i) we include covariates and random effects described in Section \ref{sec:ovenbirdanalysis}, and (ii) we perform only a single replicate, because each gamma model takes an average of 5 days to fit.

For the first two simulation studies, we choose true parameter values so that: (i) the expected detection probability is 0.80, (ii) the expected count is the same as observed in the Ovenbird dataset, (iii) when there is a mixture component, we set $\gamma = 0.65$, meaning that 35\% of individuals are immediately detected, (iv) in non-peaked models, 70\% of \textit{detected} individuals are observed during the first two minutes, and (v) in peaked models, the detection mode for `hard to detect' individuals occurs at 5 minutes.  For Sim 3, we set true parameter values for each dataset equal to the median posterior values obtained when fitting the same TTDD model to the Ovenbird data.  Since all models applied to the Ovenbird data resulted in non-peaked model fits, we simulated peaked dataset variants by: (i) starting with posterior median Ovenbird estimates, (ii) setting the mixing parameter to $\gamma = 0.65$, and (iii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, which on average results in peak detection near 5 minutes and 75-85\% detection across all sites.  Covariates and random effect levels are taken directly from the Ovenbird dataset.  Unlike Sims 1 \& 2, overall detection probabilities in Sim 3 are not the same for each dataset.

\subsection{Model fitting}

We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.

For each intercept-only simulation model, we instantiated a single chain running 60,000 iterations.  For the Ovenbird analysis and covariate simulations, we ran chains for 250,000-375,000 iterations.  We discarded half of the iterations as warmup.  We reran models if the effective sample size of for any root parameter was below 1000.  We monitored convergence of the MCMC chains using Geweke diagnostics.  For most models, we accepted the Stan defaults for generating dispersed initial values.  However, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from the posterior means of preliminary short-run models (Ovenbird analysis, incorrect-family simulations) or from true parameter values (correct-family simulations).

\subsection{Model assessment, inference, and between-model comparison}

To assess goodness of fit, we rely on posterior predictive checks \citep{Gelman1996}/adam{Rubin1984?}.  Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  In particular, we are concerned with the goodness of fit for the TTDD distribution, so we define a check statistic for each observation interval:
\[D(n_i) = \sum\limits_{sr} n_{sri} \big/ \sum\limits_{sri}^{} n_{sri}\]
which is the proportion of the total count that occurs during interval $i$ marginally across all surveys.  The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.  Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.  We simulate replicate datasets at each iteration of our retained MCMC samples.

We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{sri}n_{sri}\big/$ $\sum\limits_{sri}N_{sri}$.  Because some posterior distributions are skewed, we emphasize posterior medians.  Within simulation studies, we compare model estimates to true $p^{(det)}$ values by coverage and by posterior p-values --- the proportion of MCMC samples that have a smaller value than the true value.  We do not focus on estimation of individual detection parameters, since different sets of parameters can yield similar density functions for time-to-detection conditional upon being detected: $f_T(t\;|\;t<C_I)$.

We compare different model fits by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  For the full model with random effects, which is a missing data model, we use formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.

\section{Results}\label{sec:results}

\subsection{Simulation studies}

\subsubsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}

In the first simulation study, we simulated intercept-only datasets from 14 TTDDs and fit each with both the correctly specified model and one that was correct except for the mixture component (Table \ref{tbl:sim1}).  When data derive from a mixture model, then inference models that lack a mixture component misestimate the detection probability $p^{(det)}$ badly in every case.  This is particularly egregious when datasets feature peaked mixtures; then model coverage is always negligible.  Conversely, when data derive from a non-mixture model, then inference models that include a mixture component perform just as well as the true model.  Indeed, for non-peaked non-mixture data, the mixture inference model provides more accurate posterior estimates of $p^{(det)}$ than the non-mixture inference model.  The one exception is the exponential case, where both models estimate $p^{(det)}$ accurately, but the mixture inference model yields credible intervals about 25\% wider than those from its mixture counterpart.

% Possible other comments
%(1) For non-peaked data, the true inference model is biased low\\
%(2) Mixture inference models generate higher estimates than non-mixture models\\
%(3) True models are generally good\\
%(4) Estimates of non-mixture peaked data are much more precise\\
%(5) Note that plots of TTDD at the median are indistinguishable from the truth except in the blatant fail scenario

\begin{table}[ht]\centering\small
\begin{tabular}{ll|l|cccc|cccc||c}
 \multicolumn{3}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
 \multicolumn{3}{c|}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
  \hline
  \hline
 \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\ 
 & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
 & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
  \cline{3-12}
 & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
  \cline{3-12}
 & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
 & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
 & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
  \hline
\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\ 
 & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
 & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
  \cline{3-12}
 & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
  \cline{3-12}
 & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
 & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
 & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
   \hline
\end{tabular}
\caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim 1).  In all cases, the model family matches the dataset family.  Med $p$: mean value across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  Q($p$): proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value on average.  50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  $\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model. \adam{Thoughts on use of bold/italics to highlight some results?}}
\end{table}

\subsubsection{Constant vs. non-constant detection}\label{sec:family}

In the second simulation study, we simulated intercept-only datasets from 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, Weibull, and lognormal families (Table \ref{tbl:sim2}).  Mixture exponential inference models returned accurate estimates of $p^{(det)}$ for mixture exponential data and had 95\% credible intervals from one-third to one-half as wide as from other inference models.  However, exponential mixture models yielded biased estimates and poor coverage for non-exponential data, overestimating detection for non-peaked datasets and underestimating it from peaked datasets.  Lognormal mixture models underestimated detection from all datasets; when modeling lognormal data, they were competitive with other models in terms of accuracy and coverage but were not superior.  Mixture gamma and Weibull models provided solid coverage for 90\% credible intervals with all datasets.  Gamma models were more accurate than others for non-peaked mixture datasets; gamma and Weibull seem equally accurate for peaked mixture datasets.  95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.

\begin{table}[ht]
\small\centering
\begin{tabular}{l|l|ccccc|ccccc}
 \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
 \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff \\ 
  \hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\ 
&   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
&   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
\cline{2-12}
&   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
\cline{2-12}
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
&   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
&   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
   \hline
\end{tabular}
\vspace{0.5cm}\\
\begin{tabular}{l|l|ccccc|ccccc}
 \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
 \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff \\ 
  \hline
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\ 
&   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
&   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
\cline{2-12}
&   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
\cline{2-12}
\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
&   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
&   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
   \hline
\end{tabular}
\caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim 2).   Med $p$: mean value across simulations of the posterior median of $p^{(det)}$.  Q($p$): mean proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value (0.80).  50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  $\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
\end{table}

\subsubsection{Simulations with covariates and random effects}\label{sec:simfull}

In the third simulation study, we generated datasets from models including covariates and random effects for both detection and abundance.  Results echoed those of the first two simulation studies, suggesting that the addition of covariates and random effects does not alter basic model behavior.  When data came from a mixture, non-mixture inference models generally led to biased estimates of detection while mixture inference models did not (Figure \ref{pdet_cater_correct}).  When data did not come from a mixture, then: (i) credible intervals for peaked data were similar for both inference model types, (ii) mixture model estimates of $p^{(det)}$ were higher for non-peaked data, and (iii) mixture model credible intervals were about 30\% wider for truly exponential data.

\begin{figure}[h!]\centering
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
\caption{\label{pdet_cater_correct} Sim 3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.  Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  Each column presents one family of simulated dataset.  Upper plots show non-mixture datasets; lower plots show mixture datasets.  'X' marks the expected marginal probability of detection based on true parameter values.}
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
\caption{\label{pdet_cater_family}  Sim 3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$. All data and inference models include mixtures.  Each plot presents one simulated dataset.  'X' marks the expected marginal probability of detection based on true parameter values.}
\end{figure}

As in Sim 2, for mixture exponential datasets only, mixture exponential inference models in Sim 3 returned accurate estimates of $p^{(det)}$ and had credible intervals from one-fourth to one-half as wide as from other inference models (Figure \ref{pdet_cater_family}).  However, exponential mixture models were generally biased high for non-peaked datasets and low for peaked datasets.  Lognormal mixture models tended to underestimate detection from non-lognormal datasets.  Gamma models were more accurate than other models for non-peaked mixture datasets; gamma and Weibull both seem equally accurate for peaked mixture datasets.  95\% credible intervals for gamma models were 15-35\% narrower than for both Weibull and lognormal models.

In many situations, individual parameter estimates are of interest.  For every dataset we simulated, estimates of abundance parameters were virtually the same regardless of the TTDD employed.

Estimates of detection parameters varied by context.  In comparisons of mixture and non-mixture models within the same TTDD family (a la Sim 1), estimates were similar except when data derived from peaked mixture TTDDs.  In that case, mixture models yielded narrower credible intervals than non-mixture models --- except in the exponential case, where non-mixture exponential models were overly precise.  In comparisons of models across families (a la Sim 2), estimates across gamma, lognormal, and Weibull mixture inference models are always quite similar to one another.  Credible intervals from exponential mixture models are wider for peaked data but narrower for exponential and non-peaked data.

% The following plots of posteriors for abundance and detection parameters are just too large (and must be shrunk too small to make fit)
%\begin{figure}[h!]\centering
%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select6.pdf}
%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select2.pdf}
%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select13.pdf}
%\end{figure}

\subsubsection{Model fit}\label{sec:modelfit}

Across all simulations, posterior predictive checks of by-interval counts provided only limited ability to identify poor model fits.  For peaked and exponential mixture datasets (Sims 1 \& 3), posterior predictive checks clearly showed lack of fit for non-mixture inference models (Figure \ref{postpredmix}).  For gamma, lognormal, and Weibull mixture datasets fit with an exponential inference model, second-interval posterior predictive values near to one often pointed a model misfit (Sims 2 \& 3).  These are all cases where true data distribution has one more mode than the inference model can accommodate; in these cases, the second interval is hardest to fit, because of the mixture formulation of the models.  For all other models, posterior predictive checks did not consistently indicate lack of fit.

\adam{I'm guessing I should not include the following: (1) my posterior predictive pvalues are not independent; (2) for instance, for the first interval, they stay very close to 0.5-0.6 --- the models are effective at getting the first interval right; (3) likewise, a post. p-value $>$ 0.8 during the 2nd interval likely shows model misfit; all of which raises the possibility that (4) multiple simulation under the null could provide a reference distribution of posterior p-values.}

\begin{figure}[h!]\centering
\includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{sri}^{(rep)} > p_{sri}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim 1).  Trendlines show mean posterior predictive p-values.  Dashed lines show central 95\% intervals.  P-values near 0 or 1 potentially indicate poor model fit.}
\end{figure}

DIC showed a limited ability to select among models with different TTDDs.  For peaked and exponential mixture datasets (Sims 1 \& 3), DIC regularly favored a mixture inference model over a non-mixture model (differences were always greater than 12).  Also, when peaked gamma or lognormal mixtures were the true data model (Sim 2), DIC signalled against fitting an exponential mixture model (differences greater than 5).  These are the same scenarios where posterior predictive checks highlighted model misfit.

DIC comparisons of TTDD family in the single replicate from Sim 3 were variable.  For peaked mixture datasets, correct models were favored over mixture exponential inference models by at least 10 in gamma and lognormal cases and by 4 in the Weibull case.  For the exponential mixture datasets, the exponential inference models was favored by at least 6 over all other models.  For non-peaked mixture datasets, models were generally within 4 units of one another --- with the exception that the exponential mixture model was favored by 8 units for the lognormal mixture dataset.  Considering that exponential models were biased for detection in non-peaked models (Figure \ref{pdet_cater_family}), we have little confidence in our DIC results for non-peaked models.  DIC performance in Sim 3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects, but also because DIC is calculated according to a different method.

\section{Ovenbird analysis}

We fit mixture and non-mixture models from each family of TTDD to the Ovenbird dataset.  Posterior predictive check p-values by interval (Table \ref{tbl:ovencounts}) show a clear lack of fit for the straight exponential model over the first three intervals.  Two features of the data proved difficult to fit for all of the models: an increase in counts during the fourth interval, and a decrease during the final interval.  Mixture models were better able to accommodate these features than non-mixture models.

\begin{table}[ht]
\centering
\begin{tabular}{lccccccccc}
  \hline
Minutes & 0-2 & 2-3 & 3-4 & 4-5 & 5-6 & 6-7 & 7-8 & 8-9 & 9-10 \\ 
Count & 596 & 69 & 62 & 68 & 43 & 33 & 35 & 27 & 14 \\ 
\hline
\multicolumn{10}{l}{Posterior predictive check $p$-values by model:}\\
\hline
Expo & 0.000 & 1.000 & 1.000 & 0.347 & 0.590 & 0.372 & 0.007 & 0.006 & 0.190 \\ 
  Gamma & 0.582 & 0.820 & 0.394 & 0.006 & 0.362 & 0.637 & 0.241 & 0.519 & 0.977 \\ 
  LogNormal & 0.571 & 0.943 & 0.471 & 0.005 & 0.322 & 0.554 & 0.165 & 0.407 & 0.974 \\ 
  Weibull & 0.582 & 0.862 & 0.401 & 0.005 & 0.359 & 0.616 & 0.221 & 0.509 & 0.982 \\ 
  Exponential Mix & 0.533 & 0.804 & 0.577 & 0.021 & 0.501 & 0.668 & 0.190 & 0.338 & 0.918 \\ 
  Gamma Mix & 0.593 & 0.720 & 0.431 & 0.015 & 0.439 & 0.670 & 0.246 & 0.476 & 0.947 \\ 
  LogNormal Mix & 0.576 & 0.806 & 0.508 & 0.016 & 0.421 & 0.627 & 0.197 & 0.413 & 0.951 \\ 
  Weibull Mix & 0.596 & 0.734 & 0.411 & 0.013 & 0.421 & 0.663 & 0.252 & 0.492 & 0.958 \\ 
   \hline
\end{tabular}
\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive check p-values by observation interval for each model across all surveys.}
\end{table}

Ovenbird results reflect several patterns that were observed in the simulation studies for both exponential and non-peaked data.  Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $p^{(det)}$, and the log10 number of Ovenbirds uncounted.  Shape parameter estimates from mixture models favor non-peaked distributions --- i.e., detection rates are higher earlier in the observation period --- however, in mixture gamma and Weibull models, an exponential TTDD model ($\alpha=1$ or $k=1$) is well within 95\% credible intervals.  These pattern is similar to our simulations involving non-peaked mixture datasets, especially exponential and lognormal.

Abundance parameter estimates are virtually the same for all TTDD models.  Detection parameter estimates are similar by mixture type across gamma, Weibull, and lognormal TTDD models.  Credible intervals from exponential mixture models are more precise than for non-exponential mixture models.  Within mixture type, estimates of detection probability across all surveys, $p^{(det)}$, are lower and much more uncertain for gamma, Weibull, and lognormal models than for exponential.

DIC calculations prefer the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively), and these differences closely mirror those from the exponential mixture dataset in Sim 3.  However, Sim 3 demonstrated that DIC for non-peaked mixture data may not be reliable; indeed, for the simulated non-peak lognormal mixture dataset, the exponential mixture TTDD was favored by over a 4-unit difference; yet, its posterior 95\% credible interval for detection probability did not contained the true value, while it was within a central 60\% credible interval for all three other mixture models.  In short, we DIC may not be trustworthy in this situation.

Credible intervals for two of the abundance parameters do not contain zero, thereby suggesting notable effects.  Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% it's non-logged level.  Credible intervals for Site Age indicate that each decade of age increases abundance from 1.5-13\%.

\begin{figure}[h!]
\includegraphics[width=0.98\textwidth]{OVEN/oven_sum/OVEN_posteriors.pdf}
\caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to the Ovenbird data.  Abundance and detection parameters are identified by (A) and (D), respectively.  Gamma is the mixing parameter.  Alpha, k, and sigma\_det are shape parameters for gamma, Weibull, and lognormal distributions, respectively.  p\_det and log10(Uncounted) are the detection probabilty and the number of birds present but not counted across all surveys.  Black bars are 95\% credible intervals, orange bars are 50\% credible intervals, and black dots are posterior medians.}
\end{figure}

\section{Discussion}

Historically, analysis of removal-sampled data assumes constant detection rates throughout the sampling period.  This assumption provides an intuitive null model and is analytically tractable.  However, although analysis of removal models has grown more sophisticated in recent years, the constant-detection assumption nonetheless continues unquestioned.  In our study, we have formulated a model for non-constant detection patterns using a time-to-event approach within a hierarchical N-mixture framework.  Our results demonstrate that, for better or worse, the assumption of constant detection is a very strong assumption.  When it is correct, it is the most accurate and precise of TTDD models, but when detection rates are not constant, it results in significant bias.  Models with non-constant TTDDs, on the other hand, can adapt to a wider variety of detection patterns.

We modeled two varieties of heterogeneous detection in conjunction with one other: (i) time-varying detection rates, as modeled by non-exponential TTDDs, and (ii) detection heterogeneity among subgroups of individuals, as modeled by mixture TTDDs.  The exponential TTDD, even when modeled with a mixture component, results in biased estimates of detection when detection actually varies with time.  Gamma, Weibull, and lognormal TTDDs, with their extra parameter, are more accurate in handling non-constant detection data, though they are less precise when estimating data with constant detection.

The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Etterson2009, Reidy2011}\adam{Alldredge, Farnsworth, Pledger, Reidy verified}.  The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.

In a limited set of situations, posterior predictive checks and DIC can identify when non-mixture or exponential TTDDs inadequately describe the marginal pattern of detections over time.  But in many situations, neither tool points a clear way to select the correct family of TTDD.  Given the bias that can result from assuming constant detection or omitting a mixture component, conservatism dictates that non-constant TTDDs with a mixture should be preferred.  Among the TTDDs we considered, the mixture gamma TTDD appears to provide the best combination of accuracy and precision for non-peaked data.  For peaked data, gamma and Weibull models seem about equally as good.  MCMC sampling from gamma models required from 5-10 the run time as lognormal and Weibull models.

If the estimation of effect parameters is the primary interest, then our results suggest that the exact choice of TTDD may not be important.  Abundance effect estimates are similar regardless of the chosen TTDD, while detection effect estimates are similar for all mixture non-exponential TTDDs.  These finding may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.

Because our focus is detection, we assumed Poisson-distributed abundance for simplicity.  Abundance distributions that have been used to account for overdispersion include negative binomial, zero-inflated Poisson, and Poisson with survey-level random effects.  The pattern of counts in our Ovenbird data actually reflect underdispersion, so a Conway-Maxwell Poisson distribution, which can model both overdispersion and underdispersion, may be more appropriate \citep{Wu2015}\adam{apparently, Sellers et al. (in Wu) has a comprehensive overview of the CMP.}

% Overview: Wu15
% ZIP: Etterson, Solymos12
% NB: Royle04(?)

At present, Stan does not sample discrete parameters, so the ability to marginalize out the discrete-valued latent variables $N_{sr}$, as we did, can greatly facilitate MCMC sampling.

% \adam{I have contemplated on and off including a discussion of broad priors and the limit behavior as $\varphi \to 0$, but I think it's not really on topic.  Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}

% All estimates are based on extrapolation of the observed detections\\
% -- a) explore the tail more?  I think that'd require a lot of data\\
% -- b) requires extended observation periods or full detection data (Alldredge?)\\
% same goes for sample size

% Really, I have totally ignored the discussion of the true value of p, which I have tended to simulate in a fairly high range

To improve our choice of TTDD and/or its parameters, it makes sense to obtain more data in one of a few ways.  One approach is to collect complete detection history records, not just first time-to-detection observations \citep{Alldredge2007}.  This may not be feasible in studies like MNFB, where many species are observed.  Another approach is to conduct a longer survey that better assesses the tail distribution of the TTDD; however, the longer the survey is, the greater the risk that individuals enter/depart the study area or are double-counted, which violates removing sampling assumptions \citep{LeeMarsden2008, Reidy2011}.  A third idea is to obtain more precise time-to-detection data, even exact time-to-detection data.  In some early explorations of our model considering just an exponential mixture TTDD, we found that credible intervals for Ovenbird detection probability were 20-25\% narrower using 9-interval surveys compared to more traditional 3-interval surveys (0-3, 3-5, and 5-10 minutes).  For all of the above, a small pilot study may be a practical starting point.

With the increasing use of microphone arrays for automated detection \adam{refs}, it is tempting to think TTDD models may be valuable for microphone-collected data.  This should be the case when variations in detection result solely from organismal behaviors, such as bout singing, or if we wish to model individual random effects in the detection model.  However, some explanations for non-constant detection are explicitly related to the observation process --- either effects of the observer on animal behavior or variations in observer effort caused by survey methodology.  The assumption of constant detection seems more plausible in studies where observers are absent, though detection heterogeneity may still exist across subgroups of the population.

Application of TTDD modeling within the N-mixture framework should be straight-forward for other time-to-detection datasets, such as full detection-history and double-observer.  The key difference will be in the data model, in that responses will take a different form than multinomially distributed first times to detection.  But the underlying abundance and detection models remain the same.

The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}.  In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.  I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  See also McCallum 2005.}.  In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.

The time-to-detection model as presented here in our study does not incorporate distance.  Consequently, our definition of a TTDD necessarily represents an averaging across distance classes and a consequent loss of information if distance data are available.  We can modify our model to account for distance.  To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify Equation \ref{eq:pdet} to reflect the detection distance to individual $q$: $p_{srq}^{(det)} = g(d_q) F_T^A(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.

However, our analysis makes possible a different strategy that is more consistent with a time-to-event conceptualization of the problem.  Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{srq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{sr}^A(t)$ and the perceptibility function $g(d/`8-_q)$.  So, $h_{srq}(t) = g(d_q) h_{sr}^A(t)$.  From standard survival analysis results, $p_{srq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{sr}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) \right]^{g(d_q)}$.

In short, the time-to-event conceptualization frames perceptibility as a rate-of-availability multiplier rather than as a probability-of-availability multiplier.  In a context where rates of availability are allowed to vary from individual to individual, this makes an intuitive sense --- a bird that sings twenty times during the observation period should have its overall detection probability discounted much less than a bird at the same distance which only sings once.  This is an area of ongoing exploration.

% Skeptics of removal sampling: Johnson08, EffordDawson09, Reidy11

%\subsection{Lit Notes on Distance and Poisson Abundance}
%--- Borchers is all about this\\
%--- Farnsworth2005, EffordDawson2009, Amundson2014\\
%----- particularly noticeable with finite mixture models\\
%- Oedekoven2013 is all about covariates on distance sampling curve\\
%- necessity of using distance, too, is backed by Solymos2013, (Buckland et al 2001, 2004), Farnsworth2005, EffordDawson2009...\\
%
%Poisson assumption (do I want to discuss underdispersion?\\
%--- Denes et al 2015 discuss how Poisson assumption leads to bias in the absence of detection heterog (ref to Solymos et al 2012) or when inflated zeros are present (Joseph et al 2009) [with the obvious remedies being: NegBin and ZIP, respectively]... but is that negated by our underdispersion?\\
%
%`Unmodeled individual heterogeneity causes population size to be underestimated using capture-recapture and removal methods' (from Efford and Dawson, who cite Otis et al. 1978)\\

% Ideas for the future
% (i) Distance-removal time-to-event model
% (ii) Fit more species from this dataset over full survey duration; maybe some model selection (even Bayesian)
% (iii) Take a stab at modeling habitat
% (iv) Model for non-independent singing?
% (v) Model for density-dependent singing? (only valid, I think, if we have lots of underdispersion from(ii))
% (vi) I know I've  been warned off it, but automated detection intrigues me
% (vii) Further testing of existing model: sensitivity to priors, performance for varying detectability levels, sample size, # intervals
% (viii) Incorporate spatial random effects.  This work has already been done, just not for TTDD models, but the real spatial dependency should be in abundance, not detection
% (ix) Comparison of methods -- just looking at my lit review, there are gadzooks strategies that people have proposed.  Combining/comparing them seems maybe informative?  To what degree has this already been done?

\bibliography{masterbib}
\bibliographystyle{biom}

\end{document}

### 2016-03-02 Draft.tex
Comments to self inserted.

@@
 First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.
-Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Refs}.
+Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Amundson2014}.
 Such an approach is not possible when detection rates are not constant.
+\adam{This statement is false.
+It has been done, but the form of the non-constancy must be proscribed.
+See Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples}.
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
-Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection models can only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.
+Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.
 
@@
 But the underlying abundance and detection models remain the same.
+\adam{Time-heterogeneous models have been done for mark-recapture (Farcomeni and Scacciatelli 2013, who reference Hwang and Chao 2002).}
 
-The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}.
+The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.
 In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.

### 2016-03-03 Draft.tex
Routine additions.

@@
 
-However, the assumption of a constant detection rate throughout the observation period is not always justified.
+However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}.
 Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.
 The act of observation may affect animal behavior.
-Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002}.
+Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.
 Survey methodology itself can affect detection rates.
-In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time.
+In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.
 Observers may also become distracted as the survey progresses \citep{Johnson2008}.
@@
 Such an approach is not possible when detection rates are not constant.
-\adam{This statement is false.
-It has been done, but the form of the non-constancy must be proscribed.
-See Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples}.
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
 Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.
+
+\adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.
+In the fisheries world, models of time-varying catchability have been implemented -- see Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples.
+Also, Alldredge 2007 implements a complete-history of detection model (technically mark-recapture, not removal) that estimates interval-specific detection probabilities (complete with covariates and detection heterogeneity).
+However, it cannot be applied in a removal context, because removal data contains only first times to detection, and the Alldredge model requires subsequent observations -- removal models compensate by specifying a TTDD, thus enabling us to extrapolate beyond the observation period.}
 
@@
 
-The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Etterson2009, Reidy2011}\adam{Alldredge, Farnsworth, Pledger, Reidy verified}.
+The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Reidy2011}.
 The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.

### 2016-03-17 Draft.tex
Use \begin and \end abstract
Replace left quotes with `

@@
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
+\newcommand{\jarad}[1]{{\color{red} #1}}
 \newenvironment{indpar}[1]%
@@
 
-\section{Abstract}
+\begin{abstract}
 
@@
 Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.
-Under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals.
+\jarad{in the abstract focus on what we did} Under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals.
 We instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.
@@
 
+\end{abstract}
+
 \section{Introduction}\label{sec:intro}
 
-\adam{WRITE A SPELLBINDING OPENING PARAGRAPH}
+\adam{WRITE A SPELLBINDING OPENING PARAGRAPH}\jarad{no need to aim so high}
 
@@
 
-Increasing hazard functions result in a probability distribution function (pdf) for $T$ that has a mode away from zero; we term these TTDDs 'peaked'.
-We contrast these to decreasing hazard functions with a mode at zero, which we term 'non-peaked'.
+Increasing hazard functions result in a probability distribution function (pdf) for $T$ that has a mode away from zero; we term these TTDDs `peaked'.
+We contrast these to decreasing hazard functions with a mode at zero, which we term `non-peaked'.
 For our purposes, the exponential TTDD is neither peaked nor non-peaked.
@@
 Upper plots show non-mixture datasets; lower plots show mixture datasets.
-'X' marks the expected marginal probability of detection based on true parameter values.}
+`X' marks the expected marginal probability of detection based on true parameter values.}
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
@@
 Each plot presents one simulated dataset.
-'X' marks the expected marginal probability of detection based on true parameter values.}
+`X' marks the expected marginal probability of detection based on true parameter values.}
 \end{figure}

### 2016-03-17 Draft.tex
Add commments about organization of the manuscript

@@
      {\end{list}}
+
+\newcommand{\vn}{\textbf{n}}
 
@@
 
+\jarad{The introduction needs to end earlier.
+It should be about 1-2 pages covering what has previously been done and outline our approach.
+So this should probably start a new section and before this there should be an outline of our manuscript.
+}
+
 \subsection{TTDDs as behavior and methodology}
@@
 \includegraphics[width=0.68\textwidth]{Hazards.pdf}
-\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
+\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.
+\jarad{Are these setup to have the same mean and variance or anything like that?}
 \end{figure}
@@
 Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$.
-We present the model for times to detection censored into $i=1,\dotso,I$ intervals, but it is straight-forward to simplify the model for exact time-to-detection data.
-Our data therefore take the form of counts indexed by site, time period, and interval: $n_{sri}$.
-For any given survey (uniquely defined by site and time period), we write the vector of counts across all $I$ intervals as $\textbf{n}_{sr}$.
+\jarad{So time periods are R because we already use t for intervals?
+Argh!}  We present the model for times to detection censored into $i=1,\dotso,I$ intervals \jarad{Why don't we present the exact time-to-detection model first and then censoring it?}, but it is straight-forward to simplify the model for exact time-to-detection data.
+We denote the number of individual observed at site $s$, time $r$, and in interval $i$ as $n_{sri}$ and define the total number of individuals observed during survey as $\vn_{sr} = (\sum_{i=1}^I n_{sri})n_{sr1},\ldots,n_{srI})$.
 
@@
 
+\jarad{We should try to avoid using subsubsections. Perhaps we should provide a data section and then a model section?}
+
 \subsubsection{Data Model}
@@
 The data model is then:
-
 \begin{align}\label{eq:datamodel}
-\textbf{n}_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
+\vn_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
 n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
 \end{align}
+
+\jarad{Pedogogically, I would introduce the binomial model first and then the multinomial model. Or we could introduce a Multinomial with parameter $N_{sr}$ and an additional bin for non-detection.}
 
@@
 \subsubsection{Detection Model}\label{sec:detectionmodel}
+
+\jarad{I think you should introduce just the exponential model first with no covariates.
+Thus, relative to what has been done before, you are only adding in the time-to-detection distribution.
+Later you can introduce other TTDDs and covariates.}
+
 We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by means of the partitioned cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$. 
+\jarad{Do we want to introduce explanatory variables already? Perhaps it would be smoother to wait until later to introduce them.}
 Here, $T$ is a random variable for the first time at which an individual is detected, $\textbf{X}_{sr}^D$ and $\textbf{Z}_{sr}^D$ are detection-related covariates and random effect factor levels to be used in a mixed effects linear regression, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.
@@
 The first four TTDDs we employ are:
+\jarad{You don't need an enumerate here. Would it be easier to introduce the exponential first and then the other 3 altogether later?}
 \begin{enumerate}
@@
 \end{alignat}
+\jarad{Are these necessary?}
 where $\boldsymbol{\beta}^D$ is a vector of fixed effects, and $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.
@@
 For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.
-Conditioned on an individual being detected, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.
+Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.
 In summary, the detection model is:
@@
 (\ref{eq:datamodel}).
-Instead of modeling the interval-specific counts $\textbf{n}_{sr}$, we model the times to detection directly.
+Instead of modeling the interval-specific counts $\vn_{sr}$, we model the times to detection directly.
 Defining $\textbf{T}_{sr}$ as a vector of all observed detection times for a given survey, and continuing to treat all detections as independent, we write
@@
 \subsubsection{Abundance Model}
+
+\jarad{Again, I would introduce this without covariates or random effects first.
+
 For simplicity, we model abundance at each site during each year $N_{sr}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{sr}$, though any non-negative integer-valued distribution may be used.
@@
 
+\jarad{Put a discussion of priors in a new section.
+You should introduce the general form of the priors with specific values and specify the specific values in the simulation study and real data analysis sections.}
+
 Since we use this hierarchical model in a Bayesian context, the complete model formulation requires priors for $\text{Intercept}^D, \boldsymbol{\beta}^D, \sigma_{D}^2, \boldsymbol{\theta}, \text{Intercept}^A, \boldsymbol{\beta}^A, \text{ and } \sigma_{A}^2$, such as we specify in Section \ref{sec:priors}.
 
+\jarad{You need to include a section about the MCMC, e.g. Stan/JAGS how to monitor convergence, etc.}
+
 \subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
+
+\jarad{This should not be a model (or methods section).
+I think you should put a Data section earlier that introduces the data collection methods.
+Then the rest of this section should go much later, after the simulation study, in a real data analysis section.
+}
 
@@
 All priors are independent.
+\jarad{This takes up too much space.}
 \begin{align*}
@@
 \end{align*}
+
+\jarad{New section for simulation studies with the results.}
 

### 2016-03-17 Draft.tex
Ensure the file compiles

@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.68\textwidth]{Hazards.pdf}
-\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.
-\jarad{Are these setup to have the same mean and variance or anything like that?}
+\includegraphics[width=0.68\textwidth]{Hazards}
+\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.
+\jarad{Are these setup to have the same mean and variance or anything like that?}}
+\label{HazardFxns}
 \end{figure}
@@
 
-\jarad{Again, I would introduce this without covariates or random effects first.
+\jarad{Again, I would introduce this without covariates or random effects first.}
 

### 2016-03-22 Draft.tex
Add more reorganization comments
Implementing most of Philip's comments from 3/17.

@@
 
-Removal sampling is a survey methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
+Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
 Assuming that the probability of capturing a not-yet-trapped individual is the same for all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time \citep{Moran1951, Zippin1958, Seber1982}.
@@
 The act of observation may affect animal behavior.
-Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues from the studied species \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.
+Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.
 Survey methodology itself can affect detection rates.
-In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.
 Observers may also become distracted as the survey progresses \citep{Johnson2008}.
 And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.
+In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.
 
-In this study, we model scenarios where detection rates are not constant.
+In this study, we develop a model for scenarios where detection rates are not constant.
 We model time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function $F_T(t)$.
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).
-The instantaneous detection rate, a.k.a.
-hazard function $h(t)$, can be found from the TTDD and vice versa, by observing $h(t) = f_T(t) / (1-F_T(t))$.
+The instantaneous detection rate or hazard function $h(t)$ can be found from the TTDD as $h(t) = f_T(t) / (1-F_T(t))$.
 
-Our model is different from most survival analysis in that the number of individuals $N$ present at a survey is unknown.
-To address this data truncation, we embed the TTDD in a hierarchical framework for multinomial counts sometimes termed an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.
-An N-mixture model treats abundance as a latent variable modeled according to a Poisson or other discrete distribution, and it models detection probabilities independently.
-Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014}.
+Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown.
+We embed the TTDD in a hierarchical framework for multinomial counts using an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.
 For our purposes, the N-mixture framework provides three clear benefits.
@@
 Third, for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
+The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.
+Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014}.
 
-The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.
-First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.
-Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Amundson2014}.
+Framing a model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.
+First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}, whereas most existing studies instead construct regression models on the logit of the equal-interval detection probability \adam{Amundson2014 and others?}.
 Such an approach is not possible when detection rates are not constant.
@@
 
-A TTDD expresses the range of detection patterns possible in an analysis.
-The most basic is an exponential TTDD, which imposes a constant detection rate.
-If species behavior and/or observation methodology may cause a systematic non-constant detection regime, then we should choose a TTDD accordingly.
-For example, we may wish to model an observer effect where: (i) the observer's arrival either suppresses or stimulates detectable cues from organisms at a site, but (ii) as organisms acclimate to the observer, cues gradually return to an undisturbed baseline of constant detection.
-As illustrated by the hazard functions in Figure \ref{HazardFxns}, a gamma TTDD would be appropriate for this scenario.
+A TTDD defines the range of detection patterns possible in an analysis.
+The most basic is an exponential TTDD, which imposes a constant first detection rate.
+Choosing another TTDD can allow for a systematic non-constant detection regime.
+For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate (Figure \ref{HazardFxns}).
 
@@
 
-One issue related to non-constant detection is detection heterogeneity.
-Heterogeneity across population subgroups can be modeled using a mixture of detection distributions \citep{Pledger2000}\adam{longer history here...
-do I even need a reference?
-According to NorrisPollock1996, this goes all the way back to Burnham1972}.
-In particular, \citet{Farnsworth2002} introduce a partial finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
+We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.
+We reserve `detection heterogeneity' for differences across population subgroups.
+It is often modeled using a mixture of detection distributions.
+\citet{Farnsworth2002} introduce a partial finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
 Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.
-A similar marginal detection pattern can be achieved by modeling non-constant detection using a non-peaked TTDD.
-Detectability models constructed to address heterogeneity may provide reasonable estimates when detection rates change for behavioral reasons and vice versa \citep{Mantyniemi2005}.
+We use `non-constant detection' when detection rates for an organism change with time.
+Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
 
@@
 
-We construct a hierarchical model of removal-sampled point-counts patterned after an N-mixture framework \citep{Wyatt2002, Royle2004NMixture}.
-We treat the hierarchical model in three pieces: (a) a multinomial data model, (b) a detection model based upon the partitioned cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a multinomial data model, (b) a detection model based upon the partitioned cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.
 Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
@@
 where for each survey, $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$ is the number of individuals counted across all intervals, $N_{sr}$ is the total number of individuals present (counted and uncounted), and $p_{sr}^{(det)}$ is the probability that an individual is detected at any time during the observation period.
-The number of uncounted individuals at any survey can be written: $n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)}$.
+The number of uncounted individuals at any survey is then: $n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)}$.
 
@@
 
-We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by means of the partitioned cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$. 
+We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by partitioning the cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$. 
 \jarad{Do we want to introduce explanatory variables already? Perhaps it would be smoother to wait until later to introduce them.}
@@
 
+\subsubsection{Priors}\label{sec:priors}
+Our goal in selecting priors is to be non-informative within a reasonable range of values.
+Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.
+Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.
+Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).
+Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.
+For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.
+All priors are independent.
+\jarad{This takes up too much space.}
+\begin{align*}
+\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
+\text{Intercept}^A &\sim N(1, 1)\\
+\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
+\boldsymbol{\beta}^A &\sim N(0, 1)\\
+\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
+\gamma &\sim \text{Beta}(1,1)
+\end{align*}
+
 \subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
@@
 
-We fit eight models to the Ovenbird data, once using each of the time-to-detection distributions described in Section \ref{sec:detectionmodel}.
+We fit eight models to the Ovenbird data representing each of the time-to-detection distributions described in Section \ref{sec:detectionmodel}.
 
-\subsubsection{Priors}\label{sec:priors}
-Our goal in selecting priors is to be non-informative within a reasonable range of values.
-Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.
-Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.
-Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).
-Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.
-For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.
-All priors are independent.
-\jarad{This takes up too much space.}
-\begin{align*}
-\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
-\text{Intercept}^A &\sim N(1, 1)\\
-\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
-\boldsymbol{\beta}^A &\sim N(0, 1)\\
-\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
-\gamma &\sim \text{Beta}(1,1)
-\end{align*}
+\subsection{Model fitting}
+
+We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.
+
+For each intercept-only simulation model, we instantiated a single chain running 60,000 iterations.
+For the Ovenbird analysis and covariate simulations, we ran chains for 250,000-375,000 iterations.
+We discarded half of the iterations as warmup and then thinned by 1 in 10.
+We reran models if the effective sample size for any parameter was below 1000.
+We monitored convergence of the MCMC chains using Geweke diagnostics.
+For most models, we accepted the Stan defaults for generating dispersed initial values.
+However, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from the posterior means of preliminary short-run models (Ovenbird analysis, incorrect-family simulations) or from true parameter values (correct-family simulations).
 
@@
 
-We conduct three simulation studies, the first two of which implement intercept-only models.
-The objectives for the studies are to assess: (i) the importance of modeling detection heterogeneity via a partial finite-mixture component, (ii) the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to a dataset simulated from a Weibull TTDD), and (iii) model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.
+We conduct three simulation studies to assess: (Sim 1) the importance of modeling detection heterogeneity via a partial finite-mixture component, (Sim 2) the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD), and (Sim 3) model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.
 Because model behavior may differ between datasets with peaked and non-peaked TTDDs, we simulate both kinds of dataset for gamma, Weibull, and lognormal models.
@@
 
-\subsection{Model fitting}
+\subsection{Model assessment, inference, and between-model comparison}
 
-We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.
-
-For each intercept-only simulation model, we instantiated a single chain running 60,000 iterations.
-For the Ovenbird analysis and covariate simulations, we ran chains for 250,000-375,000 iterations.
-We discarded half of the iterations as warmup.
-We reran models if the effective sample size of for any root parameter was below 1000.
-We monitored convergence of the MCMC chains using Geweke diagnostics.
-For most models, we accepted the Stan defaults for generating dispersed initial values.
-However, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from the posterior means of preliminary short-run models (Ovenbird analysis, incorrect-family simulations) or from true parameter values (correct-family simulations).
-
-\subsection{Model assessment, inference, and between-model comparison}
+\jarad{This is also part of methods. Or perhaps make it part of evaluation of simulation studies.}
 
@@
 The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.
+We simulate replicate datasets and calculate the check statistic at each iteration of our MCMC samples.
 Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
-We simulate replicate datasets at each iteration of our retained MCMC samples.
 
@@
 \begin{table}[ht]\centering\small
-\begin{tabular}{ll|l|cccc|cccc||c}
- \multicolumn{3}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
- \multicolumn{3}{c|}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+\begin{tabular}{l|l|l|l|cccc|cccc||c}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
+ \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
   \hline
   \hline
- \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\
- & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
- & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
-  \cline{3-12}
- & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
-  \cline{3-12}
- & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
- & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
- & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
-  \hline
-\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\
- & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
- & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
-  \cline{3-12}
- & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
-  \cline{3-12}
- & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
- & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
- & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\
+ & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
+ & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
+  \cline{3-13}
+ & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
+  \cline{3-13}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
+ & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
+ & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
+  \cline{2-13}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\
+ & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
+ & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
+  \cline{3-13}
+ & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
+  \cline{3-13}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
+ & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
+ & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
    \hline
@@
  \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
- \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff \\ 
+ \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
   \hline
@@
  \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
- \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff & Med. $p$ & Q($p$) & 50\% & 90\% & DIC Diff \\ 
+ \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
   \hline
@@
 
-In many situations, individual parameter estimates are of interest.  For every dataset we simulated, estimates of abundance parameters were virtually the same regardless of the TTDD employed.
-
+For every dataset we simulated, estimates of abundance covariate coefficients were virtually the same regardless of the TTDD employed.
 Estimates of detection parameters varied by context.
@@
 Two features of the data proved difficult to fit for all of the models: an increase in counts during the fourth interval, and a decrease during the final interval.
-Mixture models were better able to accommodate these features than non-mixture models.
+Mixture models were better able to fit these features than non-mixture models.
 
@@
 
-Abundance parameter estimates are virtually the same for all TTDD models.
+Abundance covariate coefficient estimates are virtually the same for all TTDD models.
 Detection parameter estimates are similar by mixture type across gamma, Weibull, and lognormal TTDD models.
@@
 However, Sim 3 demonstrated that DIC for non-peaked mixture data may not be reliable; indeed, for the simulated non-peak lognormal mixture dataset, the exponential mixture TTDD was favored by over a 4-unit difference; yet, its posterior 95\% credible interval for detection probability did not contained the true value, while it was within a central 60\% credible interval for all three other mixture models.
-In short, we DIC may not be trustworthy in this situation.
+In short, DIC may not be trustworthy in this situation.
 
 Credible intervals for two of the abundance parameters do not contain zero, thereby suggesting notable effects.
-Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% it's non-logged level.
+Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% of the abundance for unlogged sites.
 Credible intervals for Site Age indicate that each decade of age increases abundance from 1.5-13\%.
@@
 This may not be feasible in studies like MNFB, where many species are observed.
-Another approach is to conduct a longer survey that better assesses the tail distribution of the TTDD; however, the longer the survey is, the greater the risk that individuals enter/depart the study area or are double-counted, which violates removing sampling assumptions \citep{LeeMarsden2008, Reidy2011}.
+Another approach is to conduct a longer survey that better assesses the tail distribution of the TTDD; however, the longer the survey is, the greater the risk that individuals enter/depart the study area or are double-counted, which violates removal sampling assumptions \citep{LeeMarsden2008, Reidy2011}.
 A third idea is to obtain more precise time-to-detection data, even exact time-to-detection data.

### 2016-03-23 Draft.tex
Implementing some of Jarad's comments.

@@
 \usepackage{color}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{caption}
@@
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-\newcommand{\jarad}[1]{{\color{red} #1}}
+\newcommand{\jarad}[1]{{\color{Orange} #1}}
 \newenvironment{indpar}[1]%
@@
 Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.
-The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate.
-However, the assumption of constant detection rates during surveys is often not justified.
-Detection rates can be influenced by organismal behaviors, including responses to observer presence and heterogeneity among population subgroups, and also by survey methods that affect observer effort.
+The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption is often not justified.
+Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.
 Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.
-\jarad{in the abstract focus on what we did} Under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals.
-We instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.
+Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.
 Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.
@@
 
-Increasing hazard functions result in a probability distribution function (pdf) for $T$ that has a mode away from zero; we term these TTDDs `peaked'.
+Increasing hazard functions result in a pdf for $T$ that has a mode away from zero; we term these TTDDs `peaked'.
 We contrast these to decreasing hazard functions with a mode at zero, which we term `non-peaked'.
@@
 \includegraphics[width=0.68\textwidth]{Hazards}
-\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.
-\jarad{Are these setup to have the same mean and variance or anything like that?}}
+\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
 \label{HazardFxns}
@@
 
-Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$.
-\jarad{So time periods are R because we already use t for intervals?
-Argh!}  We present the model for times to detection censored into $i=1,\dotso,I$ intervals \jarad{Why don't we present the exact time-to-detection model first and then censoring it?}, but it is straight-forward to simplify the model for exact time-to-detection data.
-We denote the number of individual observed at site $s$, time $r$, and in interval $i$ as $n_{sri}$ and define the total number of individuals observed during survey as $\vn_{sr} = (\sum_{i=1}^I n_{sri})n_{sr1},\ldots,n_{srI})$.
+Our model is formulated for single-visit removal-sampled point-count surveys at a set of sites $s=1,\dotso,S$ repeated across distinct sampling periods $r=1,\dotso,R$.
+We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.
+For exact data, we use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey, each with a time to detection $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.
+For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.
+The number of animals counted during any interval is $n_{sri}$, the vector of counts for a survey is written $\vn_{sr}$, and $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$.
 
@@
 
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a multinomial data model, (b) a detection model based upon the partitioned cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon the cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.
 Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
 
-\jarad{We should try to avoid using subsubsections. Perhaps we should provide a data section and then a model section?}
+\subsection{Data Model}
+We assume that counts from different surveys are conditionally independent.
+We also assume that the times to detection for all individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.
+The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.
+If the total number of animals present at the survey is $N_{sr}$, then the data model is:
+\begin{align}\label{eq:datamodel}
+n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)\\
+\label{eq:Tsrq}T_{srq} &\sim f_{T|det}(t|t_{srq}<C_I)
+\end{align}
+When detection times are censored, we replace Eq.
+\ref{eq:Tsrq} with $\vn_{sr} \sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)$, where $\textbf{p}_{sr}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{srq}<C_I)$ at the interval cutoff times.
+It follows that $\sum_{i=1}^I p_{sri} = 1$.
 
-\subsubsection{Data Model}
-We assume that counts from different surveys are conditionally independent.
-We also assume that the time-to-detection is independent, identically distributed for all individuals of a species at any given survey.
-Given that an organism is detected, we denote the probability it is first detected during interval $i$ as $p_{sri}$ such that $\sum_{i=1}^I p_{sri} = 1$.
-We write $\textbf{p}_{sr}$ to denote the vector of interval-specific detection probabilities at any given survey.
-The data model is then:
-\begin{align}\label{eq:datamodel}
-\vn_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
-n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
-\end{align}
-
-\jarad{Pedogogically, I would introduce the binomial model first and then the multinomial model. Or we could introduce a Multinomial with parameter $N_{sr}$ and an additional bin for non-detection.}
-
-where for each survey, $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$ is the number of individuals counted across all intervals, $N_{sr}$ is the total number of individuals present (counted and uncounted), and $p_{sr}^{(det)}$ is the probability that an individual is detected at any time during the observation period.
-The number of uncounted individuals at any survey is then: $n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)}$.
-
-\subsubsection{Detection Model}\label{sec:detectionmodel}
+\subsection{Detection Model}\label{sec:detectionmodel}
 
@@
 
-For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.
 Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.
@@
 
-\subsubsection{Abundance Model}
+\subsection{Abundance Model}
 
@@
 
-\subsubsection{Priors}\label{sec:priors}
+\subsection{Priors}\label{sec:priors}
 Our goal in selecting priors is to be non-informative within a reasonable range of values.
@@
 
-\subsubsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}
+\subsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}
 
@@
 
-\subsubsection{Constant vs. non-constant detection}\label{sec:family}
+\subsection{Constant vs. non-constant detection}\label{sec:family}
 
@@
 
-\subsubsection{Simulations with covariates and random effects}\label{sec:simfull}
+\subsection{Simulations with covariates and random effects}\label{sec:simfull}
 
@@
 
-\subsubsection{Model fit}\label{sec:modelfit}
+\subsection{Model fit}\label{sec:modelfit}
 

### 2016-03-23 Draft.tex
Insert line break at the end of each sentence.  (By find-and-replacing all period-double-spaces with an added line break... may not be 100% effective).

@@
 We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  
-For exact data, we use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey, each with a time to detection $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.
+We use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey.  
+For exact data, times to detection are $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.  
 For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  
@@
 
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon the cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon a TTDD, and (c) an abundance model based upon Poisson regression.
 Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
@@
 We assume that counts from different surveys are conditionally independent.  
-We also assume that the times to detection for all individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.
+We also assume that the times to detection for all individuals in any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.  
 The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.
-If the total number of animals present at the survey is $N_{sr}$, then the data model is:
+Given $N_{sr}$ total number animals present at a survey, the data model is:
 \begin{align}\label{eq:datamodel}
@@
 
-To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq.
-(\ref{eq:datamodel}).
+To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq. (\ref{eq:datamodel}).  
 Instead of modeling the interval-specific counts $\vn_{sr}$, we model the times to detection directly.  
@@
 \]
-The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.
+The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  
+Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.
 
@@
 \end{tabular}
-\caption{\label{tbl:sim1} Summary of mixture vs.
-non-mixture model fits (Sim 1).
+\caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim 1).  
 In all cases, the model family matches the dataset family.  
@@
 50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
-\adam{Thoughts on use of bold/italics to highlight some results?}}
+$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model. \adam{Thoughts on use of bold/italics to highlight some results?}}
 \end{table}
@@
 % \adam{I have contemplated on and off including a discussion of broad priors and the limit behavior as $\varphi \to 0$, but I think it's not really on topic.  
-Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}
+% Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}
 
@@
 In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.
-I'm  beginning to think this may have too-numerous antecedents to cite...
-in Reidy, the earliest reference is Marsh and Sinclair 1989.
+I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  
 See also McCallum 2005.}.  
@@
 % (vii) Further testing of existing model: sensitivity to priors, performance for varying detectability levels, sample size, # intervals
-% (viii) Incorporate spatial random effects.  This work has already been done, just not for TTDD models, but the real spatial dependency should be in abundance, not detection
+% (viii) Incorporate spatial random effects.  
+This work has already been done, just not for TTDD models, but the real spatial dependency should be in abundance, not detection
 % (ix) Comparison of methods -- just looking at my lit review, there are gadzooks strategies that people have proposed.  
-Combining/comparing them seems maybe informative?
-To what degree has this already been done?
+Combining/comparing them seems maybe informative?  To what degree has this already been done?
 

### 2016-03-27 Draft.tex
More model streamlining.  (1) Changed the {sr} subscripting to just {s}.  (2) Seriously shortening the model elucidation.

@@
 
-Our model is formulated for single-visit removal-sampled point-count surveys at a set of sites $s=1,\dotso,S$ repeated across distinct sampling periods $r=1,\dotso,R$.  
+Our model is formulated for single-visit removal-sampled point-count surveys $s=1,\dotso,S$.  
 We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  
-We use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey.  
-For exact data, times to detection are $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.  
+We use $n_{s}^{(obs)}$ to denote the number of individuals observed during a survey.  
+For exact data, times to detection are $t_{sq}$, where $q = 1,\dotso,n_{s}^{(obs)}$ indexes observed individuals at a survey.  
 For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  
-The number of animals counted during any interval is $n_{sri}$, the vector of counts for a survey is written $\vn_{sr}$, and $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$.
+The number of animals counted during any interval is $n_{si}$, the vector of counts for a survey is written $\vn_{s}$, and $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$.
 
@@
 
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon a TTDD, and (c) an abundance model based upon Poisson regression.
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and time-to-detection pieces, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued distribution.
 Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
@@
 We assume that counts from different surveys are conditionally independent.  
-We also assume that the times to detection for all individuals in any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.  
-The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.
-Given $N_{sr}$ total number animals present at a survey, the data model is:
+We define $T_{sq}$ as a random variable for the first time at which an individual is detected.
+We assume that detection times for individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$, which may depend on covariates.  
+The probability of detecting an animal is thus $p_{s}^{(det)} = Pr(t_{sq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) / F_T(C_I)$.
+Given $N_{s}$ total number animals present at a survey, the data model is:
 \begin{align}\label{eq:datamodel}
-n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)\\
-\label{eq:Tsrq}T_{srq} &\sim f_{T|det}(t|t_{srq}<C_I)
+n_{s}^{(obs)} &\sim \mbox{Binomial}\left(N_{s}, p_{s}^{(det)}\right)\\
+\label{eq:Tsrq}T_{sq} &\sim f_{T|det}(t|t_{sq}<C_I)
 \end{align}
 When detection times are censored, we replace Eq.
-\ref{eq:Tsrq} with $\vn_{sr} \sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)$, where $\textbf{p}_{sr}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{srq}<C_I)$ at the interval cutoff times.
-It follows that $\sum_{i=1}^I p_{sri} = 1$.
+\ref{eq:Tsrq} with $\vn_{s} \sim \mbox{Multinomial}\left(n_{s}^{(obs)}, \textbf{p}_{s}\right)$, where $\textbf{p}_{s}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{sq}<C_I)$ at the interval cutoff times.
+It follows that $\sum_{i=1}^I p_{si} = 1$.
 
@@
 
-\jarad{I think you should introduce just the exponential model first with no covariates.
-Thus, relative to what has been done before, you are only adding in the time-to-detection distribution.
-Later you can introduce other TTDDs and covariates.}
+The two main elements of the detection model for $T_{sq}$ are: (i) the distributional form and (ii) the expectation function.
+$T_{sq}$ should be based on non-negative continuous distributions featuring some variety of rate parameter than can be estimated via regression.  
+In this study, we adopt TTDDs for $T_{sq}$ using the following: Exponential($\varphi_{s}$), Gamma($\alpha, \varphi_{s}$), Weibull($k, \varphi_{s}$), and Lognormal($\mu_{s}, \sigma_{det}^2$).  
+Here, $\varphi_{s}$ and $\mu_{s}$ are survey-specific detection rate parameters, and we consider the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ to be constant across all surveys.
 
-We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by partitioning the cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$. 
-\jarad{Do we want to introduce explanatory variables already? Perhaps it would be smoother to wait until later to introduce them.}
-Here, $T$ is a random variable for the first time at which an individual is detected, $\textbf{X}_{sr}^D$ and $\textbf{Z}_{sr}^D$ are detection-related covariates and random effect factor levels to be used in a mixed effects linear regression, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.
-In this study, we consider eight different forms for the TTDD.  
-The first four TTDDs we employ are:
-\jarad{You don't need an enumerate here. Would it be easier to introduce the exponential first and then the other 3 altogether later?}
-\begin{enumerate}
-\item $T_{srq} \overset{ind}{\sim} \text{Exponential}(\varphi_{sr})$
-\item $T_{srq} \overset{ind}{\sim} \text{Gamma}(\alpha, \varphi_{sr})$
-\item $T_{srq} \overset{ind}{\sim} \text{Weibull}(k, \varphi_{sr})$
-\item $T_{srq} \overset{ind}{\sim} \text{Lognormal}(\mu_{sr}, \sigma_{det}^2)$
-\end{enumerate}
-where $q = 1,\dotso,N_{sr}$ indexes individuals present at survey $sr$, and $\varphi_{sr}$ is always a survey-specific detection rate parameter.  
-We treat the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ as constant across all surveys.
+We can model detection heterogeneity by defining $T_{sq}$ as a mixture of TTDD components.  
+In this study, in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a component from an exponential, gamma, Weibull, or lognormal distribution.
+We define a constant mixing parameter $\gamma \in [0,1]$ for the proportion of animals not distributed according to the point mass.
 
-The other four TTDDs we use are partial finite-mixtures of a point mass at zero paired with each of the above distributions.  
-They can be written: $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$ where $\gamma \in (0,1)$ is a mixing parameter and $F_{T1}(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$ is one of the four above distributions.
+We model the expectation function via mixed effects linear regression.  
+For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{s}$.  
+In so doing, we equivalently model the log of the expected time-to-detection, which is inversely related to the detection rate.  
+For the lognormal distribution, we fit a regression to $-\mu_{s}$.  
+%Thus, the link functions in our detection model are:
+%\begin{alignat}{3}
+%&\text{Exponential:\;} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}])\\
+%&\text{Gamma:} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \log(\alpha)\\
+%&\text{Weibull:}  &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
+%&\text{Lognormal:} &&-\mu_{s} &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \sigma_{det}^2/2
+%\end{alignat}
+
+Combining the distribution and expectation function for $T_{s}$, we can describe a partial-mixture TTDD cdf as:
+\[F_T(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})\]
+where: $F_{T1}$ is the cdf of the non-point-mass component, $\textbf{X}_{sr}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{sr}^D$ specify random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.
+$F_T(t|\cdot)$ becomes a non-mixture TTDD if we fix $\gamma=1$.
 Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.
-
-For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{sr}$ via mixed effects regression using survey-specific covariates.  
-In so doing, we are equivalently modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  
-For the lognormal distribution, we fit a regression to $-\mu_{sr}$.  
-Thus, the link functions in our detection model are:
-\begin{alignat}{3}
-&\text{Exponential:\;} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}])\\
-&\text{Gamma:} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log(\alpha)\\
-&\text{Weibull:}  &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
-&\text{LogNormal:} &&-\mu_{sr} &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \sigma_{det}^2/2
-\end{alignat}
-\jarad{Are these necessary?}
-where $\boldsymbol{\beta}^D$ is a vector of fixed effects, and $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.
-When the TTDD is a mixture distribution, we model $F_{T1}$ portion of the mixture according to the above link functions and treat the mixing parameter $\gamma$ as constant across all surveys.
-
-Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.
-In summary, the detection model is:
-\begin{align}
-p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
-\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
-F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
-\end{align}
-
-To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq. (\ref{eq:datamodel}).  
-Instead of modeling the interval-specific counts $\vn_{sr}$, we model the times to detection directly.  
-Defining $\textbf{T}_{sr}$ as a vector of all observed detection times for a given survey, and continuing to treat all detections as independent, we write
-\[
-Pr\left(\textbf{T}_{sr}|n_{sr}^{(obs)}\right) = \prod\limits_{q=1}^{n_{sr}^{(obs)}} f_T(t_{srq}|t_{srq}<C_I)
-\]
-The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  
-Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.
 
@@
 
-For simplicity, we model abundance at each site during each year $N_{sr}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{sr}$, though any non-negative integer-valued distribution may be used.
-We can decompose $N_{sr}$ into observed and unobserved portions.  
+For simplicity, we model abundance at each site during each year $N_{s}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{s}$, though any non-negative integer-valued distribution may be used.
+We can decompose $N_{s}$ into observed and unobserved portions.  
 The unobserved portion follows a binomial-Poisson hierarchy, resulting in Equation \ref{eq:binPois}.  
-We model expected abundance $\lambda_{sr}$ using a mixed effects linear regression with a log link.
+We model expected abundance $\lambda_{s}$ using a mixed effects linear regression with a log link.
 
 \begin{align}
-\label{eq:binPois} n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)} &\sim \mbox{Poisson}\left(\lambda_{sr}(1-p_{sr}^{(det)})\right)\\
-\log (\lambda_{sr}) &= \text{Intercept}^A + \textbf{X}_{sr}^A\boldsymbol{\beta}^A + \textbf{Z}_{sr}^A\boldsymbol{\xi}^A
+\label{eq:binPois} n_{s}^{(unobs)} = N_{s} - n_{s}^{(obs)} &\sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)\\
+\log (\lambda_{s}) &= \text{Intercept}^A + \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A
 \end{align}
@@
 Our goal in selecting priors is to be non-informative within a reasonable range of values.  
-Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  
-Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  
+Priors for Intercept$^D$ are centered so that the median detection probability is $p_{s}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  
+Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{s}^{(det)} \in (0.01, 1.0)$.  
 Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
@@
 In particular, we are concerned with the goodness of fit for the TTDD distribution, so we define a check statistic for each observation interval:
-\[D(n_i) = \sum\limits_{sr} n_{sri} \big/ \sum\limits_{sri}^{} n_{sri}\]
+\[D(n_i) = \sum\limits_{s} n_{si} \big/ \sum\limits_{si}^{} n_{si}\]
 which is the proportion of the total count that occurs during interval $i$ marginally across all surveys.  
@@
 
-We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{sri}n_{sri}\big/$ $\sum\limits_{sri}N_{sri}$.  
+We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{si}n_{si}\big/$ $\sum\limits_{si}N_{si}$.  
 Because some posterior distributions are skewed, we emphasize posterior medians.  
@@
 \includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
-\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{sri}^{(rep)} > p_{sri}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim 1).
+\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim 1).
 Trendlines show mean posterior predictive p-values.  
@@
 
-At present, Stan does not sample discrete parameters, so the ability to marginalize out the discrete-valued latent variables $N_{sr}$, as we did, can greatly facilitate MCMC sampling.
+At present, Stan does not sample discrete parameters, so the ability to marginalize out the discrete-valued latent variables $N_{s}$, as we did, can greatly facilitate MCMC sampling.
 
@@
 We can modify our model to account for distance.  
-To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify Equation \ref{eq:pdet} to reflect the detection distance to individual $q$: $p_{srq}^{(det)} = g(d_q) F_T^A(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.
+To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify Equation \ref{eq:pdet} to reflect the detection distance to individual $q$: $p_{sq}^{(det)} = g(d_q) F_T^A(C_I|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.
 
@@
 Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  
-In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{srq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{sr}^A(t)$ and the perceptibility function $g(d/`8-_q)$.
-So, $h_{srq}(t) = g(d_q) h_{sr}^A(t)$.  
-From standard survival analysis results, $p_{srq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{sr}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) \right]^{g(d_q)}$.
+In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d/`8-_q)$.
+So, $h_{sq}(t) = g(d_q) h_{s}^A(t)$.  
+From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) \right]^{g(d_q)}$.
 

### 2016-03-28 Draft.tex
Revisions based on Jarad's comments about strucutre: model fitting, data, priors, etc.  The last structure issue yet to be resolved is the Simulations and Sim Results section.

@@
 
-\section{Model}\label{sec:model}
+\section{Methods}\label{sec:model}
 
-\subsection{Data structure}\label{sec:data}
+\subsection{Data}\label{sec:data}
+
+Our analysis is motivated by point-count surveys conducted in Chippewa National Forest over a six-year period from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  
+The MNFB design is described in \citet{Hanowski1995}, and we summarize only essential details here.  
+Single-visit time-to-first-detection point-count surveys are conducted by trained observers at a fixed set of sampling locations once annually.  
+Survey durations are 10 minutes, with times to first detection for over 50 species being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
+The relatively fine-scale interval sizes should provide more information about TTDDs than the three-interval methodology (0-3, 3-5, and 5-10 minutes) that characterizes many avian point count surveys \citep{Ralph1995}.
+There is no prescribed settling-in period between the observer's arrival and the time they begin observing.
 
@@
 We can model detection heterogeneity by defining $T_{sq}$ as a mixture of TTDD components.  
-In this study, in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a component from an exponential, gamma, Weibull, or lognormal distribution.
+In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a component from an exponential, gamma, Weibull, or lognormal distribution.
 We define a constant mixing parameter $\gamma \in [0,1]$ for the proportion of animals not distributed according to the point mass.
@@
 
-\jarad{Again, I would introduce this without covariates or random effects first.}
-
-For simplicity, we model abundance at each site during each year $N_{s}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{s}$, though any non-negative integer-valued distribution may be used.
-We can decompose $N_{s}$ into observed and unobserved portions.  
-The unobserved portion follows a binomial-Poisson hierarchy, resulting in Equation \ref{eq:binPois}.  
-We model expected abundance $\lambda_{s}$ using a mixed effects linear regression with a log link.
-
-\begin{align}
-\label{eq:binPois} n_{s}^{(unobs)} = N_{s} - n_{s}^{(obs)} &\sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)\\
-\log (\lambda_{s}) &= \text{Intercept}^A + \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A
-\end{align}
-where $\textbf{X}^A$ is a matrix of covariates associated with the vector of fixed effects $\boldsymbol{\beta}^A$, and $\textbf{Z}^A$ is a matrix identifying factor levels for $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ random effects.
-
-\jarad{Put a discussion of priors in a new section.
-You should introduce the general form of the priors with specific values and specify the specific values in the simulation study and real data analysis sections.}
-
-Since we use this hierarchical model in a Bayesian context, the complete model formulation requires priors for $\text{Intercept}^D, \boldsymbol{\beta}^D, \sigma_{D}^2, \boldsymbol{\theta}, \text{Intercept}^A, \boldsymbol{\beta}^A, \text{ and } \sigma_{A}^2$, such as we specify in Section \ref{sec:priors}.
-
-\jarad{You need to include a section about the MCMC, e.g. Stan/JAGS how to monitor convergence, etc.}
-
-\subsection{Priors}\label{sec:priors}
-Our goal in selecting priors is to be non-informative within a reasonable range of values.  
-Priors for Intercept$^D$ are centered so that the median detection probability is $p_{s}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  
-Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{s}^{(det)} \in (0.01, 1.0)$.  
-Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
-Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  
-For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  
-All priors are independent.
-\jarad{This takes up too much space.}
-\begin{align*}
-\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
-\text{Intercept}^A &\sim N(1, 1)\\
-\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
-\boldsymbol{\beta}^A &\sim N(0, 1)\\
-\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
-\gamma &\sim \text{Beta}(1,1)
-\end{align*}
+For simplicity, we model total abundance at each survey $N_{s}$ using a Poisson distribution, though any non-negative integer-valued distribution may be used.  
+We model the expected survey abundance $\lambda_{s}$ via mixed effects linear regression with a log link: $\log (\lambda_{s}) = \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A$.
+As with the detection model, $\textbf{X}_{sr}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{sr}^A$ specify random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
+We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)$.
 
 \subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
+To demonstrate our model, we fit counts of Ovenbirds (\textit{Seiurus aurocapilla}) from the MNFB.
+We restricted survey sites in our analysis to those of the most abundant habitat type (sawtimber red pine); in so doing, we focused modeling efforts upon the detection model and skirted the complication of modeling abundance across habitat types.
+We further constrained analysis to 65 sites with a site origin year between 1870-1970.  
+In total, we used 381 site-year surveys in our analysis with a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
 
-\jarad{This should not be a model (or methods section).
-I think you should put a Data section earlier that introduces the data collection methods.
-Then the rest of this section should go much later, after the simulation study, in a real data analysis section.
-}
-
-We analyze point-count surveys conducted in Chippewa National Forest over a six-year period from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  
-The MNFB design is described in \citet{Hanowski1995}, and we only summarize relevant details here.  
-Point-count time-to-detection surveys are conducted by trained observers at a fixed set of sampling locations once annually.  
-Forest stands follow a stratified random design, with three survey sites in each stand.  
-Survey durations are 10 minutes, with times to detection being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
-We restrict survey sites in our analysis to those of the most abundant habitat type (sawtimber red pine); in so doing, we focus modeling efforts upon the detection model and skirt the complication of modeling abundance across habitat types.
-We further constrain analysis to 65 sites with a site origin year between 1870-1970.  
-In total, we use 381 site-year surveys in our analysis with a total of 947 Ovenbirds counted, and a maximum of eight birds counted in any single survey.
-
-For the abundance half of our model, we use four covariates plus two random effects.  
-The covariates are: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
-We associate random effects with each survey year and each stand.  
-For the detection half of our model, we use covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.
-69\% of surveys in our dataset involve observers in their first year at MNFB.  
+For the abundance half of our model, we used four covariates plus two random effects.  
+The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
+We associated random effects with each survey year and each stand, with most stands consisting of three survey sites.  
+For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.
+69\% of surveys in our dataset involved observers in their first year at MNFB.  
 Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  
-Finally, we associate random effects with each observer.  
-We center and standardize all continuous covariates prior to fitting models.
+We associated random effects with each observer.  
+We centered and standardized all continuous covariates prior to fitting models.
 
@@
 
-We fit eight models to the Ovenbird data representing each of the time-to-detection distributions described in Section \ref{sec:detectionmodel}.
+\subsection{Model fitting}\adam{Looking at other papers, the following all tend to get bundled into the same paragraph or two.  It should follow presentation of our data and of our simulation scheme.}
 
-\subsection{Model fitting}
+We fit the models by MCMC sampling in the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
+For intercept-only simulations, we instantiated a single chain running 60,000 iterations; for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.  
+We discarded half of the iterations as warmup and then thinned by 1 in 10.  
+We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}.  
+We reran models if the effective sample size for any parameter was below 1000.  
+For most models, we accepted Stan defaults for dispersed initial values; however, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
 
-We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.
+We selected priors to be non-informative within a reasonable range of values.  
+Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of 1: (i) median detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
+Normal priors for Intercept$^A$ centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
+Because covariates were centered and scaled to mean zero and variance one, normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
+For all shape and variance parameters, we chose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a Uniform(0,1) prior.  
+All priors were independent.
 
-For each intercept-only simulation model, we instantiated a single chain running 60,000 iterations.  
-For the Ovenbird analysis and covariate simulations, we ran chains for 250,000-375,000 iterations.  
-We discarded half of the iterations as warmup and then thinned by 1 in 10.  
-We reran models if the effective sample size for any parameter was below 1000.  
-We monitored convergence of the MCMC chains using Geweke diagnostics.  
-For most models, we accepted the Stan defaults for generating dispersed initial values.  
-However, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from the posterior means of preliminary short-run models (Ovenbird analysis, incorrect-family simulations) or from true parameter values (correct-family simulations).
+To assess goodness of fit, we relied on posterior predictive checks \citep{Gelman1996}.  
+Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  
+In particular, we are concerned with the goodness of fit for the TTDD distribution, so we defined a check statistic for each observation interval:
+\[D(n_i) = \sum\limits_{s} n_{si} \big/ \sum\limits_{si} n_{si}\]
+which is the proportion of the total count that occurs during interval $i$ marginally across all surveys.  
+The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.  
+We simulated replicate datasets and calculated the check statistic at each iteration of our MCMC sampling.  
+Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
 
-\jarad{New section for simulation studies with the results.}
+We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{s}n_{s}^{(obs)}\big/$ $\sum\limits_{s}N_{s}$.  
+Within simulation studies, we compared model estimates to true $p^{(det)}$ values by percent coverage and by posterior p-values --- the proportion of MCMC samples having a smaller value than the true value.
 
-\subsection{Simulation studies}
+We compared different models by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
+For the full model with random effects, which is a missing data model, we used formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
+
+\section{Simulation studies}
+
+\jarad{Not really Jarad's comment.  But he left notes that there should be a separate Sim section with results bundled within it.}
 
@@
 Unlike Sims 1 \& 2, overall detection probabilities in Sim 3 are not the same for each dataset.
-
-\subsection{Model assessment, inference, and between-model comparison}
-
-\jarad{This is also part of methods. Or perhaps make it part of evaluation of simulation studies.}
-
-To assess goodness of fit, we rely on posterior predictive checks \citep{Gelman1996}/adam{Rubin1984?}.  
-Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  
-In particular, we are concerned with the goodness of fit for the TTDD distribution, so we define a check statistic for each observation interval:
-\[D(n_i) = \sum\limits_{s} n_{si} \big/ \sum\limits_{si}^{} n_{si}\]
-which is the proportion of the total count that occurs during interval $i$ marginally across all surveys.  
-The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.  
-We simulate replicate datasets and calculate the check statistic at each iteration of our MCMC samples.  
-Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
-
-We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{si}n_{si}\big/$ $\sum\limits_{si}N_{si}$.  
-Because some posterior distributions are skewed, we emphasize posterior medians.  
-Within simulation studies, we compare model estimates to true $p^{(det)}$ values by coverage and by posterior p-values --- the proportion of MCMC samples that have a smaller value than the true value.
-We do not focus on estimation of individual detection parameters, since different sets of parameters can yield similar density functions for time-to-detection conditional upon being detected: $f_T(t\;|\;t<C_I)$.
-
-We compare different model fits by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
-For the full model with random effects, which is a missing data model, we use formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
-
-\section{Results}\label{sec:results}
-
-\subsection{Simulation studies}
 
@@
 DIC performance in Sim 3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects, but also because DIC is calculated according to a different method.
+
+\section{Results}\label{sec:results}
 

### 2016-03-31 Draft.tex
Edits to simulation methods, including changing verbs to past tense.

@@
 
-We conduct three simulation studies to assess: (Sim 1) the importance of modeling detection heterogeneity via a partial finite-mixture component, (Sim 2) the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD), and (Sim 3) model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.
-Because model behavior may differ between datasets with peaked and non-peaked TTDDs, we simulate both kinds of dataset for gamma, Weibull, and lognormal models.  
-For all families, we simulate both mixture and non-mixture datasets.  
-In total, we simulate 14 different dataset types.  
-Counts are censored into nine-intervals as with the Ovenbird dataset.
+We conducted three simulation studies to assess: [Sim 1] the performance of modeling detection heterogeneity via a partial finite-mixture TTDD relative to a non-mixture TTDD, [Sim 2] the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD), and [Sim 3] model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.
+Our simulations and models were from exponential, gamma, lognormal, and Weibull TTDD families.  
+Because model behavior may differ between peaked and non-peaked datasets, we simulated both for all TTDD families except exponential.  
+Thus, combined with partial finite-mixtures and non-mixtures, we simulated datasets from 14 different TTDDs in total.  
+Counts were censored into nine-intervals as with the Ovenbird dataset.
 
-In the first simulation study (Sim 1), for all 14 dataset types, we fit models with both mixture and non-mixture TTDD where the family is correctly specified.  
-In Sim 2, we consider only mixture TTDD datasets, and we fit them with mixture TTDD models from all four families.  
-For both Sim 1 and Sim 2, we simulate and model 16 replicates.  
-In Sim 3, we repeat both the first two studies, but: (i) we include covariates and random effects described in Section \ref{sec:ovenbirdanalysis}, and (ii) we perform only a single replicate, because each gamma model takes an average of 5 days to fit.
+In the first simulation study (Sim 1), for all 14 dataset types, we fit models with both mixture and non-mixture TTDD where the family was correctly specified.  
+In Sim 2, we considered only mixture TTDD datasets, and we fit them with mixture TTDD models from all four families.  
+For both Sim 1 and Sim 2, we simulated and modeled 16 replicates.  
+In Sim 3, we repeated both the first two studies, but: (i) we included covariates and random effects described in Section \ref{sec:ovenbirdanalysis}, and (ii) we performed only a single replicate, because gamma models took an average of 5 days to fit.
 
-For the first two simulation studies, we choose true parameter values so that: (i) the expected detection probability is 0.80, (ii) the expected count is the same as observed in the Ovenbird dataset, (iii) when there is a mixture component, we set $\gamma = 0.65$, meaning that 35\% of individuals are immediately detected, (iv) in non-peaked models, 70\% of \textit{detected} individuals are observed during the first two minutes, and (v) in peaked models, the detection mode for `hard to detect' individuals occurs at 5 minutes.
-For Sim 3, we set true parameter values for each dataset equal to the median posterior values obtained when fitting the same TTDD model to the Ovenbird data.  
-Since all models applied to the Ovenbird data resulted in non-peaked model fits, we simulated peaked dataset variants by: (i) starting with posterior median Ovenbird estimates, (ii) setting the mixing parameter to $\gamma = 0.65$, and (iii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, which on average results in peak detection near 5 minutes and 75-85\% detection across all sites.
-Covariates and random effect levels are taken directly from the Ovenbird dataset.  
-Unlike Sims 1 \& 2, overall detection probabilities in Sim 3 are not the same for each dataset.
+Estimates from the Ovenbird analysis were used to inform simulation parameters except as follows.  
+For Sim1 and Sim2, for consistency across simulations, we chose true parameter values so that: (i) the expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in non-peaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+Since all models applied to the Ovenbird data resulted in non-peaked model fits, for Sim3 we simulated peaked dataset variants by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, which resulted on average in peak detection near 5 minutes and 75-85\% detection across all sites.
+Unlike Sim1 and Sim2, overall detection probabilities in Sim 3 were not the same for each dataset.
 

### 2016-04-07 Draft.tex
Inching my way through Sim Results.  Complete reorganization of the first three paragraphs.  Lots of changes to verb tense.

@@
 
-Our analysis is motivated by point-count surveys conducted in Chippewa National Forest over a six-year period from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  
+Our analysis is motivated by point-count surveys conducted in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  
 The MNFB design is described in \citet{Hanowski1995}, and we summarize only essential details here.  
-Single-visit time-to-first-detection point-count surveys are conducted by trained observers at a fixed set of sampling locations once annually.  
-Survey durations are 10 minutes, with times to first detection for over 50 species being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
+Single-visit time-to-first-detection point-count surveys were conducted by trained observers at a fixed set of sampling locations once annually.  
+Survey durations were 10 minutes, with times to first detection for over 50 species being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
 The relatively fine-scale interval sizes should provide more information about TTDDs than the three-interval methodology (0-3, 3-5, and 5-10 minutes) that characterizes many avian point count surveys \citep{Ralph1995}.
-There is no prescribed settling-in period between the observer's arrival and the time they begin observing.
+There was no prescribed settling-in period between the observer's arrival and the time they began observing.
 
@@
 For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  
-The number of animals counted during any interval is $n_{si}$, the vector of counts for a survey is written $\vn_{s}$, and $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$.
+The number of animals counted during any interval is $n_{si}$ such that $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and the vector of counts across all intervals at a survey is written $\vn_{s}$.
 
@@
 The two main elements of the detection model for $T_{sq}$ are: (i) the distributional form and (ii) the expectation function.
-$T_{sq}$ should be based on non-negative continuous distributions featuring some variety of rate parameter than can be estimated via regression.  
+$T_{sq}$ should be based on non-negative continuous distributions, preferrably featuring a rate parameter than can be estimated via regression.  
 In this study, we adopt TTDDs for $T_{sq}$ using the following: Exponential($\varphi_{s}$), Gamma($\alpha, \varphi_{s}$), Weibull($k, \varphi_{s}$), and Lognormal($\mu_{s}, \sigma_{det}^2$).  
@@
 
-We can model detection heterogeneity by defining $T_{sq}$ as a mixture of TTDD components.  
+We can model detection heterogeneity by defining $F_T(t)$ as a mixture of TTDD components.  
 In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a component from an exponential, gamma, Weibull, or lognormal distribution.
@@
 We model the expectation function via mixed effects linear regression.  
-For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{s}$.  
-In so doing, we equivalently model the log of the expected time-to-detection, which is inversely related to the detection rate.  
-For the lognormal distribution, we fit a regression to $-\mu_{s}$.  
+For exponential, gamma, and Weibull TTDDs, we model: $\log(\varphi_{s}) = \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D$, where $\textbf{X}_{s}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{s}^D$ specify random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.
+This regression is equivalent to modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  
+For the lognormal distribution, we fit the regression model to $-\mu_{s}$.  
 %Thus, the link functions in our detection model are:
@@
 \[F_T(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})\]
-where: $F_{T1}$ is the cdf of the non-point-mass component, $\textbf{X}_{sr}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{sr}^D$ specify random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.
-$F_T(t|\cdot)$ becomes a non-mixture TTDD if we fix $\gamma=1$.
+where $F_{T1}$ is the cdf of the non-point-mass component and $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$.  
 Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.
@@
 
-For simplicity, we model total abundance at each survey $N_{s}$ using a Poisson distribution, though any non-negative integer-valued distribution may be used.  
+For simplicity, we model total abundance at each survey $N_{s}$ using a Poisson distribution, though other non-negative integer-valued distributions may be used.  
 We model the expected survey abundance $\lambda_{s}$ via mixed effects linear regression with a log link: $\log (\lambda_{s}) = \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A$.
-As with the detection model, $\textbf{X}_{sr}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{sr}^A$ specify random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
+As with the detection model, $\textbf{X}_{s}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{s}^A$ specify random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
 We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)$.
@@
 
-\subsection{Model fitting}\adam{Looking at other papers, the following all tend to get bundled into the same paragraph or two.  It should follow presentation of our data and of our simulation scheme.}
-
 We fit the models by MCMC sampling in the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
@@
 We selected priors to be non-informative within a reasonable range of values.  
-Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of 1: (i) median detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
-Normal priors for Intercept$^A$ centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
-Because covariates were centered and scaled to mean zero and variance one, normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
+Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of 1: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
+Normal priors for Intercept$^A$ were centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
+Normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
 For all shape and variance parameters, we chose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a Uniform(0,1) prior.  
@@
 We compared different models by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
-For the full model with random effects, which is a missing data model, we used formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
+For the full model with random effects (Ovenbirds and Sim3), which is a missing data model, we used formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
 
@@
 
-\jarad{Not really Jarad's comment.  But he left notes that there should be a separate Sim section with results bundled within it.}
+We conducted three simulation studies to explore the behavior of models with non-constant TTDDs.
+In Sim1, we compared models with detection heterogeneity (via a partial finite-mixture TTDD) versus models without detection heterogeneity.
+We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
+Because model behavior may differ between peaked and non-peaked datasets, we simulated both for each TTDD family except exponential.  
+Thus, combining mixtures, `peakedness' and TTDD families, we simulated datasets from 14 different TTDDs in total.  
+In Sim2, we examined the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
+Based on findings from Sim1, we considered only mixture TTDD datasets, and we fit each with mixture TTDD models from all four families.  
+In Sim3, we modeled a more complete scenario where both data generation and model fitting included covariates and random effects for abundance and detection.
 
-We conducted three simulation studies to assess: [Sim 1] the performance of modeling detection heterogeneity via a partial finite-mixture TTDD relative to a non-mixture TTDD, [Sim 2] the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD), and [Sim 3] model performance in a more complete scenario when both data generation and model fitting include covariates and random effects for abundance and detection.
-Our simulations and models were from exponential, gamma, lognormal, and Weibull TTDD families.  
-Because model behavior may differ between peaked and non-peaked datasets, we simulated both for all TTDD families except exponential.  
-Thus, combined with partial finite-mixtures and non-mixtures, we simulated datasets from 14 different TTDDs in total.  
+We used posterior median estimates from the Ovenbird analysis to inform simulation parameters except as follows.  
+For Sim1 and Sim2, we chose true parameter values so that: (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in non-peaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+For Sim3, all parameters used for non-peaked datasets came from the Ovenbird analysis, which produced only non-peaked model fits.
+In order to simulate peaked dataset variants, we modified the Ovenbird parameters by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, resulting on average in a detection mode near 5 minutes and 75-85\% overall detection across all sites.
+Unlike Sim1 and Sim2, detection probabilities in Sim3 were not the same for all datasets.
 Counts were censored into nine-intervals as with the Ovenbird dataset.  
+For both Sim1 and Sim2, we simulated 16 replicates, but for Sim3 we generated only a single replicate, because gamma models took an average of 5 days to fit.
 
-In the first simulation study (Sim 1), for all 14 dataset types, we fit models with both mixture and non-mixture TTDD where the family was correctly specified.  
-In Sim 2, we considered only mixture TTDD datasets, and we fit them with mixture TTDD models from all four families.  
-For both Sim 1 and Sim 2, we simulated and modeled 16 replicates.  
-In Sim 3, we repeated both the first two studies, but: (i) we included covariates and random effects described in Section \ref{sec:ovenbirdanalysis}, and (ii) we performed only a single replicate, because gamma models took an average of 5 days to fit.
+YOU ARE HERE.  ADD A COMMENT ABOUT ABUNDANCE ESTIMATION.
 
-Estimates from the Ovenbird analysis were used to inform simulation parameters except as follows.  
-For Sim1 and Sim2, for consistency across simulations, we chose true parameter values so that: (i) the expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in non-peaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-Since all models applied to the Ovenbird data resulted in non-peaked model fits, for Sim3 we simulated peaked dataset variants by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, which resulted on average in peak detection near 5 minutes and 75-85\% detection across all sites.
-Unlike Sim1 and Sim2, overall detection probabilities in Sim 3 were not the same for each dataset.
-
-\subsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}
+\subsection{Sim1 Results: mixture vs. non-mixture TTDDs}\label{sec:mixture}
 
 In the first simulation study, we simulated intercept-only datasets from 14 TTDDs and fit each with both the correctly specified model and one that was correct except for the mixture component (Table \ref{tbl:sim1}).
-When data derive from a mixture model, then inference models that lack a mixture component misestimate the detection probability $p^{(det)}$ badly in every case.  
-This is particularly egregious when datasets feature peaked mixtures; then model coverage is always negligible.  
-Conversely, when data derive from a non-mixture model, then inference models that include a mixture component perform just as well as the true model.  
-Indeed, for non-peaked non-mixture data, the mixture inference model provides more accurate posterior estimates of $p^{(det)}$ than the non-mixture inference model.  
-The one exception is the exponential case, where both models estimate $p^{(det)}$ accurately, but the mixture inference model yields credible intervals about 25\% wider than those from its mixture counterpart.
+When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $p^{(det)}$ badly in every case.  
+This was particularly egregious when datasets feature peaked mixtures, where model coverage was always negligible.  
+Conversely, when data derived from a non-mixture model, then inference models incorporating a mixture component performed as well as the true model.  
+Indeed, for non-peaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $p^{(det)}$ than the non-mixture inference model.  
+The one exception was the exponential case, where both models estimated $p^{(det)}$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
+In most cases, model comparisons by DIC did not demonstrate a clear preference for the mixture model, even though non-mixture models displayed stronger bias and poorer coverage.
 
@@
 \end{tabular}
-\caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim 1).  
+\caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim1).  
 In all cases, the model family matches the dataset family.  
@@
 
-\subsection{Constant vs. non-constant detection}\label{sec:family}
+\subsection{Sim2: Constant vs. non-constant detection TTDDs}\label{sec:family}
 
@@
 \end{tabular}
-\caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim 2).  
+\caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim2).  
  Med $p$: mean value across simulations of the posterior median of $p^{(det)}$.  
@@
 
-\subsection{Simulations with covariates and random effects}\label{sec:simfull}
+\subsection{Sim3: with covariates and random effects}\label{sec:simfull}
 
@@
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim 3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.  
+\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.  
 Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
@@
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim 3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.
+\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.
 All data and inference models include mixtures.
@@
 
-As in Sim 2, for mixture exponential datasets only, mixture exponential inference models in Sim 3 returned accurate estimates of $p^{(det)}$ and had credible intervals from one-fourth to one-half as wide as from other inference models (Figure \ref{pdet_cater_family}).
+As in Sim2, for mixture exponential datasets only, mixture exponential inference models in Sim3 returned accurate estimates of $p^{(det)}$ and had credible intervals from one-fourth to one-half as wide as from other inference models (Figure \ref{pdet_cater_family}).
 However, exponential mixture models were generally biased high for non-peaked datasets and low for peaked datasets.  
@@
 Estimates of detection parameters varied by context.  
-In comparisons of mixture and non-mixture models within the same TTDD family (a la Sim 1), estimates were similar except when data derived from peaked mixture TTDDs.  
+In comparisons of mixture and non-mixture models within the same TTDD family (a la Sim1), estimates were similar except when data derived from peaked mixture TTDDs.  
 In that case, mixture models yielded narrower credible intervals than non-mixture models --- except in the exponential case, where non-mixture exponential models were overly precise.  
-In comparisons of models across families (a la Sim 2), estimates across gamma, lognormal, and Weibull mixture inference models are always quite similar to one another.  
+In comparisons of models across families (a la Sim2), estimates across gamma, lognormal, and Weibull mixture inference models are always quite similar to one another.  
 Credible intervals from exponential mixture models are wider for peaked data but narrower for exponential and non-peaked data.
@@
 \includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
-\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim 1).
+\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim1).
 Trendlines show mean posterior predictive p-values.  
@@
 For peaked and exponential mixture datasets (Sims 1 \& 3), DIC regularly favored a mixture inference model over a non-mixture model (differences were always greater than 12).  
-Also, when peaked gamma or lognormal mixtures were the true data model (Sim 2), DIC signalled against fitting an exponential mixture model (differences greater than 5).  
+Also, when peaked gamma or lognormal mixtures were the true data model (Sim2), DIC signalled against fitting an exponential mixture model (differences greater than 5).  
 These are the same scenarios where posterior predictive checks highlighted model misfit.
 
-DIC comparisons of TTDD family in the single replicate from Sim 3 were variable.  
+DIC comparisons of TTDD family in the single replicate from Sim3 were variable.  
 For peaked mixture datasets, correct models were favored over mixture exponential inference models by at least 10 in gamma and lognormal cases and by 4 in the Weibull case.  
@@
 Considering that exponential models were biased for detection in non-peaked models (Figure \ref{pdet_cater_family}), we have little confidence in our DIC results for non-peaked models.  
-DIC performance in Sim 3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects, but also because DIC is calculated according to a different method.
+DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects, but also because DIC is calculated according to a different method.
 
@@
 
-DIC calculations prefer the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively), and these differences closely mirror those from the exponential mixture dataset in Sim 3.
-However, Sim 3 demonstrated that DIC for non-peaked mixture data may not be reliable; indeed, for the simulated non-peak lognormal mixture dataset, the exponential mixture TTDD was favored by over a 4-unit difference; yet, its posterior 95\% credible interval for detection probability did not contained the true value, while it was within a central 60\% credible interval for all three other mixture models.
+DIC calculations prefer the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively), and these differences closely mirror those from the exponential mixture dataset in Sim3.
+However, Sim3 demonstrated that DIC for non-peaked mixture data may not be reliable; indeed, for the simulated non-peak lognormal mixture dataset, the exponential mixture TTDD was favored by over a 4-unit difference; yet, its posterior 95\% credible interval for detection probability did not contained the true value, while it was within a central 60\% credible interval for all three other mixture models.
 In short, DIC may not be trustworthy in this situation.

### 2016-04-13 Draft.tex
Many changes over the past week.  Mainly to the simulations and results sections.

@@
 Choosing another TTDD can allow for a systematic non-constant detection regime.  
-For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate (Figure \ref{HazardFxns}).
+For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
 
 Increasing hazard functions result in a pdf for $T$ that has a mode away from zero; we term these TTDDs `peaked'.  
-We contrast these to decreasing hazard functions with a mode at zero, which we term `non-peaked'.  
-For our purposes, the exponential TTDD is neither peaked nor non-peaked.
+We contrast these to decreasing hazard functions with a mode at zero, which we term `nonpeaked'.  
+For our purposes, the exponential TTDD is neither peaked nor nonpeaked.
 
@@
 \adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
-
-\begin{figure}[h!]\centering
-\includegraphics[width=0.68\textwidth]{Hazards}
-\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
-\label{HazardFxns}
-\end{figure}
 
@@
 Single-visit time-to-first-detection point-count surveys were conducted by trained observers at a fixed set of sampling locations once annually.  
-Survey durations were 10 minutes, with times to first detection for over 50 species being censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
-The relatively fine-scale interval sizes should provide more information about TTDDs than the three-interval methodology (0-3, 3-5, and 5-10 minutes) that characterizes many avian point count surveys \citep{Ralph1995}.
+Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
+These relatively fine-scale intervals should provide more information about TTDDs than the three-interval methodology (0-3, 3-5, and 5-10 minutes) used in many avian point count surveys \citep{Ralph1995}.
 There was no prescribed settling-in period between the observer's arrival and the time they began observing.
@@
 
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and time-to-detection pieces, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued distribution.
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and observed times to detection, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued count distribution.
 Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
@@
 We assume that detection times for individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$, which may depend on covariates.  
-The probability of detecting an animal is thus $p_{s}^{(det)} = Pr(t_{sq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) / F_T(C_I)$.
-Given $N_{s}$ total number animals present at a survey, the data model is:
+The probability of detection for an animal is thus $p_{s}^{(det)} = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) \textbf{I}(t<C_I) / F_T(C_I)$.
+Denoting $N_{s}$ as the total number animals present at a survey, the data model is:
 \begin{align}\label{eq:datamodel}
@@
 In this study, we adopt TTDDs for $T_{sq}$ using the following: Exponential($\varphi_{s}$), Gamma($\alpha, \varphi_{s}$), Weibull($k, \varphi_{s}$), and Lognormal($\mu_{s}, \sigma_{det}^2$).  
-Here, $\varphi_{s}$ and $\mu_{s}$ are survey-specific detection rate parameters, and we consider the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ to be constant across all surveys.
+Here, $\varphi_{s}$ and $\mu_{s}$ are survey-specific detection rate parameters, and we consider the shape parameters $\alpha$, $k$, and $\sigma_{det}^2$ to be constant across all surveys.
 
 We can model detection heterogeneity by defining $F_T(t)$ as a mixture of TTDD components.  
-In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a component from an exponential, gamma, Weibull, or lognormal distribution.
+In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a continuous component from an exponential, gamma, Weibull, or lognormal distribution.
 We define a constant mixing parameter $\gamma \in [0,1]$ for the proportion of animals not distributed according to the point mass.
@@
 We model the expectation function via mixed effects linear regression.  
-For exponential, gamma, and Weibull TTDDs, we model: $\log(\varphi_{s}) = \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D$, where $\textbf{X}_{s}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{s}^D$ specify random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.
+For exponential, gamma, and Weibull TTDDs, we model: $\log(\varphi_{s}) = \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D$, where $\textbf{X}_{s}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{s}^D$ specifies random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, and $\sigma_{D[i]}^2$ denotes the variance associated with $\xi_i^D$.
 This regression is equivalent to modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  
-For the lognormal distribution, we fit the regression model to $-\mu_{s}$.  
+For the lognormal distribution, we fit the same regression model to $-\mu_{s}$.  
 %Thus, the link functions in our detection model are:
@@
 \[F_T(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})\]
-where $F_{T1}$ is the cdf of the non-point-mass component and $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$.  
+where $F_{T1}$ is the cdf of the continuous component and $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$.  
 Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.
@@
 We model the expected survey abundance $\lambda_{s}$ via mixed effects linear regression with a log link: $\log (\lambda_{s}) = \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A$.
-As with the detection model, $\textbf{X}_{s}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{s}^A$ specify random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
+As with the detection model, $\textbf{X}_{s}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{s}^A$ specifies random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
 We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)$.
@@
 
+We selected priors to be non-informative within a reasonable range of values.  
+Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of one: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
+Normal priors for Intercept$^A$ were centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
+Normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
+For all shape and variance parameters, we chose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a Uniform(0,1) prior.  
+All priors were independent.
+
 % Effects: * - means signif
@@
 
+\subsection{Model Fitting}
+
 We fit the models by MCMC sampling in the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
@@
 
-We selected priors to be non-informative within a reasonable range of values.  
-Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of 1: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
-Normal priors for Intercept$^A$ were centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
-Normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
-For all shape and variance parameters, we chose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a Uniform(0,1) prior.  
-All priors were independent.
-
 To assess goodness of fit, we relied on posterior predictive checks \citep{Gelman1996}.  
@@
 The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.  
+Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
 We simulated replicate datasets and calculated the check statistic at each iteration of our MCMC sampling.
-Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
 
@@
 We conducted three simulation studies to explore the behavior of models with non-constant TTDDs.
-In Sim1, we compared models with detection heterogeneity (via a partial finite-mixture TTDD) versus models without detection heterogeneity.
+In Sim1, we compared models with detection heterogeneity via a partial finite-mixture TTDD versus models without detection heterogeneity.
 We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
-Because model behavior may differ between peaked and non-peaked datasets, we simulated both for each TTDD family except exponential.  
+Because model behavior may differ between peaked and nonpeaked datasets, we simulated both for each TTDD family except exponential.  
 Thus, combining mixtures, `peakedness' and TTDD families, we simulated datasets from 14 different TTDDs in total.  
-In Sim2, we examined the impact of misspecifying the TTDD family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
+In Sim2, we examined the impact of misspecifying the $F_{T1}$ family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
+Of particular interest were scenarios with a time-varying $F_{T1}$ component in the data but a time-constant component in the inference model, and vice versa.
 Based on findings from Sim1, we considered only mixture TTDD datasets, and we fit each with mixture TTDD models from all four families.  
@@
 We used posterior median estimates from the Ovenbird analysis to inform simulation parameters except as follows.  
-For Sim1 and Sim2, we chose true parameter values so that: (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in non-peaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-For Sim3, all parameters used for non-peaked datasets came from the Ovenbird analysis, which produced only non-peaked model fits.
+For Sim1 and Sim2, we chose true parameter values so that: (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+For Sim3, all parameters used for nonpeaked datasets came from the Ovenbird analysis.
 In order to simulate peaked dataset variants, we modified the Ovenbird parameters by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, resulting on average in a detection mode near 5 minutes and 75-85\% overall detection across all sites.
-Unlike Sim1 and Sim2, detection probabilities in Sim3 were not the same for all datasets.
+Unlike Sim1 and Sim2, expected detection probabilities in Sim3 were not the same for all datasets.
 Counts were censored into nine-intervals as with the Ovenbird dataset.  
@@
 
-YOU ARE HERE.  ADD A COMMENT ABOUT ABUNDANCE ESTIMATION.
-
 \subsection{Sim1 Results: mixture vs. non-mixture TTDDs}\label{sec:mixture}
 
-In the first simulation study, we simulated intercept-only datasets from 14 TTDDs and fit each with both the correctly specified model and one that was correct except for the mixture component (Table \ref{tbl:sim1}).
+In Sim1, we simulated intercept-only datasets from 14 TTDDs and fit each with two models from the correct TTDD family: one allowing for detection heterogeneity in the form of a mixture component and one not (Table \ref{tbl:sim1}).
 When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $p^{(det)}$ badly in every case.  
-This was particularly egregious when datasets feature peaked mixtures, where model coverage was always negligible.  
-Conversely, when data derived from a non-mixture model, then inference models incorporating a mixture component performed as well as the true model.  
-Indeed, for non-peaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $p^{(det)}$ than the non-mixture inference model.  
-The one exception was the exponential case, where both models estimated $p^{(det)}$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
-In most cases, model comparisons by DIC did not demonstrate a clear preference for the mixture model, even though non-mixture models displayed stronger bias and poorer coverage.
+This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
+Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
+Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $p^{(det)}$ than did the true non-mixture inference model.  
+The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $p^{(det)}$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
+Mixture model posterior medians for $p^{(det)}$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
+
+\adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
+Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $p^{(det)}$ above and shared the same coverage rates.  
+Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
+Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
+For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
+For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
+
+Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
+DIC did reveal a preference for mixture models with exponential and peaked datasets.
+In these cases, posterior predictive checks clearly showed poor fit from the non-mixture models with p-values of 1.00 in the second and sometimes third observation intervals.
+These p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
 
 % Possible other comments
-%(1) For non-peaked data, the true inference model is biased low\\
+%(1) For nonpeaked data, the true inference model is biased low\\
 %(2) Mixture inference models generate higher estimates than non-mixture models\\
@@
 \caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim1).  
-In all cases, the model family matches the dataset family.  
-Med $p$: mean value across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  
-Q($p$): proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value on average.  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value.  
 50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model. \adam{Thoughts on use of bold/italics to highlight some results?}}
+$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.}
 \end{table}
 
-\subsection{Sim2: Constant vs. non-constant detection TTDDs}\label{sec:family}
+\subsection{Sim2: constant vs. non-constant detection mixture TTDDs}\label{sec:family}
 
-In the second simulation study, we simulated intercept-only datasets from 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, Weibull, and lognormal families (Table \ref{tbl:sim2}).
-Mixture exponential inference models returned accurate estimates of $p^{(det)}$ for mixture exponential data and had 95\% credible intervals from one-third to one-half as wide as from other inference models.
-However, exponential mixture models yielded biased estimates and poor coverage for non-exponential data, overestimating detection for non-peaked datasets and underestimating it from peaked datasets.  
-Lognormal mixture models underestimated detection from all datasets; when modeling lognormal data, they were competitive with other models in terms of accuracy and coverage but were not superior.  
-Mixture gamma and Weibull models provided solid coverage for 90\% credible intervals with all datasets.  
-Gamma models were more accurate than others for non-peaked mixture datasets; gamma and Weibull seem equally accurate for peaked mixture datasets.  
-95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
+In Sim2, we simulated intercept-only datasets from the 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families (Table \ref{tbl:sim2}).
+Mixture exponential models returned more accurate estimates of $p^{(det)}$ than other models when data came from an exponential mixture, and 95\% credible intervals for $p^{(det)}$ were from one-third to one-half as wide.
+However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.
+Differences among non-constant TTDDs were less pronounced.  
+Gamma models were more accurate than others for nonpeaked mixture datasets.  
+Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
+Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
+
+Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
+For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
+So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
+For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
+
+In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.
+For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.
 
 \begin{table}[ht]
-\small\centering
-\begin{tabular}{l|l|ccccc|ccccc}
- \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
- \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+\footnotesize\centering
+\begin{tabular}{l|l|l|ccccc|ccccc}
+ \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
   \hline
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\ 
-&   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
-&   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
-\cline{2-12}
-&   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
-\cline{2-12}
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
-&   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
-&   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\
+& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
+& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
+\cline{2-13}
+& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
+\cline{2-13}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
+& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
+& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
    \hline
@@
 \vspace{0.5cm}\\
-\begin{tabular}{l|l|ccccc|ccccc}
- \multicolumn{2}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
- \multicolumn{2}{c}{Dataset} & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+\begin{tabular}{l|l|l|ccccc|ccccc}
+ \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
   \hline
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\ 
-&   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
-&   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
-\cline{2-12}
-&   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
-\cline{2-12}
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
-&   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
-&   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\
+& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
+& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
+\cline{2-13}
+& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
+\cline{2-13}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
+& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
+& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
    \hline
@@
 \caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim2).  
- Med $p$: mean value across simulations of the posterior median of $p^{(det)}$.  
-Q($p$): mean proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value (0.80).  
+Med $p$: average across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value.  
 50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  
@@
 
-\subsection{Sim3: with covariates and random effects}\label{sec:simfull}
+\subsection{Sim3: including covariates and random effects}\label{sec:simfull}
 
-In the third simulation study, we generated datasets from models including covariates and random effects for both detection and abundance.  
-Results echoed those of the first two simulation studies, suggesting that the addition of covariates and random effects does not alter basic model behavior.  
-When data came from a mixture, non-mixture inference models generally led to biased estimates of detection while mixture inference models did not (Figure \ref{pdet_cater_correct}).  
-When data did not come from a mixture, then: (i) credible intervals for peaked data were similar for both inference model types, (ii) mixture model estimates of $p^{(det)}$ were higher for non-peaked data, and (iii) mixture model credible intervals were about 30\% wider for truly exponential data.
+In Sim3, we generated a single replicate of datasets from models including covariates and random effects for both detection and abundance.  
+Results echoed those of the first two simulation studies, suggesting that the addition of covariates and random effects did not alter basic model behavior.
+These results included the detection bias of non-mixture models used to fit mixture datasets (Figure \ref{pdet_cater_correct}), detection bias of exponential mixture models used to fit data mixtures with time-varying $F_{T1}$ components (Figure \ref{pdet_cater_family}), and more precise and accurate performance of exponential mixture models when data were from exponential mixtures.
+
+Abundance estimates mirrored the first two simulation studies except: (i) credible intervals across all mixture models were on average 50\% narrower than in Sim2, and bias in posterior median abundance from exponential model fits of peaked datasets was roughly halved relative to Sim2, though it was still worse than gamma model fits by about 35\% of actual abundance.
+These results indicate that covariates provided significant information for reducing uncertainty.
 
@@
 
-As in Sim2, for mixture exponential datasets only, mixture exponential inference models in Sim3 returned accurate estimates of $p^{(det)}$ and had credible intervals from one-fourth to one-half as wide as from other inference models (Figure \ref{pdet_cater_family}).
-However, exponential mixture models were generally biased high for non-peaked datasets and low for peaked datasets.  
-Lognormal mixture models tended to underestimate detection from non-lognormal datasets.  
-Gamma models were more accurate than other models for non-peaked mixture datasets; gamma and Weibull both seem equally accurate for peaked mixture datasets.  
-95\% credible intervals for gamma models were 15-35\% narrower than for both Weibull and lognormal models.
+For every dataset we simulated, estimates and credible intervals for abundance covariate coefficients $\beta^A$ were virtually the same regardless of the inference TTDD employed.  
+Estimates of detection parameters $\beta^D$ varied by context.  
+Comparisons of mixture and non-mixture models (like in Sim1) produced similar $\beta^D$ estimates except when fit to mixture peaked datasets, where models without a mixture component produced credible intervals twice as wide as those from mixture models with no benefit in terms of coverage.
+Comparing constant-detection and time-varying models (like in Sim2), time-varying mixture TTDD models generally produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and narrower; they also had lower coverage across all $\beta^D$ parameters (88.5\% vs.
+94-97\%).
+Differences in estimates of $\beta^D$ parameters to some degree reflected differing estimation of the $\gamma$ mixing parameter -- the exponential model underestimated $\gamma$ with nonpeaked datasets and overestimated it in peaked datasets.
+Many of the above patterns for nonpeaked datasets can be noted in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
 
-For every dataset we simulated, estimates of abundance covariate coefficients were virtually the same regardless of the TTDD employed.  
-Estimates of detection parameters varied by context.  
-In comparisons of mixture and non-mixture models within the same TTDD family (a la Sim1), estimates were similar except when data derived from peaked mixture TTDDs.  
-In that case, mixture models yielded narrower credible intervals than non-mixture models --- except in the exponential case, where non-mixture exponential models were overly precise.  
-In comparisons of models across families (a la Sim2), estimates across gamma, lognormal, and Weibull mixture inference models are always quite similar to one another.  
-Credible intervals from exponential mixture models are wider for peaked data but narrower for exponential and non-peaked data.
+Results from DIC and posterior predictive checks suggested mixture models over non-mixture models for peaked data (DIC differences $>10$) and time-varying mixture TTDDs over the exponential mixture for gamma and lognormal peaked data ($>8$).
+DIC preferred exponential mixtures to time-varying mixtures for exponential mixture data ($>6$); however, it also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.
+In this last case, the posterior 95\% credible interval for $p^{(det)}$ from the exponential mixture model did not contain the true value of $p^{(det)}$, while the true value was within a central 60\% credible interval for all three other mixture models.
+In conjunction with Sim2 results, this last finding calls into question the utility of DIC for nonpeaked datasets.
+DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
 
-% The following plots of posteriors for abundance and detection parameters are just too large (and must be shrunk too small to make fit)
+%\adam{I'm guessing I should not include the following: (1) my posterior predictive pvalues are not independent; (2) for instance, for the first interval, they stay very close to 0.5-0.6 --- the models are effective at getting the first interval right; (3) likewise, a post.
+p-value $>$ 0.8 during the 2nd interval likely shows model misfit; all of which raises the possibility that (4) multiple simulation under the null could provide a reference distribution of posterior p-values.}
+
+% Posterior predictive p-values for Sim1
 %\begin{figure}[h!]\centering
-%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select6.pdf}
-%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select2.pdf}
-%\includegraphics[width=0.32\textwidth]{Sims/SimFull/Posteriors_select13.pdf}
+%\includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
+%\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim1).
+%Trendlines show mean posterior predictive p-values.  
+%Dashed lines show central 95\% intervals.  
+%P-values near 0 or 1 potentially indicate poor model fit.}
 %\end{figure}
 
-\subsection{Model fit}\label{sec:modelfit}
-
-Across all simulations, posterior predictive checks of by-interval counts provided only limited ability to identify poor model fits.  
-For peaked and exponential mixture datasets (Sims 1 \& 3), posterior predictive checks clearly showed lack of fit for non-mixture inference models (Figure \ref{postpredmix}).  
-For gamma, lognormal, and Weibull mixture datasets fit with an exponential inference model, second-interval posterior predictive values near to one often pointed a model misfit (Sims 2 \& 3).  
-These are all cases where true data distribution has one more mode than the inference model can accommodate; in these cases, the second interval is hardest to fit, because of the mixture formulation of the models.
-For all other models, posterior predictive checks did not consistently indicate lack of fit.
-
-\adam{I'm guessing I should not include the following: (1) my posterior predictive pvalues are not independent; (2) for instance, for the first interval, they stay very close to 0.5-0.6 --- the models are effective at getting the first interval right; (3) likewise, a post.
-p-value $>$ 0.8 during the 2nd interval likely shows model misfit; all of which raises the possibility that (4) multiple simulation under the null could provide a reference distribution of posterior p-values.}
-
-\begin{figure}[h!]\centering
-\includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
-\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim1).
-Trendlines show mean posterior predictive p-values.  
-Dashed lines show central 95\% intervals.  
-P-values near 0 or 1 potentially indicate poor model fit.}
-\end{figure}
-
-DIC showed a limited ability to select among models with different TTDDs.  
-For peaked and exponential mixture datasets (Sims 1 \& 3), DIC regularly favored a mixture inference model over a non-mixture model (differences were always greater than 12).  
-Also, when peaked gamma or lognormal mixtures were the true data model (Sim2), DIC signalled against fitting an exponential mixture model (differences greater than 5).  
-These are the same scenarios where posterior predictive checks highlighted model misfit.
-
-DIC comparisons of TTDD family in the single replicate from Sim3 were variable.  
-For peaked mixture datasets, correct models were favored over mixture exponential inference models by at least 10 in gamma and lognormal cases and by 4 in the Weibull case.  
-For the exponential mixture datasets, the exponential inference models was favored by at least 6 over all other models.  
-For non-peaked mixture datasets, models were generally within 4 units of one another --- with the exception that the exponential mixture model was favored by 8 units for the lognormal mixture dataset.
-Considering that exponential models were biased for detection in non-peaked models (Figure \ref{pdet_cater_family}), we have little confidence in our DIC results for non-peaked models.  
-DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects, but also because DIC is calculated according to a different method.
-
-\section{Results}\label{sec:results}
-
-\section{Ovenbird analysis}
+\section{Ovenbird analysis}\label{sec:ovenbirds}
 
 We fit mixture and non-mixture models from each family of TTDD to the Ovenbird dataset.  
-Posterior predictive check p-values by interval (Table \ref{tbl:ovencounts}) show a clear lack of fit for the straight exponential model over the first three intervals.  
-Two features of the data proved difficult to fit for all of the models: an increase in counts during the fourth interval, and a decrease during the final interval.  
-Mixture models were better able to fit these features than non-mixture models.
+Posterior predictive p-values by interval (Table \ref{tbl:ovencounts}) showed clear lack of fit for the non-mixture exponential model over the first three intervals.
 
@@
   Weibull & 0.582 & 0.862 & 0.401 & 0.005 & 0.359 & 0.616 & 0.221 & 0.509 & 0.982 \\ 
+\hline
   Exponential Mix & 0.533 & 0.804 & 0.577 & 0.021 & 0.501 & 0.668 & 0.190 & 0.338 & 0.918 \\ 
@@
 \end{tabular}
-\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive check p-values by observation interval for each model across all surveys.}
+\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive p-values -- Pr$\left(D(n_i)^{(rep)} > D(n_i)^{(obs)}\right)$ -- by observation interval for each model across all surveys.
+P-values near zero or one may indicate poor model fit.}
 \end{table}
 
-Ovenbird results reflect several patterns that were observed in the simulation studies for both exponential and non-peaked data.  
 Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $p^{(det)}$, and the log10 number of Ovenbirds uncounted.  
-Shape parameter estimates from mixture models favor non-peaked distributions --- i.e., detection rates are higher earlier in the observation period --- however, in mixture gamma and Weibull models, an exponential TTDD model ($\alpha=1$ or $k=1$) is well within 95\% credible intervals.
-These pattern is similar to our simulations involving non-peaked mixture datasets, especially exponential and lognormal.
-
-Abundance covariate coefficient estimates are virtually the same for all TTDD models.  
-Detection parameter estimates are similar by mixture type across gamma, Weibull, and lognormal TTDD models.  
-Credible intervals from exponential mixture models are more precise than for non-exponential mixture models.  
-Within mixture type, estimates of detection probability across all surveys, $p^{(det)}$, are lower and much more uncertain for gamma, Weibull, and lognormal models than for exponential.
-
-DIC calculations prefer the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively), and these differences closely mirror those from the exponential mixture dataset in Sim3.
-However, Sim3 demonstrated that DIC for non-peaked mixture data may not be reliable; indeed, for the simulated non-peak lognormal mixture dataset, the exponential mixture TTDD was favored by over a 4-unit difference; yet, its posterior 95\% credible interval for detection probability did not contained the true value, while it was within a central 60\% credible interval for all three other mixture models.
-In short, DIC may not be trustworthy in this situation.
+Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $p^{(det)}$, and total abundance across models, Ovenbird data most likely arose from a exponential or nonpeaked mixture TTDD.
+As in Sim3, abundance covariate coefficient estimates were virtually the same across all TTDD models.
 
@@
 
-\begin{figure}[h!]
-\includegraphics[width=0.98\textwidth]{OVEN/oven_sum/OVEN_posteriors.pdf}
+DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).
+These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
+
+The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a partial finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with partial finite-mixtures involving a time-varying TTDD.
+Simulation results demonstrate that the exponential mixture model fits exponential mixture data well, but it does not fit time-varying mixture data well.
+At this point, we have no reliable tool for differentiating field data between the two TTDD types.
+
+\begin{figure}[h!]\centering
+\includegraphics[width=0.88\textwidth]{OVEN/oven_sum/OVEN_posteriors.pdf}
 \caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to the Ovenbird data.  
@@
 Given the bias that can result from assuming constant detection or omitting a mixture component, conservatism dictates that non-constant TTDDs with a mixture should be preferred.  
-Among the TTDDs we considered, the mixture gamma TTDD appears to provide the best combination of accuracy and precision for non-peaked data.  
+Among the TTDDs we considered, the mixture gamma TTDD appears to provide the best combination of accuracy and precision for nonpeaked data.  
 For peaked data, gamma and Weibull models seem about equally as good.  

### 2016-04-16 Draft.tex
Removed one-sentence coverage comparison from Sim3 (it was overreach).

@@
 Comparisons of mixture and non-mixture models (like in Sim1) produced similar $\beta^D$ estimates except when fit to mixture peaked datasets, where models without a mixture component produced credible intervals twice as wide as those from mixture models with no benefit in terms of coverage.
-Comparing constant-detection and time-varying models (like in Sim2), time-varying mixture TTDD models generally produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and narrower; they also had lower coverage across all $\beta^D$ parameters (88.5\% vs.
-94-97\%).
+Comparing constant-detection to time-varying models (like in Sim2), time-varying mixture TTDD models produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and either narrower (exponential and nonpeaked data) or wider (peaked data).
 Differences in estimates of $\beta^D$ parameters to some degree reflected differing estimation of the $\gamma$ mixing parameter -- the exponential model underestimated $\gamma$ with nonpeaked datasets and overestimated it in peaked datasets.
@@
 However, although analysis of removal models has grown more sophisticated in recent years, the constant-detection assumption nonetheless continues unquestioned.  
-In our study, we have formulated a model for non-constant detection patterns using a time-to-event approach within a hierarchical N-mixture framework.  
+We have formulated a model for non-constant detection using a time-to-event approach within a hierarchical N-mixture framework.  
 Our results demonstrate that, for better or worse, the assumption of constant detection is a very strong assumption.  

### 2016-05-12 Draft.tex
Corrected 'partial finite-mixture'. Changed nomenclature for F_T(t|...). Changed T_{sq} to T_{sq}|det.

@@
 It is often modeled using a mixture of detection distributions.  
-\citet{Farnsworth2002} introduce a partial finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
+\citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
  Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
@@
 In this study, we construct a time-to-detection N-mixture model for estimating abundance from removal-sampled point-count surveys.  
-We explore model behavior through a series of simulations testing: (i) how well constant-detection models fit non-constant detection data and vice versa, (ii) the utility of a partial finite-mixtures to model detection heterogeneity involving a non-constant detection subgroup, and (iii) model performance with covariates and random effects affecting both abundance and detection.
+We explore model behavior through a series of simulations testing: (i) how well constant-detection models fit non-constant detection data and vice versa, (ii) the utility of finite-mixtures to model detection heterogeneity involving a non-constant detection subgroup, and (iii) model performance with covariates and random effects affecting both abundance and detection.
 Finally, we apply our model to Ovenbird data from the Minnesota Forest Breeding Bird Project.
@@
 n_{s}^{(obs)} &\sim \mbox{Binomial}\left(N_{s}, p_{s}^{(det)}\right)\\
-\label{eq:Tsrq}T_{sq} &\sim f_{T|det}(t|t_{sq}<C_I)
+\label{eq:Tsrq}T_{sq}|det &\sim f_{T|det}(t|t_{sq}<C_I)
 \end{align}
@@
 We can model detection heterogeneity by defining $F_T(t)$ as a mixture of TTDD components.  
-In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider partial finite-mixtures of a point mass during the first survey interval plus a continuous component from an exponential, gamma, Weibull, or lognormal distribution.
-We define a constant mixing parameter $\gamma \in [0,1]$ for the proportion of animals not distributed according to the point mass.
+In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider finite-mixtures of a point mass during the first survey interval plus a continuous component from an exponential, gamma, Weibull, or lognormal distribution.
 
@@
 
-Combining the distribution and expectation function for $T_{s}$, we can describe a partial-mixture TTDD cdf as:
-\[F_T(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})\]
-where $F_{T1}$ is the cdf of the continuous component and $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$.  
+Combining the distribution and expectation function for $T_{s}$, we can describe a mixture TTDD cdf as:
+\[F_T(t|\varphi_s, \boldsymbol{\theta}, \gamma) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\varphi_s, \boldsymbol{\theta})\]
+where $F_{T1}$ is the cdf of the continuous component, $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$, and $\gamma \in [0,1]$ is the proportion of animals not distributed according to the point mass.
 Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.
@@
 We conducted three simulation studies to explore the behavior of models with non-constant TTDDs.
-In Sim1, we compared models with detection heterogeneity via a partial finite-mixture TTDD versus models without detection heterogeneity.
+In Sim1, we compared models with detection heterogeneity via a finite-mixture TTDD versus models without detection heterogeneity.
 We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
@@
 
-The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a partial finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with partial finite-mixtures involving a time-varying TTDD.
+The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with finite-mixtures involving a time-varying TTDD.
 Simulation results demonstrate that the exponential mixture model fits exponential mixture data well, but it does not fit time-varying mixture data well.
@@
 We can modify our model to account for distance.  
-To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify Equation \ref{eq:pdet} to reflect the detection distance to individual $q$: $p_{sq}^{(det)} = g(d_q) F_T^A(C_I|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.
+To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify the equation for detection probability to reflect the detection distance to individual $q$: $p_{sq}^{(det)} = g(d_q) F_T^A(C_I|\varphi_s, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.
 
@@
 Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  
-In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d/`8-_q)$.
+In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d_q)$.
 So, $h_{sq}(t) = g(d_q) h_{s}^A(t)$.  
-From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\textbf{X}_{s}^D, \textbf{Z}_{s}^D, \boldsymbol{\theta}) \right]^{g(d_q)}$.
+From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\varphi_s, \boldsymbol{\theta}) \right]^{g(d_q)}$.
 

### 2016-05-12 Draft.tex
A few revisions to introduction for readability and flow.  A few comments to self added elsewhere.

@@
 
-\adam{WRITE A SPELLBINDING OPENING PARAGRAPH}\jarad{no need to aim so high}
+Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
+Removal sampling provide adapted to point counts provides one established methodology for estimating detection probabilities.
+% cost-effective in that it requires neither multiple observers nor multiple site visits.
+Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
+Assuming that individual capture rates remain constant over all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time .
+\citet{Farnsworth2002} adapted removal sampling to point-count surveys, suggesting that if: (i) observers record only the first time-to-detection for each animal, and (ii) times to first detection are interval-censored, then data will be similar in for to traditional removal sampled data.
+In the analysis of such data, the observation period is subdivided into unit intervals of equal duration, each of which is equivalent to a trapping session.
 
@@
 
-Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.  
-Assuming that the probability of capturing a not-yet-trapped individual is the same for all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time \citep{Moran1951, Zippin1958, Seber1982}.
-\citet{Farnsworth2002} adapted removal sampling to avian point-count surveys, suggesting that observers record just the first time-to-detection for each bird and mentally remove it thereafter.  
-Analysis proceeds by subdividing the observation period into intervals of equal duration and assuming that the rate of detection is the same at all times (i.e., following a Poisson process); then, each interval can be treated as a trapping event, and the interval-specific probability of detection can be estimated as with conventional removal sampling.
-
-However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}.  
-Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.  
+However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}. \adam{Others, like Mantyniemi.}  
+Animal behaviors such as intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011, Borcherspre-print}.  
 The act of observation may affect animal behavior.  
-Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  
-Survey methodology itself can affect detection rates.  
+The presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  
+Survey methodology itself can affect recording of detection times.  
+Observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
 Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
-And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
 In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.
@@
 Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014}.
+
+%Application of this methodology generally assumes that organisms at a survey site are detected at a constant rate \citep{Farnsworth2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}; however, this assumption is often not justified in the field.
+%Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.
 
@@
 Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.
+% "It is tempting to choose classes of models that yield narrower confidence intervals for N; however, Coull's (1997) extensive simulation studies (summarized by Coull and Agresti (1999)) clearly reveal that this behavior can produce erroneous inferences." -- Dorazio and Royle 2003
 
@@
 I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  
+Ditto Amundson.
 See also McCallum 2005.}.  

### 2016-06-16 Draft.tex
Edit abstract

@@
 The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption is often not justified.
-Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.  
-Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  
-Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.
+%Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.  
+%Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  
+We consider a class of N-mixture models that allow for detection heterogeniety through a mixture component, flexible two-parameter time-to-detection distributions (TTDDs), and allows for fixed and random effects for both abundance and detection.
+%Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution (TTDD) embedded within a hierarchical N-mixture framework.
 Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
-We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  
-Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.
-Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.
+We find that TTDDs that include a mixture component resulting in increased probability of detection in the initial observation period are necessary when the data have a chance of arising from distributions of this nature.
+In addition, the flexibility offered by two-parameter distributions, e.g. gamma, can outperform exponential based models when the truth is exponential or non-exponential.
+Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection and demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs and that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
+
+% We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  
+% Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.
+% Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.
 

### 2016-06-16 Draft.tex
Rewrite part of the methods section
Very much a work in progress

@@
 
+\newpage
 \section{Introduction}\label{sec:intro}
@@
 Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
-Removal sampling provide adapted to point counts provides one established methodology for estimating detection probabilities.
+Removal sampling, where individuals are solely counted on their first capture, provides one established methodology for estimating detection probabilities \citep{Farnsworth2002}.
 % cost-effective in that it requires neither multiple observers nor multiple site visits.
-Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
-Assuming that individual capture rates remain constant over all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time .
-\citet{Farnsworth2002} adapted removal sampling to point-count surveys, suggesting that if: (i) observers record only the first time-to-detection for each animal, and (ii) times to first detection are interval-censored, then data will be similar in for to traditional removal sampled data.
-In the analysis of such data, the observation period is subdivided into unit intervals of equal duration, each of which is equivalent to a trapping session.
-
+%Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
+%Assuming that individual capture rates remain constant over all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time .
+%\citet{Farnsworth2002} adapted removal sampling to point-count surveys, suggesting that if: (i) observers record only the first time-to-detection for each animal, and (ii) times to first detection are interval-censored, then data will be similar in for to traditional removal sampled data.
+%In the analysis of such data, the observation period is subdivided into unit intervals of equal duration, each of which is equivalent to a trapping session.
 %Farnsworth's removal assumptions:\\
@@
 % Johnson2008: all birds present are available for detection
+A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007}.
+%\adam{Others, like Mantyniemi.}
+In particular, animal behaviors, e.g.
+intermittent singing in birds and frogswhales \citep{Scott2005, Diefenbach2007, Reidy2011}, observer behaviors, e.g.
+pre-counting and disturbing species \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007, LeeMarsden2008, Johnson2008}, and survey methodology, e.g.
+lack of settling in period \jarad{what else were you thinking?
+ref?}, can all lead to detection heterogeneity \citep{Farnsworth2005}.
+%Survey methodology itself can affect recording of detection times.  
+% Observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
+% Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
+%In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time .
 
-However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}. \adam{Others, like Mantyniemi.}  
-Animal behaviors such as intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011, Borcherspre-print}.  
-The act of observation may affect animal behavior.  
-The presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  
-Survey methodology itself can affect recording of detection times.  
-Observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
-Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
-In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.
+A common assumption to deal with some of the issues above is to increase the probability of detection in the initial observation period \citep{Etterson2009} \jarad{Adam: add other refs}, although this is not yet the standard \citep{Amundson2014} \jarad{Adam: add other refs}.
+In this manuscript, we consider the choice of whether to include a mixture component, but we also consider a broader class of non-constant distributions.
+%In this study, we develop a model for scenarios where detection rates are not constant. 
+In particular, we consider time-to-detection as done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function $F_T(t)$.
+We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).
 
-In this study, we develop a model for scenarios where detection rates are not constant.  
-We model time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function $F_T(t)$.
-We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
-The instantaneous detection rate or hazard function $h(t)$ can be found from the TTDD as $h(t) = f_T(t) / (1-F_T(t))$.
-
-Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown.  
+Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown and may be the primary quantity of interest.  
 We embed the TTDD in a hierarchical framework for multinomial counts using an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.  
-For our purposes, the N-mixture framework provides three clear benefits.  
-First, its multinomial data framework accords with the interval censored data collection that is customary in point-count surveys \citep{Ralph1995}.  
-Second, the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}.  
-Third, for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).  
+For our purposes, the N-mixture framework provides three clear benefits: 1) its multinomial data framework accords with the interval censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
 The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.  
@@
 First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}, whereas most existing studies instead construct regression models on the logit of the equal-interval detection probability \adam{Amundson2014 and others?}.
-Such an approach is not possible when detection rates are not constant.  
+The later is not possible when detection rates are not constant.  
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
 Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.
+% 
+% \adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.  
+% In the fisheries world, models of time-varying catchability have been implemented -- see Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples.  
+% Also, Alldredge 2007 implements a complete-history of detection model (technically mark-recapture, not removal) that estimates interval-specific detection probabilities (complete with covariates and detection heterogeneity).
+% However, it cannot be applied in a removal context, because removal data contains only first times to detection, and the Alldredge model requires subsequent observations -- removal models compensate by specifying a TTDD, thus enabling us to extrapolate beyond the observation period.}
 
-\adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.  
-In the fisheries world, models of time-varying catchability have been implemented -- see Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples.  
-Also, Alldredge 2007 implements a complete-history of detection model (technically mark-recapture, not removal) that estimates interval-specific detection probabilities (complete with covariates and detection heterogeneity).
-However, it cannot be applied in a removal context, because removal data contains only first times to detection, and the Alldredge model requires subsequent observations -- removal models compensate by specifying a TTDD, thus enabling us to extrapolate beyond the observation period.}
-
-\jarad{The introduction needs to end earlier.
-It should be about 1-2 pages covering what has previously been done and outline our approach.
-So this should probably start a new section and before this there should be an outline of our manuscript.
-}
-
-\subsection{TTDDs as behavior and methodology}
-
-A TTDD defines the range of detection patterns possible in an analysis.  
-The most basic is an exponential TTDD, which imposes a constant first detection rate.  
-Choosing another TTDD can allow for a systematic non-constant detection regime.  
-For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
-
-Increasing hazard functions result in a pdf for $T$ that has a mode away from zero; we term these TTDDs `peaked'.  
-We contrast these to decreasing hazard functions with a mode at zero, which we term `nonpeaked'.  
-For our purposes, the exponential TTDD is neither peaked nor nonpeaked.
-
-We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.  
-We reserve `detection heterogeneity' for differences across population subgroups.  
-It is often modeled using a mixture of detection distributions.  
-\citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
- Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
-We use `non-constant detection' when detection rates for an organism change with time.  
-Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
-
-\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
+\jarad{Need a roadmap. Provide section number in the paragraph below.}
 
@@
 
-\section{Methods}\label{sec:model}
+\section{Interval-censored avian point counts}\label{sec:data}
 
-\subsection{Data}\label{sec:data}
+Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citet{Hanowski1995}.  
+Survey sites were randomly selected from sawtimber red pine stands with no recent logging activity such that each stand had up to three sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
+The data include 65 sites and a total of 381 with site specific variables including site age,  stock density, an indicator of select-/partial-cut logging during the 1990s.
 
-Our analysis is motivated by point-count surveys conducted in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB).  
-The MNFB design is described in \citet{Hanowski1995}, and we summarize only essential details here.  
-Single-visit time-to-first-detection point-count surveys were conducted by trained observers at a fixed set of sampling locations once annually.  
-Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals -- a two-minute interval followed by eight one-minute intervals.  
-These relatively fine-scale intervals should provide more information about TTDDs than the three-interval methodology (0-3, 3-5, and 5-10 minutes) used in many avian point count surveys \citep{Ralph1995}.
-There was no prescribed settling-in period between the observer's arrival and the time they began observing.
+%with a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
 
-Our model is formulated for single-visit removal-sampled point-count surveys $s=1,\dotso,S$.  
-We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  
-We use $n_{s}^{(obs)}$ to denote the number of individuals observed during a survey.  
-For exact data, times to detection are $t_{sq}$, where $q = 1,\dotso,n_{s}^{(obs)}$ indexes observed individuals at a survey.  
-For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  
-The number of animals counted during any interval is $n_{si}$ such that $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and the vector of counts across all intervals at a survey is written $\vn_{s}$.
+\jarad{Adam: How many observers}
 
-\subsection{Framework}
+%The MNFB design is described in , and we summarize only essential details here.  
+Single-visit (per year) point-count surveys were conducted by trained observers at each site once annually (weather permitting).  
+Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
+During each survey, the Julian date, time of day, and temperature were recorded.
 
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and observed times to detection, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued count distribution.
-Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
+\section{Time-to-detection N-mixture models}
 
-\subsection{Data Model}
-We assume that counts from different surveys are conditionally independent.  
-We define $T_{sq}$ as a random variable for the first time at which an individual is detected.
-We assume that detection times for individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$, which may depend on covariates.  
-The probability of detection for an animal is thus $p_{s}^{(det)} = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) \textbf{I}(t<C_I) / F_T(C_I)$.
-Denoting $N_{s}$ as the total number animals present at a survey, the data model is:
-\begin{align}\label{eq:datamodel}
-n_{s}^{(obs)} &\sim \mbox{Binomial}\left(N_{s}, p_{s}^{(det)}\right)\\
-\label{eq:Tsrq}T_{sq}|det &\sim f_{T|det}(t|t_{sq}<C_I)
-\end{align}
-When detection times are censored, we replace Eq.
-\ref{eq:Tsrq} with $\vn_{s} \sim \mbox{Multinomial}\left(n_{s}^{(obs)}, \textbf{p}_{s}\right)$, where $\textbf{p}_{s}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{sq}<C_I)$ at the interval cutoff times.
-It follows that $\sum_{i=1}^I p_{si} = 1$.
+Before considering the interval censoring and the explanatory variables, we first consider the scenario of exact time to detections with no explanatory variables. 
+We then incorporate interval censoring and follow with inclusion of fixed and random effects for abundance and detection.
 
-\subsection{Detection Model}\label{sec:detectionmodel}
+\subsection{Exact time to detection}
+
+Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
+Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function $F_T(t)$ and probability density function $f_T(t)$. 
+If the times to first detection are censored at $C$, due a finite survey length of $C$, the probability of detection for each individual is $p^{(det)}=F_T(C)$ and the conditional distribution of observed detection times is $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$.
+Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, p^{(det)}\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
+where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
+The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
+
+A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e.
+$T_{sb}\stackrel{ind}{\sim} \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e.
+$h(t) = \lambda$.
+Choosing another TTDD can allow for a systematic non-constant detection regime. 
+For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
+In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis.
+
+With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $p^{(det)}$. 
+With an estimate and uncertainty for $p^{(det)}$, we are in a scenario of a binomial model with unknown $N$ and ``known'' $p$ and thus can estimate $N$.
+In these models, estimates of $N_s$ are unstable unless $p^{(det)}>0.5$.
+One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} Po(\lambda)$. 
+Although alternative distributions could be considered, e.g.
+negative binomial, our experience with avian point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here.
+\jarad{Adam: add refs$.}
+
+\subsection{Interval-censored times to detection}
+
+Due to the harried process of avian point counts, times to first detection are typically not recorded exactly, but are instead censored into $I$ intervals. 
+Let $C_i$ for $i=1,\dots,I$ indicate the right endpoint of the $i$th interval then $C_I$ is the total survey duration and, letting $C_0=0$, the $i$th interval is $(C_{i-1},C_{i}]$. 
+Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \sim \mbox{Multinomial}\left(n_{s}^{(obs)}, \textbf{p}_{s}\right)$, where $\textbf{p}_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
+
+A common observation in avian point counts are increased observations in the first interval relative to an exponential distribution.
+To accommodate this empirical observation, many models include a term to increase the probability of observing 
+\jarad{Adam: add refs.}
+
+\subsection{Incorporating detection explanatory variables}\label{sec:detectionmodel}
 
@@
 
+We distinguish two categories of TTDDs: peaked and nonpeaked. 
+A peaked TTDD has an increasing hazard function and therefore a pdf whose mode is not at zero. 
+A nonpeaked TTDD has a decreasing hazard function and therefore a pdf whose mode is at zero.
+Thus, the exponential TTDD is neither peaked nor nonpeaked.
+
+We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.  
+We reserve `detection heterogeneity' for differences across population subgroups.  
+It is often modeled using a mixture of detection distributions.  
+\citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
+ Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
+We use `non-constant detection' when detection rates for an organism change with time.  
+Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
+
+\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
+
+\section{Methods}\label{sec:model}
+
+\subsection{Framework}
+
+We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and observed times to detection, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued count distribution.
+Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
+
+\subsection{Data Model}
+We assume that counts from different surveys are conditionally independent.  
+We define $T_{sq}$ as a random variable for the first time at which an individual is detected.
+We assume that detection times for individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$, which may depend on covariates.  
+The probability of detection for an animal is thus $p_{s}^{(det)} = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) \textbf{I}(t<C_I) / F_T(C_I)$.
+
+When detection times are censored, we replace equation \ref{eq:Tsrq} with 
+It follows that $\sum_{i=1}^I p_{si} = 1$.
+
 \subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
+
 To demonstrate our model, we fit counts of Ovenbirds (\textit{Seiurus aurocapilla}) from the MNFB.

### 2016-06-17 Draft.tex
Fix capitalization in OVEN_*P*
closes #3
Edit through methods
although I have ignored DIC and posterior predictive checks which I think we ought to remove

@@
 \newcommand{\vn}{\textbf{n}}
+\newcommand{\vp}{\textbf{p}}
+\newcommand{\vX}{\textbf{X}}
+\newcommand{\vZ}{\textbf{Z}}
+\newcommand{\vbeta}{\boldsymbol{\beta}}
+\newcommand{\vxi}{\boldsymbol{\xi}}
+
+\newcommand{\Exp}{\mbox{Exp}}
+\newcommand{\Ga}{\mbox{Ga}}
+\newcommand{\We}{\mbox{We}}
+\newcommand{\LN}{\mbox{LN}}
+\newcommand{\Po}{\mbox{Po}}
+\newcommand{\Mult}{\mbox{Mult}}
 
@@
 
+\jarad{Put information about Ovenbird here?}
+
+a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
+
 \section{Time-to-detection N-mixture models}
 
-Before considering the interval censoring and the explanatory variables, we first consider the scenario of exact time to detections with no explanatory variables. 
+Before considering interval censoring and explanatory variables, we first consider the scenario of exact time to detections with no explanatory variables. 
 We then incorporate interval censoring and follow with inclusion of fixed and random effects for abundance and detection.
@@
 Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
-Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function $F_T(t)$ and probability density function $f_T(t)$. 
-If the times to first detection are censored at $C$, due a finite survey length of $C$, the probability of detection for each individual is $p^{(det)}=F_T(C)$ and the conditional distribution of observed detection times is $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$.
+Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$. 
+If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $p^{(det)}=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
 Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, p^{(det)}\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
@@
 In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
+For consistency that will be clear when including explanatory variables, we use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.
 
-With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $p^{(det)}$. 
-With an estimate and uncertainty for $p^{(det)}$, we are in a scenario of a binomial model with unknown $N$ and ``known'' $p$ and thus can estimate $N$.
+With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $p^{(det)}$ and its uncertainty. 
+With an estimate and uncertainty for $p^{(det)}$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
 In these models, estimates of $N_s$ are unstable unless $p^{(det)}>0.5$.
-One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} Po(\lambda)$. 
+One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} \Po(\lambda)$. 
+We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \Po\left(\lambda(1-p^{(det)})\right)$.
 Although alternative distributions could be considered, e.g.
 negative binomial, our experience with avian point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here.
-\jarad{Adam: add refs$.}
+\jarad{Adam: add refs.}
 
@@
 Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
-Assuming independence amongst individuals and sites, we have $\vn_{s} \sim \mbox{Multinomial}\left(n_{s}^{(obs)}, \textbf{p}_{s}\right)$, where $\textbf{p}_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \sim \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
 
 A common observation in avian point counts are increased observations in the first interval relative to an exponential distribution.
-To accommodate this empirical observation, many models include a term to increase the probability of observing 
-\jarad{Adam: add refs.}
+To accommodate this empirical observation, many models include a parameter to increase the probability of observing individuals in the first interval. \jarad{Adam: add refs.}
+We augment our model with a mixing parameter $\gamma\in[0,1]$ which updates the following definitions
+$f_{T|det}(t) = f_T(t)/[\gamma+F_T(C)]$\ and $p_{s1} = \gamma + F_{T|det}(C_1)$.
+When including this parameter and these modifications, we refer to the model as a ``mixture'' model. 
+If $\gamma=0$, then the non-mixture model is recovered.
 
-\subsection{Incorporating detection explanatory variables}\label{sec:detectionmodel}
+\subsection{Incorporating explanatory variables}\label{sec:covariates}
 
-The two main elements of the detection model for $T_{sq}$ are: (i) the distributional form and (ii) the expectation function.
-$T_{sq}$ should be based on non-negative continuous distributions, preferrably featuring a rate parameter than can be estimated via regression.  
-In this study, we adopt TTDDs for $T_{sq}$ using the following: Exponential($\varphi_{s}$), Gamma($\alpha, \varphi_{s}$), Weibull($k, \varphi_{s}$), and Lognormal($\mu_{s}, \sigma_{det}^2$).  
-Here, $\varphi_{s}$ and $\mu_{s}$ are survey-specific detection rate parameters, and we consider the shape parameters $\alpha$, $k$, and $\sigma_{det}^2$ to be constant across all surveys.
+As discussed in Section \ref{sec:data}, explanatory variables are available for sites and for surveys. 
+Generally, we suspect that site variables, e.g. habitat, will affect abundance and survey variables, e.g. time of day, will affect detection probability. 
+Thus, we allow for incorporating explanatory variables on both the abundance and detection.
 
-We can model detection heterogeneity by defining $F_T(t)$ as a mixture of TTDD components.  
-In this study, we use interval-censored observations, so in addition to the four above TTDDs (with no mixture component), we consider finite-mixtures of a point mass during the first survey interval plus a continuous component from an exponential, gamma, Weibull, or lognormal distribution.
+To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ via a mixed effects linear regression with a log link, i.e.
+$\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.
 
-We model the expectation function via mixed effects linear regression.  
-For exponential, gamma, and Weibull TTDDs, we model: $\log(\varphi_{s}) = \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D$, where $\textbf{X}_{s}^D$ are covariates, $\boldsymbol{\beta}^D$ is a vector of fixed effects, $\textbf{Z}_{s}^D$ specifies random effect levels, $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, and $\sigma_{D[i]}^2$ denotes the variance associated with $\xi_i^D$.
-This regression is equivalent to modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  
-For the lognormal distribution, we fit the same regression model to $-\mu_{s}$.  
+To incorporate explanatory variables on detection probability, we let the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
+Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \overset{ind}{\sim} N(0,\\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
+
+\subsection{Estimation}
+
+For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicity. 
+\begin{subequations}
+\begin{eqnarray} \label{eq:model}
+n_s^{(obs)} &\sim& Po(\lambda_s) \\
+\vn_s &\sim& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
+p_{s1} &=& \gamma + F_{T|det}(C_1|\alpha,\varphi_s),\, p_{si} = F_{T|det}(C_i|\alpha,\varphi_s)-F_{T|det}(C_{i-1}|\alpha,\varphi_s)\\
+\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2) \\
+\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \overset{ind}{\sim} N(0,\tau_{D[j]}^2)
+\end{eqnarray}
+\end{subequations}
+
+\jarad{May need to remove subnumbers, but I'll wait to see if we need them.}
+
+We adopt a Bayesian approach and therefore require a prior over the model parameters.
+To ease construction of a default prior for this model, we standardize all explanatory variables and then construct priors to be diffuse within a reasonable range of values.
+Prior mean and standard deviation (sd) for the abundance intercept was set at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
+Prior mean and sd for the detection intercept were chosen so that, based on an intercept-only non-mixture model with $\alpha=1$: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
+Normal priors for fixed effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
+All standard deviations and $\alpha$ were given half-Cauchy priors with location 0 and scale 1 for the untruncated Cauchy and Unif(0,1) prior on the mixture parameter $\gamma$ in the mixture models.
+All scalar parameters were assumed independent \emph{a priori}.
+
+\jarad{Adam: I appreciate that you provided the justification for the priors, but we should also include the exact vaules where possible.}
+
+We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
+We discarded half of the iterations as warmup and then thinned by 10.  
+We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}. \jarad{Adam: what did you do with the Geweke diagnostics?}  
+We reran models if the effective sample size for any parameter was below 1000.  
+The number of iterations used depended on the model and is detailed later. 
+For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed \jarad{Adam: how did they fail?} unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
+\jarad{Adam: this doesn't make sense.
+Why would the means from these preliminary short-run models be any better than letting these preliminary short-run models continue?
+You claim seems to be that letting these preliminary short-runs continue would mean that those models would fail.}
+\jarad{Adam: if you are using Geweke diagnostics, then it would be better for your to start from somewhere "good". For fixed effects, you could start at 0, mixing parameter at 0.5, etc.}
+
+\jarad{Adam: we should include the Stan models in the supplement.}
+
 %Thus, the link functions in our detection model are:
 %\begin{alignat}{3}
-%&\text{Exponential:\;} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}])\\
-%&\text{Gamma:} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \log(\alpha)\\
-%&\text{Weibull:}  &&\log(\varphi_{s}) &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
-%&\text{Lognormal:} &&-\mu_{s} &&= \text{Intercept}^D + \textbf{X}_{s}^D\boldsymbol{\beta}^D + \textbf{Z}_{s}^D\boldsymbol{\xi}^D = -\log(E[T_{sq}]) + \sigma_{det}^2/2
+%&\text{Exponential:\;} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}])\\
+%&\text{Gamma:} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \log(\alpha)\\
+%&\text{Weibull:}  &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
+%&\text{Lognormal:} &&-\mu_{s} &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \sigma_{det}^2/2
 %\end{alignat}
 
-Combining the distribution and expectation function for $T_{s}$, we can describe a mixture TTDD cdf as:
-\[F_T(t|\varphi_s, \boldsymbol{\theta}, \gamma) = (1-\gamma)\mathbb{I}(t>0) + \gamma F_{T1}(t|\varphi_s, \boldsymbol{\theta})\]
-where $F_{T1}$ is the cdf of the continuous component, $\boldsymbol{\theta}$ is a vector of parameters appropriate to $F_{T1}$, and $\gamma \in [0,1]$ is the proportion of animals not distributed according to the point mass.
-Note that parameters in a mixture TTDD do not share the exact same interpretation as equivalent parameters from a non-mixture TTDD, since they only pertain to the $\gamma$ proportion of individuals that follow $F_{T1}$.
+%\subsection{Abundance Model}
 
-\subsection{Abundance Model}
+% We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.  
+% We reserve `detection heterogeneity' for differences across population subgroups.  
+% It is often modeled using a mixture of detection distributions.  
+% \citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
+%  Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
+% We use `non-constant detection' when detection rates for an organism change with time.  
+% Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
 
-For simplicity, we model total abundance at each survey $N_{s}$ using a Poisson distribution, though other non-negative integer-valued distributions may be used.  
-We model the expected survey abundance $\lambda_{s}$ via mixed effects linear regression with a log link: $\log (\lambda_{s}) = \textbf{X}_{s}^A\boldsymbol{\beta}^A + \textbf{Z}_{s}^A\boldsymbol{\xi}^A$.
-As with the detection model, $\textbf{X}_{s}^A$ are covariates, $\boldsymbol{\beta}^A$ is a vector of fixed effects, $\textbf{Z}_{s}^A$ specifies random effect levels, and $\xi_i^A \overset{ind}{\sim} N(0,\sigma_{A[i]}^2)$ are random effects.
-We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \mbox{Poisson}\left(\lambda_{s}(1-p_{s}^{(det)})\right)$.
+%\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
 
 We distinguish two categories of TTDDs: peaked and nonpeaked.
-A peaked TTDD has an increasing hazard function and therefore a pdf whose mode is not at zero. 
-A nonpeaked TTDD has a decreasing hazard function and therefore a pdf whose mode is at zero.
-Thus, the exponential TTDD is neither peaked nor nonpeaked.
-
-We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.  
-We reserve `detection heterogeneity' for differences across population subgroups.  
-It is often modeled using a mixture of detection distributions.  
-\citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
- Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
-We use `non-constant detection' when detection rates for an organism change with time.  
-Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
-
-\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
-
-\section{Methods}\label{sec:model}
-
-\subsection{Framework}
-
-We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a data model incorporating binomial detection and observed times to detection, (b) a detection model based upon a TTDD, and (c) an abundance model based upon an integer-valued count distribution.
-Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.
-
-\subsection{Data Model}
-We assume that counts from different surveys are conditionally independent.  
-We define $T_{sq}$ as a random variable for the first time at which an individual is detected.
-We assume that detection times for individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$, which may depend on covariates.  
-The probability of detection for an animal is thus $p_{s}^{(det)} = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{sq}<C_I) = f_T(t) \textbf{I}(t<C_I) / F_T(C_I)$.
-
-When detection times are censored, we replace equation \ref{eq:Tsrq} with 
-It follows that $\sum_{i=1}^I p_{si} = 1$.
-
-\subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
-
-To demonstrate our model, we fit counts of Ovenbirds (\textit{Seiurus aurocapilla}) from the MNFB.
-We restricted survey sites in our analysis to those of the most abundant habitat type (sawtimber red pine); in so doing, we focused modeling efforts upon the detection model and skirted the complication of modeling abundance across habitat types.
-We further constrained analysis to 65 sites with a site origin year between 1870-1970.  
-In total, we used 381 site-year surveys in our analysis with a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
-
-For the abundance half of our model, we used four covariates plus two random effects.  
-The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
-We associated random effects with each survey year and each stand, with most stands consisting of three survey sites.  
-For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.
-69\% of surveys in our dataset involved observers in their first year at MNFB.  
-Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  
-We associated random effects with each observer.  
-We centered and standardized all continuous covariates prior to fitting models.
-
-We selected priors to be non-informative within a reasonable range of values.  
-Normal priors for the detection model intercept term, Intercept$^D$, were chosen so that, based on an intercept-only non-mixture model with scale parameter of one: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
-Normal priors for Intercept$^A$ were centered at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
-Normal priors for effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
-For all shape and variance parameters, we chose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a Uniform(0,1) prior.  
-All priors were independent.
+\jarad{Adam: I don't like the previous definition because log-normal never has a peak at zero.
+In addition, if the peak occurs before $C_1$ then we really aren't interested in it.}
+A peaked TTDD has a mode greater than $C_1$ while a non-peaked TTDD has a mode less than $C_1$. \jarad{Adam: do we want a figure demonstrating this point?}
 
@@
 \subsection{Model Fitting}
-
-We fit the models by MCMC sampling in the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
-For intercept-only simulations, we instantiated a single chain running 60,000 iterations; for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.  
-We discarded half of the iterations as warmup and then thinned by 1 in 10.  
-We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}.  
-We reran models if the effective sample size for any parameter was below 1000.  
-For most models, we accepted Stan defaults for dispersed initial values; however, gamma and Weibull models sometimes failed unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
 
@@
 \section{Simulation studies}
+
+For intercept-only simulations, we instantiated a single chain running 60,000 iterations; for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.
 
@@
 
+\subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
+
+For the abundance half of our model, we used four covariates plus two random effects.  
+The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
+We associated random effects with each survey year and each stand, with most stands consisting of three survey sites.  
+For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.
+69\% of surveys in our dataset involved observers in their first year at MNFB.  
+Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  
+We associated random effects with each observer.  
+We centered and standardized all continuous covariates prior to fitting models.
+
 We fit mixture and non-mixture models from each family of TTDD to the Ovenbird dataset.  
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.88\textwidth]{OVEN/oven_sum/OVEN_posteriors.pdf}
+\includegraphics[width=0.88\textwidth]{OVEN/oven_sum/OVEN_Posteriors.pdf}
 \caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to the Ovenbird data.  

### 2016-06-17 Draft.tex
Edit simulation study section
work in progress

@@
 
-\subsection{Model Fitting}
+\subsection{Model diagnostics}
 
 To assess goodness of fit, we relied on posterior predictive checks \citep{Gelman1996}.  
-Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  
+%Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  
 In particular, we are concerned with the goodness of fit for the TTDD distribution, so we defined a check statistic for each observation interval:
@@
 
-For intercept-only simulations, we instantiated a single chain running 60,000 iterations; for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.
+We conducted three simulation studies to explore the behavior of models with non-constant TTDDs.
+The first study compares mixture vs non-mixture models.
+The second study compares the time to detection distribution families.
+In the first two studies, we utilized intercept only models to focus attention on the time to detection distribution. 
+For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families.
 
-We conducted three simulation studies to explore the behavior of models with non-constant TTDDs.
-In Sim1, we compared models with detection heterogeneity via a finite-mixture TTDD versus models without detection heterogeneity.
-We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
-Because model behavior may differ between peaked and nonpeaked datasets, we simulated both for each TTDD family except exponential.  
-Thus, combining mixtures, `peakedness' and TTDD families, we simulated datasets from 14 different TTDDs in total.  
-In Sim2, we examined the impact of misspecifying the $F_{T1}$ family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
-Of particular interest were scenarios with a time-varying $F_{T1}$ component in the data but a time-constant component in the inference model, and vice versa.
-Based on findings from Sim1, we considered only mixture TTDD datasets, and we fit each with mixture TTDD models from all four families.  
-In Sim3, we modeled a more complete scenario where both data generation and model fitting included covariates and random effects for abundance and detection.
-
-We used posterior median estimates from the Ovenbird analysis to inform simulation parameters except as follows.  
-For Sim1 and Sim2, we chose true parameter values so that: (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-For Sim3, all parameters used for nonpeaked datasets came from the Ovenbird analysis.
-In order to simulate peaked dataset variants, we modified the Ovenbird parameters by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, resulting on average in a detection mode near 5 minutes and 75-85\% overall detection across all sites.
-Unlike Sim1 and Sim2, expected detection probabilities in Sim3 were not the same for all datasets.
-Counts were censored into nine-intervals as with the Ovenbird dataset.  
-For both Sim1 and Sim2, we simulated 16 replicates, but for Sim3 we generated only a single replicate, because gamma models took an average of 5 days to fit.
+% In Sim1, we compared models with detection heterogeneity via a finite-mixture TTDD versus models without detection heterogeneity.
+% We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
+% Because model behavior may differ between peaked and nonpeaked datasets, we simulated both for each TTDD family except exponential.  
+% Thus, combining mixtures, `peakedness' and TTDD families, we simulated datasets from 14 different TTDDs in total.  
+% In Sim2, we examined the impact of misspecifying the $F_{T1}$ family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
+% Of particular interest were scenarios with a time-varying $F_{T1}$ component in the data but a time-constant component in the inference model, and vice versa.
+% Based on findings from Sim1, we considered only mixture TTDD datasets, and we fit each with mixture TTDD models from all four families.  
+% In Sim3, we modeled a more complete scenario where both data generation and model fitting included covariates and random effects for abundance and detection.  
+% 
+% We used posterior median estimates from the Ovenbird analysis to inform simulation parameters except as follows.  
+% For Sim1 and Sim2,   
+% For Sim3, all parameters used for nonpeaked datasets came from the Ovenbird analysis.
+% In order to simulate peaked dataset variants, we modified the Ovenbird parameters by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, resulting on average in a detection mode near 5 minutes and 75-85\% overall detection across all sites.
+% Unlike Sim1 and Sim2, expected detection probabilities in Sim3 were not the same for all datasets.
+% Counts were censored into nine-intervals as with the Ovenbird dataset.  
+% For both Sim1 and Sim2, we simulated 16 replicates, but for Sim3 we generated only a single replicate, because gamma models took an average of 5 days to fit.
 
@@
 
-In Sim1, we simulated intercept-only datasets from 14 TTDDs and fit each with two models from the correct TTDD family: one allowing for detection heterogeneity in the form of a mixture component and one not (Table \ref{tbl:sim1}).
+In Sim1, we simulated 16 intercept-only datasets from 14 TTDDs. 
+We chose true parameter values so that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+\jarad{Adam: we need specific values here, perhaps in the supplement.}
+
+Each dataset was fit with two models from the correct TTDD family: mixture and non-mixture. 
+Using Stan, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+
 When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $p^{(det)}$ badly in every case.  
@@
 \section{Ovenbird analysis}\label{sec:ovenbirds}
+
+for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.
 

### 2016-06-17 Draft.tex
Update Sim2 discussion

@@
 \newcommand{\Mult}{\mbox{Mult}}
+
+\newcommand{\pdet}{p^{(det)}}
 
@@
 Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$. 
-If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $p^{(det)}=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
-Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, p^{(det)}\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
+If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
+Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
 where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
@@
 
-With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $p^{(det)}$ and its uncertainty. 
-With an estimate and uncertainty for $p^{(det)}$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
-In these models, estimates of $N_s$ are unstable unless $p^{(det)}>0.5$.
+With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
+With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
+In these models, estimates of $N_s$ are unstable unless $\pdet>0.5$.
 One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} \Po(\lambda)$. 
-We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \Po\left(\lambda(1-p^{(det)})\right)$.
+We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \Po\left(\lambda(1-\pdet)\right)$.
 Although alternative distributions could be considered, e.g.
@@
 
-We focus inference on the overall detection probability across all surveys: $p^{(det)} = \sum\limits_{s}n_{s}^{(obs)}\big/$ $\sum\limits_{s}N_{s}$.  
-Within simulation studies, we compared model estimates to true $p^{(det)}$ values by percent coverage and by posterior p-values --- the proportion of MCMC samples having a smaller value than the true value.
+We focus inference on the overall detection probability across all surveys: $\pdet = \sum\limits_{s}n_{s}^{(obs)}\big/$ $\sum\limits_{s}N_{s}$.  
+Within simulation studies, we compared model estimates to true $\pdet$ values by percent coverage and by posterior p-values --- the proportion of MCMC samples having a smaller value than the true value.
 
@@
 For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families. 
+In all simulation studies, we focus on accuracy in estimation of $\pdet$ which then translates into estimation of abundance.
 
@@
 
-In Sim1, we simulated 16 intercept-only datasets from 14 TTDDs. 
+In Sim1, we simulated 16 intercept-only datasets from each of 14 TTDDs. 
 We chose true parameter values so that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-\jarad{Adam: we need specific values here, perhaps in the supplement.}
+\jarad{Adam: we need specific values here for $\alpha$ and $\varphi$ (or $\beta_0^D$). If it is too much to include here put it in the supplement.}
 
-Each dataset was fit with two models from the correct TTDD family: mixture and non-mixture. 
-Using Stan, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+Each dataset was fit with two models: mixture and non-mixture version of the distribution family, e.g. exponential, used to simulate the data.
+We ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
 
-When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $p^{(det)}$ badly in every case.  
-This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
-Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
-Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $p^{(det)}$ than did the true non-mixture inference model.  
-The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $p^{(det)}$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
-Mixture model posterior medians for $p^{(det)}$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
+Table \ref{tbl:sim1} provides a summary of the mixture and non-mixture models ability to capture the detection probability. 
+When a non-mixture model is used to simulate the data (top half of the figure), there is no clearly discernable differences between the ability of a non-mixture or mixture model to capture.
+Similarly DIC cannot clearly determine that a non-mixture model should be preferred over a mixture model in this scenario.
 
-\adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
-Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $p^{(det)}$ above and shared the same coverage rates.  
-Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
-Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
-For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
-For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
+When a mixture model is used to simulate the data (top half of the figure), there is clearly a benefit to using a mixture model.
+For the non-mixture models, the credible interval coverage is near zero for most of the models with the exponential model overestimating $\pdet$ and the other models underestimating.
+While DIC could not easily differentiate a mixture from a non-mixture model for inference when a non-peaked simulation model was used, DIC clearly preferred mixture models when the data arose from an exponential distribution or any of the peaked distributions.
 
-Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
+These results support the use of a mixture model unless you are absolutely sure your data arose from a non-mixture distribution.
+
+% When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $\pdet$ badly in every case.  
+% This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
+% Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
+% Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $\pdet$ than did the true non-mixture inference model.  
+% The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $\pdet$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
+% Mixture model posterior medians for $\pdet$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
+
+% \adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
+% \jarad{We shouldn't be talking about results that we are not showing. The results should either be in the manuscript or the supplement.}
+% Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $\pdet$ above and shared the same coverage rates.  
+% Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
+% Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
+% For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
+% For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
+
+% Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
 DIC did reveal a preference for mixture models with exponential and peaked datasets.
@@
 \begin{table}[ht]\centering\small
-\begin{tabular}{l|l|l|l|cccc|cccc||c}
+\begin{tabular}{l|l|l|l|cccc|cccc||r}
  \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
@@
 \end{tabular}
-\caption{\label{tbl:sim1} Summary of mixture vs. non-mixture model fits (Sim1).  
+\caption{Summary of mixture vs. non-mixture model fits (Sim1).  
 In all cases, the inference model family matches the dataset family.  
-Med $p$: average across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.}
+Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
+\jarad{Adam: we should consider splitting this table in two.
+The main point is what happens when the data arise from a mixture model but a non-mixture distribution is used.
+So the bottom half could be included in the main body while the top half could be put in the supplement.}
+}
+\label{tbl:sim1}
 \end{table}
@@
 In Sim2, we simulated intercept-only datasets from the 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families (Table \ref{tbl:sim2}).
-Mixture exponential models returned more accurate estimates of $p^{(det)}$ than other models when data came from an exponential mixture, and 95\% credible intervals for $p^{(det)}$ were from one-third to one-half as wide.
+Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.
 However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.
@@
 \caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim2).  
-Med $p$: average across simulations of the posterior median of $p^{(det)}$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $p^{(det)}$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $p^{(det)}$ lies within the appropriate credible interval.  
+Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
 $\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
@@
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.  
+\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
 Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
@@
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $p^{(det)}$.
+\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
 All data and inference models include mixtures.
@@
 DIC preferred exponential mixtures to time-varying mixtures for exponential mixture data ($>6$); however, it also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.
-In this last case, the posterior 95\% credible interval for $p^{(det)}$ from the exponential mixture model did not contain the true value of $p^{(det)}$, while the true value was within a central 60\% credible interval for all three other mixture models.
+In this last case, the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
 In conjunction with Sim2 results, this last finding calls into question the utility of DIC for nonpeaked datasets.
@@
 
-Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $p^{(det)}$, and the log10 number of Ovenbirds uncounted.  
-Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $p^{(det)}$, and total abundance across models, Ovenbird data most likely arose from a exponential or nonpeaked mixture TTDD.
+Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $\pdet$, and the log10 number of Ovenbirds uncounted.  
+Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from a exponential or nonpeaked mixture TTDD.
 As in Sim3, abundance covariate coefficient estimates were virtually the same across all TTDD models.
@@
 Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
-The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  
+The probability of detection is thus formulated as availability times perceptibility: $\pdet = P_aP_b$.  
 Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.

### 2016-06-17 Draft.tex
Edit through Sim2

@@
 \newcommand{\pdet}{p^{(det)}}
+\newcommand{\ind}{\stackrel{ind}{\sim}}
 
@@
 
-\jarad{Need a roadmap. Provide section number in the paragraph below.}
-
-In this study, we construct a time-to-detection N-mixture model for estimating abundance from removal-sampled point-count surveys.  
-We explore model behavior through a series of simulations testing: (i) how well constant-detection models fit non-constant detection data and vice versa, (ii) the utility of finite-mixtures to model detection heterogeneity involving a non-constant detection subgroup, and (iii) model performance with covariates and random effects affecting both abundance and detection.
-Finally, we apply our model to Ovenbird data from the Minnesota Forest Breeding Bird Project.
+Section \label{sec:data} provides a description of the interval-censored time-to-detection avian point count data under consideration.
+Section \label{sec:model} introduces an N-mixture model for estimating abundance from removal-sampled point-count surveys and variety of TTDDs.
+In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
+Section \label{sec:sim} provides three simulation studies to assess the impact of TTDD choice on detection probability. 
+Section \label{sec:ovenbirds} analyzes an Ovenbird data set under the different TTDD to determine the impact this choice has on detection probability and therefore the estimate of abundance.
+Finaly, section \label{sec:discuss} discusses
 
@@
 
-\section{Time-to-detection N-mixture models}
+\section{Time-to-detection N-mixture models} \label{sec:model}
 
@@
 If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
-Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \stackrel{ind}{\sim} \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \stackrel{ind}{\sim} f_{T|det}(t_{sb})$
+Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \ind f_{T|det}(t_{sb})$
 where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
@@
 
-A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e.
-$T_{sb}\stackrel{ind}{\sim} \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e.
-$h(t) = \lambda$.
+A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e. $T_{sb}\ind \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \lambda$.
 Choosing another TTDD can allow for a systematic non-constant detection regime. 
@@
 In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-For consistency that will be clear when including explanatory variables, we use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.
+We use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.
+The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$.
 
@@
 In these models, estimates of $N_s$ are unstable unless $\pdet>0.5$.
-One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\stackrel{ind}{\sim} \Po(\lambda)$. 
-We can decompose $N_{s}$ into observed and unobserved portions; the latent unobserved portion follows a binomial-Poisson hierarchy, with the result that $n_{s}^{(unobs)} \sim \Po\left(\lambda(1-\pdet)\right)$.
+One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$. 
+With this assumption, we can decompose $N_{s}$ into observed and unobserved portions: $n^{(obs)} \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
 Although alternative distributions could be considered, e.g.
@@
 Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
-Assuming independence amongst individuals and sites, we have $\vn_{s} \sim \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
 
@@
 To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ via a mixed effects linear regression with a log link, i.e.
-$\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.
+$\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \ind N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.
 
 To incorporate explanatory variables on detection probability, we let the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
-Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \overset{ind}{\sim} N(0,\\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
+Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
 
@@
 
-For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicity. 
-\begin{subequations}
-\begin{eqnarray} \label{eq:model}
-n_s^{(obs)} &\sim& Po(\lambda_s) \\
-\vn_s &\sim& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
+For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit.
+
+\begin{eqnarray*} 
+n_s^{(obs)} &\ind& Po(\lambda_s) \\
+\vn_s &\ind& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
 p_{s1} &=& \gamma + F_{T|det}(C_1|\alpha,\varphi_s),\, p_{si} = F_{T|det}(C_i|\alpha,\varphi_s)-F_{T|det}(C_{i-1}|\alpha,\varphi_s)\\
-\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \overset{ind}{\sim} N(0,\sigma_{A[j]}^2) \\
-\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \overset{ind}{\sim} N(0,\tau_{D[j]}^2)
-\end{eqnarray}
-\end{subequations}
-
-\jarad{May need to remove subnumbers, but I'll wait to see if we need them.}
-
+\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \ind N(0,\sigma_{A[j]}^2) \\
+\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \ind N(0,\tau_{D[j]}^2)
+\end{eqnarray*}
 We adopt a Bayesian approach and therefore require a prior over the model parameters.
@@
 
+\jarad{Adam: do we want posterior predictive pvalues at all?}
+
 To assess goodness of fit, we relied on posterior predictive checks \citep{Gelman1996}.  
@@
 
+\jarad{Adam: do we want to remove DIC?}
+
 We compared different models by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
@@
 
-\section{Simulation studies}
+\section{Simulation studies} \label{sec:sim}
 
@@
 
-\subsection{Sim1 Results: mixture vs. non-mixture TTDDs}\label{sec:mixture}
+\subsection{Mixture versus non-mixture TTDDs}\label{sec:mixture}
 
@@
 Similarly DIC cannot clearly determine that a non-mixture model should be preferred over a mixture model in this scenario.
-
-When a mixture model is used to simulate the data (top half of the figure), there is clearly a benefit to using a mixture model.
-For the non-mixture models, the credible interval coverage is near zero for most of the models with the exponential model overestimating $\pdet$ and the other models underestimating.
-While DIC could not easily differentiate a mixture from a non-mixture model for inference when a non-peaked simulation model was used, DIC clearly preferred mixture models when the data arose from an exponential distribution or any of the peaked distributions.
-
-These results support the use of a mixture model unless you are absolutely sure your data arose from a non-mixture distribution.
-
-% When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $\pdet$ badly in every case.  
-% This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
-% Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
-% Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $\pdet$ than did the true non-mixture inference model.  
-% The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $\pdet$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
-% Mixture model posterior medians for $\pdet$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
-
-% \adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
-% \jarad{We shouldn't be talking about results that we are not showing. The results should either be in the manuscript or the supplement.}
-% Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $\pdet$ above and shared the same coverage rates.  
-% Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
-% Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
-% For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
-% For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
-
-% Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
-DIC did reveal a preference for mixture models with exponential and peaked datasets.
-In these cases, posterior predictive checks clearly showed poor fit from the non-mixture models with p-values of 1.00 in the second and sometimes third observation intervals.
-These p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
-
-% Possible other comments
-%(1) For nonpeaked data, the true inference model is biased low\\
-%(2) Mixture inference models generate higher estimates than non-mixture models\\
-%(3) True models are generally good\\
-%(4) Estimates of non-mixture peaked data are much more precise\\
-%(5) Note that plots of TTDD at the median are indistinguishable from the truth except in the blatant fail scenario
 
@@
 
-\subsection{Sim2: constant vs. non-constant detection mixture TTDDs}\label{sec:family}
+When a mixture model is used to simulate the data (top half of the figure), there is clearly a benefit to using a mixture model.
+For the non-mixture models, the credible interval coverage is near zero for most of the models with the exponential model overestimating $\pdet$ and the other models underestimating.
+While DIC could not easily differentiate a mixture from a non-mixture model for inference when a non-peaked simulation model was used, DIC clearly preferred mixture models when the data arose from an exponential distribution or any of the peaked distributions.
 
-In Sim2, we simulated intercept-only datasets from the 7 different mixture TTDD models and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families (Table \ref{tbl:sim2}).
-Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.
-However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.
-Differences among non-constant TTDDs were less pronounced.  
-Gamma models were more accurate than others for nonpeaked mixture datasets.  
-Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
-Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
+These results support the use of a mixture model unless you are absolutely sure your data arose from a non-mixture distribution.
 
-Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
-For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
-So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
-For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
+% When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $\pdet$ badly in every case.  
+% This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
+% Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
+% Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $\pdet$ than did the true non-mixture inference model.  
+% The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $\pdet$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
+% Mixture model posterior medians for $\pdet$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
 
-In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.
-For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.
+% \adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
+% \jarad{We shouldn't be talking about results that we are not showing. The results should either be in the manuscript or the supplement.}
+% Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $\pdet$ above and shared the same coverage rates.  
+% Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
+% Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
+% For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
+% For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
+
+% Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
+DIC did reveal a preference for mixture models with exponential and peaked datasets.
+In these cases, posterior predictive checks clearly showed poor fit from the non-mixture models with p-values of 1.00 in the second and sometimes third observation intervals.
+These p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
+
+% Possible other comments
+%(1) For nonpeaked data, the true inference model is biased low\\
+%(2) Mixture inference models generate higher estimates than non-mixture models\\
+%(3) True models are generally good\\
+%(4) Estimates of non-mixture peaked data are much more precise\\
+%(5) Note that plots of TTDD at the median are indistinguishable from the truth except in the blatant fail scenario
+
+\subsection{Constant vs. non-constant detection mixture TTDDs}\label{sec:family}
+
+The previous section adressed model mis-specification in terms of the mixture component. 
+Now we turn to model misspecification of the distribution family. 
+We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using settings that are the same as in the previous section and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
+Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+
+Table \ref{tbl:sim2} provides a summary of the ability of each model to capture the detection probability. 
+Each quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
+The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak since the two parameters, exponential rate and mixing parameter, do not provide enough flexibility for the mixture exponential distribution to adequately model times to detection that have an initial increase and delayed mode.
+In this situation, the exponential model underestimates the actual detection probability.
+In contrast, if there is no delayed peak, the exponential model typically overestimates the actual detection probability.
 
@@
 \end{tabular}
-\caption{\label{tbl:sim2} Summary of mixture inference models of all families fit to mixture datasets (Sim2).  
+\caption{Summary of mixture inference models of all families fit to mixture datasets.  
 Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
@@
 $\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
+\label{tbl:sim2}
 \end{table}
+
+When comparing the different three-parameter TTDDs, model misspecification is not as serious an issue because these models can adequately account for the interval-censored times to detection. 
+Nonetheless, it appears the gamma mixture model is better able to account for data from lognormal or Weibull mixture models. 
+In the gamma quadrant, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility.
+In contrast, when the lognormal model is used for inference, it performed worse when the data arose from a gamma or Weibull and, when the Weibull model was used for inference, it performed worse when the data arose from a lognormal.
+
+Other than providing evidence against the exponential mixture model, DIC was not helpful in distinguishing amongst these models.
+
+% Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.
+% However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.
+% Differences among non-constant TTDDs were less pronounced.  
+% Gamma models were more accurate than others for nonpeaked mixture datasets.  
+% Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
+% Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
+% 
+% Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
+% For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
+% So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
+% For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
+% 
+% In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.
+For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.
 
@@
 
-In Sim3, we generated a single replicate of datasets from models including covariates and random effects for both detection and abundance.  
-Results echoed those of the first two simulation studies, suggesting that the addition of covariates and random effects did not alter basic model behavior.
+The previous sections studying the effect of the time to detection assumption in the context of no explanatory variables.
+Here we incorporate fixed and random effects for abundance and detection. 
+For each of the 14 combinations of TTDDs \jarad{Adam: what did we do? i don't remember how the data sets were constructed and therefore how to interpret the figures.}
+We chose values for fixed effects and random effect variances similar to those found in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
+Due to the computation time involved in estimating models with these fixed and random effects, we simulated two data sets: mixture and non-mixture.
+
+To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.
+
+Figure \ref{pdet_cater_correct}
+
 These results included the detection bias of non-mixture models used to fit mixture datasets (Figure \ref{pdet_cater_correct}), detection bias of exponential mixture models used to fit data mixtures with time-varying $F_{T1}$ components (Figure \ref{pdet_cater_family}), and more precise and accurate performance of exponential mixture models when data were from exponential mixtures.
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discuss}
 

### 2016-06-19 Draft.tex
Abstract edits (Adam)

@@
 %Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  
-We consider a class of N-mixture models that allow for detection heterogeniety through a mixture component, flexible two-parameter time-to-detection distributions (TTDDs), and allows for fixed and random effects for both abundance and detection.
+We consider a class of N-mixture model that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
+%We consider a class of N-mixture models that allow for detection heterogeniety through a mixture component, flexible two-parameter time-to-detection distributions (TTDDs), and allows for fixed and random effects for both abundance and detection.
 %Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution (TTDD) embedded within a hierarchical N-mixture framework.
 Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
-We find that TTDDs that include a mixture component resulting in increased probability of detection in the initial observation period are necessary when the data have a chance of arising from distributions of this nature.
-In addition, the flexibility offered by two-parameter distributions, e.g. gamma, can outperform exponential based models when the truth is exponential or non-exponential.
-Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection and demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs and that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
-
+We specifically explore two-parameter families of TTDDs (e.g. gamma) that can additionally include a mixture component to model increased probability of detection in the initial observation period.
+We find that modeling a TTDD by using a mixture component is necessary when data have a chance of arising from a distribution of this nature.
+In addition, two-parameter distributions can outperform exponential-based models both when the truth is exponential or non-exponential.
+Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
+We demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs but that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
+%We find that modeling a TTDD by using a mixture component for increased probability of detection in the initial observation period is necessary when data have a chance of arising from distributions of this nature.
+%In addition, the flexibility offered by two-parameter distributions, e.g. gamma, can outperform exponential based models when the truth is exponential or non-exponential.
+%Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection and demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs and that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
 % We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  
@@
 
-{\bf Keywords:} abundance; availability; N-mixture model; survival analysis; Bayesian
+{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Bayesian
 

### 2016-06-20 Draft.tex
Revisions to Introduction.

@@
 %\adam{Others, like Mantyniemi.}
-In particular, animal behaviors, e.g.
+In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
+lack of settling in period \citep{LeeMarsden2008, Johnson2008}, can all lead to time-varying rates of detection.
+%In particular, animal behaviors, e.g.
 intermittent singing in birds and frogswhales \citep{Scott2005, Diefenbach2007, Reidy2011}, observer behaviors, e.g.
@@
 
-A common assumption to deal with some of the issues above is to increase the probability of detection in the initial observation period \citep{Etterson2009} \jarad{Adam: add other refs}, although this is not yet the standard \citep{Amundson2014} \jarad{Adam: add other refs}.
-In this manuscript, we consider the choice of whether to include a mixture component, but we also consider a broader class of non-constant distributions.
-%In this study, we develop a model for scenarios where detection rates are not constant. 
-In particular, we consider time-to-detection as done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function $F_T(t)$.
+In this manuscript, we develop a model for scenarios where detection rates are not constant over time. 
+We consider the first time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
+One common strategy to deal with data that do not fit a constant-detection assumption is to model increased detection probability in the initial observation period via a mixture component \citep{Farnsworth2002, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
+We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
 
@@
 We embed the TTDD in a hierarchical framework for multinomial counts using an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.  
-For our purposes, the N-mixture framework provides three clear benefits: 1) its multinomial data framework accords with the interval censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
+For our purposes, the N-mixture framework provides three clear benefits: 1) its multinomial data framework accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
 The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.  
@@
 Framing a model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  
-First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}, whereas most existing studies instead construct regression models on the logit of the equal-interval detection probability \adam{Amundson2014 and others?}.
-The later is not possible when detection rates are not constant.  
+First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}, whereas most existing studies instead construct regression models on the logit of the equal-interval detection probability.
+The latter is not possible when detection rates are not constant.  
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
@@
 
-Section \label{sec:data} provides a description of the interval-censored time-to-detection avian point count data under consideration.
-Section \label{sec:model} introduces an N-mixture model for estimating abundance from removal-sampled point-count surveys and variety of TTDDs.
+Section \ref{sec:data} provides a description of the interval-censored time-to-detection avian point count data under consideration.
+Section \ref{sec:model} introduces an N-mixture model with a generically defined TTDD for estimating abundance from removal-sampled point-count surveys.
 In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
-Section \label{sec:sim} provides three simulation studies to assess the impact of TTDD choice on detection probability. 
-Section \label{sec:ovenbirds} analyzes an Ovenbird data set under the different TTDD to determine the impact this choice has on detection probability and therefore the estimate of abundance.
-Finaly, section \label{sec:discuss} discusses
+Section \ref{sec:sim} provides three simulation studies to assess the impact of TTDD choice on estimated detection probability. 
+Section \ref{sec:ovenbirds} analyzes an Ovenbird data set under different TTDDs to determine the impact of this choice on estimated detection probability and therefore estimated abundance.
+Finally, section \ref{sec:discuss} discusses 
+\adam{The previous sentence ends mid-thought.  Is this for the Discussion section?}
 

### 2016-07-12 Draft.tex
Large scale changes and comments upon everything up through Sim3.

@@
 
-Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citet{Hanowski1995}.  
-Survey sites were randomly selected from sawtimber red pine stands with no recent logging activity such that each stand had up to three sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
-The data include 65 sites and a total of 381 with site specific variables including site age,  stock density, an indicator of select-/partial-cut logging during the 1990s.
+Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.  
+Survey sites were selected from sawtimber red pine stands with no recent logging activity.  
+Each stand had up to four sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
+The data include 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, an indicator of select-/partial-cut logging during the 1990s.
 
-%with a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
-
-\jarad{Adam: How many observers}
-
-%The MNFB design is described in , and we summarize only essential details here.  
 Single-visit (per year) point-count surveys were conducted by trained observers at each site once annually (weather permitting).  
+Fourteen different observers conducted surveys during the study period.  
 Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
@@
 
-\section{Time-to-detection N-mixture models} \label{sec:model}
+\section{Continuous time-to-detection N-mixture models} \label{sec:model}
 
-Before considering interval censoring and explanatory variables, we first consider the scenario of exact time to detections with no explanatory variables. 
+Before considering interval censoring and explanatory variables, we first present the scenario of exact time to detections with no explanatory variables. 
 We then incorporate interval censoring and follow with inclusion of fixed and random effects for abundance and detection.
@@
 Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
-Assuming all individuals are observed at a common rate, then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$. 
-If the times to first detection are censored at $C$, due to a finite survey length of $C$, the probability of detection for each individual is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
-Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming individuals are independent, we have $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb} \ind f_{T|det}(t_{sb})$
-where $T_{sb}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
+Assuming all individuals are observed according to a common time-to-detection distribution (TTDD), then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
+If the times to first detection are truncated at $C$, due to a finite survey length of $C$, the probability of detection for each individual present at the survey site is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
+Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming detection times for individuals are independent, we have $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb|det} \ind f_{T|det}(t_{sb})$
+where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
 The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
 
-A common choice for time-to-detection distribution (TTDD) is an exponential distribution, i.e. $T_{sb}\ind \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \lambda$.
+\adam{Thoughts on the above paragraph, recognizing that it is one of the most finicky in the whole paper: (i) Do the first few sentences make clear enough the distinction between $t \in (0, \infty)$ and $t|det \in (0,C)$?
+(ii) Is $T_{sb|det} \ind f_{T|det}(t_{sb})$ correct notation?
+It feels odd to me to have a pdf on the right side of that tilde, but maybe it's just inexperience.
+(iii) You originally wrote ``all individuals are observed according to a common rate".
+Due to heterogeneity, I replaced ``rate" with ``time-to-detection distribution", though I'd be fine with ``hazard function".
+But, in truth, even that's not true -- due to heterogeneity and the mixture formulation.
+What we're actually doing is modeling all individual birds from the marginal TTDD across group-specific hazard rates...
+rather a mouthful.
+(iv) I liked the structure that was imparted in the earlier draft by having a data model, abundance model, and detection model.}
+
+A common choice for TTDD is an exponential distribution, i.e. $T_{sb}\ind \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \varphi$.
 Choosing another TTDD can allow for a systematic non-constant detection regime. 
@@
 In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-We use the following parameterization: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\varphi \mathrm{\Gamma}(1+1/\alpha)$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\varphi+\alpha/2}$.
+To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\alpha/2}/\varphi$.
+\adam{Expressing the LN in rate parameterization is a little unconventional...
+We may prefer to be more explicit about the relationship between $\mu$ and $\varphi$ by writing: $T \sim LN(-\log(\varphi), \alpha)$.}
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$.
 
-With exact times to first detection, these times could be used to estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
+Given exact times to first detection, we can estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
 With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
-In these models, estimates of $N_s$ are unstable unless $\pdet>0.5$.
-One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$. 
+In these models, point estimates of $N_s$ are unstable unless $\pdet>0.4$ \citep{Olkin1981}.
+One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$ \citep{Raftery1988, Royle2004NMixture}. 
 With this assumption, we can decompose $N_{s}$ into observed and unobserved portions: $n^{(obs)} \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
@@
 \jarad{Adam: add refs.}
+\adam{The underdispersion finding comes from our own modeling of the Ovenbird dataset.
+I recall no discussion of underdispersion in the literature.
+The primary source for abundance distributions removal models is Etterson.}
 
@@
 
-A common observation in avian point counts are increased observations in the first interval relative to an exponential distribution.
-To accommodate this empirical observation, many models include a parameter to increase the probability of observing individuals in the first interval. \jarad{Adam: add refs.}
+It is common in avian point counts to observe increased detections in the first interval relative to an exponential distribution.
+To accommodate this empirical observation, many models of interval-censored removal times include a parameter to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
 We augment our model with a mixing parameter $\gamma\in[0,1]$ which updates the following definitions
@@
 If $\gamma=0$, then the non-mixture model is recovered.
+\adam{Three things that we'll probably want to talk through: (i) you've switched the mixture components -- my $\gamma$ has become (1-$\gamma$) in the new nomenclature; (ii) I feel that $f_T(t)$ and $F_T(t)$ should describe the entire TTDD and that a different function nomenclature should be used for the `hard to detect' birds.
+(iii) This dates back to my original draft...
+but this particular form of mixture -- with the point mass -- is really specific to interval-censoring.
+It feels incongruous to be mixing a point-mass and a continuous distribution.
+However, writing it propoerly depends upon nomenclature decisions.}
 
@@
 
-For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit.
+For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit. \adam{Which equation?}
 
 \begin{eqnarray*}
-n_s^{(obs)} &\ind& Po(\lambda_s) \\
+n_s^{(obs)} &\ind& Po(\lambda_s \pdet_s) \\
 \vn_s &\ind& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
@@
 To ease construction of a default prior for this model, we standardize all explanatory variables and then construct priors to be diffuse within a reasonable range of values.
-Prior mean and standard deviation (sd) for the abundance intercept was set at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
-Prior mean and sd for the detection intercept were chosen so that, based on an intercept-only non-mixture model with $\alpha=1$: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
+Normal prior mean and standard deviation (sd) for the abundance intercept was set at a median abundance of 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  
+Normal prior mean and sd for the detection intercept were chosen so that, based on an intercept-only non-mixture model with $\alpha=1$: (i) median prior detection probability was $p_{s}^{(det)} = 0.50$, and (ii) 95\% of the prior detection probability was within $p_{s}^{(det)} \in (0.01, 1.0)$.
 Normal priors for fixed effect parameters were centered at zero with standard deviations matching the appropriate intercept term.  
-All standard deviations and $\alpha$ were given half-Cauchy priors with location 0 and scale 1 for the untruncated Cauchy and Unif(0,1) prior on the mixture parameter $\gamma$ in the mixture models.
+All standard deviations and $\alpha$ were given half-Cauchy priors with location 0 and scale 1 for the untruncated Cauchy, and the mixture parameter $\gamma$ was assigned a Unif(0,1) prior in mixture models.
 All scalar parameters were assumed independent \emph{a priori}.
-
-\jarad{Adam: I appreciate that you provided the justification for the priors, but we should also include the exact vaules where possible.}
 
@@
 We discarded half of the iterations as warmup and then thinned by 10.  
-We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}. \jarad{Adam: what did you do with the Geweke diagnostics?}  
+We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}.
+\jarad{Adam: what did you do with the Geweke diagnostics?} \adam{I'm unclear what you intend with this question.
+I used them to monitor convergence.
+Any time a chains raised a Geweke flag, I re-ran the model.}
 We reran models if the effective sample size for any parameter was below 1000.  
 The number of iterations used depended on the model and is detailed later. 
-For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed \jarad{Adam: how did they fail?} unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
+For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed \jarad{Adam: how did they fail?}\adam{I don't remember the error message, but Stan failed with no output generated.} unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
 \jarad{Adam: this doesn't make sense.
 Why would the means from these preliminary short-run models be any better than letting these preliminary short-run models continue?
-You claim seems to be that letting these preliminary short-runs continue would mean that those models would fail.}
-\jarad{Adam: if you are using Geweke diagnostics, then it would be better for your to start from somewhere "good". For fixed effects, you could start at 0, mixing parameter at 0.5, etc.}
+You claim seems to be that letting these preliminary short-runs continue would mean that those models would fail.}\adam{I have brilliantly deleted the original files, but my memory suggests that models tended to fail early.
+Some short-run models failed quickly, too, but if I ran a model for 1000 iterations, the results provided stable initial values which I could subsequently used whenever I tweaked the models and had to rerun everything.
+I'm thinking maybe these sentences can just be dropped.}
 
@@
 We distinguish two categories of TTDDs: peaked and nonpeaked.  
-\jarad{Adam: I don't like the previous definition because log-normal never has a peak at zero.
-In addition, if the peak occurs before $C_1$ then we really aren't interested in it.}
-A peaked TTDD has a mode greater than $C_1$ while a non-peaked TTDD has a mode less than $C_1$. \jarad{Adam: do we want a figure demonstrating this point?}
+A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
+\jarad{Adam: do we want a figure demonstrating this point?}\adam{I don't think a figure is required.}
 
@@
 
-\jarad{Adam: do we want posterior predictive pvalues at all?}
+\jarad{Adam: do we want posterior predictive pvalues at all?}\adam{Interesting question.
+I don't think they're essential to our model.
+I am open to omitting them.
+However...
+Amundson et al.
+are vague about the posterior p-values -- ``we used Bayesian P values generated from the posterior predictive distributions to assess goodness-of-fit" -- they don't state what aspect of fit they tested -- it looks like maybe they use the model-fit value of $\pdet$ as their statistic?!?
+Anyhow, I suspect they misapplied the tool, and I somewhat wanted to counteract that with an example of a better application of it.
+Also, I think it's important to show that we checked some measure of fit.}
 
@@
 
-\jarad{Adam: do we want to remove DIC?}
+\jarad{Adam: do we want to remove DIC?}\adam{As with posterior predictive checks, I can live without DIC, which has been nothing but headaches.}
 
@@
 
-% In Sim1, we compared models with detection heterogeneity via a finite-mixture TTDD versus models without detection heterogeneity.
-% We simulated both mixture and non-mixture datasets from exponential, gamma, lognormal, and Weibull TTDD families.  
-% Because model behavior may differ between peaked and nonpeaked datasets, we simulated both for each TTDD family except exponential.  
-% Thus, combining mixtures, `peakedness' and TTDD families, we simulated datasets from 14 different TTDDs in total.  
-% In Sim2, we examined the impact of misspecifying the $F_{T1}$ family (e.g., fitting an exponential TTDD model to data generated from a Weibull TTDD).  
-% Of particular interest were scenarios with a time-varying $F_{T1}$ component in the data but a time-constant component in the inference model, and vice versa.
-% Based on findings from Sim1, we considered only mixture TTDD datasets, and we fit each with mixture TTDD models from all four families.  
-% In Sim3, we modeled a more complete scenario where both data generation and model fitting included covariates and random effects for abundance and detection.  
-% 
-% We used posterior median estimates from the Ovenbird analysis to inform simulation parameters except as follows.  
-% For Sim1 and Sim2,   
-% For Sim3, all parameters used for nonpeaked datasets came from the Ovenbird analysis.
-% In order to simulate peaked dataset variants, we modified the Ovenbird parameters by:  (i) setting $\gamma = 0.65$, and (ii) scaling the Intercept$^D$ and $\sigma_D^2$ terms by 0.25 and 1.1, resulting on average in a detection mode near 5 minutes and 75-85\% overall detection across all sites.
-% Unlike Sim1 and Sim2, expected detection probabilities in Sim3 were not the same for all datasets.
-% Counts were censored into nine-intervals as with the Ovenbird dataset.  
-% For both Sim1 and Sim2, we simulated 16 replicates, but for Sim3 we generated only a single replicate, because gamma models took an average of 5 days to fit.
-
 \subsection{Mixture versus non-mixture TTDDs}\label{sec:mixture}
@@
 In Sim1, we simulated 16 intercept-only datasets from each of 14 TTDDs. 
-We chose true parameter values so that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-\jarad{Adam: we need specific values here for $\alpha$ and $\varphi$ (or $\beta_0^D$). If it is too much to include here put it in the supplement.}
+We chose true parameter values (see Supplement) so that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+These values are largely consistent with estimates of Ovenbird populations from Section \ref{sec:ovenbirds}.
 
@@
 
-Table \ref{tbl:sim1} provides a summary of the mixture and non-mixture models ability to capture the detection probability. 
-When a non-mixture model is used to simulate the data (top half of the figure), there is no clearly discernable differences between the ability of a non-mixture or mixture model to capture.
+Table \ref{tbl:sim1} provides a summary of the mixture and non-mixture models' ability to capture the detection probability. 
+When a non-mixture model is used to simulate the data (top half of the table), there are no clearly discernable differences between the ability of a non-mixture or mixture model to capture.
+\adam{For nonpeaked non-mixture data, the mixture model is less biased for lognormal and Weibull scenarios, and CIs are 15-40\% narrower.}
 Similarly DIC cannot clearly determine that a non-mixture model should be preferred over a mixture model in this scenario.
@@
 So the bottom half could be included in the main body while the top half could be put in the supplement.}
+\adam{The obvious concern for non-mixture data is that the mixture model might be more uncertain than a non-mixture model for non-mixture data.}
 }
@@
 
-When a mixture model is used to simulate the data (top half of the figure), there is clearly a benefit to using a mixture model.
-For the non-mixture models, the credible interval coverage is near zero for most of the models with the exponential model overestimating $\pdet$ and the other models underestimating.
+When a mixture model is used to simulate the data (bottom half of the table), there is clearly a benefit to using a mixture inference model.  
+For the non-mixture models, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ and the other models underestimating. 
 While DIC could not easily differentiate a mixture from a non-mixture model for inference when a non-peaked simulation model was used, DIC clearly preferred mixture models when the data arose from an exponential distribution or any of the peaked distributions.
@@
 Now we turn to model misspecification of the distribution family. 
-We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using settings that are the same as in the previous section and fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
-Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
+%Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
 
@@
 Each quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
-The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak since the two parameters, exponential rate and mixing parameter, do not provide enough flexibility for the mixture exponential distribution to adequately model times to detection that have an initial increase and delayed mode.
+The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak, because the two parameters (rate and mixing parameter) do not provide enough flexibility for the mixture exponential distribution to adequately fit a TTDD with both an initial increase and a delayed mode.
 In this situation, the exponential model underestimates the actual detection probability.
-In contrast, if there is no delayed peak, the exponential model typically overestimates the actual detection probability.
+In contrast, if the simulated data are non-peaked, the exponential model typically overestimates the actual detection probability. 
+\adam{For exponential data, the exponential model performs better.  I feel this should be noted.}
 
@@
 Nonetheless, it appears the gamma mixture model is better able to account for data from lognormal or Weibull mixture models. 
-In the gamma quadrant, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility.
+In the gamma quadrant of Table \ref{tbl:sim2}, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility.
 In contrast, when the lognormal model is used for inference, it performed worse when the data arose from a gamma or Weibull and, when the Weibull model was used for inference, it performed worse when the data arose from a lognormal.
+Other than sometimes providing evidence against the exponential mixture model, DIC was not helpful in distinguishing amongst these models. 
+\adam{I'd be okay with jettisoning this paragraph beyond the first sentence.  Conversely, rather than praising gamma, we could discourage lognormal.}
 
-Other than providing evidence against the exponential mixture model, DIC was not helpful in distinguishing amongst these models.
+These results support the use of a non-constant rate mixture TTDD model unless you are absolutely sure your data arose from an exponential mixture TTDD.
 
@@
 
-The previous sections studying the effect of the time to detection assumption in the context of no explanatory variables.
+\adam{Here's a brazen idea...
+let's omit Sim3.
+We can reference it in the Ovenbird analysis, saying `we did a preliminary simulation study of the model with covariates and random effects, and results were consistent with Sim1, Sim2, and the Ovenbird findings...'.
+What do you think?}
+
+The previous sections studyied the effect of the time to detection assumption in the context of no explanatory variables.
 Here we incorporate fixed and random effects for abundance and detection. 
-For each of the 14 combinations of TTDDs \jarad{Adam: what did we do? i don't remember how the data sets were constructed and therefore how to interpret the figures.}
-We chose values for fixed effects and random effect variances similar to those found in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Due to the computation time involved in estimating models with these fixed and random effects, we simulated two data sets: mixture and non-mixture.
+For each of the 14 TTDDs, we simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
+Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as in Sim1 and Sim2, (ii) using covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
+(See supplement for actual parameter values).
+Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
 
@@
 
-Figure \ref{pdet_cater_correct}
+Patterns in posterior estimates of $\pdet_s$ with respect to mixture and family TTDD forms were the same in Sim3 as in Sim1 and Sim2 (Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family}.
 
-These results included the detection bias of non-mixture models used to fit mixture datasets (Figure \ref{pdet_cater_correct}), detection bias of exponential mixture models used to fit data mixtures with time-varying $F_{T1}$ components (Figure \ref{pdet_cater_family}), and more precise and accurate performance of exponential mixture models when data were from exponential mixtures.
+Possible results to focus upon:\\
+\begin{enumerate}
+\item Precision of estimates of $\pdet$ relative to Sim1 and Sim2 (they're narrower in Sim3)
+% These results indicate that covariates provided significant information for reducing uncertainty.
 
-Abundance estimates mirrored the first two simulation studies except: (i) credible intervals across all mixture models were on average 50\% narrower than in Sim2, and bias in posterior median abundance from exponential model fits of peaked datasets was roughly halved relative to Sim2, though it was still worse than gamma model fits by about 35\% of actual abundance.
-These results indicate that covariates provided significant information for reducing uncertainty.
+\item $\beta^A$ estimates are pretty much invariant to choice of TTDD
+%For every dataset we simulated, estimates and credible intervals for abundance covariate coefficients $\beta^A$ were virtually the same regardless of the inference TTDD employed.
+
+\item $\beta^D$ estimates from exponential mixtures are more precise than for other families.  The coverage situation is unclear based on size of sim study.
+%Estimates of detection parameters $\beta^D$ varied by context.  
+%Comparisons of mixture and non-mixture models (like in Sim1) produced similar $\beta^D$ estimates except when fit to mixture peaked datasets, where models without a mixture component produced credible intervals twice as wide as those from mixture models with no benefit in terms of coverage.
+%Comparing exponential to other TTDD family models (as in Sim2), time-varying mixture TTDD models produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and either narrower (exponential and nonpeaked data) or wider (peaked data).
+%Differences in estimates of $\beta^D$ parameters to some degree reflected differing estimation of the $\gamma$ mixing parameter -- the exponential model underestimated $\gamma$ with nonpeaked datasets and overestimated it in peaked datasets.
+%Many of the above patterns for nonpeaked datasets can be noted in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
+
+\item DIC can be effective in the same scenarios as Sim1 and Sim2, but it can also be misleading (provide stark example).  This sets the stage for the Ovenbird results.
+%Results from DIC and posterior predictive checks suggested mixture models over non-mixture models for peaked data (DIC differences $>10$) and time-varying mixture TTDDs over the exponential mixture for gamma and lognormal peaked data ($>8$).
+%DIC preferred exponential mixtures to time-varying mixtures for exponential mixture data ($>6$); however, it also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.
+%In this last case, the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
+%In conjunction with Sim2 results, this last finding calls into question the utility of DIC for nonpeaked datasets.
+%DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
+
+\end{enumerate}
+
+\adam{There is a question as to what Figures/Tables would effectively express the above points.
+I think the cleanest approach would be to pick 'representative' datasets and models.
+Or, in my first draft, I simply referenced the Ovenbird output to illustrate several of the points...}
 
@@
 \end{figure}
-
-For every dataset we simulated, estimates and credible intervals for abundance covariate coefficients $\beta^A$ were virtually the same regardless of the inference TTDD employed.  
-Estimates of detection parameters $\beta^D$ varied by context.  
-Comparisons of mixture and non-mixture models (like in Sim1) produced similar $\beta^D$ estimates except when fit to mixture peaked datasets, where models without a mixture component produced credible intervals twice as wide as those from mixture models with no benefit in terms of coverage.
-Comparing constant-detection to time-varying models (like in Sim2), time-varying mixture TTDD models produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and either narrower (exponential and nonpeaked data) or wider (peaked data).
-Differences in estimates of $\beta^D$ parameters to some degree reflected differing estimation of the $\gamma$ mixing parameter -- the exponential model underestimated $\gamma$ with nonpeaked datasets and overestimated it in peaked datasets.
-Many of the above patterns for nonpeaked datasets can be noted in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
-
-Results from DIC and posterior predictive checks suggested mixture models over non-mixture models for peaked data (DIC differences $>10$) and time-varying mixture TTDDs over the exponential mixture for gamma and lognormal peaked data ($>8$).
-DIC preferred exponential mixtures to time-varying mixtures for exponential mixture data ($>6$); however, it also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.
-In this last case, the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
-In conjunction with Sim2 results, this last finding calls into question the utility of DIC for nonpeaked datasets.
-DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
-
-%\adam{I'm guessing I should not include the following: (1) my posterior predictive pvalues are not independent; (2) for instance, for the first interval, they stay very close to 0.5-0.6 --- the models are effective at getting the first interval right; (3) likewise, a post.
-p-value $>$ 0.8 during the 2nd interval likely shows model misfit; all of which raises the possibility that (4) multiple simulation under the null could provide a reference distribution of posterior p-values.}
-
-% Posterior predictive p-values for Sim1
-%\begin{figure}[h!]\centering
-%\includegraphics[width=0.78\textwidth]{Sims/SimZero/PostPredP_corr.pdf}
-%\caption{\label{postpredmix} Posterior predictive p-values by interval -- Pr$\left(p_{si}^{(rep)} > p_{si}\right)$ -- for peaked mixture datasets fit with non-mixture inference models of the correct TTDD family over 16 replicates (Sim1).
-%Trendlines show mean posterior predictive p-values.  
-%Dashed lines show central 95\% intervals.  
-%P-values near 0 or 1 potentially indicate poor model fit.}
-%\end{figure}
 
@@
 Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.
-% "It is tempting to choose classes of models that yield narrower confidence intervals for N; however, Coull's (1997) extensive simulation studies (summarized by Coull and Agresti (1999)) clearly reveal that this behavior can produce erroneous inferences." -- Dorazio and Royle 2003
+% ``It is tempting to choose classes of models that yield narrower confidence intervals for N; however, Coull's (1997) extensive simulation studies (summarized by Coull and Agresti (1999)) clearly reveal that this behavior can produce erroneous inferences." -- Dorazio and Royle 2003
 
@@
 
+\section{Supplement}
+
+\begin{table}[!ht]
+\centering
+\begin{tabular}{lll|ccc}
+  \hline
+Family & Mixture & Peaked & $\gamma$ & $\varphi$ & $\alpha$ \\ 
+  \hline
+Exponential & Non-mixture & &  & -1.827 &  \\ 
+  & Mixture & & 0.65 & -2.138 &  \\ 
+  \hline
+  Gamma & Non-mixture & Non-peaked &  & -3.279 & 0.257 \\ 
+  & & Peaked &  & -0.746 & 3.371 \\ 
+  & Mixture & Non-peaked & 0.65 & -2.773 & 0.577 \\ 
+  & & Peaked & 0.65 & -1.210 & 2.491 \\ 
+  \hline
+  Lognormal & Non-mixture & Non-peaked &  & -0.341 & 2.330 \\ 
+  & & Peaked &  & -1.872 & 0.512 \\ 
+  & Mixture & Non-peaked & 0.65 & -1.462 & 1.674 \\ 
+  & & Peaked & 0.65 & -1.992 & 0.618 \\ 
+  \hline
+  Weibull & Non-mixture & Non-peaked &  & -1.165 & 0.418 \\ 
+  & & Peaked &  & -2.042 & 1.829 \\ 
+  & Mixture & Non-peaked & 0.65 & -2.063 & 0.687 \\ 
+  & & Peaked & 0.65 & -2.201 & 1.621 \\ 
+   \hline
+\end{tabular}
+\caption{Table of parameter values used to generate data for Sim1 and Sim2 where $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lllrrrrrrrrrrrrrrrr}
+  \hline
+Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
+  \hline
+Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
+   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
+  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
+   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
+   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
+   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
+  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
+   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
+   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
+   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
+  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
+   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
+   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
+   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
+   \hline
+\end{tabular}
+\caption{Table of parameter values used to generate data for Sim3 where $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
+\end{table}
+
 \bibliography{masterbib}

### 2016-07-14 Draft.tex
Add title page
Revised first paragraph of model (it does simplify things to start with N_s... good suggestion). Wrote results for Sim3. Wrote results for Ovenbirds.

@@
 
+\title{Hierarchical model assessing the impacts of time to detection distribution assumptions on detection probability estimation}
+
+\author{Adam Martin-Schwarze, Jarad Niemi, and Philip Dixon}
+
 \begin{document}
+
+\maketitle
+\newpage
 
@@
 Fourteen different observers conducted surveys during the study period.  
+69\% of surveys in our dataset involved observers in their first year at MNFB.  
 Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
@@
 
-Suppose at each survey location $s$ ($s=1,\dots,S$), the exact time to first detection $t_{sb}$ of individual $b$ (for bird, $b=1,\dots,n_s^{(obs)}$) is recorded.
-Assuming all individuals are observed according to a common time-to-detection distribution (TTDD), then $T_{sb}$ is a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
-If the times to first detection are truncated at $C$, due to a finite survey length of $C$, the probability of detection for each individual present at the survey site is $\pdet=F_T(C)$ and the conditional distribution of observed detection times has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$.
-Denoting $N_{s}\ge n_s^{(obs)}$ as the total number of individuals present at a survey and assuming detection times for individuals are independent, we have $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $T_{sb|det} \ind f_{T|det}(t_{sb})$
-where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
+Suppose that, for each survey $s$ ($s=1,\dots,S$), $N_s$ individuals are present.  
+Imagine an observer could remain at the survey location until every individual is detected, marking the time to detection $t_{sb}$ (for bird, $b=1,\dots,N_s$) for each.
+Then, assuming detection times for all individuals at a survey are independent, identically distributed according to a common time-to-detection distribution (TTDD), we define $T_{sb}$ as a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
+However, in reality, times to first detection are truncated due to a finite survey length of $C$, meaning that each individual has a detection probability $\pdet=F_T(C)$.  
+The conditional distribution of observed detection times then has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$. 
+We model the number of individuals for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$.
+%$T_{sb|det} \ind f_{T|det}(t_{sb})$ where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$. 
 The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
-
-\adam{Thoughts on the above paragraph, recognizing that it is one of the most finicky in the whole paper: (i) Do the first few sentences make clear enough the distinction between $t \in (0, \infty)$ and $t|det \in (0,C)$?
-(ii) Is $T_{sb|det} \ind f_{T|det}(t_{sb})$ correct notation?
-It feels odd to me to have a pdf on the right side of that tilde, but maybe it's just inexperience.
-(iii) You originally wrote ``all individuals are observed according to a common rate".
-Due to heterogeneity, I replaced ``rate" with ``time-to-detection distribution", though I'd be fine with ``hazard function".
-But, in truth, even that's not true -- due to heterogeneity and the mixture formulation.
-What we're actually doing is modeling all individual birds from the marginal TTDD across group-specific hazard rates...
-rather a mouthful.
-(iv) I liked the structure that was imparted in the earlier draft by having a data model, abundance model, and detection model.}
+\adam{I liked the structure that was imparted in the earlier draft by having a data model, abundance model, and detection model.}
 
@@
 Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
+For simplicity, we model the shape parameter $\alpha$ as constant across sites.
 
@@
 Each dataset was fit with two models: mixture and non-mixture version of the distribution family, e.g. exponential, used to simulate the data.
-We ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+We ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effective samples for all parameters.
 
@@
 Now we turn to model misspecification of the distribution family. 
-We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
+For Sim2, we simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
 %Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
@@
 
-\subsection{Sim3: including covariates and random effects}\label{sec:simfull}
+\subsection{Models including covariates and random effects}\label{sec:simfull}
 
-\adam{Here's a brazen idea...
-let's omit Sim3.
-We can reference it in the Ovenbird analysis, saying `we did a preliminary simulation study of the model with covariates and random effects, and results were consistent with Sim1, Sim2, and the Ovenbird findings...'.
-What do you think?}
-
-The previous sections studyied the effect of the time to detection assumption in the context of no explanatory variables.
-Here we incorporate fixed and random effects for abundance and detection. 
-For each of the 14 TTDDs, we simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as in Sim1 and Sim2, (ii) using covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
-(See supplement for actual parameter values).
+The previous sections studied effects of time to detection assumptions in the context of no explanatory variables.
+In Sim3, we incorporate fixed and random effects for abundance and detection. 
+We simulated data from each of the 7 mixture TTDDs and fit models from exponential, gamma, lognormal, and Weibull mixture models.  
+We simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
+Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in Sim1 and Sim2, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
+(See Table \ref{tbl:sim3} in supplement for actual parameter values).
 Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
-
 To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.
 
-Patterns in posterior estimates of $\pdet_s$ with respect to mixture and family TTDD forms were the same in Sim3 as in Sim1 and Sim2 (Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family}.
+Patterns in posterior estimates of $\pdet_s$ with respect to mixture and family TTDD forms were the same in Sim3 as in Sim1 and Sim2 -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family} in the Supplement).
+Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (see Supplemental figure).  
+Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
+It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.  
+The above patterns can also be discerned in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
 
-Possible results to focus upon:\\
-\begin{enumerate}
-\item Precision of estimates of $\pdet$ relative to Sim1 and Sim2 (they're narrower in Sim3)
-% These results indicate that covariates provided significant information for reducing uncertainty.
-
-\item $\beta^A$ estimates are pretty much invariant to choice of TTDD
-%For every dataset we simulated, estimates and credible intervals for abundance covariate coefficients $\beta^A$ were virtually the same regardless of the inference TTDD employed.
-
-\item $\beta^D$ estimates from exponential mixtures are more precise than for other families.  The coverage situation is unclear based on size of sim study.
-%Estimates of detection parameters $\beta^D$ varied by context.  
-%Comparisons of mixture and non-mixture models (like in Sim1) produced similar $\beta^D$ estimates except when fit to mixture peaked datasets, where models without a mixture component produced credible intervals twice as wide as those from mixture models with no benefit in terms of coverage.
-%Comparing exponential to other TTDD family models (as in Sim2), time-varying mixture TTDD models produced similar posteriors of $\beta^D$ to one another, but posteriors from the exponential mixture model were both location shifted and either narrower (exponential and nonpeaked data) or wider (peaked data).
-%Differences in estimates of $\beta^D$ parameters to some degree reflected differing estimation of the $\gamma$ mixing parameter -- the exponential model underestimated $\gamma$ with nonpeaked datasets and overestimated it in peaked datasets.
-%Many of the above patterns for nonpeaked datasets can be noted in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
-
-\item DIC can be effective in the same scenarios as Sim1 and Sim2, but it can also be misleading (provide stark example).  This sets the stage for the Ovenbird results.
-%Results from DIC and posterior predictive checks suggested mixture models over non-mixture models for peaked data (DIC differences $>10$) and time-varying mixture TTDDs over the exponential mixture for gamma and lognormal peaked data ($>8$).
-%DIC preferred exponential mixtures to time-varying mixtures for exponential mixture data ($>6$); however, it also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.
+%DIC can be effective in the same scenarios as Sim1 and Sim2, but it can also be misleading.  
+%For peaked mixture data, mixture models were preferred over non-mixture models (DIC differences $>10$), and time-varying mixture TTDDs were preferred over the exponential mixture TTDD for gamma and lognormal peaked data ($>8$).
+%For exponential mixture data, DIC preferred the exponential mixture TTDD to time-varying mixture TTDDs.  
+%However, DIC also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.  
 %In this last case, the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
@@
 
-\end{enumerate}
+\section{Ovenbird analysis}\label{sec:ovenbirds}
 
-\adam{There is a question as to what Figures/Tables would effectively express the above points.
-I think the cleanest approach would be to pick 'representative' datasets and models.
-Or, in my first draft, I simply referenced the Ovenbird output to illustrate several of the points...}
+We fit the Ovenbird dataset with exponential, gamma, lognormal, and Weibull mixture models.  
+For the abundance half of our model, we used four covariates plus two random effects.  
+The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density was over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
+We associated random effects with each survey year and each stand.
+For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve.
+We associated random effects with each observer.  
+Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  
+We centered and standardized all continuous covariates prior to fitting models.
+We ran chains 250,000-375,000 iterations; Geweke diagnostics showed no indication of lack of fit, and effecive sample sizes were over 1000 for all parameters.
+
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{lccccccccc}
+%  \hline
+%Minutes & 0-2 & 2-3 & 3-4 & 4-5 & 5-6 & 6-7 & 7-8 & 8-9 & 9-10 \\ 
+%Count & 596 & 69 & 62 & 68 & 43 & 33 & 35 & 27 & 14 \\ 
+%\hline
+%\multicolumn{10}{l}{Posterior predictive check $p$-values by model:}\\
+%\hline
+%Expo & 0.000 & 1.000 & 1.000 & 0.347 & 0.590 & 0.372 & 0.007 & 0.006 & 0.190 \\ 
+%  Gamma & 0.582 & 0.820 & 0.394 & 0.006 & 0.362 & 0.637 & 0.241 & 0.519 & 0.977 \\ 
+%  LogNormal & 0.571 & 0.943 & 0.471 & 0.005 & 0.322 & 0.554 & 0.165 & 0.407 & 0.974 \\ 
+%  Weibull & 0.582 & 0.862 & 0.401 & 0.005 & 0.359 & 0.616 & 0.221 & 0.509 & 0.982 \\ 
+%\hline
+%  Exponential Mix & 0.533 & 0.804 & 0.577 & 0.021 & 0.501 & 0.668 & 0.190 & 0.338 & 0.918 \\ 
+%  Gamma Mix & 0.593 & 0.720 & 0.431 & 0.015 & 0.439 & 0.670 & 0.246 & 0.476 & 0.947 \\ 
+%  LogNormal Mix & 0.576 & 0.806 & 0.508 & 0.016 & 0.421 & 0.627 & 0.197 & 0.413 & 0.951 \\ 
+%  Weibull Mix & 0.596 & 0.734 & 0.411 & 0.013 & 0.421 & 0.663 & 0.252 & 0.492 & 0.958 \\ 
+%   \hline
+%\end{tabular}
+%\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive p-values -- Pr$\left(D(n_i)^{(rep)} > D(n_i)^{(obs)}\right)$ -- by observation interval for each model across all surveys.
+P-values near zero or one may indicate poor model fit.}
+%\end{table}
+
+Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $\pdet$, and the log10 number of Ovenbirds uncounted.  
+Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from an exponential or nonpeaked mixture TTDD.
+
+As in Sim3, abundance covariate coefficient estimates were virtually the same across all TTDD models.  
+95\% credible intervals for two of the abundance parameters do not contain zero, thereby suggesting notable effects.  
+Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% of the abundance for unlogged sites.  
+Credible intervals for Site Age indicate that each decade of age increases abundance from 1.5-13\%.
+Credible intervals for detection parameters do not indicate significant effects for any of the predictors.
+
+%DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).  
+%These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
+
+In spite of the similarity of effect parameter estimates, the posterior distributions for detection probability and uncounted abundance differ greatly between the exponential and non-exponential models.
+It is clear that the assumption of constant detection leads to much higher and more precise estimates of detection than would be obtained if we are unwilling to make that assumption.
+%The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with finite-mixtures involving a time-varying TTDD.
 
 \begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-Each column presents one family of simulated dataset.  
-Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
-All data and inference models include mixtures.
-Each plot presents one simulated dataset.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\end{figure}
-
-\section{Ovenbird analysis}\label{sec:ovenbirds}
-
-for the Ovenbird analysis and simulations with covariates, chains were 250,000-375,000 iterations.
-
-\subsection{Ovenbird analysis}\label{sec:ovenbirdanalysis}
-
-For the abundance half of our model, we used four covariates plus two random effects.  
-The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density is over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
-We associated random effects with each survey year and each stand, with most stands consisting of three survey sites.  
-For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve \citep{Alldredge2007}.
-69\% of surveys in our dataset involved observers in their first year at MNFB.  
-Preliminary model fits did not support the inclusion of quadratic terms for any detection covariates.  
-We associated random effects with each observer.  
-We centered and standardized all continuous covariates prior to fitting models.
-
-We fit mixture and non-mixture models from each family of TTDD to the Ovenbird dataset.  
-Posterior predictive p-values by interval (Table \ref{tbl:ovencounts}) showed clear lack of fit for the non-mixture exponential model over the first three intervals.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lccccccccc}
-  \hline
-Minutes & 0-2 & 2-3 & 3-4 & 4-5 & 5-6 & 6-7 & 7-8 & 8-9 & 9-10 \\ 
-Count & 596 & 69 & 62 & 68 & 43 & 33 & 35 & 27 & 14 \\ 
-\hline
-\multicolumn{10}{l}{Posterior predictive check $p$-values by model:}\\
-\hline
-Expo & 0.000 & 1.000 & 1.000 & 0.347 & 0.590 & 0.372 & 0.007 & 0.006 & 0.190 \\ 
-  Gamma & 0.582 & 0.820 & 0.394 & 0.006 & 0.362 & 0.637 & 0.241 & 0.519 & 0.977 \\ 
-  LogNormal & 0.571 & 0.943 & 0.471 & 0.005 & 0.322 & 0.554 & 0.165 & 0.407 & 0.974 \\ 
-  Weibull & 0.582 & 0.862 & 0.401 & 0.005 & 0.359 & 0.616 & 0.221 & 0.509 & 0.982 \\ 
-\hline
-  Exponential Mix & 0.533 & 0.804 & 0.577 & 0.021 & 0.501 & 0.668 & 0.190 & 0.338 & 0.918 \\ 
-  Gamma Mix & 0.593 & 0.720 & 0.431 & 0.015 & 0.439 & 0.670 & 0.246 & 0.476 & 0.947 \\ 
-  LogNormal Mix & 0.576 & 0.806 & 0.508 & 0.016 & 0.421 & 0.627 & 0.197 & 0.413 & 0.951 \\ 
-  Weibull Mix & 0.596 & 0.734 & 0.411 & 0.013 & 0.421 & 0.663 & 0.252 & 0.492 & 0.958 \\ 
-   \hline
-\end{tabular}
-\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive p-values -- Pr$\left(D(n_i)^{(rep)} > D(n_i)^{(obs)}\right)$ -- by observation interval for each model across all surveys.
-P-values near zero or one may indicate poor model fit.}
-\end{table}
-
-Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $\pdet$, and the log10 number of Ovenbirds uncounted.  
-Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from a exponential or nonpeaked mixture TTDD.
-As in Sim3, abundance covariate coefficient estimates were virtually the same across all TTDD models.
-
-Credible intervals for two of the abundance parameters do not contain zero, thereby suggesting notable effects.  
-Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% of the abundance for unlogged sites.  
-Credible intervals for Site Age indicate that each decade of age increases abundance from 1.5-13\%.
-
-DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).
-These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
-
-The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with finite-mixtures involving a time-varying TTDD.
-Simulation results demonstrate that the exponential mixture model fits exponential mixture data well, but it does not fit time-varying mixture data well.
-At this point, we have no reliable tool for differentiating field data between the two TTDD types.
-
-\begin{figure}[h!]\centering
-\includegraphics[width=0.88\textwidth]{OVEN/oven_sum/OVEN_Posteriors.pdf}
-\caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to the Ovenbird data.  
+\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors.pdf}
+\caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to Ovenbird data.  
 Abundance and detection parameters are identified by (A) and (D), respectively.  
 Gamma is the mixing parameter.  
-Alpha, k, and sigma\_det are shape parameters for gamma, Weibull, and lognormal distributions, respectively.  
+Shape(Alpha) are shape parameters for gamma, Weibull, and lognormal distributions.  
 p\_det and log10(Uncounted) are the detection probabilty and the number of birds present but not counted across all surveys.  
@@
 
+\begin{appendix}
 \section{Supplement}
@@
 \end{tabular}
-\caption{Table of parameter values used to generate data for Sim1 and Sim2 where $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
+\caption{Table of parameter values used to generate data for: (i) the mixture vs.
+non-mixture simulation, and (ii) the constant vs.
+non-constant rate simulation.
+Here $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
 \end{table}
@@
 \end{tabular}
-\caption{Table of parameter values used to generate data for Sim3 where $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
+\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.
+Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
 \end{table}
+
+\begin{figure}[h!]\centering
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
+\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
+Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
+Each column presents one family of simulated dataset.  
+Upper plots show non-mixture datasets; lower plots show mixture datasets.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
+\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
+All data and inference models include mixtures.
+Each plot presents one simulated dataset.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\end{figure}
+
+\end{appendix}
 

### 2016-07-15 Draft.tex
Implemented several changes to the model/methods that we discussed on Tuesday.

@@
 \newcommand{\ind}{\stackrel{ind}{\sim}}
+\newcommand{\Fm}{F_T^{(M)}}
 
@@
 To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\alpha/2}/\varphi$.
-\adam{Expressing the LN in rate parameterization is a little unconventional...
-We may prefer to be more explicit about the relationship between $\mu$ and $\varphi$ by writing: $T \sim LN(-\log(\varphi), \alpha)$.}
+This parameterization of the lognormal relates to the standard ($\mu, \sigma^2$) parameterization by $\varphi = \exp(-\mu)$ and $\alpha = \sigma^2$.  %$\mu = -\log(\varphi)$
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$.
 
-Given exact times to first detection, we can estimate the parameters in a TTDD and thus estimate $\pdet$ and its uncertainty. 
+Using maximum likelihood, we can estimate the parameters in a TTDD from exact times to first detection and thus estimate $\pdet$ and its uncertainty.
 With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
@@
 Although alternative distributions could be considered, e.g.
-negative binomial, our experience with avian point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here.
-\jarad{Adam: add refs.}
-\adam{The underdispersion finding comes from our own modeling of the Ovenbird dataset.
-I recall no discussion of underdispersion in the literature.
-The primary source for abundance distributions removal models is Etterson.}
+negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here.
 
@@
 Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
-Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and is calculated from the TTDD: $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
 
 It is common in avian point counts to observe increased detections in the first interval relative to an exponential distribution.
-To accommodate this empirical observation, many models of interval-censored removal times include a parameter to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
-We augment our model with a mixing parameter $\gamma\in[0,1]$ which updates the following definitions
-$f_{T|det}(t) = f_T(t)/[\gamma+F_T(C)]$\ and $p_{s1} = \gamma + F_{T|det}(C_1)$.
+To accommodate this empirical observation, many models of interval-censored removal times define a TTDD that includes a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
+We define a TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time distribution $\Fm(t)$ such that $F_T(t) = (1-\gamma) + \gamma \Fm(t)$ for $t>0$.
+%\adam{This formulation became messy, but I think we can omit it.} The updated interval-specific detection probabilities become: $p_{s1} = [1 - \gamma + \gamma\Fm(C_1)]/F_T(C_I)$ and $p_{si} = \gamma[\Fm(C_i)-\Fm(C_{i-1})]/F_T(C_I)$, for $i=2\dots I$.
 When including this parameter and these modifications, we refer to the model as a ``mixture'' model. 
-If $\gamma=0$, then the non-mixture model is recovered.
-\adam{Three things that we'll probably want to talk through: (i) you've switched the mixture components -- my $\gamma$ has become (1-$\gamma$) in the new nomenclature; (ii) I feel that $f_T(t)$ and $F_T(t)$ should describe the entire TTDD and that a different function nomenclature should be used for the `hard to detect' birds.
-(iii) This dates back to my original draft...
-but this particular form of mixture -- with the point mass -- is really specific to interval-censoring.
-It feels incongruous to be mixing a point-mass and a continuous distribution.
-However, writing it propoerly depends upon nomenclature decisions.}
+If $\gamma=1$, then the non-mixture model is recovered.
 
@@
 
-For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit. \adam{Which equation?}
-
-\begin{eqnarray*}
-n_s^{(obs)} &\ind& Po(\lambda_s \pdet_s) \\
-\vn_s &\ind& Mult(n_s^{(obs)}, \vp_s), \vp_s = (p_{s1},\dots,p_{sI}) \\
-p_{s1} &=& \gamma + F_{T|det}(C_1|\alpha,\varphi_s),\, p_{si} = F_{T|det}(C_i|\alpha,\varphi_s)-F_{T|det}(C_{i-1}|\alpha,\varphi_s)\\
-\log(\lambda_s) &=& \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A,\, \xi_j^A \ind N(0,\sigma_{A[j]}^2) \\
-\log(\varphi_{s}) &=& \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D,\, \xi_j^D \ind N(0,\tau_{D[j]}^2)
-\end{eqnarray*}
+For ease of reference, the final full model is provided in equation \eqref{eq:model} where the conditioning of the TTDD cdf on $\alpha$ and $\varphi_s$ is made explicit.
+\begin{align}
+n_s^{(obs)} &\ind \Po(\lambda_s \pdet_s) \nonumber\\
+\vn_s &\ind \Mult(n_s^{(obs)}, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
+p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
+p_{si} &= \left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
+\log(\lambda_s) &= \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A; \qquad \xi_j^A \ind N(0,\sigma_{A[j]}^2) \nonumber\\
+\label{eq:model}\log(\varphi_{s}) &= \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D; \qquad \xi_j^D \ind N(0,\tau_{D[j]}^2)
+\end{align}
 We adopt a Bayesian approach and therefore require a prior over the model parameters.
@@
 We discarded half of the iterations as warmup and then thinned by 10.  
-We monitored convergence of the MCMC chains using Geweke diagnostics \citep{Geweke1991}.
-\jarad{Adam: what did you do with the Geweke diagnostics?} \adam{I'm unclear what you intend with this question.
-I used them to monitor convergence.
-Any time a chains raised a Geweke flag, I re-ran the model.}
+We monitored convergence of the MCMC chains using Geweke z-score diagnostics \citep{Geweke1991} and reran models if lack of convergence was indicated by a non-normal distribution of the z-scores. 
 We reran models if the effective sample size for any parameter was below 1000.  
 The number of iterations used depended on the model and is detailed later. 
-For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed \jarad{Adam: how did they fail?}\adam{I don't remember the error message, but Stan failed with no output generated.} unless we specified initial values, which we supplied from true parameter values when known or from the posterior means of preliminary short-run models.
-\jarad{Adam: this doesn't make sense.
-Why would the means from these preliminary short-run models be any better than letting these preliminary short-run models continue?
-You claim seems to be that letting these preliminary short-runs continue would mean that those models would fail.}\adam{I have brilliantly deleted the original files, but my memory suggests that models tended to fail early.
-Some short-run models failed quickly, too, but if I ran a model for 1000 iterations, the results provided stable initial values which I could subsequently used whenever I tweaked the models and had to rerun everything.
-I'm thinking maybe these sentences can just be dropped.}
+For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed to run unless care was taken in the specification of initial values.
 
@@
 A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
-\jarad{Adam: do we want a figure demonstrating this point?}\adam{I don't think a figure is required.}
 

### 2016-07-22 Draft.tex
Mid-stream: expunging DIC.  Also, other minor changes and removals of comments.

@@
 
+\newif\ifdic   % Type \ifdic at the beginning of any optional section... use \fi to denote the end of the optional section
+  % See: http://tex.stackexchange.com/questions/33576/conditional-typesetting-build
+%\dictrue         % comment out if I want to hide DIC and model fit paragraphs
+
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-\newcommand{\jarad}[1]{{\color{Orange} #1}}
+\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
 \newenvironment{indpar}[1]%
@@
 The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
-\adam{I liked the structure that was imparted in the earlier draft by having a data model, abundance model, and detection model.}
 
@@
 Although alternative distributions could be considered, e.g.
-negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed and thus we will use the Poisson assumption here.
+negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed, and thus we will use the Poisson assumption here.
 
@@
 It is common in avian point counts to observe increased detections in the first interval relative to an exponential distribution.
-To accommodate this empirical observation, many models of interval-censored removal times define a TTDD that includes a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
-We define a TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time distribution $\Fm(t)$ such that $F_T(t) = (1-\gamma) + \gamma \Fm(t)$ for $t>0$.
-%\adam{This formulation became messy, but I think we can omit it.} The updated interval-specific detection probabilities become: $p_{s1} = [1 - \gamma + \gamma\Fm(C_1)]/F_T(C_I)$ and $p_{si} = \gamma[\Fm(C_i)-\Fm(C_{i-1})]/F_T(C_I)$, for $i=2\dots I$.
-When including this parameter and these modifications, we refer to the model as a ``mixture'' model. 
-If $\gamma=1$, then the non-mixture model is recovered.
+To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
+We specify a mixture TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time detection distribution $\Fm(t)$.  
+The mixture TTDD cdf is defined: $F_T(t) = (1-\gamma) + \gamma \Fm(t)$ for $t>0$.
+If $\gamma=1$, the non-mixture model is recovered.
 
@@
 
-% Effects: * - means signif
-% Observer variation - Farnsworth2002, LinkSauer98, LinkSauer97, Dief07 cites Sauer94 & Dief03, Simons et al03
-% First-year observer - LinkSauer98
-% Time of day - Farnsworth2002, Soly13*, Amundson
-% Season (jdate) - Farnsworth2002, Soly13*, Amundson, Dief07
-% Quadratic effects - Soly13
-% Tree cover (density) - Soly13
-% Year (on detection!) - Reidy11, see also Norvell03
-% See Table 1 of Johnson2008, see McShea and Rappole, lit review from Warren2013, Rosenstock02
-
+\ifdic\
 \subsection{Model diagnostics}
-
-\jarad{Adam: do we want posterior predictive pvalues at all?}\adam{Interesting question.
-I don't think they're essential to our model.
-I am open to omitting them.
-However...
-Amundson et al.
-are vague about the posterior p-values -- ``we used Bayesian P values generated from the posterior predictive distributions to assess goodness-of-fit" -- they don't state what aspect of fit they tested -- it looks like maybe they use the model-fit value of $\pdet$ as their statistic?!?
-Anyhow, I suspect they misapplied the tool, and I somewhat wanted to counteract that with an example of a better application of it.
-Also, I think it's important to show that we checked some measure of fit.}
 
@@
 
-\jarad{Adam: do we want to remove DIC?}\adam{As with posterior predictive checks, I can live without DIC, which has been nothing but headaches.}
-
 We compared different models by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
 For the full model with random effects (Ovenbirds and Sim3), which is a missing data model, we used formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
+\fi
 
@@
 \adam{For nonpeaked non-mixture data, the mixture model is less biased for lognormal and Weibull scenarios, and CIs are 15-40\% narrower.}
-Similarly DIC cannot clearly determine that a non-mixture model should be preferred over a mixture model in this scenario.
 
+\ifdic
 \begin{table}[ht]\centering\small
@@
 $\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
-\jarad{Adam: we should consider splitting this table in two.
+}
+\label{tbl:sim1}
+\end{table}
+
+\else
+
+\begin{table}[ht]\centering\small
+\begin{tabular}{l|l|l|l|cccc|cccc}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+ \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+  \hline
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 \\
+ & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00  \\ 
+ & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 \\ 
+  \cline{3-12}
+ & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88  \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 \\ 
+ & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 \\ 
+ & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 \\ 
+  \cline{2-12}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 \\
+ & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 \\ 
+ & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
+  \cline{3-12}
+ & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
+ & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 \\ 
+ & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture vs. non-mixture model fits (Sim1).  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+\jarad{We should consider splitting this table in two.
 The main point is what happens when the data arise from a mixture model but a non-mixture distribution is used.
@@
 \end{table}
+\fi
 
@@
 For the non-mixture models, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ and the other models underestimating.
-While DIC could not easily differentiate a mixture from a non-mixture model for inference when a non-peaked simulation model was used, DIC clearly preferred mixture models when the data arose from an exponential distribution or any of the peaked distributions.
 
@@
 
-% When data derived from a mixture model, then inference models lacking a mixture component misestimated the detection probability $\pdet$ badly in every case.  
-% This was particularly egregious for datasets featuring exponential and peaked mixtures, where non-mixture model coverage was always negligible.  
-% Conversely, when data derived from a non-mixture model, inference models incorporating a mixture component performed as well as the true model.  
-% Indeed, for nonpeaked non-mixture data, the mixture inference model provided more accurate posterior estimates of $\pdet$ than did the true non-mixture inference model.  
-% The one case where the non-mixture model performed better was the non-mixture exponential dataset, where both inference models estimated $\pdet$ accurately, but the mixture inference model yielded credible intervals about 25\% wider than those from its mixture counterpart.
-% Mixture model posterior medians for $\pdet$ were always lower than non-mixture posterior medians for time-varying datasets and higher for exponential TTDDs.
-
-% \adam{Ack!  I'm having a Dickens of a time abbreviating the following paragraph.  I could use a hand.... maybe keep just the first sentence?}
-% \jarad{We shouldn't be talking about results that we are not showing. The results should either be in the manuscript or the supplement.}
-% Because abundance estimates are inversely related to detection estimates, they mirrored the observations for $\pdet$ above and shared the same coverage rates.  
 % Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
@@
 % For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
-
-% Neither DIC nor posterior predictive checks indicated a clear preference for mixture models with nonpeaked data, even though non-mixture models displayed stronger bias and poorer coverage.
-DIC did reveal a preference for mixture models with exponential and peaked datasets.
-In these cases, posterior predictive checks clearly showed poor fit from the non-mixture models with p-values of 1.00 in the second and sometimes third observation intervals.
-These p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
 
@@
 
+\ifdic
 \begin{table}[ht]
@@
 \end{table}
+
+\else
+
+\begin{table}[ht]
+\footnotesize\centering
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\%  & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 0.76 & 0.41 & 0.69 & 1.00 \\
+& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 0.85 & 0.72 & 0.44 & 0.69 \\ 
+& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.79 & 0.49 & 0.69 & 1.00 \\ 
+\cline{2-11}
+& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & 0.68 & 0.24 & 0.25 & 0.81 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
+& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 0.84 & 0.73 & 0.44 & 0.81 \\ 
+& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 0.66 & 0.15 & 0.19 & 0.75 \\
+   \hline
+\end{tabular}
+\vspace{0.5cm}\\
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.67 & 0.24 & 0.31 & 1.00 \\
+& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & 0.77 & 0.48 & 0.44 & 1.00 \\ 
+& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
+\cline{2-11}
+& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & 0.63 & 0.20 & 0.25 & 0.75 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & 0.79 & 0.52 & 0.69 & 0.88 \\ 
+& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & 0.89 & 0.88 & 0.12 & 0.69 \\ 
+& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture inference models of all families fit to mixture datasets.  
+Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
+\label{tbl:sim2}
+\end{table}
+\fi
 
@@
 
+\ifdic
+\subsection{Model Fits}\label{sec:fits}
+
+In Sim1, DIC and posterior predictive checks could not easily differentiate a mixture from a non-mixture model for inference when a nonpeaked simulation model was used, even though non-mixture models displayed strong bias and poor coverage.
+Both tools clearly preferred mixture models when the data arose from a mixture exponential distribution or any of the peaked mixture distributions. 
+In these cases, posterior predictive p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
+\fi
+
 \section{Ovenbird analysis}\label{sec:ovenbirds}
@@
 We ran chains 250,000-375,000 iterations; Geweke diagnostics showed no indication of lack of fit, and effecive sample sizes were over 1000 for all parameters.
+
+% Effects: * - means signif
+% Observer variation - Farnsworth2002, LinkSauer98, LinkSauer97, Dief07 cites Sauer94 & Dief03, Simons et al03
+% First-year observer - LinkSauer98
+% Time of day - Farnsworth2002, Soly13*, Amundson
+% Season (jdate) - Farnsworth2002, Soly13*, Amundson, Dief07
+% Quadratic effects - Soly13
+% Tree cover (density) - Soly13
+% Year (on detection!) - Reidy11, see also Norvell03
+% See Table 1 of Johnson2008, see McShea and Rappole, lit review from Warren2013, Rosenstock02
 

### 2016-07-23 Draft.tex
DIC moved to a conditional block for exclusion (or inclusion). \gamma added to model summary fo p_{si}

@@
 p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
-p_{si} &= \left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
+p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
 \log(\lambda_s) &= \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A; \qquad \xi_j^A \ind N(0,\sigma_{A[j]}^2) \nonumber\\
@@
 Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
 \label{tbl:sim2}
@@
 In contrast, when the lognormal model is used for inference, it performed worse when the data arose from a gamma or Weibull and, when the Weibull model was used for inference, it performed worse when the data arose from a lognormal.
-Other than sometimes providing evidence against the exponential mixture model, DIC was not helpful in distinguishing amongst these models. 
 \adam{I'd be okay with jettisoning this paragraph beyond the first sentence.  Conversely, rather than praising gamma, we could discourage lognormal.}
@@
 
-% Mixture exponential models returned more accurate estimates of $\pdet$ than other models when data came from an exponential mixture, and 95\% credible intervals for $\pdet$ were from one-third to one-half as wide.
-% However, mixture exponential models yielded notably biased estimates with poor coverage for all non-exponential data, overestimating detection for nonpeaked datasets and underestimating it for peaked datasets.
-% Differences among non-constant TTDDs were less pronounced.  
 % Gamma models were more accurate than others for nonpeaked mixture datasets.  
@@
 % Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
-% For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had bias of 5-20\% for Weibull and 13-23\% for lognormal.
-% So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
+% For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had positive bias of 5-20\% for Weibull and 13-23\% for lognormal.
+% So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate (but in the opposite direction) with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
 % For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
-% 
-% In most cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage suggested one was inferior.
-For  exponential mixture fits of gamma and lognormal peaked datasets, DIC did show a clear difference, and posterior predictive checks revealed lack-of-fit with p-values of 1.00 in the second observation interval.
 
@@
 
-%DIC can be effective in the same scenarios as Sim1 and Sim2, but it can also be misleading.  
-%For peaked mixture data, mixture models were preferred over non-mixture models (DIC differences $>10$), and time-varying mixture TTDDs were preferred over the exponential mixture TTDD for gamma and lognormal peaked data ($>8$).
-%For exponential mixture data, DIC preferred the exponential mixture TTDD to time-varying mixture TTDDs.  
-%However, DIC also preferred the exponential mixture to time-varying mixtures (by 4-8) for lognormal nonpeaked data.  
-%In this last case, the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
-%In conjunction with Sim2 results, this last finding calls into question the utility of DIC for nonpeaked datasets.
-%DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
-
 \ifdic
@@
 
-In Sim1, DIC and posterior predictive checks could not easily differentiate a mixture from a non-mixture model for inference when a nonpeaked simulation model was used, even though non-mixture models displayed strong bias and poor coverage.
+In Sim1, DIC and posterior predictive checks could not easily differentiate a mixture from a non-mixture model for inference when a nonpeaked simulation model was used, even though non-mixture models displayed bias and poor coverage.
 Both tools clearly preferred mixture models when the data arose from a mixture exponential distribution or any of the peaked mixture distributions. 
 In these cases, posterior predictive p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
+
+In Sim2, DIC did show a clear difference for exponential mixture fits of gamma and lognormal peaked datasets, and posterior predictive checks revealed lack-of-fit in the same scenarios, yielding p-values of 1.00 in the second observation interval.
+However, in all other cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage across simulations suggested one was inferior.
+
+For Sim3, DIC was effective in the same scenarios where it was effective in Sim1 and Sim2.  
+Additionally, for exponential mixture data, DIC preferred the exponential mixture TTDD to time-varying mixture TTDDs.  
+Unfortunately, DIC could also be misleading.  
+DIC favored the exponential mixture to time-varying mixtures (by 4-8) when the data were in fact lognormal nonpeaked data.  
+Yet the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
+In conjunction with the ineffectiveness of DIC for nonpeaked data in Sim2, this last finding calls into question the utility of DIC for nonpeaked datasets.
+
+DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
 \fi
@@
 
-%DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).  
-%These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
+\ifdic
+DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).  
+These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
+\fi
 
@@
 
+\ifdic
 In a limited set of situations, posterior predictive checks and DIC can identify when non-mixture or exponential TTDDs inadequately describe the marginal pattern of detections over time.  
 But in many situations, neither tool points a clear way to select the correct family of TTDD.  
+\fi
 Given the bias that can result from assuming constant detection or omitting a mixture component, conservatism dictates that non-constant TTDDs with a mixture should be preferred.  

### 2016-07-24 Draft.tex
Revisions to Discussion.  Still needs the last couple of paragraphs on continuous-time modeling and removal-distance sampling.

@@
 A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007}.
-%\adam{Others, like Mantyniemi.}
+\adam{I need to add others, like Mantyniemi.}
 In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
@@
 
-\jarad{Adam: we should include the Stan models in the supplement.}
+\jarad{We should include the Stan models in the supplement.}
 
@@
 When comparing the different three-parameter TTDDs, model misspecification is not as serious an issue because these models can adequately account for the interval-censored times to detection. 
-Nonetheless, it appears the gamma mixture model is better able to account for data from lognormal or Weibull mixture models. 
+Nonetheless, it appears the gamma mixture model is better able to account for data from most non-peaked datasets, while gamma and Weibull mixture models have comparable performance across peaked datasets.
 In the gamma quadrant of Table \ref{tbl:sim2}, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility.
-In contrast, when the lognormal model is used for inference, it performed worse when the data arose from a gamma or Weibull and, when the Weibull model was used for inference, it performed worse when the data arose from a lognormal.
-\adam{I'd be okay with jettisoning this paragraph beyond the first sentence.  Conversely, rather than praising gamma, we could discourage lognormal.}
+In contrast, when the lognormal model is used for inference, it performs worse across nearly all data types.
+However, gamma models required 9-11 times the computer run time as did lognormal and Weibull models, and almost 25 times that of exponential models.
 
@@
 
-Historically, analysis of removal-sampled data assumes constant detection rates throughout the sampling period.  
-This assumption provides an intuitive null model and is analytically tractable.  
-However, although analysis of removal models has grown more sophisticated in recent years, the constant-detection assumption nonetheless continues unquestioned.  
-We have formulated a model for non-constant detection using a time-to-event approach within a hierarchical N-mixture framework.  
-Our results demonstrate that, for better or worse, the assumption of constant detection is a very strong assumption.  
-When it is correct, it is the most accurate and precise of TTDD models, but when detection rates are not constant, it results in significant bias.  
-Models with non-constant TTDDs, on the other hand, can adapt to a wider variety of detection patterns.
+Historically, analysis of removal-sampled point-count survey data assumes constant detection rates throughout the sampling period.  
+We have formulated a model for these data that allows for non-constant detection rates.
+Our model adopts a time-to-event approach within a hierarchical N-mixture framework, and it allows times to first detection to be modeled according to flexibly defined TTDD families.
+Our results show that non-constant TTDDs can return reasonable estimates of detection probabilities across a variety of time-to-detection data patterns, whereas traditional constant-rate TTDDs return biased and overly precise estimates when data deviate from the constant-rate assumption.
+Because the exponential TTDD is a special case of both gamma and Weibull TTDDs, we can interpret the differences in estimation between models as resulting from the information conveyed by the assumption of constant detection.
+Phrased differently, in the Bayesian context, the exponential model is equivalent to a gamma or Weibull model where the prior for the shape parameter $\alpha$ is defined: Pr($\alpha=1) = 1$; and the differences between the models results from the strength of this prior.
+Among the TTDDs modeled in our study, the gamma mixture TTDD gave the most accurate estimates across data distributions, though the Weibull mixture TTDD was comparable for peaked datasets.
 
-We modeled two varieties of heterogeneous detection in conjunction with one other: (i) time-varying detection rates, as modeled by non-exponential TTDDs, and (ii) detection heterogeneity among subgroups of individuals, as modeled by mixture TTDDs.
-The exponential TTDD, even when modeled with a mixture component, results in biased estimates of detection when detection actually varies with time.  
-Gamma, Weibull, and lognormal TTDDs, with their extra parameter, are more accurate in handling non-constant detection data, though they are less precise when estimating data with constant detection.
+We have additionally demonstrated for non-constant models the utility of using a mixture TTDD formulation.
+Inference models with a mixture component are accurate whether the data have a mixture or not, whereas inference models without the mixture are badly biased when the data do feature a mixture.  
+Mixture models even outperform non-mixture models when the data are non-mixture and nonpeaked.
+%For a discussion of mixtures as heterogeneity, see Pledger 2000, dating all the way back to Otis 1978.
 
-The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Reidy2011}.  
-The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  
-Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  
-Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  
-Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  
-Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.
-% ``It is tempting to choose classes of models that yield narrower confidence intervals for N; however, Coull's (1997) extensive simulation studies (summarized by Coull and Agresti (1999)) clearly reveal that this behavior can produce erroneous inferences." -- Dorazio and Royle 2003
+If the estimation of effect parameters and the roles of explanatory variables are the primary interest, then our results suggest that the exact choice of TTDD may not be important.  
+Abundance effect estimates are similar regardless of the chosen TTDD.
+Detection effect estimates, while conditional on the mixing parameter $\gamma$, are similar across all mixture non-exponential TTDDs.  
+These finding may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.
 
@@
 \fi
-Given the bias that can result from assuming constant detection or omitting a mixture component, conservatism dictates that non-constant TTDDs with a mixture should be preferred.  
-Among the TTDDs we considered, the mixture gamma TTDD appears to provide the best combination of accuracy and precision for nonpeaked data.  
-For peaked data, gamma and Weibull models seem about equally as good.  
-MCMC sampling from gamma models required from 5-10 the run time as lognormal and Weibull models.
 
-If the estimation of effect parameters is the primary interest, then our results suggest that the exact choice of TTDD may not be important.  
-Abundance effect estimates are similar regardless of the chosen TTDD, while detection effect estimates are similar for all mixture non-exponential TTDDs.  
-These finding may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.
-
-Because our focus is detection, we assumed Poisson-distributed abundance for simplicity.  
-Abundance distributions that have been used to account for overdispersion include negative binomial, zero-inflated Poisson, and Poisson with survey-level random effects.  
-The pattern of counts in our Ovenbird data actually reflect underdispersion, so a Conway-Maxwell Poisson distribution, which can model both overdispersion and underdispersion, may be more appropriate \citep{Wu2015}\adam{apparently, Sellers et al.
-(in Wu) has a comprehensive overview of the CMP.}
-
-% Overview: Wu15
-% ZIP: Etterson, Solymos12
-% NB: Royle04(?)
-
-At present, Stan does not sample discrete parameters, so the ability to marginalize out the discrete-valued latent variables $N_{s}$, as we did, can greatly facilitate MCMC sampling.
-
-% \adam{I have contemplated on and off including a discussion of broad priors and the limit behavior as $\varphi \to 0$, but I think it's not really on topic.  
 % Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}
@@
 
-% Really, I have totally ignored the discussion of the true value of p, which I have tended to simulate in a fairly high range
+% Really, I have totally ignored the discussion of the true value of p, which I have tended to simulate at a fairly high value
 
-To improve our choice of TTDD and/or its parameters, it makes sense to obtain more data in one of a few ways.  
-One approach is to collect complete detection history records, not just first time-to-detection observations \citep{Alldredge2007}.  
+If the estimation of abundance is the primary interest, then we may reasonably ask whether removal sampled point-count surveys are really the data we want.  
+The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would only be detected if the observation period lasted longer.  
+As such, it is entirely based upon extrapolation and, in terms of our methodology, correctly estimating the probability in the tail of the TTDD based on recorded observations.
+The constant detection assumption constrains uncertainty in that extrapolated tail but at the potential cost of bias.
+When we allow for non-constant detection, estimation of the tail probability becomes more variable.
+In theory, estimation of the unsampled tail probability can be improved by conducting longer observations, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
+Without extending the observation period, we can still obtain more information by recording complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}.
 This may not be feasible in studies like MNFB, where many species are observed.
-Another approach is to conduct a longer survey that better assesses the tail distribution of the TTDD; however, the longer the survey is, the greater the risk that individuals enter/depart the study area or are double-counted, which violates removal sampling assumptions \citep{LeeMarsden2008, Reidy2011}.
-A third idea is to obtain more precise time-to-detection data, even exact time-to-detection data.  
-In some early explorations of our model considering just an exponential mixture TTDD, we found that credible intervals for Ovenbird detection probability were 20-25\% narrower using 9-interval surveys compared to more traditional 3-interval surveys (0-3, 3-5, and 5-10 minutes).
-For all of the above, a small pilot study may be a practical starting point.
 
-With the increasing use of microphone arrays for automated detection \adam{refs}, it is tempting to think TTDD models may be valuable for microphone-collected data.  
-This should be the case when variations in detection result solely from organismal behaviors, such as bout singing, or if we wish to model individual random effects in the detection model.  
-However, some explanations for non-constant detection are explicitly related to the observation process --- either effects of the observer on animal behavior or variations in observer effort caused by survey methodology.
-The assumption of constant detection seems more plausible in studies where observers are absent, though detection heterogeneity may still exist across subgroups of the population.
+%With the increasing use of microphone arrays for automated detection \adam{refs}, it is tempting to think TTDD models may be valuable for microphone-collected data.  
+%This should be the case when variations in detection result solely from organismal behaviors, such as bout singing, or if we wish to model individual random effects in the detection model.  
+%However, some explanations for non-constant detection are explicitly related to the observation process --- either effects of the observer on animal behavior or variations in observer effort caused by survey methodology.
+%The assumption of constant detection seems more plausible in studies where observers are absent, though detection heterogeneity may still exist across subgroups of the population.
 
-Application of TTDD modeling within the N-mixture framework should be straight-forward for other time-to-detection datasets, such as full detection-history and double-observer.  
-The key difference will be in the data model, in that responses will take a different form than multinomially distributed first times to detection.  
-But the underlying abundance and detection models remain the same.  
-\adam{Time-heterogeneous models have been done for mark-recapture (Farcomeni and Scacciatelli 2013, who reference Hwang and Chao 2002).}
+%\adam{Time-heterogeneous models have been done for mark-recapture (Farcomeni and Scacciatelli 2013, who reference Hwang and Chao 2002).}
 
-The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.
+The time-heterogeneous continuous time-to-detection approach in our model mirrors continuous-time work in mark-and-recapture \citep{HwangChao2002, FarcomeniScacciatelli2013, Borchers2014} and removal-distance sampling \adam{How do I cite Borchers' preprint?}; however, it is a departure from the probability-of-detection approach that predominates removal-sampled point-count surveys
+
+\citep{Farnsworth2002, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}
+
+\citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.  
 In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.

### 2016-07-24 Draft.tex
Finished Discussion.  Added some citations.

@@
 \adam{I need to add others, like Mantyniemi.}
-In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
-lack of settling in period \citep{LeeMarsden2008, Johnson2008}, can all lead to time-varying rates of detection.
-%In particular, animal behaviors, e.g.
-intermittent singing in birds and frogswhales \citep{Scott2005, Diefenbach2007, Reidy2011}, observer behaviors, e.g.
-pre-counting and disturbing species \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007, LeeMarsden2008, Johnson2008}, and survey methodology, e.g.
-lack of settling in period \jarad{what else were you thinking?
-ref?}, can all lead to detection heterogeneity \citep{Farnsworth2005}.
-%Survey methodology itself can affect recording of detection times.  
-% Observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
-% Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
-%In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time .
+In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Otis1978, Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
+saturation or lack of settling in period \citep{Petit1995, LeeMarsden2008, Johnson2008}, can all lead to time-varying rates of detection.
 
@@
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
-One common strategy to deal with data that do not fit a constant-detection assumption is to model increased detection probability in the initial observation period via a mixture component \citep{Farnsworth2002, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
+One common strategy to deal with data that do not fit a constant-detection assumption is to model increased detection probability in the initial observation period via a mixture component \citep{Farnsworth2002, Farnsworth2005, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
 We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
@@
 The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.  
-Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014}.
+Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}.
 
@@
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
-Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.
-% 
+Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine equal-duration intervals \citep{Reidy2011, Amundson2014}.
+
 % \adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.  
@@
 It is common in avian point counts to observe increased detections in the first interval relative to an exponential distribution.
+This is often understood to reflect unmodeled detection heterogeneity across behavioral groups in the study population.
+Failure to account for such heterogeneity in the constant-detection scenario leads to negative bias in abundance estimates \citep{Otis1978}.
 To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
@@
 
-% Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}
-
-% All estimates are based on extrapolation of the observed detections\\
-% -- a) explore the tail more?  I think that'd require a lot of data\\
-% -- b) requires extended observation periods or full detection data (Alldredge?)\\
-% same goes for sample size
+% Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
 
@@
 The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would only be detected if the observation period lasted longer.  
-As such, it is entirely based upon extrapolation and, in terms of our methodology, correctly estimating the probability in the tail of the TTDD based on recorded observations.
-The constant detection assumption constrains uncertainty in that extrapolated tail but at the potential cost of bias.
-When we allow for non-constant detection, estimation of the tail probability becomes more variable.
+As such, it is entirely based upon extrapolation and correctly estimating the probability in the tail of the TTDD based on recorded observations.
+An assumption of constant detection constrains uncertainty in the extrapolated tail but at the cost of potential bias.
+When we allow for non-constant detection, estimation of the tail probability becomes so variable that we were only able to estimate Ovenbird abundance within an order of magnitude.
 In theory, estimation of the unsampled tail probability can be improved by conducting longer observations, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
-Without extending the observation period, we can still obtain more information by recording complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}.
-This may not be feasible in studies like MNFB, where many species are observed.
+Without extending the observation period, an alternate to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB, where many species are observed.
 
@@
 
-%\adam{Time-heterogeneous models have been done for mark-recapture (Farcomeni and Scacciatelli 2013, who reference Hwang and Chao 2002).}
+Point-count survey data often include the recorded distance between observer and detected organism.
+Because our focus has been on modeling variations in detection rates during the survey period, we have not incorporated distance into our model.
+Consequently, our application of a TTDD represents an averaging across distance classes, which induces systematic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.  
+To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier, as is done in \adam{How do I cite Borchers' preprint?}.  
+This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Amundson2014}.
+The differences between these implementations may impact estimates of detection and abundance, especially in the presence of behavioral heterogeneity in availability rates across subgroups of the study population.
+This is an area of ongoing exploration.
+%Our adoption of a time-heterogeneous continuous time-to-detection approach in our model mirrors continuous time work in mark-and-recapture \citep{HwangChao2002, FarcomeniScacciatelli2013, Borchers2014}.
 
-The time-heterogeneous continuous time-to-detection approach in our model mirrors continuous-time work in mark-and-recapture \citep{HwangChao2002, FarcomeniScacciatelli2013, Borchers2014} and removal-distance sampling \adam{How do I cite Borchers' preprint?}; however, it is a departure from the probability-of-detection approach that predominates removal-sampled point-count surveys
+%Examples of the interval-censored approach.
+%\citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011, Solymos2013, Amundson2014, Reidy2016}.
 
-\citep{Farnsworth2002, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}
-
-\citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.  
-In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.
-I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  
-Ditto Amundson.
-See also McCallum 2005.}.  
-In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
-Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
-The probability of detection is thus formulated as availability times perceptibility: $\pdet = P_aP_b$.  
-Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.
-
-The time-to-detection model as presented here in our study does not incorporate distance.  
-Consequently, our definition of a TTDD necessarily represents an averaging across distance classes and a consequent loss of information if distance data are available.  
-We can modify our model to account for distance.  
-To be consistent with earlier distance-removal studies \citep{Farnsworth2005, Amundson2014}, we could: (i) define a time-to-\textit{availability} distribution $F_T^A(t)$ in the same manner that we have defined a TTDD in this study, (ii) modify the equation for detection probability to reflect the detection distance to individual $q$: $p_{sq}^{(det)} = g(d_q) F_T^A(C_I|\varphi_s, \boldsymbol{\theta})$, and (iii) tweak our data model accordingly.
-
-However, our analysis makes possible a different strategy that is more consistent with a time-to-event conceptualization of the problem.  
-Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  
-In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d_q)$.
-So, $h_{sq}(t) = g(d_q) h_{s}^A(t)$.  
-From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\varphi_s, \boldsymbol{\theta}) \right]^{g(d_q)}$.
-
-In short, the time-to-event conceptualization frames perceptibility as a rate-of-availability multiplier rather than as a probability-of-availability multiplier.  
-In a context where rates of availability are allowed to vary from individual to individual, this makes an intuitive sense --- a bird that sings twenty times during the observation period should have its overall detection probability discounted much less than a bird at the same distance which only sings once.
-This is an area of ongoing exploration.
+%In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.
+%\adam{I'm  beginning to think this may have too-numerous antecedents to cite... in Amundson and in Reidy, the earliest reference is Marsh and Sinclair 1989.  See also McCallum 2005.}.  
+%In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
+%Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
+%The probability of detection is thus formulated as availability times perceptibility: $\pdet = P_aP_b$.  
+%However, our analysis makes possible a different strategy that is more consistent with a time-to-event conceptualization of the problem.  
+%Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  
+%In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d_q)$.
+%So, $h_{sq}(t) = g(d_q) h_{s}^A(t)$.  
+%From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\varphi_s, \boldsymbol{\theta}) \right]^{g(d_q)}$.
+%In short, the time-to-event conceptualization frames perceptibility as a rate-of-availability multiplier rather than as a probability-of-availability multiplier.  
+%In a context where rates of availability are allowed to vary from individual to individual, this makes an intuitive sense --- a bird that sings twenty times during the observation period should have its overall detection probability discounted much less than a bird at the same distance which only sings once.
 
@@
 
-%\subsection{Lit Notes on Distance and Poisson Abundance}
-%--- Borchers is all about this\\
-%--- Farnsworth2005, EffordDawson2009, Amundson2014\\
-%----- particularly noticeable with finite mixture models\\
-%- Oedekoven2013 is all about covariates on distance sampling curve\\
-%- necessity of using distance, too, is backed by Solymos2013, (Buckland et al 2001, 2004), Farnsworth2005, EffordDawson2009...\\
-%
-%Poisson assumption (do I want to discuss underdispersion?\\
-%--- Denes et al 2015 discuss how Poisson assumption leads to bias in the absence of detection heterog (ref to Solymos et al 2012) or when inflated zeros are present (Joseph et al 2009) [with the obvious remedies being: NegBin and ZIP, respectively]...
-but is that negated by our underdispersion?\\
-%
-%`Unmodeled individual heterogeneity causes population size to be underestimated using capture-recapture and removal methods' (from Efford and Dawson, who cite Otis et al. 1978)\\
+We recommend that time-heterogeneous detection rates be explicitly modeled in analyses involving removal-sampled point-count survey data where estimation of abundance is a primary objective.  
+The assumption of constant detection, while computationally simple and reasonable as a null model, proves to be rather informative and can result in pronounced bias.  
+Meanwhile, the causes of non-constant detection -- i.e., observer effects on behavior and systematic variations in observer effort -- are both plausible and not trivially discounted.  
+It would be nice if the data itself could inform us whether constant detection is a reasonable assumption; however, our preliminary efforts to diagnose this assumption using deviance information criterion (DIC) and posterior predictive check statistics have led to weak and sometimes erroneous findings.
+Development of such a diagnostic tool would be useful, but given the limitations of first time-to-detection data, we are not confident a reliable tool could be easily developed.  
+We believe that more informative data collection, such as complete time-to-detection histories and microphone arrays, offer more effective tools for time-to-event modeling going forward.
 
-% Ideas for the future
-% (i) Distance-removal time-to-event model
-% (ii) Fit more species from this dataset over full survey duration; maybe some model selection (even Bayesian)
-% (iii) Take a stab at modeling habitat
-% (iv) Model for non-independent singing?
-% (v) Model for density-dependent singing? (only valid, I think, if we have lots of underdispersion from(ii))
-% (vi) I know I've  been warned off it, but automated detection intrigues me
-% (vii) Further testing of existing model: sensitivity to priors, performance for varying detectability levels, sample size, # intervals
-% (viii) Incorporate spatial random effects.  
-This work has already been done, just not for TTDD models, but the real spatial dependency should be in abundance, not detection
-% (ix) Comparison of methods -- just looking at my lit review, there are gadzooks strategies that people have proposed.  
-Combining/comparing them seems maybe informative?  To what degree has this already been done?
+\adam{Two topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree:\\
+\begin{enumerate}
+\item DIC and posterior predictive checks
+\item Selection of priors
+\item Values for $\pdet$ used in simulations
+\end{enumerate}
+}
 
+\newpage
 \begin{appendix}
 \section{Supplement}
+
+\begin{figure}[h!]\centering
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
+\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
+Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
+Each column presents one family of simulated dataset.  
+Upper plots show non-mixture datasets; lower plots show mixture datasets.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
+\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
+All data and inference models include mixtures.
+Each plot presents one simulated dataset.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\end{figure}
 
@@
 
-\begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-Each column presents one family of simulated dataset.  
-Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
-All data and inference models include mixtures.
-Each plot presents one simulated dataset.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\end{figure}
+\end{appendix}
 
-\end{appendix}
+\newpage
 

### 2016-08-05 Draft.tex
Make some edits

@@
 \adam{The previous sentence ends mid-thought.  Is this for the Discussion section?}
+\jarad{yes}
 
@@
 Survey sites were selected from sawtimber red pine stands with no recent logging activity.  
-Each stand had up to four sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
+Each stand had up to four \jarad{I thought it was 3} sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
 The data include 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, an indicator of select-/partial-cut logging during the 1990s.
+
+\jarad{did we use unlimited distance point counts or did we use $<$100m?}
 
@@
 
-\subsection{Interval-censored times to detection}
+\subsection{Interval-censored times to detection} \label{s:interval}
 
@@
 
-To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ via a mixed effects linear regression with a log link, i.e.
+To incorporate explanatory variables on abundance, we model the expected survey abundance $\lambda_{s}$ with log-linear mixed effects, i.e.
 $\log (\lambda_{s}) = \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A$ where $\vX_{s}^A$ are explanatory variables, $\vbeta^A$ is a vector of fixed effects, $\vZ_{s}^A$ specifies random effect levels, and $\xi_j^A \ind N(0,\sigma_{A[j]}^2)$ are random effects where $A[j]$ assigns the appropriate variance for the $j$th abundance random effect.
 
-To incorporate explanatory variables on detection probability, we let the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
+To incorporate explanatory variables on detection probability, we let the continuous portion of the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
 Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
-For simplicity, we model the shape parameter $\alpha$ as constant across sites.
+For simplicity, we assume the shape parameter $\alpha$ as constant across sites.
 
@@
 \vn_s &\ind \Mult(n_s^{(obs)}, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
-p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
-p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I) \nonumber\\
+p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/F_T(C_I|\alpha,\varphi_s) \nonumber\\
+p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I|\alpha,\varphi_s) \nonumber\\
 \log(\lambda_s) &= \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A; \qquad \xi_j^A \ind N(0,\sigma_{A[j]}^2) \nonumber\\
@@
 
-We distinguish two categories of TTDDs: peaked and nonpeaked.  
-A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
-
 \ifdic\
@@
 The first study compares mixture vs non-mixture models.
-The second study compares the time to detection distribution families.
-In the first two studies, we utilized intercept only models to focus attention on the time to detection distribution. 
+The second study compares the TTDD families.
+In the first two studies, we utilized intercept only models to focus attention on the TTDD. 
 For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families. 
 In all simulation studies, we focus on accuracy in estimation of $\pdet$ which then translates into estimation of abundance.
+For the simulation studies following analysis we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.
+A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
 
@@
 
-In Sim1, we simulated 16 intercept-only datasets from each of 14 TTDDs. 
-We chose true parameter values (see Supplement) so that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-These values are largely consistent with estimates of Ovenbird populations from Section \ref{sec:ovenbirds}.
+To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 16 intercept-only datasets from each of 14 TTDDs: each combination of peaked/non-peaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal where all exponential models are non-peaked.
+We chose true parameter values (see Supplement \jarad{ADAM: what section in the supplement}) to mimic values from the Ovenbird danalysis see Section \ref{sec:ovenbirds}.
+In particular, we set parameters such that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+\jarad{ADAM: How many years and sites?}
 
 Each dataset was fit with two models: mixture and non-mixture version of the distribution family, e.g. exponential, used to simulate the data.
-We ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effective samples for all parameters.
+For each dataset-analysis combination, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effective samples for all parameters.
 
-Table \ref{tbl:sim1} provides a summary of the mixture and non-mixture models' ability to capture the detection probability. 
-When a non-mixture model is used to simulate the data (top half of the table), there are no clearly discernable differences between the ability of a non-mixture or mixture model to capture.
-\adam{For nonpeaked non-mixture data, the mixture model is less biased for lognormal and Weibull scenarios, and CIs are 15-40\% narrower.}
+We summarized the analysis be calculating the average across the simulations of the posterior median of $\pdet$ (Med.
+$p$), the average proportion of the posterior distribution of $\pdet$ larger than the truth (Q($p$)), and average coverage for 50\% and 90\% credible intervals.
+If the analyses are providing a reasonable estimate of $\pdet$, we would expect Med.
+$p$ to be around the true value of 0.8, Q($p$) to be around 0.5 indicating that half of the posterior is above and below the true value of 0.8, and the coverage to be close to their credibility.
+Table \ref{tbl:sim1} provides a summary of these quantities.
+When a mixture model is used to simulate the data (bottom half of the table), there is clearly a benefit to using a mixture model for inference.  
+Using a non-mixture model for inference, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ (Q($p$) $\approx 1$) and the other models underestimating.
+When a non-mixture model is used to simulate the data (top half of the table), there are no clearly discernable differences between the ability of a non-mixture or mixture model to capture $\pdet$. 
+For nonpeaked non-mixture data (top 4 lines), the mixture model is less biased for lognormal and Weibull scenarios (Q($p$)$\approx 0.5$), and CIs are 15-40\% narrower (results not shown).
+These results support the general default use of a mixture model over a non-mixture model.
 
@@
 \end{tabular}
-\caption{Summary of mixture vs. non-mixture model fits (Sim1).  
+\caption{Summary of mixture vs. non-mixture model fits.  
 In all cases, the inference model family matches the dataset family.  
@@
 50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-\jarad{We should consider splitting this table in two.
+% \jarad{We should consider splitting this table in two.
 The main point is what happens when the data arise from a mixture model but a non-mixture distribution is used.
 So the bottom half could be included in the main body while the top half could be put in the supplement.}
-\adam{The obvious concern for non-mixture data is that the mixture model might be more uncertain than a non-mixture model for non-mixture data.}
+% \adam{The obvious concern for non-mixture data is that the mixture model might be more uncertain than a non-mixture model for non-mixture data.}
 }
@@
 \fi
-
-When a mixture model is used to simulate the data (bottom half of the table), there is clearly a benefit to using a mixture inference model.  
-For the non-mixture models, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ and the other models underestimating.
-
-These results support the use of a mixture model unless you are absolutely sure your data arose from a non-mixture distribution.
 
@@
 Now we turn to model misspecification of the distribution family. 
-For Sim2, we simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
+We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
 %Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
@@
 \adam{For exponential data, the exponential model performs better.  I feel this should be noted.}
+\jarad{go for it}
 
@@
 In contrast, when the lognormal model is used for inference, it performs worse across nearly all data types.
-However, gamma models required 9-11 times the computer run time as did lognormal and Weibull models, and almost 25 times that of exponential models.
+However, gamma models required $\sim 10$ times the computational time as did lognormal and Weibull models and almost 25 times that of exponential models.
 
-These results support the use of a non-constant rate mixture TTDD model unless you are absolutely sure your data arose from an exponential mixture TTDD.
+%These results support the use of a non-constant rate mixture TTDD model unless you are absolutely sure your data arose from an exponential mixture TTDD.
 
@@
 The previous sections studied effects of time to detection assumptions in the context of no explanatory variables.
-In Sim3, we incorporate fixed and random effects for abundance and detection. 
+We now incorporate fixed and random effects for abundance and detection. 
 We simulated data from each of the 7 mixture TTDDs and fit models from exponential, gamma, lognormal, and Weibull mixture models.  
 We simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in Sim1 and Sim2, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
-(See Table \ref{tbl:sim3} in supplement for actual parameter values).
+Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
+See Table \ref{tbl:sim3} in supplement for actual parameter values.
 Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
@@
 
-Patterns in posterior estimates of $\pdet_s$ with respect to mixture and family TTDD forms were the same in Sim3 as in Sim1 and Sim2 -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family} in the Supplement).
-Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (see Supplemental figure).  
+Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions, see Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family} in the Supplement.
+Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed, see Supplemental figure.  
 Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
-It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.  
+%It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.  
 The above patterns can also be discerned in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
+
+\jarad{need to define what we mean by $\pdet$ here.}
+
+\jarad{where's the supplement? I think we should just refer to to the Figures on the supplement just like we refer to papers in the main manuscript and then interpret them.}
 
@@
 The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density was over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
-We associated random effects with each survey year and each stand.
+We associated random effects with each survey year and each stand \jarad{did we define stand previously?}.
 For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve.
@@
 
-Figure \ref{ovenposteriors} presents caterpillar plots of posterior estimates for model parameters, overall detection probability $\pdet$, and the log10 number of Ovenbirds uncounted.  
-Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from an exponential or nonpeaked mixture TTDD.
+Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the log10 number of uncounted Ovenbirds.  
+Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remain relatively large.
+%Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from an exponential or nonpeaked mixture TTDD.
 
-As in Sim3, abundance covariate coefficient estimates were virtually the same across all TTDD models.  
-95\% credible intervals for two of the abundance parameters do not contain zero, thereby suggesting notable effects.  
+As in the final simulation study, abundance covariate coefficient estimates were virtually the same across all models.  
+95\% credible intervals for two of the abundance parameters (site age and logging) do not contain zero, thereby suggesting notable effects.  
 Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% of the abundance for unlogged sites.  
-Credible intervals for Site Age indicate that each decade of age increases abundance from 1.5-13\%.
-Credible intervals for detection parameters do not indicate significant effects for any of the predictors.
+Credible intervals for site age coefficient indicate that each decade of age increases abundance from 1.5-13\%.
+Credible intervals for detection parameters do not indicate significant effects for any of the included predictors.
 
@@
 \includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors.pdf}
-\caption{\label{ovenposteriors} Caterpillar plots of posterior parameter estimates for models fit to Ovenbird data.  
-Abundance and detection parameters are identified by (A) and (D), respectively.  
-Gamma is the mixing parameter.  
-Shape(Alpha) are shape parameters for gamma, Weibull, and lognormal distributions.  
-p\_det and log10(Uncounted) are the detection probabilty and the number of birds present but not counted across all surveys.  
-Black bars are 95\% credible intervals, orange bars are 50\% credible intervals, and black dots are posterior medians.}
+\caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.
+Posteriors are also available for the number of uncounted individuals and the overall probability of detection across all sites.}
 \end{figure}
+
+\jarad{ADAM: please change facet names to use Greek letters for $\gamma$ and $\alpha$ and the standard notation we have been using for $\pdet$.}
 
@@
 
-Historically, analysis of removal-sampled point-count survey data assumes constant detection rates throughout the sampling period.  
-We have formulated a model for these data that allows for non-constant detection rates.
+%Historically, analysis of removal-sampled point-count survey data assumes constant detection rates throughout the sampling period.  
+We formulated a model for these data that allows for non-constant detection rates.
 Our model adopts a time-to-event approach within a hierarchical N-mixture framework, and it allows times to first detection to be modeled according to flexibly defined TTDD families.
@@
 Because the exponential TTDD is a special case of both gamma and Weibull TTDDs, we can interpret the differences in estimation between models as resulting from the information conveyed by the assumption of constant detection.
-Phrased differently, in the Bayesian context, the exponential model is equivalent to a gamma or Weibull model where the prior for the shape parameter $\alpha$ is defined: Pr($\alpha=1) = 1$; and the differences between the models results from the strength of this prior.
+%Phrased differently, in the Bayesian context, the exponential model is equivalent to a gamma or Weibull model where the prior for the shape parameter $\alpha$ is defined: Pr($\alpha=1) = 1$; and the differences between the models results from the strength of this prior.
 Among the TTDDs modeled in our study, the gamma mixture TTDD gave the most accurate estimates across data distributions, though the Weibull mixture TTDD was comparable for peaked datasets.

### 2016-08-09 Draft.tex
Add some additional comments

@@
 a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
+
+\jarad{Abstract the data}
 
@@
 
-\jarad{need to define what we mean by $\pdet$ here.}
+\jarad{need to define what we mean by $\pdet$ here since it's definition has changed.}
 
-\jarad{where's the supplement? I think we should just refer to to the Figures on the supplement just like we refer to papers in the main manuscript and then interpret them.}
+\jarad{I think we should just refer to to the Figures on the supplement just like we refer to papers in the main manuscript and then interpret them, e.g. Figure S1 shows ...}
 
@@
 The covariates were: (a) site age, (b) survey year, (c) an indicator of whether the site stock density was over 70\%, and (d) an indicator of whether the site experienced select-/partial-cut logging during the 1990s.
-We associated random effects with each survey year and each stand \jarad{did we define stand previously?}.
+We associated random effects with each survey year and each stand.
 For the detection half of our model, we used covariates for: (a) Julian date, (b) time of day, (c) temperature, (d) an indicator of whether it is the observer's first year in the database, and (e) an interaction between (a) and (d) to approximate a new observer's learning curve.

### 2016-08-13 Draft.tex
Supplement, Stan models, and some updates to the Draft.

@@
 \usepackage{setspace}
+\usepackage{xr}
+\externaldocument[S-]{Supplement}
 \doublespacing
@@
 A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007}.  
-\adam{I need to add others, like Mantyniemi.}
 In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Otis1978, Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
@@
 Section \ref{sec:ovenbirds} analyzes an Ovenbird data set under different TTDDs to determine the impact of this choice on estimated detection probability and therefore estimated abundance.
-Finally, section \ref{sec:discuss} discusses 
-\adam{The previous sentence ends mid-thought.  Is this for the Discussion section?}
-\jarad{yes}
 
-\section{Interval-censored avian point counts}\label{sec:data}
+\section{Interval-censored point counts}\label{sec:data}
 
-Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.  
-Survey sites were selected from sawtimber red pine stands with no recent logging activity.  
-Each stand had up to four \jarad{I thought it was 3} sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
-The data include 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, an indicator of select-/partial-cut logging during the 1990s.
-
-\jarad{did we use unlimited distance point counts or did we use $<$100m?}
+Our analysis is motivated by avian point-count surveys of over 50 species in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.
+However, the model we describe is appropriate for point-count surveys of any species, so long as time-to-first-detection data are available.  
+For our analysis, we focused on Ovenbird counts selected from one habitat type: sawtimber red pine stands with no recent logging activity.  
+Each stand had up to four sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
+The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s.
 
@@
 69\% of surveys in our dataset involved observers in their first year at MNFB.  
-Survey durations were 10 minutes, with times to first detection for over 50 species censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
+Survey durations were 10 minutes, with times to first detection censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
 During each survey, the Julian date, time of day, and temperature were recorded.
 
-\jarad{Put information about Ovenbird here?}
-
-a total of 947 Ovenbirds counted and a maximum of eight birds counted in any single survey.
-
-\jarad{Abstract the data}
+A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
+The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
 
@@
 
+\subsection{Detection heterogeneity across subgroups} \label{s:subgroups}
+
 It is common in avian point counts to observe increased detections in the first interval relative to an exponential distribution.
@@
 Failure to account for such heterogeneity in the constant-detection scenario leads to negative bias in abundance estimates \citep{Otis1978}.
-To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Etterson2009, Reidy2011}.
+To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Alldredge2007, Etterson2009, Reidy2011}.
 We specify a mixture TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time detection distribution $\Fm(t)$.  
@@
 To incorporate explanatory variables on detection probability, we let the continuous portion of the TTDD depend on the explanatory variables through the now site-specific parameter $\varphi_s$. 
-Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\tau_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
+Specifically, we model $\log(\varphi_{s}) = \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D$, where $\vX_{s}^D$ are explanatory variables, $\vbeta^D$ is a vector of fixed effects, $\vZ_{s}^D$ specifies random effect levels, $\xi_j^D \ind N(0,\sigma_{D[j]}^2)$ are random effects where $D[j]$ assigns the appropriate variance for the $j$th detection random effect.
 For simplicity, we assume the shape parameter $\alpha$ as constant across sites.
@@
 \vn_s &\ind \Mult(n_s^{(obs)}, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
-p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/F_T(C_I|\alpha,\varphi_s) \nonumber\\
-p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/F_T(C_I|\alpha,\varphi_s) \nonumber\\
+\pdet_s &= F_T(C_I|\alpha,\varphi_s)\\
+p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/\pdet_s \nonumber\\
+p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/\pdet_s \nonumber\\
 \log(\lambda_s) &= \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A; \qquad \xi_j^A \ind N(0,\sigma_{A[j]}^2) \nonumber\\
-\label{eq:model}\log(\varphi_{s}) &= \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D; \qquad \xi_j^D \ind N(0,\tau_{D[j]}^2)
+\label{eq:model}\log(\varphi_{s}) &= \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D; \qquad \xi_j^D \ind N(0,\sigma_{D[j]}^2)
 \end{align}
@@
 We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
+Stan model code and \texttt{rstan} code to run the models can be found in supplemental materials (Section \ref{S-s:models}).  
 We discarded half of the iterations as warmup and then thinned by 10.  
@@
 For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed to run unless care was taken in the specification of initial values.
-
-\jarad{We should include the Stan models in the supplement.}
 
@@
 
-To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 16 intercept-only datasets from each of 14 TTDDs: each combination of peaked/non-peaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal where all exponential models are non-peaked.
-We chose true parameter values (see Supplement \jarad{ADAM: what section in the supplement}) to mimic values from the Ovenbird danalysis see Section \ref{sec:ovenbirds}.
+To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 16 intercept-only datasets from each of 14 TTDDs: each combination of peaked/non-peaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal, where all exponential models are non-peaked.
+We chose the number of surveys (381) and true parameter values (supplemental table \ref{S-simvals}) to mimic values from the Ovenbird analysis (Section \ref{sec:ovenbirds}).  
 In particular, we set parameters such that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
-\jarad{ADAM: How many years and sites?}
 
@@
 In contrast, if the simulated data are non-peaked, the exponential model typically overestimates the actual detection probability. 
-\adam{For exponential data, the exponential model performs better.  I feel this should be noted.}
-\jarad{go for it}
+However, exponential model estimates are both less biased and more precise when the data actually do derive from an exponential mechanism.
 
@@
 Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
-See Table \ref{tbl:sim3} in supplement for actual parameter values.
+See Table \ref{S-tbl:sim3} in supplement for actual parameter values.
 Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
 To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.
+Because of the difficulty in integrating random effects over all sites, approximate posterior distributions for the study-wide marginal $\pdet$ were obtained by simulating data from each MCMC iteration and calculating the proportion simulated Ovenbirds that were observed.
 
-Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions, see Figures \ref{pdet_cater_correct} and \ref{pdet_cater_family} in the Supplement.
+Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
 Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed, see Supplemental figure.  
+\adam{Create Supplemental Figure!}
 Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
@@
 The above patterns can also be discerned in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
-
-\jarad{need to define what we mean by $\pdet$ here since it's definition has changed.}
-
-\jarad{I think we should just refer to to the Figures on the supplement just like we refer to papers in the main manuscript and then interpret them, e.g. Figure S1 shows ...}
 
@@
 
-\newpage
-\begin{appendix}
-\section{Supplement}
-
-\begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-Each column presents one family of simulated dataset.  
-Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
-All data and inference models include mixtures.
-Each plot presents one simulated dataset.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\end{figure}
-
-\begin{table}[!ht]
-\centering
-\begin{tabular}{lll|ccc}
-  \hline
-Family & Mixture & Peaked & $\gamma$ & $\varphi$ & $\alpha$ \\ 
-  \hline
-Exponential & Non-mixture & &  & -1.827 &  \\ 
-  & Mixture & & 0.65 & -2.138 &  \\ 
-  \hline
-  Gamma & Non-mixture & Non-peaked &  & -3.279 & 0.257 \\ 
-  & & Peaked &  & -0.746 & 3.371 \\ 
-  & Mixture & Non-peaked & 0.65 & -2.773 & 0.577 \\ 
-  & & Peaked & 0.65 & -1.210 & 2.491 \\ 
-  \hline
-  Lognormal & Non-mixture & Non-peaked &  & -0.341 & 2.330 \\ 
-  & & Peaked &  & -1.872 & 0.512 \\ 
-  & Mixture & Non-peaked & 0.65 & -1.462 & 1.674 \\ 
-  & & Peaked & 0.65 & -1.992 & 0.618 \\ 
-  \hline
-  Weibull & Non-mixture & Non-peaked &  & -1.165 & 0.418 \\ 
-  & & Peaked &  & -2.042 & 1.829 \\ 
-  & Mixture & Non-peaked & 0.65 & -2.063 & 0.687 \\ 
-  & & Peaked & 0.65 & -2.201 & 1.621 \\ 
-   \hline
-\end{tabular}
-\caption{Table of parameter values used to generate data for: (i) the mixture vs.
-non-mixture simulation, and (ii) the constant vs.
-non-constant rate simulation.
-Here $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lllrrrrrrrrrrrrrrrr}
-  \hline
-Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
-  \hline
-Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
-   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
-  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
-   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
-   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
-   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
-  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
-   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
-   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
-   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
-  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
-   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
-   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
-   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
-   \hline
-\end{tabular}
-\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.
-Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
-\end{table}
-
-\end{appendix}
-
-\newpage
-
 \bibliography{masterbib}

### 2016-08-13 Supplement.tex
Supplement, Stan models, and some updates to the Draft.

\documentclass[useAMS,usenatbib,referee,12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage[round]{natbib}
\usepackage{caption}
\usepackage{tabularx, booktabs}
\usepackage{float}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{appendix}

\newcommand{\pdet}{p^{(det)}}
\renewcommand{\thesection}{S-\arabic{section}}   % Label section with 'S-#'
\renewcommand{\thetable}{S-\arabic{table}}
\renewcommand{\thefigure}{S-\arabic{figure}}

\begin{document}

\section{Tables and Figures}

\begin{figure}[h!]\centering
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
Each column presents one family of simulated dataset.  
Upper plots show non-mixture datasets; lower plots show mixture datasets.  
`X' marks the expected marginal probability of detection based on true parameter values.}
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$. All data and inference models include mixtures.  
Each plot presents one simulated dataset.  
`X' marks the expected marginal probability of detection based on true parameter values.}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lll|ccc}
  \hline
Family & Mixture & Peaked & $\gamma$ & $\varphi$ & $\alpha$ \\ 
  \hline
Exponential & Non-mixture & &  & -1.827 &  \\ 
  & Mixture & & 0.65 & -2.138 &  \\ 
  \hline
  Gamma & Non-mixture & Non-peaked &  & -3.279 & 0.257 \\ 
  & & Peaked &  & -0.746 & 3.371 \\ 
  & Mixture & Non-peaked & 0.65 & -2.773 & 0.577 \\ 
  & & Peaked & 0.65 & -1.210 & 2.491 \\ 
  \hline
  Lognormal & Non-mixture & Non-peaked &  & -0.341 & 2.330 \\ 
  & & Peaked &  & -1.872 & 0.512 \\ 
  & Mixture & Non-peaked & 0.65 & -1.462 & 1.674 \\ 
  & & Peaked & 0.65 & -1.992 & 0.618 \\ 
  \hline
  Weibull & Non-mixture & Non-peaked &  & -1.165 & 0.418 \\ 
  & & Peaked &  & -2.042 & 1.829 \\ 
  & Mixture & Non-peaked & 0.65 & -2.063 & 0.687 \\ 
  & & Peaked & 0.65 & -2.201 & 1.621 \\ 
   \hline
\end{tabular}
\caption{\label{simvals}Table of parameter values used to generate data for: (i) the mixture vs. non-mixture simulation, and (ii) the constant vs. non-constant rate simulation.  Here $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrrrrrrrrrrrr}
  \hline
Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\ 
  \hline
Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
   \hline
\end{tabular}
\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.  Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
\end{table}

\small
\section{Stan Models}\label{s:models}
\subsection{Exponential Model}
\lstinputlisting[breaklines]{./OVEN/Expo_model.stan}
\subsection{Exponential Mixture Model}
\lstinputlisting[breaklines]{./OVEN/ExpoMix.stan}
\subsection{Gamma Model}
\lstinputlisting[breaklines]{./OVEN/Gamma_model.stan}
\subsection{Gamma Mixture Model}
\lstinputlisting[breaklines]{./OVEN/GammaMix.stan}
\subsection{Lognormal Model}
\lstinputlisting[breaklines]{./OVEN/LogNormal_model.stan}
\subsection{Lognormal Mixture Model}
\lstinputlisting[breaklines]{./OVEN/LogNormalMix.stan}
\subsection{Weibull Model}
\lstinputlisting[breaklines]{./OVEN/Weibull_model.stan}
\subsection{Weibull Mixture Model}
\lstinputlisting[breaklines]{./OVEN/WeibullMix.stan}

\end{document}

### 2016-08-17 Draft.tex
Comment removals.  Figure added to Supplement.  Updates to existing figures.

@@
 50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-% \jarad{We should consider splitting this table in two.
-The main point is what happens when the data arise from a mixture model but a non-mixture distribution is used.
-So the bottom half could be included in the main body while the top half could be put in the supplement.}
-% \adam{The obvious concern for non-mixture data is that the mixture model might be more uncertain than a non-mixture model for non-mixture data.}
 }
@@
 To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.
-Because of the difficulty in integrating random effects over all sites, approximate posterior distributions for the study-wide marginal $\pdet$ were obtained by simulating data from each MCMC iteration and calculating the proportion simulated Ovenbirds that were observed.
+Because of the difficulty in integrating random effects over all sites, approximate posterior distributions for the study-wide marginal $\pdet$ were obtained by simulating data from each MCMC sample and calculating the proportion of simulated Ovenbirds that were observed.
 
 Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
-Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed, see Supplemental figure.  
-\adam{Create Supplemental Figure!}
+Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (Section \ref{S-s:sim3}).  
 Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
@@
 
-\jarad{ADAM: please change facet names to use Greek letters for $\gamma$ and $\alpha$ and the standard notation we have been using for $\pdet$.}
-
 \section{Discussion} \label{sec:discuss}
@@
 
-If the estimation of abundance is the primary interest, then we may reasonably ask whether removal sampled point-count surveys are really the data we want.  
+If the estimation of abundance is the primary interest, then the choice of TTDD has large consequences, and we may reasonably ask whether removal sampled point-count surveys are really the data we want.
 The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would only be detected if the observation period lasted longer.  
 As such, it is entirely based upon extrapolation and correctly estimating the probability in the tail of the TTDD based on recorded observations.
-An assumption of constant detection constrains uncertainty in the extrapolated tail but at the cost of potential bias.
-When we allow for non-constant detection, estimation of the tail probability becomes so variable that we were only able to estimate Ovenbird abundance within an order of magnitude.
+An assumption of constant detection places constraints  on the amount of uncertainty in the extrapolated tail but at the cost of potentially sizable bias.
+When we allow for non-constant detection, estimation of the tail probability becomes less certiain.
 In theory, estimation of the unsampled tail probability can be improved by conducting longer observations, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
@@
 
-%With the increasing use of microphone arrays for automated detection \adam{refs}, it is tempting to think TTDD models may be valuable for microphone-collected data.  
+%With the increasing use of microphone arrays for automated detection, it is tempting to think TTDD models may be valuable for microphone-collected data.  
 %This should be the case when variations in detection result solely from organismal behaviors, such as bout singing, or if we wish to model individual random effects in the detection model.  
@@
 
-%In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.
-%\adam{I'm  beginning to think this may have too-numerous antecedents to cite... in Amundson and in Reidy, the earliest reference is Marsh and Sinclair 1989.  See also McCallum 2005.}.  
+%In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}...
+citations too numerous to enumerate.
 %In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
@@
 
-\adam{Two topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree:\\
+\adam{Topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree):
 \begin{enumerate}
-\item DIC and posterior predictive checks
-\item Selection of priors
+\item Selection of priors (especially for intercepts; also, the possibility of a spike-and-slab for $\alpha$)
 \item Values for $\pdet$ used in simulations

### 2016-08-17 Supplement.tex
Comment removals.  Figure added to Supplement.  Updates to existing figures.

@@
 \usepackage{verbatim}
-\usepackage{appendix}
+\usepackage{graphicx}
 
@@
 
-\section{Tables and Figures}
+\section{Figures}
 
@@
 
+\subsection{Posteriors from simulations including covariates and random effects}\label{s:sim3}
+\includegraphics[width=1.0\textwidth, page=1]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=2]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=3]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=4]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=5]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=6]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=7]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=8]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=9]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=10]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=11]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=12]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=13]{Sims/SimFull/Posteriors_all_simdat.pdf}
+\includegraphics[width=1.0\textwidth, page=14]{Sims/SimFull/Posteriors_all_simdat.pdf}
+
+\section{Tables}
 \begin{table}[!ht]

### 2016-08-19 Supplement.tex
Remove Stan models from supplement.  It occurs to me that the .stan files stand alone just fine and don't need to be embedded in a .tex file.  Fixes #7.
Added captions to all figures.  Rotated and split the large table of simulation parameter values.

--- 2016-08-13 Supplement.tex
@@
 \usepackage{tabularx, booktabs}
-\usepackage{float}
+\usepackage{float, placeins}
 \usepackage{adjustbox}
@@
 \usepackage{verbatim}
-\usepackage{appendix}
+\usepackage{graphicx}
+\usepackage{rotating}
 
 \newcommand{\pdet}{p^{(det)}}
+\newcommand{\capt}[1]{Caterpillar plots of posterior parameter estimates from all models fit to the simulated #1 dataset (from simulations involving covariates and random effects).
+Black and orange lines depict central 95\% and 50\% credible intervals, respectively.
+The black dot is the posterior meadian.
+The blue 'X' is the true parameter value and is positioned on the correctly specified model.
+$\beta$'s are fixed effect parameters, $\sigma$'s are random effect standard deviations, and $\gamma$ is a mixing parameter.
+Parameters designated with an 'A' are abundance parameters; those designated with a 'D' are detection parameters.}
+\newcommand{\sfig}[2]{\begin{figure}
+  \includegraphics[width=1.0\textwidth, page=#2]{Sims/SimFull/Posteriors_all_simdat.pdf}
+  \caption{\capt{#1}}
+  \end{figure}
+}
 \renewcommand{\thesection}{S-\arabic{section}}   % Label section with 'S-#'
@@
 
-\section{Tables and Figures}
-
-\begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-Each column presents one family of simulated dataset.  
-Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
-All data and inference models include mixtures.
-Each plot presents one simulated dataset.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
-\end{figure}
-
+\section{Tables}
 \begin{table}[!ht]
@@
 
-\begin{table}[ht]
+\begin{sidewaystable}[h!]\footnotesize
 \centering
-\begin{tabular}{lllrrrrrrrrrrrrrrrr}
+\begin{tabular}{lll|rrrr|rrrrr}
   \hline
-Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
+Family & Mixture & Peak & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$  \\
   \hline
-Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
-   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
-  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
-   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
-   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
-   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
-  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
-   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
-   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
-   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
-  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
-   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
-   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
-   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
+Exponential & Non-mixture & &0.122&0.002&0.094&-1.039&-0.01&-0.056&0.118&0.249&0.164 \\
+& Mixture & &0.124&0.012&0.064&-1.07&-0.068&-0.146&0.03&0.312&0.27 \\
+Gamma & Non-mixture & Non-peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
+& & Peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
+& Mixture & Non-peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
+& & Peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
+Lognormal & Non-mixture & Non-peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
+& & Peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
+& Mixture & Non-peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
+& & Peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
+Weibull & Non-mixture & Non-peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
+& & Peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
+& Mixture & Non-peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
+& & Peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
+  \hline
+\end{tabular}\\
+\vspace{1.25cm}
+\begin{tabular}{ll|lrrrrrrr}
+  \hline
+Family & Mixture & Peak & Intercept$^A$ & Intercept$^D$ & $\sigma_{A1}$ & $\sigma_{A2}$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
+\hline
+Exponential & Non-mixture & &0.975&-1.146&0.139&0.094&0.242& & \\
+& Mixture & &1.097&-1.97&0.13&0.091&0.259&0.647& \\
+Gamma & Non-mixture & Non-peaked &1.383&-3.941&0.121&0.072&0.39& & 0.316 \\
+& & Peaked &1.134&-0.746&0.121&0.072&0.39& & 3.371 \\
+& Mixture & Non-peaked &1.215&-2.7&0.121&0.079&0.332&0.79& 0.597 \\
+& & Peaked &1.134&-1.21&0.121&0.079&0.332&0.65& 2.491 \\
+Lognormal & Non-mixture & Non-peaked &1.613&-2.444&0.151&0.077&0.353& & 3.287 \\
+& & Peaked &1.134&-1.872&0.151&0.077&0.353& & 0.512 \\
+& Mixture & Non-peaked &1.282&-2.051&0.14&0.079&0.312&0.706& 1.467 \\
+& & Peaked &1.134&-1.992&0.14&0.079&0.312&0.65& 0.618 \\
+Weibull & Non-mixture & Non-peaked &1.57&-3.137&0.131&0.074&0.383& & 0.387 \\
+& & Peaked &1.134&-2.042&0.131&0.074&0.383& & 1.829 \\
+& Mixture & Non-peaked &1.294&-2.354&0.122&0.078&0.344&0.814& 0.653 \\
+& & Peaked &1.134&-2.201&0.122&0.078&0.344&0.65& 1.621 \\
    \hline
@@
 Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
-\end{table}
+\end{sidewaystable}
+%\begin{sidewaystable}[ht]
+%\centering
+%\begin{tabular}{lllrrrrrrrrrrrrrrrr}
+%  \hline
+%Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
+%  \hline
+%Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
+%   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
+%  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
+%   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
+%   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
+%   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
+%  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
+%   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
+%   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
+%   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
+%  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
+%   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
+%   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
+%   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
+%   \hline
+%\end{tabular}
+%\begin{tabular}
+%\end{tabular}
+%\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.
+Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
+%\end{sidewaystable}
+\FloatBarrier
 
-\small
-\section{Stan Models}\label{s:models}
-\subsection{Exponential Model}
-\lstinputlisting[breaklines]{./OVEN/Expo_model.stan}
-\subsection{Exponential Mixture Model}
-\lstinputlisting[breaklines]{./OVEN/ExpoMix.stan}
-\subsection{Gamma Model}
-\lstinputlisting[breaklines]{./OVEN/Gamma_model.stan}
-\subsection{Gamma Mixture Model}
-\lstinputlisting[breaklines]{./OVEN/GammaMix.stan}
-\subsection{Lognormal Model}
-\lstinputlisting[breaklines]{./OVEN/LogNormal_model.stan}
-\subsection{Lognormal Mixture Model}
-\lstinputlisting[breaklines]{./OVEN/LogNormalMix.stan}
-\subsection{Weibull Model}
-\lstinputlisting[breaklines]{./OVEN/Weibull_model.stan}
-\subsection{Weibull Mixture Model}
-\lstinputlisting[breaklines]{./OVEN/WeibullMix.stan}
+\section{Figures}
+
+\begin{figure}[h!]\centering
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
+\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
+Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
+Each column presents one family of simulated dataset.  
+Upper plots show non-mixture datasets; lower plots show mixture datasets.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
+\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
+All data and inference models include mixtures.
+Each plot presents one simulated dataset.  
+`X' marks the expected marginal probability of detection based on true parameter values.}
+\end{figure}
+
+\subsection{Posteriors parameter estimates from simulations including covariates and random effects}\label{s:sim3}
+\sfig{non-mixture exponential}{1}
+\sfig{exponential mixture}{2}
+\sfig{nonpeaked non-mixture gamma}{3}
+\sfig{nonpeaked gamma mixture}{4}
+\sfig{peaked non-mixture gamma}{5}
+\sfig{peaked gamma mixture}{6}
+\sfig{nonpeaked non-mixture lognormal}{7}
+\sfig{nonpeaked lognormal mixture}{8}
+\sfig{peaked non-mixture lognormal}{9}
+\sfig{peaked lognormal mixture}{10}
+\sfig{nonpeaked non-mixture Weibull}{11}
+\sfig{nonpeaked Weibull mixture}{12}
+\sfig{peaked non-mixture Weibull}{13}
+\sfig{peaked Weibull mixture}{14}
+
+%\small
+%\section{Stan Models}\label{s:models}
+%\subsection{Exponential Model}
+%\lstinputlisting[breaklines]{./OVEN/Expo_model.stan}
+%\subsection{Exponential Mixture Model}
+%\lstinputlisting[breaklines]{./OVEN/ExpoMix.stan}
+%\subsection{Gamma Model}
+%\lstinputlisting[breaklines]{./OVEN/Gamma_model.stan}
+%\subsection{Gamma Mixture Model}
+%\lstinputlisting[breaklines]{./OVEN/GammaMix.stan}
+%\subsection{Lognormal Model}
+%\lstinputlisting[breaklines]{./OVEN/LogNormal_model.stan}
+%\subsection{Lognormal Mixture Model}
+%\lstinputlisting[breaklines]{./OVEN/LogNormalMix.stan}
+%\subsection{Weibull Model}
+%\lstinputlisting[breaklines]{./OVEN/Weibull_model.stan}
+%\subsection{Weibull Mixture Model}
+%\lstinputlisting[breaklines]{./OVEN/WeibullMix.stan}
 

### 2016-08-20 Draft.tex
Remove Stan models from supplement.  It occurs to me that the .stan files stand alone just fine and don't need to be embedded in a .tex file.  Fixes #7.
Updates to bibliography and discussion.  Fixes #9.

@@
 We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
-Stan model code and \texttt{rstan} code to run the models can be found in supplemental materials (Section \ref{S-s:models}).  
+Stan model code and \texttt{rstan} code to run the models can be found in supplemental materials.  
 We discarded half of the iterations as warmup and then thinned by 10.  
@@
 
-%Historically, analysis of removal-sampled point-count survey data assumes constant detection rates throughout the sampling period.  
 We formulated a model for these data that allows for non-constant detection rates.
 Our model adopts a time-to-event approach within a hierarchical N-mixture framework, and it allows times to first detection to be modeled according to flexibly defined TTDD families.
-Our results show that non-constant TTDDs can return reasonable estimates of detection probabilities across a variety of time-to-detection data patterns, whereas traditional constant-rate TTDDs return biased and overly precise estimates when data deviate from the constant-rate assumption.
+Our results show that non-constant TTDDs can return reasonable estimates of detection probabilities across a variety of time-to-detection data patterns, whereas traditional constant-rate TTDDs return biased and overly precise estimates when data deviate from the constant-rate assumption, even when they include a mixture for heterogeneity across groups.
 Because the exponential TTDD is a special case of both gamma and Weibull TTDDs, we can interpret the differences in estimation between models as resulting from the information conveyed by the assumption of constant detection.
@@
 
-% Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
-
-% Really, I have totally ignored the discussion of the true value of p, which I have tended to simulate at a fairly high value
-
-If the estimation of abundance is the primary interest, then the choice of TTDD has large consequences, and we may reasonably ask whether removal sampled point-count surveys are really the data we want.
-The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would only be detected if the observation period lasted longer.  
+If the estimation of abundance is the primary interest, then the choice of TTDD has large consequences, and we may reasonably ask whether removal sampled point-count surveys are adequate for the purpose.
+The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would be detected if only the observation period lasted longer.  
 As such, it is entirely based upon extrapolation and correctly estimating the probability in the tail of the TTDD based on recorded observations.
@@
 When we allow for non-constant detection, estimation of the tail probability becomes less certiain.
-In theory, estimation of the unsampled tail probability can be improved by conducting longer observations, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
-Without extending the observation period, an alternate to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB, where many species are observed.
+In theory, estimation of the unsampled tail probability can be improved by conducting longer surveys, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
+Without extending the observation period, an alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB, where many species are observed.
 
@@
 %The assumption of constant detection seems more plausible in studies where observers are absent, though detection heterogeneity may still exist across subgroups of the population.
+
+Versions of time-varying models have been described for trap-based removal sampling and continuous-time capture-recapture.  
+Time variation has been modeled through a non-constant hazard function \citep{Schnute1983, HwangChao2002}, a randomly varying detection probability across trapping sessions \citep{WangLoneragan1996}, and constant detection probabilities that vary randomly from individual to individual \citep{Mantyniemi2005, Laplanche2010}.
+Most of these approaches resulted marginally in a decreasing (nonpeaked) detection function over time.  
+Their results generally echo what we have presented here.  
+\citet{Schnute1983} found that the equivalent of a mixture exponential adequately described their data.
+\citet{WangLoneragan1996}, \citet{HwangChao2002}, and \citet{Mantyniemi2005} all found constant-detection models to be flawed, producing underestimates of abundance and too-narrow error estimates; these resulted in inadequate coverage and also overstatement of effect significance.
+%Schnute: allowed time-varying detection which collapsed to a mixture model
+%Wang and Lonergan: \pi_s ~ beta for each interval at a site... not particularly applicable in an avian framework (with arbitrarily partitioned), except perhaps for bout singing?
+%- didn't have group heterogeneity
+%- constant models are overly precise (factor of 3)
+%Mantyniemi: individual-level variation in \pi_s ~ beta
+%- large number of trapping sessions required to overwhelm prior on variance (8-12)
+%- could also be viewed in terms of percent of total caught
+%- constant detection model badly biased
+%- constant detection model made indexing unreliable, since pairwise differences could be magnified
+%Laplanche: model selection with site-level abundance variation, site-level catchability variation, and individual-level catchability variation.
+%- only the abundance proved significant
+%Hwang and Chao: time-varying capture-recapture
+%- use a nonpeaked hazard function
+%- again, the constant model underestimates abundance, with coverage between 0-6\% for 95\% credible intervals (even allowing for covariate-based individual effects and behavioral responses).
 
@@
 \item Selection of priors (especially for intercepts; also, the possibility of a spike-and-slab for $\alpha$)
+% Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
 \item Values for $\pdet$ used in simulations

### 2016-08-22 OVEN/Simple Code to Run rStan.R
Code and data to fit a simple Weibull mixture model.

@@ -1,0 +1,89 @@
+# As a prerequisite, point count survey data should be organized in a dataframe saved as 'surveydata'.
+surveydata <- read.csv("surveydata.csv")
+
+########## Code to package data as a list for the model
+
+tau <- c(2,3,4,5,6,7,8,9,10)    # Observation interval cutoffs (does not include zero)
+
+### Construct covariate matrices
+# Abundance covariates: site age, year, stock density, logging indicator
+Xa <- with(surveydata, data.frame(siteage=siteage, 
+                                year=year,
+                                density=1*(stockdens==9),
+                                logged=1*(logged==1)))
+# Detection covariates: julian date, time of day, temperature, observer's first year indicator, jd_fy interaction
+Xd <- with(surveydata, data.frame(jd=jd,
+                                time=time,
+                                temp=temp,
+                                firstyr=firstyr, 
+                                jd_fy=jd*firstyr))
+
+### Construct indices for random effect covariates
+# For abundance: site, year
+ia <- with(surveydata, data.frame(yearf=as.numeric(factor(year.unscale)), 
+                                  stand=as.numeric(factor(standunique))))
+# For detection: observer
+id <- data.frame(obs=as.numeric(factor(surveydata$obs)))
+
+### Combine multiple effect types into a single vector
+
+# Abundance random effects
+lu <- function(x) length(unique(x))  # Helper function for apply statements
+n_ras <- apply(ia, 2, lu)            # Number of levels for each abundance random effect type
+va_id <- rep(1:ncol(ia), n_ras)      # Index to associate random effect levels to their random effect type
+# Renumber indices so that levels of different random effects do not share the same number
+for (j in 2:ncol(ia)) ia[,j] <- ia[,j] + cumsum(n_ras)[j-1]
+
+# Detection random effects
+n_rds <- apply(id, 2, lu)
+vd_id <- rep(1:ncol(id), n_rds)
+# Note: because we have only one detection random effect,
+#       there is no need to renumber indices cumulatively
+
+### Counts by interval
+# Here, W1...W9 are counts by observation interval
+Ival <- subset(surveydata, select=c("W1", "W2", "W3", "W4", "W5", "W6", "W7", "W8", "W9"))
+
+
+
[…]

### 2016-09-08 Draft.tex
Make minor changes
including ref to Borchers preprint

@@
 \usepackage{setspace}
+\usepackage{url}
 \usepackage{xr}
@@
 %Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  
-We consider a class of N-mixture model that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
+We consider a class of N-mixture models that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
 %We consider a class of N-mixture models that allow for detection heterogeniety through a mixture component, flexible two-parameter time-to-detection distributions (TTDDs), and allows for fixed and random effects for both abundance and detection.
@@
 Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
-We specifically explore two-parameter families of TTDDs (e.g. gamma) that can additionally include a mixture component to model increased probability of detection in the initial observation period.
+We specifically explore two-parameter families of TTDDs, e.g. gamma, that can additionally include a mixture component to model increased probability of detection in the initial observation period.
 We find that modeling a TTDD by using a mixture component is necessary when data have a chance of arising from a distribution of this nature.
@@
 
-{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Bayesian
+{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Markov chain Monte Carlo; Stan
 
@@
 In this manuscript, we develop a model for scenarios where detection rates are not constant over time. 
-We consider the first time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each animal's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
+We consider the first time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each individual's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
@@
 Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
-Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine equal-duration intervals \citep{Reidy2011, Amundson2014}.
+Indeed our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine equal-duration intervals \citep{Reidy2011, Amundson2014}.
 
@@
 Section \ref{sec:model} introduces an N-mixture model with a generically defined TTDD for estimating abundance from removal-sampled point-count surveys.
-In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
+%In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
 Section \ref{sec:sim} provides three simulation studies to assess the impact of TTDD choice on estimated detection probability. 
@@
 
-Our analysis is motivated by avian point-count surveys of over 50 species in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.
-However, the model we describe is appropriate for point-count surveys of any species, so long as time-to-first-detection data are available.  
+Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.  
+%However, the model we describe is appropriate for point-count surveys of any species, so long as time-to-first-detection data are available.  
 For our analysis, we focused on Ovenbird counts selected from one habitat type: sawtimber red pine stands with no recent logging activity.  
@@
 Single-visit (per year) point-count surveys were conducted by trained observers at each site once annually (weather permitting).  
-Fourteen different observers conducted surveys during the study period.  
-69\% of surveys in our dataset involved observers in their first year at MNFB.  
+Fourteen different observers conducted surveys during the study period and 69\% of surveys in our dataset involved observers in their first year at the MNFB.  
 Survey durations were 10 minutes, with times to first detection censored into nine intervals: a two-minute interval followed by eight one-minute intervals.  
@@
 
-A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
-The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
+% A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
+% The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
+
+While we focus on the estimation of detection probability in avian populations, the approach we describe is appropriate for point-count surveys of any species.
+The methodology allows the analysis of data with 1) recorded first (possible censored) detection of each individual, 2) site-specific explanatory variables, and 3) survey-specific explanatory variables.
 
@@
 Suppose that, for each survey $s$ ($s=1,\dots,S$), $N_s$ individuals are present.  
-Imagine an observer could remain at the survey location until every individual is detected, marking the time to detection $t_{sb}$ (for bird, $b=1,\dots,N_s$) for each.
-Then, assuming detection times for all individuals at a survey are independent, identically distributed according to a common time-to-detection distribution (TTDD), we define $T_{sb}$ as a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
-However, in reality, times to first detection are truncated due to a finite survey length of $C$, meaning that each individual has a detection probability $\pdet=F_T(C)$.  
-The conditional distribution of observed detection times then has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$ and cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$. 
-We model the number of individuals for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$.
+Imagine an observer could remain at the survey location until every individual is detected, recording the time to detection $t_{sb}$ (for bird, $b=1,\dots,N_s$) for each.
+Assuming detection times for all individuals at a survey are independent, identically distributed according to a common time-to-detection distribution (TTDD), we define $T_{sb}$ as a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
+In reality, times to first detection are often truncated due to a finite survey length of $C$, meaning that each individual has a detection probability $\pdet=F_T(C)$.  
+The conditional distribution of observed detection times then has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$, and instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
+We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
 %$T_{sb|det} \ind f_{T|det}(t_{sb})$ where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$.
-The instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
 
@@
 Let $C_i$ for $i=1,\dots,I$ indicate the right endpoint of the $i$th interval then $C_I$ is the total survey duration and, letting $C_0=0$, the $i$th interval is $(C_{i-1},C_{i}]$. 
-Finally, let $n_{si}$ be the number of animals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
+Let $n_{si}$ be the number of individuals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
 Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and is calculated from the TTDD: $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
@@
 \vn_s &\ind \Mult(n_s^{(obs)}, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
-\pdet_s &= F_T(C_I|\alpha,\varphi_s)\\
-p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/\pdet_s \nonumber\\
+\pdet_s &= F_T(C_I|\alpha,\varphi_s) \nonumber \\
+p_{s1} &= \left[(1-\gamma) + \gamma\Fm(C_1|\alpha,\varphi_s)\right]/\pdet_s \label{eq:model} \\
 p_{si} &= \gamma\left[\Fm(C_i|\alpha,\varphi_s) - \Fm(C_{i-1}|\alpha,\varphi_s)\right]/\pdet_s \nonumber\\
 \log(\lambda_s) &= \vX_{s}^A\vbeta^A + \vZ_{s}^A\vxi^A; \qquad \xi_j^A \ind N(0,\sigma_{A[j]}^2) \nonumber\\
-\label{eq:model}\log(\varphi_{s}) &= \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D; \qquad \xi_j^D \ind N(0,\sigma_{D[j]}^2)
+\log(\varphi_{s}) &= \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D; \qquad \xi_j^D \ind N(0,\sigma_{D[j]}^2) \nonumber
 \end{align}
@@
 We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
-Stan model code and \texttt{rstan} code to run the models can be found in supplemental materials.  
+Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
 We discarded half of the iterations as warmup and then thinned by 10.  
-We monitored convergence of the MCMC chains using Geweke z-score diagnostics \citep{Geweke1991} and reran models if lack of convergence was indicated by a non-normal distribution of the z-scores. 
-We reran models if the effective sample size for any parameter was below 1000.  
+We monitored convergence of the MCMC chains using Geweke z-score diagnostics \citep{Geweke1991} and reran models if lack of convergence was indicated by a non-normal distribution of the z-scores or if the effective sample size for any parameter was below 1000.
 The number of iterations used depended on the model and is detailed later. 
@@
 In all simulation studies, we focus on accuracy in estimation of $\pdet$ which then translates into estimation of abundance. 
-For the simulation studies following analysis we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.
+For the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
 A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
@@
 To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 16 intercept-only datasets from each of 14 TTDDs: each combination of peaked/non-peaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal, where all exponential models are non-peaked.
-We chose the number of surveys (381) and true parameter values (supplemental table \ref{S-simvals}) to mimic values from the Ovenbird analysis (Section \ref{sec:ovenbirds}).  
+We chose the number of surveys (381) and true parameter values (Table \ref{S-simvals}) to mimic values from the Ovenbird analysis (Section \ref{sec:ovenbirds}).  
 In particular, we set parameters such that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
@@
 
-We summarized the analysis be calculating the average across the simulations of the posterior median of $\pdet$ (Med.
+We summarized the analysis by calculating the average, across the simulations, of the posterior median of $\pdet$ (Med.
 $p$), the average proportion of the posterior distribution of $\pdet$ larger than the truth (Q($p$)), and average coverage for 50\% and 90\% credible intervals.
@@
 
-The previous section adressed model mis-specification in terms of the mixture component. 
+The previous section addressed model mis-specification in terms of the mixture component. 
 Now we turn to model misspecification of the distribution family. 
@@
 We simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
-See Table \ref{S-tbl:sim3} in supplement for actual parameter values.
+Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes, see Table \ref{S-tbl:sim3} for actual parameter values.
 Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
@@
 
+The results from this simulation are qualitatively similar to that Ovenbird analysis (Figure \ref{ovenposteriors}) and thus we only briefly review the results here and provide the corresponding figures and tables in the Supplementary Material.
 Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
@@
 %It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.
-The above patterns can also be discerned in posterior estimates from the Ovenbird data (Figure \ref{ovenposteriors}).
 
@@
 
-Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the log10 number of uncounted Ovenbirds.  
+Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the logarithm of the number of uncounted Ovenbirds.  
 Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remain relatively large.
@@
 
-As in the final simulation study, abundance covariate coefficient estimates were virtually the same across all models.  
-95\% credible intervals for two of the abundance parameters (site age and logging) do not contain zero, thereby suggesting notable effects.  
+Abundance covariate coefficient estimates were virtually the same across all models.  
+The 95\% credible intervals for two of the abundance parameters (site age and logging) do not contain zero, thereby suggesting notable effects.  
 Select- and partial-cut logging events of the 1990s depressed local Ovenbird abundance during the study perior to roughly 25-50\% of the abundance for unlogged sites.  
 Credible intervals for site age coefficient indicate that each decade of age increases abundance from 1.5-13\%.
-Credible intervals for detection parameters do not indicate significant effects for any of the included predictors.
+Credible intervals for detection parameters do not indicate significant effects, after adjusting for the other predictors, for any of the included predictors.
 
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors.pdf}
+\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors}
 \caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.
@@
 
-We formulated a model for these data that allows for non-constant detection rates.
+We formulated a model for interval-censored time to detection data that allows for non-constant detection rates.
 Our model adopts a time-to-event approach within a hierarchical N-mixture framework, and it allows times to first detection to be modeled according to flexibly defined TTDD families.
@@
 Detection effect estimates, while conditional on the mixing parameter $\gamma$, are similar across all mixture non-exponential TTDDs.  
-These finding may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.
+These findings may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.
 
@@
 In theory, estimation of the unsampled tail probability can be improved by conducting longer surveys, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
-Without extending the observation period, an alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB, where many species are observed.
+Without extending the observation period, an alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB where many species are observed simultaneously
 
@@
 Consequently, our application of a TTDD represents an averaging across distance classes, which induces systematic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.  
-To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier, as is done in \adam{How do I cite Borchers' preprint?}.  
+To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \citeauthor{borchers2015}.  
 This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Amundson2014}.
@@
 
-We recommend that time-heterogeneous detection rates be explicitly modeled in analyses involving removal-sampled point-count survey data where estimation of abundance is a primary objective.  
+We recommend that time-heterogeneous detection rates be explicitly modeled in analyses involving removal-sampled point-count survey data where estimation of detection probability or abundance is a primary objective.
 The assumption of constant detection, while computationally simple and reasonable as a null model, proves to be rather informative and can result in pronounced bias.  
@@
 
-\adam{Topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree):
-\begin{enumerate}
-\item Selection of priors (especially for intercepts; also, the possibility of a spike-and-slab for $\alpha$)
-% Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
-\item Values for $\pdet$ used in simulations
-\end{enumerate}
-}
+% \adam{Topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree):
+% \begin{enumerate}
+% \item Selection of priors (especially for intercepts; also, the possibility of a spike-and-slab for $\alpha$)
+% % Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
+% \item Values for $\pdet$ used in simulations
+% \end{enumerate}
+% }
 

### 2016-09-08 Draft.tex
Add institution
Eliminate unnecessary packages and commands

@@
-\documentclass[useAMS,usenatbib,referee,12pt]{article}
+\documentclass[12pt]{article}
 \usepackage[margin=1.0in]{geometry}
@@
 \usepackage{parskip}
-\usepackage[round]{natbib}
-\usepackage{color}
-\usepackage[dvipsnames]{xcolor}
+\usepackage{natbib}
+% \usepackage{color}
+% \usepackage[dvipsnames]{xcolor}
 \usepackage{caption}
-\usepackage{tabularx, booktabs}
-\usepackage{float}
+%\usepackage{tabularx, booktabs}
+%\usepackage{float}
 \usepackage{adjustbox}
@@
 \usepackage{xr}
+
 \externaldocument[S-]{Supplement}
+
 \doublespacing
@@
 
-\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
+% \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
+% \newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
+
 \newenvironment{indpar}[1]%
@@
 
+\providecommand{\blind}{0}
+
 \title{Hierarchical model assessing the impacts of time to detection distribution assumptions on detection probability estimation}
 
-\author{Adam Martin-Schwarze, Jarad Niemi, and Philip Dixon}
+\if0\blind
+\author{Adam Martin-Schwarze, Jarad Niemi, and Philip Dixon\\
+Department of Statistics, Iowa State University, Ames, Iowa}
+\fi
 
@@
 Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
-Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (Section \ref{S-s:sim3}).  
+Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed, see Section \ref{S-s:sim3} and figures therein.  
 Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.

### 2016-09-14 Draft.tex
Shorten title
Add keyword

@@
 
-\title{Hierarchical model assessing the impacts of time to detection distribution assumptions on detection probability estimation}
+\title{Assessing the impacts of time to detection distribution assumptions on detection probability estimation}
 
@@
 
-{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Markov chain Monte Carlo; Stan
+{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Markov chain Monte Carlo; Stan; hierarchical model
 

### 2016-09-15 Draft.tex
Add a couple of JABES references and a figure of representative posteriors of p^{(det)}.

@@
 We specifically explore two-parameter families of TTDDs, e.g. gamma, that can additionally include a mixture component to model increased probability of detection in the initial observation period.
-We find that modeling a TTDD by using a mixture component is necessary when data have a chance of arising from a distribution of this nature.
-In addition, two-parameter distributions can outperform exponential-based models both when the truth is exponential or non-exponential.
+We find that modeling a TTDD by using a two-parameter family is necessary when data have a chance of arising from a distribution of this nature.
+In addition, models with a mixture component can outperform non-mixture models even when the truth is non-mixture.  
 Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
@@
 
-{\bf Keywords:} abundance; availability; N-mixture model; point counts; removal sampling; survival analysis; Markov chain Monte Carlo; Stan; hierarchical model
+{\bf Keywords:} abundance; availability; hierarchical model; Markov chain Monte Carlo; N-mixture model; point counts; removal sampling; Stan; survival analysis
 
@@
 We embed the TTDD in a hierarchical framework for multinomial counts using an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.  
-For our purposes, the N-mixture framework provides three clear benefits: 1) its multinomial data framework accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
+For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
 The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.  
@@
 In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = e^{\alpha/2}/\varphi$.
-This parameterization of the lognormal relates to the standard ($\mu, \sigma^2$) parameterization by $\varphi = \exp(-\mu)$ and $\alpha = \sigma^2$.  %$\mu = -\log(\varphi)$
+To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = \exp(\alpha^2/2)/\varphi$.
+This parameterization of the lognormal relates to the standard ($\mu, \sigma^2$) parameterization by $\varphi = \exp(-\mu)$ and $\alpha = \sigma$.  %$\mu = -\log(\varphi)$
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$.
@@
 
-Table \ref{tbl:sim2} provides a summary of the ability of each model to capture the detection probability. 
-Each quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
+Figure \ref{fig:sim2typical} presents a representative example of posterior estimates of $\pdet$ for data and models from the exponential and gamma families.
+Table \ref{tbl:sim2} provides a more complete summary of the ability of each model to capture the detection probability. 
+Each table quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
 The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak, because the two parameters (rate and mixing parameter) do not provide enough flexibility for the mixture exponential distribution to adequately fit a TTDD with both an initial increase and a delayed mode.
@@
 However, exponential model estimates are both less biased and more precise when the data actually do derive from an exponential mechanism.
+
+\begin{figure}
+\includegraphics[width=0.95\textwidth, height=3in]{"Oral Prelim/Posteriors_Families"}
+\caption{\label{fig:sim2typical} Representative example of posterior distributions for $\pdet$.
+Data derive from mixture TTDDs of nonpeaked gamma, exponential, and peaked gamma families, and are fit with inference models from exponential and gamma families.
+The vertical dashed line presents the true value of $\pdet = 0.80$.}
+\end{figure}
 
@@
 Because our focus has been on modeling variations in detection rates during the survey period, we have not incorporated distance into our model.
-Consequently, our application of a TTDD represents an averaging across distance classes, which induces systematic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.  
-To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \citeauthor{borchers2015}.  
+Consequently, our application of a TTDD represents an averaging across distance classes, which induces systematic bias in estimates of abundance \citep{EffordDawson2009, Laake2011, Solymos2013}.  
+To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \citeauthor{Borchers2016}.  
 This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Amundson2014}.

### 2016-09-20 Draft.tex
Update to figure (black-and-white, text editing).  Change figure in draft.
Update figure caption

@@
 % \usepackage[dvipsnames]{xcolor}
-\usepackage{caption}
+\usepackage[margin=1cm]{caption}
 %\usepackage{tabularx, booktabs}
@@
 
-Figure \ref{fig:sim2typical} presents a representative example of posterior estimates of $\pdet$ for data and models from the exponential and gamma families.
-Table \ref{tbl:sim2} provides a more complete summary of the ability of each model to capture the detection probability. 
+Figure \ref{fig:sim2typical} presents a representative example of posterior distributions for $\pdet$ for data and models from the exponential and gamma mixture families. 
+In this example, the posterior distribution under an exponential inference model accurately captures the true detection probability when the simulation model is an exponential, but overestimates (underestimates) the detection probability when the simulation model has a small (large) gamma shape parameter.
+In contrast, the posterior from a gamma family, which has the additional flexibility of the shape parameter, is able to accurately capture the truth in all scenarios with increased uncertainty.
+
+To summarize these findings across the various simulation and inference models, Table \ref{tbl:sim2} provides simulation-averaged Median estimates of $\pdet$ (Med $p$), proportion of the $\pdet$ posterior larger than the truth ($Q(p)$), and 50\% and 90\% coverage proportions.
 Each table quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
@@
 
-\begin{figure}
-\includegraphics[width=0.95\textwidth, height=3in]{"Oral Prelim/Posteriors_Families"}
-\caption{\label{fig:sim2typical} Representative example of posterior distributions for $\pdet$.
+\begin{figure}\centering
+\includegraphics[width=\textwidth]{"Oral Prelim/Posteriors_Families_draft"}
+\caption{\label{fig:sim2typical} 
+Representative example of the posterior distributions for $\pdet$ simulated from nonpeaked gamma (shape$<$1), exponential (shape$=$1), and peaked gamma (shape$>$1) mixture TTDDs, fit with exponential (light solid line) and gamma (dark solid line), and compared to the true detection probability (dashed vertical line, $\pdet=0.8$).
+% Representative example of posterior distributions for $\pdet$.
 Data derive from mixture TTDDs of nonpeaked gamma, exponential, and peaked gamma families, and are fit with inference models from exponential and gamma families.
-The vertical dashed line presents the true value of $\pdet = 0.80$.}
+The vertical dashed line presents the true value of $\pdet = 0.80$.
+}
 \end{figure}

### 2016-09-20 OVEN/script.R
Rename supplementary files

@@ -1,5 +1,5 @@
 # As a prerequisite, point count survey data should be organized in a dataframe saved as 'surveydata'.
-surveydata <- read.csv("surveydata.csv")
+surveydata <- read.csv("data.csv")
 
 ########## Code to package data as a list for the model
 
@@ -66,12 +66,12 @@ dat = list(n_sites = nrow(Ival),
            tau = tau,
            y = Ival)
 
-save(dat, file="Data for rStan.Rdata")
+save(dat, file="stan_data.Rdata")
 
 
 
 ########## Run Stan code
-load("Data for rStan.Rdata")   # Load Data
+load("stan_data.Rdata")   # Load Data
 
 # Build list of variables to use.
 params <- c("intcpt_a", "ba", "intcpt_d", "bd", "sigma_a", "sigma_d", "gamma", "ra", "rd", "shape_k")

### 2016-09-21 Draft.tex
Add brief supplementary materials section
Modify caption for readability.

@@
 \caption{\label{fig:sim2typical} 
-Representative example of the posterior distributions for $\pdet$ simulated from nonpeaked gamma (shape$<$1), exponential (shape$=$1), and peaked gamma (shape$>$1) mixture TTDDs, fit with exponential (light solid line) and gamma (dark solid line), and compared to the true detection probability (dashed vertical line, $\pdet=0.8$).
-% Representative example of posterior distributions for $\pdet$.
-Data derive from mixture TTDDs of nonpeaked gamma, exponential, and peaked gamma families, and are fit with inference models from exponential and gamma families.
-The vertical dashed line presents the true value of $\pdet = 0.80$.
+Representative example of the posterior distributions for $\pdet$ from exponential (light solid line) and gamma (dark solid line) mixture models fit to data simulated from nonpeaked gamma (shape$<$1), exponential (shape$=$1), and peaked gamma (shape$>$1) mixture TTDDs, with true detection probabilities shown for comparison (dashed vertical line, $\pdet=0.8$).
 }
@@
 
+\section{Supplementary Materials}
+
+The supplementary materials include supplementary figures as well as code to fit these models.
+
 \bibliography{masterbib}

### 2016-09-22 supplemental/script.R
Changes to script.R and data.csv to make them compatible with the new code and to cut out unnecessary variables.

@@ -74,16 +74,16 @@ save(dat, file="stan_data.Rdata")
 load("stan_data.Rdata")   # Load Data
 
 # Build list of variables to use.
-params <- c("intcpt_a", "ba", "intcpt_d", "bd", "sigma_a", "sigma_d", "gamma", "ra", "rd", "shape_k")
+params <- c("intcpt_a", "ba", "intcpt_d", "bd", "sigma_a", "sigma_d", "gamma", "ra", "rd", "alpha")
 
 # MCMC variables for which to store output
-# - omitting lambda and varphi, since they can be reconstructed
+# - omitting lambda and phi, since they can be reconstructed
 # - 'yrep' are simulated counts from the posterior distribution
 # - 'unobserved' are MCMC samples of uncounted individuals = N_s-n_s^{(obs)}
 record.list <- c(params, "unobserved", "yrep")
 
 library(rstan)
-# Compile model -- takes a few minutes
-m = stan_model("WeibullMix.stan", auto_write=rstan_options("auto_write" = TRUE))
-# Perform a short 1000-iteration MCMC sampling, thinned by a factor of 20 -- takes 7.5 minutes
+# Compile model -- takes a couple of minutes
+m = stan_model("WeibullMixture.stan", auto_write=rstan_options("auto_write" = TRUE))
+# Perform a short 1000-iteration MCMC sampling, thinned by a factor of 20 -- takes several minutes
 fit <- sampling(m, data=dat, chains=1, iter=1000, thin=20, pars=record.list)

### 2017-02-15 ReviewerComments.tex
Add reviewers' comments to repository.

\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[margin=1cm]{caption}
\usepackage{adjustbox}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{xr}

\renewenvironment{quote}  % Define 
              {\list{}{\rightmargin\leftmargin}\normalfont%
               \item\relax}
              {\endlist}

\externaldocument[M-]{Draft}

%\doublespacing

\newenvironment{itquote}
{\begin{quote}\itshape}
{\end{quote}}

\newcommand{\adam}[1]{{\color{blue} ADAM: \#1}}
\newcommand{\jarad}[1]{{\color{Orange} JARAD: \#1}}

\newenvironment{indpar}[1]%
     {\begin{list}{}%
             {\setlength{\leftmargin}{\#1}}%
             \item[]%
     }
     {\end{list}}

\newcommand{\vn}{\textbf{n}}
\newcommand{\vp}{\textbf{p}}
\newcommand{\vX}{\textbf{X}}
\newcommand{\vZ}{\textbf{Z}}
\newcommand{\vbeta}{\boldsymbol{\beta}}
\newcommand{\vxi}{\boldsymbol{\xi}}

\newcommand{\Exp}{\mbox{Exp}}
\newcommand{\Ga}{\mbox{Ga}}
\newcommand{\We}{\mbox{We}}
\newcommand{\LN}{\mbox{LN}}
\newcommand{\Po}{\mbox{Po}}
\newcommand{\Mult}{\mbox{Mult}}

\newcommand{\pdet}{p^{(det)}}
\newcommand{\ind}{\stackrel{ind}{\sim}}
\newcommand{\Fm}{F_T^{(M)}}

\begin{document}
\itshape
Associate Editor:

I enjoyed reading this well-written manuscript, and believe that the modeling framework proposed is sufficiently versatile to be used in a wide range of applications. In addition to the points made by the referees, I have two comments:
\begin{enumerate}
\item Reviewer 1 is not convinced that it is helpful to make explicit the link to survival analysis. I tend to disagree with this, since I think that this is what renders the approach widely applicable and also extendable. In fact, the general idea strongly reminded me of several papers published recently by David Borchers and colleagues (using concepts from survival analysis to model detection events within distance sampling and spatially explicit capture-recapture). I was wondering if there is a common denominator between his work and the present manuscript (but the authors need not necessarily explore this).
\begin{quote}
We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
\end{quote}
\item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter. It was not clear to me why the rate parameter was chosen to be modelled as a function of covariates, as opposed to the mean of the TTDD. For example, in gamma GLMs, the mean is modelled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work). Basically I'm now wondering a) why the rate rather than the mean was modelled, and b) if that matters at all. In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used? Sorry if I'm just missing the obvious.
\end{enumerate}

\vspace{1.5cm}

Reviewer \#1:

I have included several items for the authors to consider below and provide this review as an Ecologist and potential model user.
Several pieces of information are lacking from the manuscript that will be relevant to users of this method.
\begin{enumerate}
\item What types of data are relevant to each examined distribution (e.g., gamma, Weibull)?
\item What sample sizes (i.e., number of detections) are needed to obtain reliable estimates with CIs that are not so wide as to be meaningless?
\item What are approximate computation times for models of different complexity (e.g., days? weeks?).
\end{enumerate}

Abstract

Line 14: Perhaps instead of saying its not justified, mention it can lead to bias

Line 29-31: Clarify that you conducted a simulation analysis and your results from that analysis suggest it can outperform non-mixture models.

Line 40-41.  Also worth mentioning caveats - e.g., greater imprecision and computation times.

Introduction\\
Line 36-41: I find mention of survival analyses here and later in the manuscript confusing for most readers - especially Ecologists, presuming this is the intended audience.  It may be due to the unorthodox use of time-to-detection instead of traditional time-to-event analyses often used to estimate survival.  I suggest either omitting the reference to survival analysis here and elsewhere in the manuscript, or using the time-to-event terminology in this sentence and providing citations for the cdf and pdfs outlined here for reference.  Then explain that within the manuscript you will subsequently refer to it as TTDD.

3.1 \\
Line 60: In the context of estimating abundance of wildlife, justifying the inclusion of the Weibull and lognormal distributions because they are often used in survival analysis is weak.  Are there biologically plausible scenarios in which these distributions are appropriate given this type of data?  Justify here similar to the Gamma distribution.  Otherwise, omit.

Simulation Analyses\\
Line 57-59: This is the first mention of peaked and non-peaked TTDDs.  Please define these terms and how they are manifested during point-count surveys in the Introduction - including a biological justification that warrants their inclusion in simulation studies.
As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
\begin{enumerate}
\item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\end{enumerate}

Results\\
Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).  For example, The Gamma distribution, touted as the most accurate based on the simulation study, suggests 95\% CIs of p ranges from ~ 0.3 to 1?  And subsequently CIs around N are from ~1.75 to 3.25 so we can say there are somewhere between 56 and 1,778  birds?  Is that a useful estimate?  There is a tradeoff between accuracy and precision happening here, and although precise, biased estimates are not what we want, any alternative needs to provide precision at levels that are still informative.  This is a limitation not sufficiently addressed in the Discussion.

Discussion\\
Page 21, Lines 12-16:  I do not agree with this assessment.  I think these models are appropriate for a particular subset of data - single species discrete-count surveys with a large sample size and relatively low availability (i.e., whether the animal provides a cue to an observer during the survey) where non-constant detection through time is suspected based on the behavioral ecology of the species.  As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.  Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.

\vspace{1.5cm}

Reviewer \#2:

Although many statistical solutions to known issues in abundance estimation have been proposed to date, a plethora of relevant problems remain untreated.  In this manuscript, the author(s?) put the finger on the wound, so to speak, and deal with a very relevant problem: allowing for heterogeneity in detection, or a non-constant detection rate during animal point-count surveys.  This topic is relevant because, as the authors clearly state multiple times, the statistical properties of abundance estimates are sensitive to the statistical properties of the estimates of the detection probabilities. In that sense, I commend the manuscript's pertinence: the question treated is indeed relevant and needs careful examination.
The advent of computer intensive approaches to estimate parameters in hierarchical models very quickly re-shaped the field of abundance estimation, and in less than a decade, we've moved from trying to explain a complicated natural signal with the simplest possible model to trying to explain a complicated natural signal with a 'realistic' but equally complicated sampling model. Then, the burden of the quality of the estimation is put into the process of the specification of the sampling model. However, there is little guarantee that the data, as considered in the paper, contains the necessary information to be able to reliably tease apart all the components of the statistical sampling model.  In what follows, I not only expand on this topic but I outline a few major and minor comments that I hope the author(s) will regard as useful to improve the quality of the manuscript. I will certainly recommend this manuscript for publication after the author(s) include the
modifications I request or she (he) successfully convinces me otherwise.

1. MAJOR COMMENTS

Diagnostics: Coverage. The idea of testing coverage is fantastic, but poorly implemented. First, the number of simulations (16 in Table 1) is exceedingly low to be able to reliably diagnose the patterns in coverage or the statistical explanation of these patterns. Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det). Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det) (Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is "uniformly" good?). Without variation in "true" simulated scenarios, it is easy to inadvertendly "stack the deck" in favor or against a particular combination of settings (Particularly since your "true" value of p is high: 0.8). A wide range of simulted truths is also necessary because to be useful, these methods should be widely applicable in the
tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.

Diagnostics: relation of the target parameter to other parameters in the models. Prercisely because, as the authors mention, the statistical properties of the estimator of p(det) are tightly linked with the statistical properties of the abundance estimator, then the author(s) should detail explicitly how a bad coverage in p(det), for instance, affects the estimation of abundances. The author(s) mention that bias in p translates into a bias in abundance estimation, I would like to see diagnostics for the realized N values too. And to do that, two words come to mind: profile likelihoods. Unless you are explicitly adopting a subjective Bayesian approach, you basically declare a priori ignorance for your parameters.  Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not "swamping" the priors, so to speak. Lele et al 2010 (a frequent co-author of Solymos) propose a pretty neat diagnostic tool to assess
estimability that is the by product of tricking a bayesian MCMC set up into Maximum Likelihood estimation for hierarchical models (see Lele et al 2007, 2010 and others. The keyword is "Data Cloning"). You are one step away from using Solymos and Lele's "Data Cloning" (DC) approach to get the full ML estimation working (See Lele et al 2010 and other Data Cloning papers). Now, this is relevant because using DC you can provide clear estimability diagnostics for the parameters of interest.  Are there parameters that are technically not estimable?  Given that you don't have informative priors, you are "driving in the dark" so to speak if you are not certain if some of your parameters are un-identifiable (see Lele et al 2010 and discussion in Lele and Dennis 2009).  Note that I am NOT asking you to re-do all the analysis using a ML approach, I am just suggesting the fact that using DC to get the ML estimates one can, as a very useful by product, get very neat diagnostics regarding
the estimability of your parameters.  Again, see Lele et al 2010.  Implementing DC would only imply a simple modification to the programs you already have. And by the way, Solymos has a DC package easy to use.

Generalized Pareto: In another area in biology, in evolutionary genetics, the "fitness" of individual variant strains has been modeled as coming from an exponential distribution.  For various reasons having to do with the biology of the system, such model was quickly challenged and pretty soon papers questioning the validity of this or this other model for data akin to your waiting times abounded. There is one pretty interesting paper by Beisel et al, 2007 (Genetics, 10.1534/genetics.106.068585), written towards the epiloge of such discussions, that posits that a particular parameterization of the Generalized Pareto distribution, with a single stroke, encompassed many suitable probabilistic models in the Weibull, Gumbel and Freched domain of attraction.  I wonder if the authors could write a general parameterization so that changing from one distribution to the other would simply amount to dialing a given parameter, much like Beisel et al do?  This would be just a practical
consideration that can at once deal with model selection and could speed up calculations

Writing: The multiplicity of mathematical explicit meanings for the word "mixture" may result confusing. The distinction between the mixture component a la Farnsworth et al and the N-mixture models (e.g. binomial counts with N~pois and p random) is clear to me, but it may not be for some of JABES' audience.  I suggest re-shaping the introduction to that effect. One thing that the authors could do is to take the reader by the hand by building equation 1 bit by bit: present first the simplest version of equation 1 and what is customary to date, and then add modifications/hierarchies plus text and build the model with explicit equations and little by little arrive to the final product, which is the current equation 1. Nothing better than clear simple math aided by short explanations to present a model unambiguously.

Write self-contained paragraphs, with one main idea, not two or three (and each one half developed). As a reviewer, I enjoyed verifying that the cited papers in the introduction and in the discussion were indeed meaningful (by downloading them and reading those that I was not familiar with). However, when many papers are cited, it is useful to expand in the text why the different papers are being cited.  Doing so not only clarifies your intentions, but does proper justice to the papers being referred to. It is also a useful exercise because it allows you to tell whether you have more than one clear ideas and hence, material for more than one paragraph.  For instance, I think that the paragraph in the introduction, line 34, page 3, could be expanded and/or broken into two paragraphs.  One introducing the time-to-detection as is done in survival analysis, and one presenting to the reader how data that seem not to conform to the constant-detection assumption is dealt with by
using the idea of the increased detection probability via a mixture component. These are two different ideas, hence two different paragraphs. The same approach to construct paragraphs through the text should be taken.

The authors could have crafted a beautiful and very telling graph (Figure 1) representing the multinomial intervals, the TTDD, the right censoring due to the end of the observation period, and the increased detection probability mixture component. Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.

Process vs. observation error: Finally, here's a perspective outside the author's topic, but within statistical ecology that might result in a markedly improvement of these "N-mixture" techniques, along with all of its variants. I wonder if thinking a bit more in terms of modeling the biological process behind the data can result in a better estimation of the sampling variance and as a "by-product", some novel understanding of the biology behind the data. I suppose the author(s) are familiar (or at least have seen) the multiple papers where statistical inference is done for a stochastic population dynamics model using time series of abundances while taking into account sampling error. In these settings, it is customary to deal with one observed abundance per time step. The sampling error and the ecologically-phrased variability are phrased using a hierarchical model.  When statistical inference is done for this hieararchical model, the information to be able to tease apart
the process from the observation variance lies within the structure of the temporal dependencies phrased in the model. Amazingly, with a single observation per time step, the sampling model is not ill posed as the time dependencies in the process brings about enough information in the data.  As a result, thinking of the biological process (in that case, the population dynamics model) yields better, unbiased estimates of the observation error. In your case, I cannot help but wonder if having data varying along the axis of time or space result in a much better estimation of the sampling noise. Just as exponential waiting times are, as you surely are aware of, tightly linked to continuous time, discrete state Markov models, the other TTDD models could be linked to non-Markovian processes of moving/singing/behavior.  In many instances, multiple observations in the same "point-count" are available. Anyways, this are just some thoughts rapidly put together, please comment on this
if you think it's worthwhile.

Thank you for the opportunity to comment on your paper, I hope you find my views useful.  I think yours will in the end, be a very important contribution.

\bibliography{masterbib}
\bibliographystyle{biom}

\end{document}

### 2017-02-20 Draft.tex
Remove reference to data/code in supplementary materials
I'm having trouble submitting the data/code even after rounds of emails.
Thus, I suggest we forgo at this point.
In the future, we could just link to a github repo with the materials.
Reviewers' comments and associated updates.  In process.

@@
 \usepackage{xr}
+\usepackage{lineno}
 
@@
 
-\providecommand{\blind}{0}
+\providecommand{\blind}{1}
 
@@
 Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
-The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption is often not justified.
+The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption can often lead to biased estimates.
 %Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.  
@@
 We specifically explore two-parameter families of TTDDs, e.g. gamma, that can additionally include a mixture component to model increased probability of detection in the initial observation period.
-We find that modeling a TTDD by using a two-parameter family is necessary when data have a chance of arising from a distribution of this nature.
+Based on simulation analyses, we find that modeling a TTDD by using a two-parameter family is necessary when data have a chance of arising from a distribution of this nature.
 In addition, models with a mixture component can outperform non-mixture models even when the truth is non-mixture.  
@@
 In this manuscript, we develop a model for scenarios where detection rates are not constant over time. 
-We consider the first time-to-detection as is done in survival analysis, defining a continuous random variable $T$ for each individual's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
+We analyze times to first detection as time-to-event data in a parametric survival analysis, defining a continuous random variable $T$ for each individual's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
@@
 The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s. 
+A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
+% The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
 
@@
 During each survey, the Julian date, time of day, and temperature were recorded.
-
-% A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
-% The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
 
@@
 For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
-In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
+%In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
+Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow for rates to either increase or decrease during the survey; however, the shapes of these distributions will differ, most notably in their tail probabilities.
+All three could be reasonable TTDD choices if, for instance, we believe detection rates vary from individual to individual.
+In thats case, marginal detection rates decrease over time as individuals with higher detection rates are removed from the local population.
 To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = \exp(\alpha^2/2)/\varphi$.
@@
 We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
-Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
+%Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
 We discarded half of the iterations as warmup and then thinned by 10.  
@@
 The second study compares the TTDD families.
-In the first two studies, we utilized intercept only models to focus attention on the TTDD. 
+In the first two studies, we utilized intercept only models to focus attention on robustness of the TTDD choice in the most simple of scenarios.
 For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families. 
-In all simulation studies, we focus on accuracy in estimation of $\pdet$ which then translates into estimation of abundance. 
-For the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
-A peaked TTDD has a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$).
+In all simulation studies, we focused on accuracy in estimation of $\pdet$ which then translated into estimation of abundance. 
+In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
+Detection rates $\varphi(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
+More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked.
 
@@
 
-Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the logarithm of the number of uncounted Ovenbirds.  
+Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the logarithm of total Ovenbird abundance.  
 Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remain relatively large.
@@
 
-In spite of the similarity of effect parameter estimates, the posterior distributions for detection probability and uncounted abundance differ greatly between the exponential and non-exponential models.
+In spite of the similarity of effect parameter estimates, the posterior distributions for detection probability and abundance differ greatly between the exponential and non-exponential models.  
 It is clear that the assumption of constant detection leads to much higher and more precise estimates of detection than would be obtained if we are unwilling to make that assumption.
@@
 \caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.
-Posteriors are also available for the number of uncounted individuals and the overall probability of detection across all sites.}
+Posteriors are also available for the overall probability of detection and the abundance across all sites.}
 \end{figure}
@@
 An assumption of constant detection places constraints on the amount of uncertainty in the extrapolated tail but at the cost of potentially sizable bias.
-When we allow for non-constant detection, estimation of the tail probability becomes less certiain.
+When we allow for non-constant detection, estimation of the tail probability becomes less certain.
 In theory, estimation of the unsampled tail probability can be improved by conducting longer surveys, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
@@
 To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \citeauthor{Borchers2016}.  
-This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Amundson2014}.
+This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}.
 The differences between these implementations may impact estimates of detection and abundance, especially in the presence of behavioral heterogeneity in availability rates across subgroups of the study population.
@@
 
-We recommend that time-heterogeneous detection rates be explicitly modeled in analyses involving removal-sampled point-count survey data where estimation of detection probability or abundance is a primary objective.
+We recommend that time-heterogeneous detection rates be explicitly modeled in single-species analyses involving removal-sampled point-count survey data where estimation of detection probability or abundance is a primary objective.
 The assumption of constant detection, while computationally simple and reasonable as a null model, proves to be rather informative and can result in pronounced bias.  
@@
 
-The supplementary materials include supplementary figures as well as code to fit these models.
+The supplementary materials include supplementary tables and figures.
 

### 2017-02-20 ReviewerComments.tex
Reviewers' comments and associated updates.  In process.

@@
 
-\newcommand{\adam}[1]{{\color{blue} ADAM: \#1}}
+\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
 \newcommand{\jarad}[1]{{\color{Orange} JARAD: \#1}}
@@
 We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
+Their models are for removal-distance data, whereas ours is for removal-only data.
 \end{quote}
 \item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter.
-It was not clear to me why the rate parameter was chosen to be modelled as a function of covariates, as opposed to the mean of the TTDD.
-For example, in gamma GLMs, the mean is modelled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work).
-Basically I'm now wondering a) why the rate rather than the mean was modelled, and b) if that matters at all.
+It was not clear to me why the rate parameter was chosen to be modeled as a function of covariates, as opposed to the mean of the TTDD.
+For example, in gamma GLMs, the mean is modeled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work).
+Basically I'm now wondering a) why the rate rather than the mean was modeled, and b) if that matters at all.
 In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used?
@@
 
-\vspace{1.5cm}
+\newpage
 
@@
 Line 14: Perhaps instead of saying its not justified, mention it can lead to bias
+\begin{quote}
+Done.
+\end{quote}
 
 Line 29-31: Clarify that you conducted a simulation analysis and your results from that analysis suggest it can outperform non-mixture models.
+\begin{quote}
+Done.
+\end{quote}
 
 Line 40-41.  Also worth mentioning caveats - e.g., greater imprecision and computation times.
+\begin{quote}
+\adam{Address this point}
+\end{quote}
 
@@
 Then explain that within the manuscript you will subsequently refer to it as TTDD.
+\begin{quote}
+We have explicitly defined time-to-detection data as being time-to-event data.\\
+\adam{I am unclear what the reviewer means by `providing citations for the cdf and pdfs outlined here for reference'.}
+\end{quote}
 
@@
 Otherwise, omit.
+\begin{quote}
+We have added an example where detection rates vary across individuals, causing the marginal detection rate to decline over time.
+\end{quote}
 
@@
 Please define these terms and how they are manifested during point-count surveys in the Introduction - including a biological justification that warrants their inclusion in simulation studies.
+\begin{quote}
+We have made the meaning of `peaked' and `nonpeaked' more explicit in Section \ref{M-sec:sim}.
+We do not believe this distinction requires a separate explanation in the Introduction.
+\end{quote}
+
 As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
@@
 \item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
+\begin{quote}
+We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
+However, it can be informative.
+TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
+\end{quote}
 \item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
+\begin{quote}
+\adam{Investigate more/different simulations}
+\end{quote}
 \item How might study design impact bias and precision?
@@
 Alternatively, these issues could be addressed in the Discussion.
+\begin{quote}
+\adam{We have not investigated either of these issues.
+During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
+However, I would not dream of extending those findings to other distributions.
+We obviously would prefer longer durations, but those risk the violation of closure assumptions.
+\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
+I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
+
+To the reviewer's point, we can emphasize the value of a settling-in period toward justifying the use constant-rate assumptions in analysis.}
+\end{quote}
 \item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
+\begin{quote}
+\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
+\end{quote}
 \item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.
 It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
+\begin{quote}
+We have retained just the 50\% coverage.
+\adam{Waiting until all new simulations are run.}
+\end{quote}
 \end{enumerate}
@@
 Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
+\begin{quote} Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections)\end{quote}
 For example, The Gamma distribution, touted as the most accurate based on the simulation study, suggests 95\% CIs of p ranges from ~ 0.3 to 1?
@@
 This is a limitation not sufficiently addressed in the Discussion.
+\begin{quote} 
+We think our figure may have been unclear.
+The cited credible intervals are for \textit{uncounted} birds not total abundance.
+We have now changed the figure to $\log_{10}$(Abundance) to avoid confusion.
+The gamma TTDD 95\% credible interval for total abundance is (1007, 2836).
+This two- to three-fold scale of uncertainty is commensurate with other studies cited in this manuscript \citep{Diefenbach2007, Reidy2011, Solymos2013, Amundson2014}.
+\end{quote}
 
@@
 Page 21, Lines 12-16:  I do not agree with this assessment.
-I think these models are appropriate for a particular subset of data - single species discrete-count surveys with a large sample size and relatively low availability (i.e., whether the animal provides a cue to an observer during the survey) where non-constant detection through time is suspected based on the behavioral ecology of the species.
+I think these models are appropriate for a particular subset of data - single species discrete-count surveys with a large sample size and relatively low availability (i.e., whether the animal provides a cue to an observer during the survey)
+\begin{quote}
+\adam{What is a `discrete-count survey'?  I have not found a clear reference.\\
+I accede the `large sample' point.
+But based on our findings, I'd advocate against modeling abundance/perception from any small dataset, because analysis requires either a strong assumption about detection rates or a strong prior.
+Any comment on this point would be strengthened if we ran small-dataset simulations.\\
+I think the low-availability stipulation is redundant.  If availability is really high, then a removal analysis is not necessary at all.}
+\end{quote}
+where non-constant detection through time is suspected based on the behavioral ecology of the species.
+\begin{quote}
+\adam{What I do not advocate is defaulting to the constant-rate assumption: by the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.}
+\end{quote}
+
 As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.
 Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
+\begin{quote}
+We agree.  Our current research addresses this issue in more depth.
+\end{quote}
 Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.
@@
 Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.
+\begin{quote}
+\adam{This last recommendation seems feasible, though I think our final paragraph already addesses some of these issues.}
+\end{quote}
+\vspace{1.5cm}
 
-\vspace{1.5cm}
+\newpage
 
@@
 In that sense, I commend the manuscript's pertinence: the question treated is indeed relevant and needs careful examination.
-The advent of computer intensive approaches to estimate parameters in hierarchical models very quickly re-shaped the field of abundance estimation, and in less than a decade, we've moved from trying to explain a complicated natural signal with the simplest possible model to trying to explain a complicated natural signal with a 'realistic' but equally complicated sampling model.
+The advent of computer intensive approaches to estimate parameters in hierarchical models very quickly re-shaped the field of abundance estimation, and in less than a decade, we've moved from trying to explain a complicated natural signal with the simplest possible model to trying to explain a complicated natural signal with a `realistic' but equally complicated sampling model.
 Then, the burden of the quality of the estimation is put into the process of the specification of the sampling model.
 However, there is little guarantee that the data, as considered in the paper, contains the necessary information to be able to reliably tease apart all the components of the statistical sampling model.
+\begin{quote}
+We appreciate the reviewers' insights and whole-heartedly agree.
+\end{quote}
 In what follows, I not only expand on this topic but I outline a few major and minor comments that I hope the author(s) will regard as useful to improve the quality of the manuscript.
@@
 Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det).
-Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det) (Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is "uniformly" good?).
-Without variation in "true" simulated scenarios, it is easy to inadvertendly "stack the deck" in favor or against a particular combination of settings (Particularly since your "true" value of p is high: 0.8).
-A wide range of simulted truths is also necessary because to be useful, these methods should be widely applicable in the
+Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det)
+\begin{quote}
+\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wide range or simulated $\pdet$.}
+\end{quote}
+(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
+\begin{quote}
+\adam{Bolster this with more citations?  \citet{Veech2016} argue for $p>0.5$ in the context of individual-level detection effects (no time-to-detection). 
+Plots from \citet{Field2016} suggest uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.
+Methods include removal, double-observer, distance, and multiple-visit.
+In a traditional removal-sampling context, \citet{Davis2016} report accurate $\hat{N}$ at $\pdet >0.40$, though for small abundance ($N<50$) they required $\pdet > 0.7$.}
+\end{quote}
+
+ Without variation in ``true" simulated scenarios, it is easy to inadvertendly ``stack the deck" in favor or against a particular combination of settings (Particularly since your ``true" value of p is high: 0.8).
+A wide range of simulated truths is also necessary because to be useful, these methods should be widely applicable in the
 tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
@@
 Diagnostics: relation of the target parameter to other parameters in the models.
-Prercisely because, as the authors mention, the statistical properties of the estimator of p(det) are tightly linked with the statistical properties of the abundance estimator, then the author(s) should detail explicitly how a bad coverage in p(det), for instance, affects the estimation of abundances.
+Precisely because, as the authors mention, the statistical properties of the estimator of p(det) are tightly linked with the statistical properties of the abundance estimator, then the author(s) should detail explicitly how a bad coverage in p(det), for instance, affects the estimation of abundances.
 The author(s) mention that bias in p translates into a bias in abundance estimation, I would like to see diagnostics for the realized N values too.
@@
 Unless you are explicitly adopting a subjective Bayesian approach, you basically declare a priori ignorance for your parameters.
-Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not "swamping" the priors, so to speak.
+Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not ``swamping" the priors, so to speak.
 Lele et al 2010 (a frequent co-author of Solymos) propose a pretty neat diagnostic tool to assess
 estimability that is the by product of tricking a bayesian MCMC set up into Maximum Likelihood estimation for hierarchical models (see Lele et al 2007, 2010 and others.
-The keyword is "Data Cloning").
-You are one step away from using Solymos and Lele's "Data Cloning" (DC) approach to get the full ML estimation working (See Lele et al 2010 and other Data Cloning papers).
+The keyword is ``Data Cloning").
+You are one step away from using Solymos and Lele's ``Data Cloning" (DC) approach to get the full ML estimation working (See Lele et al 2010 and other Data Cloning papers).
 Now, this is relevant because using DC you can provide clear estimability diagnostics for the parameters of interest.
 Are there parameters that are technically not estimable?
-Given that you don't have informative priors, you are "driving in the dark" so to speak if you are not certain if some of your parameters are un-identifiable (see Lele et al 2010 and discussion in Lele and Dennis 2009).
+Given that you don't have informative priors, you are ``driving in the dark" so to speak if you are not certain if some of your parameters are un-identifiable (see Lele et al 2010 and discussion in Lele and Dennis 2009).
 Note that I am NOT asking you to re-do all the analysis using a ML approach, I am just suggesting the fact that using DC to get the ML estimates one can, as a very useful by product, get very neat diagnostics regarding
@@
 
-Generalized Pareto: In another area in biology, in evolutionary genetics, the "fitness" of individual variant strains has been modeled as coming from an exponential distribution.
+Generalized Pareto: In another area in biology, in evolutionary genetics, the ``fitness" of individual variant strains has been modeled as coming from an exponential distribution.
 For various reasons having to do with the biology of the system, such model was quickly challenged and pretty soon papers questioning the validity of this or this other model for data akin to your waiting times abounded.
@@
 
-Writing: The multiplicity of mathematical explicit meanings for the word "mixture" may result confusing.
+Writing: The multiplicity of mathematical explicit meanings for the word ``mixture" may result confusing.
 The distinction between the mixture component a la Farnsworth et al and the N-mixture models (e.g.
@@
 Process vs.
-observation error: Finally, here's a perspective outside the author's topic, but within statistical ecology that might result in a markedly improvement of these "N-mixture" techniques, along with all of its variants.
-I wonder if thinking a bit more in terms of modeling the biological process behind the data can result in a better estimation of the sampling variance and as a "by-product", some novel understanding of the biology behind the data.
+observation error: Finally, here's a perspective outside the author's topic, but within statistical ecology that might result in a markedly improvement of these ``N-mixture" techniques, along with all of its variants.
+I wonder if thinking a bit more in terms of modeling the biological process behind the data can result in a better estimation of the sampling variance and as a ``by-product", some novel understanding of the biology behind the data.
 I suppose the author(s) are familiar (or at least have seen) the multiple papers where statistical inference is done for a stochastic population dynamics model using time series of abundances while taking into account sampling error.
@@
 Just as exponential waiting times are, as you surely are aware of, tightly linked to continuous time, discrete state Markov models, the other TTDD models could be linked to non-Markovian processes of moving/singing/behavior.
-In many instances, multiple observations in the same "point-count" are available.
+In many instances, multiple observations in the same ``point-count" are available.
 Anyways, this are just some thoughts rapidly put together, please comment on this

### 2017-02-21 Draft.tex
Additions to reviewer comments / modifications to draft.

@@
 In this manuscript, we develop a model for scenarios where detection rates are not constant over time. 
-We analyze times to first detection as time-to-event data in a parametric survival analysis, defining a continuous random variable $T$ for each individual's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
+We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, defining a continuous random variable $T$ for each individual's time to first detection with a probability density function (pdf) $f_T(t)$ and cumulative distribution function (cdf) $F_T(t)$.
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).  
-One common strategy to deal with data that do not fit a constant-detection assumption is to model increased detection probability in the initial observation period via a mixture component \citep{Farnsworth2002, Farnsworth2005, EffordDawson2009, Etterson2009, Reidy2011}, although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
+One common strategy to deal with data that do not fit a constant-detection assumption is to model the TTDD as a mixture of two distributions --- a continuous-time distribution and an increased detection probability in the initial observation period \citep{Farnsworth2002, Farnsworth2005, EffordDawson2009, Etterson2009, Reidy2011} --- although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
 We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
@@
 Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown and may be the primary quantity of interest.  
-We embed the TTDD in a hierarchical framework for multinomial counts using an N-mixture model \citep{Wyatt2002, Royle2004NMixture}.  
+We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture}, which references an entirely different `mixture' than the mixture TTDDs in the previous paragraph.
 For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
@@
 
+\begin{figure}\centering
+\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"}
+\caption{\label{fig:schematic} 
+Illustration of fitting mixture exponential (left) and mixture gamma (right) time-to-detection distributions (TTDDs) to multinomial removal-sampled observations (center) assuming no survey-level heterogeneity.
+The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component (light gray rectangle), which constitute $\gamma$ and $1-\gamma$ proportions of the population, respectively.
+We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving a proportion $1-\pdet$ undetected.
+}
+\end{figure}
+
 \section{Continuous time-to-detection N-mixture models} \label{sec:model}
@@
 %In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow for rates to either increase or decrease during the survey; however, the shapes of these distributions will differ, most notably in their tail probabilities.
-All three could be reasonable TTDD choices if, for instance, we believe detection rates vary from individual to individual.
-In thats case, marginal detection rates decrease over time as individuals with higher detection rates are removed from the local population.
+Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions will differ, most notably in tail probability.
+All three could be reasonable TTDD choices if, for instance, detection rates vary from individual to individual, thus leading to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population.
 To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = \exp(\alpha^2/2)/\varphi$.
@@
 With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
-In these models, point estimates of $N_s$ are unstable unless $\pdet>0.4$ \citep{Olkin1981}.
+In these models, point estimates of $N_s$ are unstable unless $\pdet>0.41$ \citep{Olkin1981}.
 One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$ \citep{Raftery1988, Royle2004NMixture}. 
@@
 To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Alldredge2007, Etterson2009, Reidy2011}.
+We refer to models with and without the mixture component as `mixture models' and `non-mixture models', respectively, which are not to be confused with the `N-mixture' model framework we employ.
 We specify a mixture TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time detection distribution $\Fm(t)$.  

### 2017-02-21 ReviewerComments.tex
Additions to reviewer comments / modifications to draft.

@@
 \begin{quote}
-\adam{Bolster this with more citations?  \citet{Veech2016} argue for $p>0.5$ in the context of individual-level detection effects (no time-to-detection). 
-Plots from \citet{Field2016} suggest uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.
-Methods include removal, double-observer, distance, and multiple-visit.
-In a traditional removal-sampling context, \citet{Davis2016} report accurate $\hat{N}$ at $\pdet >0.40$, though for small abundance ($N<50$) they required $\pdet > 0.7$.}
+It is not a hard boundary.  It results from the relative sample mean and sample variance of large samples.
+However, similar thresholds continue to surface again and again in various abundance-estimating contexts: 
+\begin{itemize}
+\item \citet{Veech2016}: suggest  $p>0.5$ in the context of individual-level detection effects (no time-to-detection data). 
+\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
+\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
+\end{itemize}
+\adam{Jarad, I am unclear on the purpose of the Olkin-citing paragraph.  I don't feel like the above arguments really merit inclusion in the text.}
 \end{quote}
@@
 Furthermore, you state that typically you will be in a case where data sets aren't large, so the information in the data is not ``swamping" the priors, so to speak.
+\begin{quote}
+We have posterior estimates of abundance from our model fits.  We have added them to the manuscript.\\
+\adam{Thoughts on how to do this:
+\begin{enumerate}
+\item Generate Tables 1 \& 2 for $\hat{N}$ as well as $\pdet$, but this gets crowded.
+\item Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$, but averaging across replicates is not a trivial issue
+\item Supplement Figure 1 with a row of representative abundance posteriors (no averaging required)
+\item Following the profile likelihood comment, generate a plot of $\hat{N}$ vs. $\pdet$ for a representative sample.
+\end{enumerate}
+}
+\end{quote}
+
 Lele et al 2010 (a frequent co-author of Solymos) propose a pretty neat diagnostic tool to assess
@@
 And by the way, Solymos has a DC package easy to use.
+\begin{quote}
+\adam{
+\begin{itemize}
+\item The Solymos package is built for BUGS and JAGS.  Most of the tools are easily hand-coded, I think.
+\item To simplify our lives, we could choose to attempt this for only quick-running models on the actual data...
+there's no reason to suspect identifiability issues for the gamma TTDD that would not appear for the Weibull or lognormal.
+\item Technical note: they suggest using narrower priors as the number of clones increases.  Likewise, we could input posterior means as initial values.
+\end{itemize}
+}
+\end{quote}
 
@@
 consideration that can at once deal with model selection and could speed up calculations
+\begin{quote}
+We appreciate the merit of the reviewer's suggestion and have added comments in the Discussion.
+We focus our comments on a possible generalized gamma TTDD, which is an umbrella distribution that includes exponential, gamma, lognormal, and Weibull distributions as special cases.
+The generalized Pareto analysis in \citet{Beisel2007} requires observations from the tail of the TTDD distribution, which we do not have.
+Additionally, its focus is on flexible domains of attraction, whereas the TTDDs in our analysis are all from the Gumbel domain of attraction.
+\adam{After researching the generalized gamma, add comments in the paragraph beginning `If the estimation of abundance is the primary interest...'.}
+\end{quote}
 
@@
 I suggest re-shaping the introduction to that effect.
+\begin{quote}
+We have inserted clarifications: (i) when we first reference the N-Mixture framework in the Introduction, and (ii) when we first use the term `mixture model' in Section \ref{M-s:subgroups}.
+\end{quote}
+
 One thing that the authors could do is to take the reader by the hand by building equation 1 bit by bit: present first the simplest version of equation 1 and what is customary to date, and then add modifications/hierarchies plus text and build the model with explicit equations and little by little arrive to the final product, which is the current equation 1.
 Nothing better than clear simple math aided by short explanations to present a model unambiguously.
+\begin{quote}
+\adam{
+This potentially requires a lot of rewiring.
+Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler version of the model without mixtures and without covariates.
+}
+\end{quote}
 
@@
 Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.
+\begin{quote}
+We have crafted a beautiful and very telling graph (Figure 1).
+\end{quote}
 

### 2017-02-21 Draft.tex
Ongoing revisions.

@@
 % Johnson2008: all birds present are available for detection
-A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007}.  
-In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Otis1978, Farnsworth2005}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
-saturation or lack of settling in period \citep{Petit1995, LeeMarsden2008, Johnson2008}, can all lead to time-varying rates of detection.
+A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007, LeeMarsden2008}.  
+In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Otis1978, Farnsworth2002}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
+saturation or lack of settling down period \citep{Petit1995, LeeMarsden2008, Johnson2008}, can all lead to time-varying rates of detection.
 
@@
 We refer to the distribution of $T$ as a time-to-detection distribution (TTDD).
-One common strategy to deal with data that do not fit a constant-detection assumption is to model the TTDD as a mixture of two distributions --- a continuous-time distribution and an increased detection probability in the initial observation period \citep{Farnsworth2002, Farnsworth2005, EffordDawson2009, Etterson2009, Reidy2011} --- although this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}.
+
+One common strategy to deal with data that do not fit a constant-detection assumption is to model the TTDD as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period \citep{Farnsworth2002, Farnsworth2005, EffordDawson2009, Etterson2009, Reidy2011}.
+However, this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}. 
 We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
+We apply the term `mixture model' to any model with a mixture TTDD.
 
 Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown and may be the primary quantity of interest.  
-We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture}, which references an entirely different `mixture' than the mixture TTDDs in the previous paragraph.
+We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture}, which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
 For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
 The N-mixture framework models abundance as a latent variable with a Poisson or other discrete distribution and independently models detection probabilities.  
-Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data while assuming constant detection rates \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}.
+Several previous studies have employed the N-mixture framework to analyze removal sampled point-count data  \citep{Royle2004Generalized, Dorazio2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}.
 
@@
 %However, the model we describe is appropriate for point-count surveys of any species, so long as time-to-first-detection data are available.  
-For our analysis, we focused on Ovenbird counts selected from one habitat type: sawtimber red pine stands with no recent logging activity.  
+For our analysis, we focus on Ovenbird counts selected from one habitat type: sawtimber red pine stands with no recent logging activity.  
 Each stand had up to four sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
 The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s. 
-A total of 947 Ovenbirds were counted with a maximum of eight birds counted in any single survey.  
+A total of 947 Ovenbirds were counted with a maximum of eight in any single survey.  
 % The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
@@
 The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component (light gray rectangle), which constitute $\gamma$ and $1-\gamma$ proportions of the population, respectively.
-We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving a proportion $1-\pdet$ undetected.
+We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving an estimated proportion $1-\pdet$ undetected.
 }
@@
 In reality, times to first detection are often truncated due to a finite survey length of $C$, meaning that each individual has a detection probability $\pdet=F_T(C)$.  
-The conditional distribution of observed detection times then has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$, and instantaneous detection rate, or hazard function, is $h(t) = f_T(t) / [1-F_T(t)]$.
-We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
+The conditional distribution of observed detection times consequently has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$, and instantaneous detection rate, or hazard function, $h(t) = f_T(t) / [1-F_T(t)]$.
 %$T_{sb|det} \ind f_{T|det}(t_{sb})$ where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$.
@@
 %In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions will differ, most notably in tail probability.
-All three could be reasonable TTDD choices if, for instance, detection rates vary from individual to individual, thus leading to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population.
+Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
+All three could be reasonable TTDD choices if, for instance, detection rates vary among individuals.
+This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population.
+
 To facilitate the later inclusion of fixed and random effects, we use the following rate-based parameterizations: $T\sim \Exp(\varphi), E[T]=1/\varphi$; $T\sim \Ga(\alpha,\varphi), E[T] = \alpha/\varphi$; $T\sim \We(\alpha,\varphi), E[T]=\mathrm{\Gamma}(1+1/\alpha)/\varphi$; and $T\sim \LN(\varphi,\alpha), E[T] = \exp(\alpha^2/2)/\varphi$.
@@
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$. 
+We employ a log link for $\varphi$.
+This makes our model equivalent to a generalized linear model with a log link on the mean detection time.
 
+We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
 Using maximum likelihood, we can estimate the parameters in a TTDD from exact times to first detection and thus estimate $\pdet$ and its uncertainty.
@@
 To accommodate this empirical observation, many models of interval-censored removal times define a TTDD with a mixture component to increase the probability of observing individuals in the first interval \citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Alldredge2007, Etterson2009, Reidy2011}.
-We refer to models with and without the mixture component as `mixture models' and `non-mixture models', respectively, which are not to be confused with the `N-mixture' model framework we employ.
 We specify a mixture TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time detection distribution $\Fm(t)$.  
@@
 %Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
-We discarded half of the iterations as warmup and then thinned by 10.  
+We discarded half of the iterations as warmup and thinned by 10.  
 We monitored convergence of the MCMC chains using Geweke z-score diagnostics \citep{Geweke1991} and reran models if lack of convergence was indicated by a non-normal distribution of the z-scores or if the effective sample size for any parameter was below 1000.
@@
 In all simulation studies, we focused on accuracy in estimation of $\pdet$ which then translated into estimation of abundance.
+
 In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
-Detection rates $\varphi(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
+Detection rates $h(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
 More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked.
@@
 We simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes, see Table \ref{S-tbl:sim3} for actual parameter values.
+Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
+See Table \ref{S-tbl:sim3} for actual parameter values.
 Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
@@
 Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
-Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed, see Section \ref{S-s:sim3} and figures therein.  
+Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (Section \ref{S-s:sim3} and figures therein).  
 Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
@@
 Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the logarithm of total Ovenbird abundance.  
-Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remain relatively large.
+Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remains relatively large.
 %Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from an exponential or nonpeaked mixture TTDD.
@@
 An assumption of constant detection places constraints on the amount of uncertainty in the extrapolated tail but at the cost of potentially sizable bias.
-When we allow for non-constant detection, estimation of the tail probability becomes less certain.
-In theory, estimation of the unsampled tail probability can be improved by conducting longer surveys, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
-Without extending the observation period, an alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB where many species are observed simultaneously
+When we allow for non-constant detection, bias is less problematic but estimation of the tail probability becomes less certain.
+
+Study design can address some issues associated with non-constant detection.
+In theory, longer surveys can improve estimation of the unsampled tail probability, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
+Observer effects on singing rates can be mitigated by introducing a settling down period but at the potential cost of a serious reduction in total observations \citep{LeeMarsden2008}.
+An alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB where many species are observed simultaneously
 

### 2017-02-21 ReviewerComments.tex
Ongoing revisions.

@@
 Sorry if I'm just missing the obvious.
+\begin{quote}
+We have added a statement clarifying that there is no mathematical difference.
+\adam{
+In response to (a), while we fit our models with common distributions for $f_T(t)$, we could alternatively begin with a detection hazard rate $h(t)$.
+Given this context, the detection rate feels more like the focus of inference than the expected time to detection.
+}
+\end{quote}
 \end{enumerate}
@@
 I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
+}
+We have added a mention of the trade-offs of a settling down period to reduce observer effects on behavior.
+\end{quote}
 
-To the reviewer's point, we can emphasize the value of a settling-in period toward justifying the use constant-rate assumptions in analysis.}
-\end{quote}
 \item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
@@
 The same approach to construct paragraphs through the text should be taken.
+\begin{quote}
+We have reviewed the entire manuscript with these points in mind.
+We have split the cited paragraph and others.
+\adam{The only times I feel I have cited many articles is when I've said ``Many authors have done this.'.}
+\end{quote}
 

### 2017-02-21 ReviewerComments.tex
Add comments

@@
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-\newcommand{\jarad}[1]{{\color{Orange} JARAD: \#1}}
+\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
 
@@
 Their models are for removal-distance data, whereas ours is for removal-only data.
+\jarad{They also don't look at other distributions for the removal portion.}
 \end{quote}
+
 \item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter.
@@
 \begin{quote}
-Done.
+Done. \jarad{Be more explicit, e.g. "We've added the following to Section X.x" ...}
 \end{quote}
@@
 \begin{quote}
-We have added an example where detection rates vary across individuals, causing the marginal detection rate to decline over time.
+We have added an example \jarad{where?} where detection rates vary across individuals, causing the marginal detection rate to decline over time.
 \end{quote}
@@
 We do not believe this distinction requires a separate explanation in the Introduction.
+\jarad{Can we reword to remove peaked and nonpeaked here?}
 \end{quote}
@@
 \item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
+\jarad{Computation time.}
 \begin{quote}
@@
 \end{quote}
+
 \item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.
@@
 \adam{Waiting until all new simulations are run.}
+\jarad{I think we can explain why here, but because the other reviewer also 
+mentioned ``poor implementation'' of coverage, we may need to address why in the manuscript.
+I think computation time is a reason.}
 \end{quote}
@@
 This two- to three-fold scale of uncertainty is commensurate with other studies cited in this manuscript \citep{Diefenbach2007, Reidy2011, Solymos2013, Amundson2014}.
+\jarad{This is the same order of magnitude and therefore is useful in my opinion.}
 \end{quote}
@@
 I think the low-availability stipulation is redundant.  If availability is really high, then a removal analysis is not necessary at all.}
+\jarad{But how do you know availability is high?}
 \end{quote}
@@
 \adam{What I do not advocate is defaulting to the constant-rate assumption: by the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.}
+\jarad{Agreed.}
 \end{quote}
@@
 \adam{This last recommendation seems feasible, though I think our final paragraph already addesses some of these issues.}
+\jarad{If we already address it, then you should just quote from the manuscript.}
 \end{quote}
@@
 \begin{quote}
-\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wide range or simulated $\pdet$.}
+\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wider range of simulated $\pdet$.}
 \end{quote}
@@
 \adam{Jarad, I am unclear on the purpose of the Olkin-citing paragraph.  I don't feel like the above arguments really merit inclusion in the text.}
+\jarad{In my opinion, if $\pdet$ is low, then information about $\pdet$ and abundance are 
+obtained through the priors and not through the data. These citations as well as the Olkin paragraph
+indicate that statistical methods cannot help us to recover when $\pdet$ is low unless we 
+are willing to make strong assumptions.}
 \end{quote}
@@
 tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
+
+\jarad{Although we should look at more simulations. The purpose of the current values
+was to give some credence to the Ovenbird analysis.}
 
@@
 }
+\jarad{I think 3.}
 \end{quote}
@@
 }
+\jarad{I'm not inclined to do this.}
 \end{quote}
@@
 }
+\jarad{But it appears the reviewer wants it up front. 
+I need to go back and reread to determine if this seems feasible. 
+I thought we did start without mixtures and without covariates.}
 \end{quote}
@@
 We have crafted a beautiful and very telling graph (Figure 1).
+\jarad{Stay away from sass.}
 \end{quote}
@@
 
+\jarad{Ran out of steam?}
+
 Thank you for the opportunity to comment on your paper, I hope you find my views useful.  I think yours will in the end, be a very important contribution.
+
+\jarad{I missed this previously.}
 

### 2017-02-24 Draft.tex
Updates.

@@
 
-\subsection{Exact time to detection}
+\subsection{Exact time to detection}\label{sec:exact_time}
 
@@
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$. 
-We employ a log link for $\varphi$.
-This makes our model equivalent to a generalized linear model with a log link on the mean detection time.
+We employ a log link for $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time.
 
@@
 In theory, longer surveys can improve estimation of the unsampled tail probability, but the longer the survey lasts, the greater the risk that individuals enter/depart the study area or are double-counted, which violates the removal sampling assumption of a closed population \citep{LeeMarsden2008, Reidy2011}.
-Observer effects on singing rates can be mitigated by introducing a settling down period but at the potential cost of a serious reduction in total observations \citep{LeeMarsden2008}.
+Observer effects on availability rates can be mitigated by introducing a settling down period but at the potential cost of a serious reduction in total observations \citep{LeeMarsden2008}.
 An alternative to removal sampling is to record complete detection records (all detections for every individual) instead of just the first \citep{Alldredge2007}; however, this may not be feasible in studies like MNFB where many species are observed simultaneously

### 2017-02-24 ReviewerComments.tex
Updates.

@@
 We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
-Their models are for removal-distance data, whereas ours is for removal-only data.
-\jarad{They also don't look at other distributions for the removal portion.}
+Interestingly, they have approached removal modeling both from distributional and from hidden Markov modeling directions.
+Our manuscript explores distinct topics in that our model is for removal-only data, whereas their models are for removal-distance data.
+Also, we explicitly investigate non-constant detection distributions, which has not been their focus.
 \end{quote}
@@
 \begin{quote}
-We have added a statement clarifying that there is no mathematical difference.
-\adam{
-In response to (a), while we fit our models with common distributions for $f_T(t)$, we could alternatively begin with a detection hazard rate $h(t)$.
-Given this context, the detection rate feels more like the focus of inference than the expected time to detection.
-}
+We have added the following in Section \ref{M-sec:exact_time} to clarify: ``We employ a log link for $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time."
+
+We feel the rate conceptualization resonates more with the framing of the problem, though perhaps the distinction is mainly aesthetic.
+Our view might be more explicit if we were to define the time-to-detection distribution from its hazard function rather than from a traditional distributional family. 
+The latter is more convenient and builds upon parametric survival modeling, but the former can also be done within our model framework.
 \end{quote}
@@
 \begin{quote}
-Done. \jarad{Be more explicit, e.g. "We've added the following to Section X.x" ...}
+We have revised the statement to read: ``The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption can often lead to biased estimates.''
 \end{quote}
@@
 \begin{quote}
-Done.
+We have prefaced the results with the clause: ``Based on simulation analyses, ..."
 \end{quote}
@@
 \begin{quote}
-We have explicitly defined time-to-detection data as being time-to-event data.\\
+We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
 \adam{I am unclear what the reviewer means by `providing citations for the cdf and pdfs outlined here for reference'.}
@@
 \begin{quote}
-We have added an example \jarad{where?} where detection rates vary across individuals, causing the marginal detection rate to decline over time.
+We have added the following sentences to \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
+All three could be reasonable TTDD choices if, for instance, detection rates vary among individuals.
+This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population."
+
+We appreciate the importance of providing biological justification for statistical assumptions.
+It is always a good foundation.
+That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
+The dynamics of the system may be more complicated than we can appreciate.
+For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
 \end{quote}
@@
 \begin{quote}
-We have made the meaning of `peaked' and `nonpeaked' more explicit in Section \ref{M-sec:sim}.
+We have given the definitions of peaked/nonpeaked their own paragraph in the Simulation Studies methodology (Section \ref{M-sec:sim}):
+``In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
+Detection rates $h(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
+More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked."
 We do not believe this distinction requires a separate explanation in the Introduction.
@@
 We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
-However, it can be informative.
+However, it is informative.
 TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
+
+On the practical side, while simulation studies are hampered by long computation times, the typical user of this model will only need to apply it to a comparatively smaller number of datasets.
 \end{quote}
@@
 \end{quote}
+
 \item How might study design impact bias and precision?
@@
 \begin{quote}
-\adam{We have not investigated either of these issues.
-During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
-However, I would not dream of extending those findings to other distributions.
-We obviously would prefer longer durations, but those risk the violation of closure assumptions.
-\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
-I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
-}
-We have added a mention of the trade-offs of a settling down period to reduce observer effects on behavior.
+We appreciate that these are important decision in the design of surveys, but we have not investigated either of these issues with respect to bias and precision.
+Not having done so, we prefer not to speculate.
+%\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
+We have separated out a paragraph in the Discussion posing some design-related questions.
+\adam{Insert a statement about the NRRI one-minute intervals relative to two- or three-interval designs?}
 \end{quote}
@@
 \adam{Waiting until all new simulations are run.}
-\jarad{I think we can explain why here, but because the other reviewer also 
-mentioned ``poor implementation'' of coverage, we may need to address why in the manuscript.
-I think computation time is a reason.}
 \end{quote}
@@
 Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
-\begin{quote} Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections)\end{quote}
+\begin{quote}
+Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections).
+We have placed this information in Section \ref{M-sec:data}: ``The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s.
+A total of 947 Ovenbirds were counted with a maximum of eight in any single survey."
+\end{quote}
+
 For example, The Gamma distribution, touted as the most accurate based on the simulation study, suggests 95\% CIs of p ranges from ~ 0.3 to 1?
@@
 \begin{quote} 
-We think our figure may have been unclear.
-The cited credible intervals are for \textit{uncounted} birds not total abundance.
+We think our figure was unclear.
+The cited credible intervals were for \textit{uncounted} birds not total abundance.
 We have now changed the figure to $\log_{10}$(Abundance) to avoid confusion.
@@
 This two- to three-fold scale of uncertainty is commensurate with other studies cited in this manuscript \citep{Diefenbach2007, Reidy2011, Solymos2013, Amundson2014}.
-\jarad{This is the same order of magnitude and therefore is useful in my opinion.}
 \end{quote}
@@
 \begin{quote}
-\adam{What I do not advocate is defaulting to the constant-rate assumption: by the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.}
-\jarad{Agreed.}
+By the precautionary principle, we believe the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.
+This is a central point in our manuscript: the constant-rate assumption should not be the default.
 \end{quote}
@@
 \begin{quote}
-We agree.  Our current research addresses this issue in more depth.
+We agree.
+This manuscript applies to removal-only analyses.
+Still, the concept should carry over to removal-distance analyses.
+The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
 \end{quote}

### 2017-02-24 ReviewerComments.tex
Last Adam updates to reviewer's comments and draft until: (i) simulations done, or (ii) Philip and Jarad review.

@@
 First, the number of simulations (16 in Table 1) is exceedingly low to be able to reliably diagnose the patterns in coverage or the statistical explanation of these patterns.
+\begin{quote}
+We have not been clear enough in stating our goals for the simulations.
+We have added the following line in Section \ref{M-sec:sim}: ``Because our main goal with these studies was to demonstrate poor performance under model misspecification, and because the computation time for sampling from some models was large, we simulated only a small number of replicates."
+\adam{Can we run more simulations?  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wider range of simulated $\pdet$.}
+\end{quote}
+
 Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det).
@@
 \end{quote}
+
 (Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
 \begin{quote}
-It is not a hard boundary.  It results from the relative sample mean and sample variance of large samples.
-However, similar thresholds continue to surface again and again in various abundance-estimating contexts: 
+It is not a hard boundary.  
+But the boundary is a property of estimating $N$ in Binomial unknown-$N$, unknown-$p$ problems.
+Olkin's calculation results from the relative sample mean and sample variance of large samples.  
+Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
 \begin{itemize}
-\item \citet{Veech2016}: suggest  $p>0.5$ in the context of individual-level detection effects (no time-to-detection data). 
+\item \citet{Veech2016}: advise caution when $p<0.5$ in the context of individual-level detection effects (no time-to-detection data). 
 \item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
@@
 \end{itemize}
-\adam{Jarad, I am unclear on the purpose of the Olkin-citing paragraph.  I don't feel like the above arguments really merit inclusion in the text.}
-\jarad{In my opinion, if $\pdet$ is low, then information about $\pdet$ and abundance are 
-obtained through the priors and not through the data. These citations as well as the Olkin paragraph
-indicate that statistical methods cannot help us to recover when $\pdet$ is low unless we 
-are willing to make strong assumptions.}
+\adam{Since I've cited the above articles, is there a reason to work them into the manuscript?  I don't feel they add to our story.}
 \end{quote}
@@
 \begin{quote}
-We have inserted clarifications: (i) when we first reference the N-Mixture framework in the Introduction, and (ii) when we first use the term `mixture model' in Section \ref{M-s:subgroups}.
+In the Introduction, we now clearly describe a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
+Later in the same paragraph, we state ``We apply the term `mixture model' to any model with a mixture TTDD."
+This is contrasted in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model \dotso, which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
 \end{quote}
@@
 This potentially requires a lot of rewiring.
-Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler version of the model without mixtures and without covariates.
+Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler summary of the model without mixtures and without covariates.
+I do think we provide the model pieces one-by-one, but we could be more heavy-handed (in a good way).
 }
@@
 We have reviewed the entire manuscript with these points in mind.
-We have split the cited paragraph and others.
-\adam{The only times I feel I have cited many articles is when I've said ``Many authors have done this.'.}
+We have split the cited paragraph and a few others.
+\adam{The only times I feel I have cited many articles en masse is when I've said ``Many authors have done this.'}
 \end{quote}
@@
 \begin{quote}
-We have crafted a beautiful and very telling graph (Figure 1).
-\jarad{Stay away from sass.}
+We have added this as Figure \ref{M-fig:schematic}.
+\adam{I have not citedthe figure in the text.  I am not convinced I need to.}
 \end{quote}
@@
 
-\jarad{Ran out of steam?}
+\begin{quote}
+\adam{
+The main application I've encountered before is \citet{Borchers2013}.
+There are technical differences between that approach and our parameteric survival analysis approach, mainly due to the continuous and interval-censored handling of time.
+For instance, I suspect that for analysis, as with \citet{Farnsworth2002} and its derivatives, the hidden Markov model would require an artificial division of the observation period into equal-intervals intervals.
+One cool feature of that approach is that it directly handles bout-singing through an availability state (which may be stochastically `available' or `not available') in conjunction with a detection model.
+I am not sure that model would serve well for a problem like ours where we are looking for systematic non-constant detection --- i.e., the Markov model would require time-dependent transition matrices.
+Additionally, at least in the case of \citet{Borchers2013}, the model utilizes separate observations of animal behaviors to estimate transitions among hidden states (e.g., moving/singing/etc.).
+In their case, they analyzed radio-tagged whales to estimate transitions for diving states.
 
-Thank you for the opportunity to comment on your paper, I hope you find my views useful.  I think yours will in the end, be a very important contribution.
-
-\jarad{I missed this previously.}
+Those are just some preliminary thoughts.
+I think the reviewer's comment is thought-provoking but not applicable to the manuscript we have crafted.}
+\end{quote}
 

### 2017-02-26 Sims/SimZero/CreateSimParameters.R
Add code to create datasets from selected parameter values.

@@ -1,0 +1,129 @@
+setwd(".../Sims/SimZero") # Run from the TimeToDetect directory
+
+########## Helper equations for each distribution (homogeneous surveys)
+### Exponential Distribution
+# First  function calculates intcpt_d to obtain a target pdet (no mixture)
+# Second function calculates intcpt_d to obtain a target pdet given gamma (mixture model)
+e.id.pdet    <- function(pdet) log(-log(1-pdet)/10)
+emix.id.pdet <- function(pdet, gamma=0.65) e.id.pdet((pdet-(1-gamma))/gamma)
+
+### Gamma Distribution
+# First  function calculates intcpt_d & alpha to obtain a target pdet and mode for f_T(t) (peaked, no mixture)
+# Second function calculates intcpt_d & alpha to obtain a target pdet and Pr(t<2)/pdet    (nonpeaked, no mixture)
+# Third  function is like the first  but conditioned on gamma 
+#         (technically solves the first function for intcpt_d & alpha to obtain target p(det|bird is hard-to-detect)
+# Fourth function is like the second but conditioned on gamma
+g.id.pdet.p <- function(pdet, mode){
+  whelp <- function(alpha, pdet, mode) abs(pdet - pgamma(10, alpha, (alpha-1)/mode))
+  alpha <- optimize(whelp, interval=c(0,10), pdet, mode)$minimum
+  int.d <- log((alpha-1)/mode)
+  return(c(int.d=int.d, alpha=alpha))
+}
+g.id.pdet.n <- function(pdet, p2){
+  # Helper fxn: for optimizing both F_T(10) = pdet AND F_T(2) = Pr(t<2)
+  whelp2 <- function(parms, pdet, p2) abs(pgamma(10,parms[1],exp(parms[2]))-pdet) +
+    abs(pgamma(2,parms[1],exp(parms[2]))/pgamma(10,parms[1],exp(parms[2]))-p2)
+  return(optim(c(2,-2), whelp2, gr=NULL, pdet, p2)$par)
+}
+gmix.id.pdet.p <- function(pdet, mode, gamma=0.65) g.id.pdet.p((pdet-(1-gamma))/gamma, mode)
[…]

### 2017-02-26 Sims/SimZero/SimulateData_zero.R

@@ -1,0 +1,87 @@
+# This code requires an input table 'Simpars' with parameters for simulations.
+# 'Simpars' is generated by the file 'CreateSimParameters.R'
+
+# Output from this code:
+# (1) In each of ZZ = 1, ..., NReps folders, it creates a file 'Sim_ZZ/SimData.Rdata'
+#     That file is a list of 1, ..., nrow(simtable) simulated datasets
+# (2) A table of the simulation parameters called 'ParValues' saved in a file called 'Sim_ZZ/TrueParams.Rdata'
+
+library(plyr)
+load("Simpars.Rdata")             # These are the hand-selected parameters for simulation
+load("../../OVEN/Data9.Rdata")    # Ovenbird data -- named 'dat' -- provides survey info and actual counts
+N <- dat9 <- uncounted <- vector('list',nrow(Simpars))
+tau <- c(0, 2:10)                 # Observation period intervals
+TTD  <- substr(Simpars$model,1,1) # TTDD distribution
+Mix  <- as.logical(substr(Simpars$model,2,2)) # Whether mixture or non-mixture
+
+##### Loop to simulated datasets from parsmeters
+# for generating multiple replicates of simulations
+
+for(Rep in 1:16){
+  for(j in 1:nrow(Simpars)){
+    ##### Start Generating Data
+    # Simulated site abundances
+    N[[j]] <- rpois(dat$n_sites, exp(Simpars$intcpt_a[j]))
+    # Storage object for site(row) x interval(col) detection probability
+    p.vector <- numeric(length(tau)-1)
+  
+    ### Calculate interval-specific detection probabilities
+    if(TTD[j]=="E"){
+      # intcpt_d = log(beta) from desired exponential
+      for(j2 in 1:9) p.vector[j2] <- pexp(tau[j2+1],exp(Simpars$intcpt_d[j])) -
+                                     pexp(tau[j2],exp(Simpars$intcpt_d[j]))
+    } 
+    if(TTD[j]=="G"){
+      alpha <- Simpars$alpha[j]
+      # intcpt_d = log(beta) from desired gamma
+      for(j2 in 1:9) p.vector[j2] <- pgamma(tau[j2+1],alpha,exp(Simpars$intcpt_d[j])) - 
[…]

### 2017-02-26 Sims/SimZero/Run_SimZero_slurm.R
Stan model code. Code to run models on slurm. S0_simtable to guide simulations.

@@ -1,0 +1,55 @@
+########## Code to fit SimZero models on slurm
+set.seed(NULL)
+bundle = 1    # How many models will be fit by a single call to this script
+
+# Code to replicate S0_sumtable according to the number of replicates
+NumReps  <- 16
+simtable <- read.csv("S0_simtable.csv")       # Load table of datasets
+simtable <- do.call('rbind', rep(list(simtable), times=NumReps))
+simtable$Rep <- rep(1:NumReps, each=nrow(simtable)/NumReps)
+
+# Collect array number from slurm and add one
+ss = Sys.getenv("SLURM_ARRAY_TASK_ID") 
+j  = as.numeric(ss) * bundle + 1
+
+## List of models to run, in case I only want to run a subset
+# analysis_dupes <- readRDS("Data/A_Dupes.rds")
+# indices <- analysis_dupes[j:(j+bundle-1)]
+
+for(indx in j:(j+bundle-1)){
+  Rep <- simtable$Rep[indx]
+  load(paste0("Sim_", Rep, "/SimData.Rdata"))  # Load all datasets in Rep.  Object name is 'dat9'
+  dataset <- dat9[[simtable$datacode[indx]]]
+  load(paste0("Sim_",Rep,"/TrueParams.Rdata")) # Load True Values associated with SimData.  Object name is 'ParValues'
+  
+  # Provide initial values.  Since data and models are often not congruent, but the 
+  # simulations use the same target values, the idea here is to use simulation parameters
+  # from the most equivalent model-related family.
+  # For example, if the data are LFT, and the model is a GammaMix, then I want the 
+  # parameters from the GTT model
+  
+  # Construct a simdat-like code for the model being fit so that I can use values from the equivalent dataset
+  modelcode <- paste0(substr(simtable$modelname,1,1), substr(simtable$modelmix,1,1), substr(simtable$simdat,3,3))
+  modelcode[modelcode=="EFT"] <- "EFF"
+  modelcode[modelcode=="ETT"] <- "ETF"
+  initlist <- ParValues[[which(modelcode[indx]==levels(simtable$simdat))]] # Initial values
+  
+  # MCMC variables to store
+  record.list <- c("intcpt_a", "intcpt_d", "uncounted", "p_global", "dev1")
+  if(simtable$modelmix[indx]) record.list <- c(record.list, "gamma")
[…]

### 2017-02-26 Sims/SimZero/SimulateData_zero.R
Tweak.

@@ -6,6 +6,8 @@
 #     That file is a list of 1, ..., nrow(simtable) simulated datasets
 # (2) A table of the simulation parameters called 'ParValues' saved in a file called 'Sim_ZZ/TrueParams.Rdata'
 
+setwd("Sims/SimZero")
+
 library(plyr)
 load("Simpars.Rdata")             # These are the hand-selected parameters for simulation
 load("../../OVEN/Data9.Rdata")    # Ovenbird data -- named 'dat' -- provides survey info and actual counts

### 2017-02-26 Sims/SimZero/CreateSimParameters.R
Untweak.

@@ -1,5 +1,3 @@
-setwd(".../Sims/SimZero") # Run from the TimeToDetect directory
-
 ########## Helper equations for each distribution (homogeneous surveys)
 ### Exponential Distribution
 # First  function calculates intcpt_d to obtain a target pdet (no mixture)

### 2017-02-26 Sims/SimZero/SimulateData_zero.R

@@ -6,8 +6,6 @@
 #     That file is a list of 1, ..., nrow(simtable) simulated datasets
 # (2) A table of the simulation parameters called 'ParValues' saved in a file called 'Sim_ZZ/TrueParams.Rdata'
 
-setwd("Sims/SimZero")
-
 library(plyr)
 load("Simpars.Rdata")             # These are the hand-selected parameters for simulation
 load("../../OVEN/Data9.Rdata")    # Ovenbird data -- named 'dat' -- provides survey info and actual counts

### 2017-02-26 Sims/SimZero/SimulateData_zero.R
Tweak.

@@ -6,6 +6,8 @@
 #     That file is a list of 1, ..., nrow(simtable) simulated datasets
 # (2) A table of the simulation parameters called 'ParValues' saved in a file called 'Sim_ZZ/TrueParams.Rdata'
 
+set.seed(NULL)
+
 library(plyr)
 load("Simpars.Rdata")             # These are the hand-selected parameters for simulation
 load("../../OVEN/Data9.Rdata")    # Ovenbird data -- named 'dat' -- provides survey info and actual counts

### 2017-02-26 OVEN/Compile_addendum.R
Forgotten file.

@@ -1,0 +1,104 @@
+### Dual-purpose function (why did I write it this way?)
+# 'simvals' is a data.frame with appropriately named columns for known parameter values
+# 'adz' is a graphical parameter indicating whether to overlay the density plot... currently disabled
+# Plots interval-censored (discrete) distributions of f(t|det)
+# AND
+# Modifies the data.frame ?!?
+
+# Trick: if I only want mixtures, change col=j to col=j/2
+plotmedians <- function(simvals, adz=T){
+  taus <- c(0,2:10)   # Interval cutoffs
+  # For plotting intervals... duplicates all interior cutoffs:
+  ends <- c(taus[1], rep(taus[-c(1,length(taus))],each=2), taus[length(taus)])
+  ys <- numeric(length(ends))
+  modcode <- simvals$modcode
+  simvals$pdet <- NA
+  simvals$indx <- simvals$modcode%%2
+  par(mfrow=c(4,4))
+  # Loop row-by-row through the table
+  # Calculate pdet from the appropriate distribution (identified by 'modcode' variable)
+  for(j in 1:nrow(simvals)){
+    plot(1, 1, type='n', xlim=c(0,10), ylim=c(0,0.5), ylab=paste0("f(t|det): ",simvals$model[j]), xlab="Time")
+    if(modcode[j] == 1) {
+      simvals$pdet[j] <- pexp(10,exp(simvals$intcpt_d[j]))
+      ys <- rep((pexp(taus[-1],exp(simvals$intcpt_d[j])) - pexp(taus[-length(taus)],exp(simvals$intcpt_d[j])))/simvals$pdet[j],each=2)
+      ys[1:2] <- 0.5*ys[1:2]
+      lines(ends, ys, col=if(simvals$indx[j]==0){"orange"}else{2}, lty=1, ylab="Expo")
+#       if(simvals$n_ints[j]==9) curve(dexp(x,exp(simvals$intcpt_d[j]))/simvals$pdet[j],
+#                                        add=adz, xlim=c(0,10), col=if(simvals$indx[j]==0){"orange"}else{2}, lty=1, ylab="Expo")
+    }
+    if(modcode[j] == 2){
+      simvals$pdet[j] <- 1-simvals$gamma[j]*(1-pexp(10,exp(simvals$intcpt_d[j])))
[…]

### 2017-02-26 Sims/SimZero/CreateSimParameters.R
Tweak.

@@ -124,4 +124,4 @@ for(j in 1:nrow(Simpars)){
 Simpars$intcpt_a <- log((1/pdet)*(947/381))
 
 
-save(Simpars, file="../Simpars.Rdata")
+save(Simpars, file="Simpars.Rdata")

### 2017-02-26 Sims/SimZero/GetTimes.R
Code to extract run times from model fits.

@@ -1,0 +1,27 @@
+##### Gets Run Times
+library(rstan)
+simtable <- read.csv("S0_simtable.csv")
+nReps    <- 1
+runtimes <- data.frame(matrix(NA, nrow=nrow(simtable), ncol=nReps))
+for(Rep in 1:nReps){
+  for(m in 1:nrow(simtable)){
+    loadfile <- paste0("Sim_", Rep, "/", simtable[m,"fname"], ".Rdata")
+    if(file.exists(loadfile)){
+      load(loadfile)
+      runtimes[m, Rep] <- sum(get_elapsed_time(eval(parse(text=as.character(simtable[m,"modeloutput"])))))
+    } else {
+      runtimes[m,Rep] <- NA}
+  }
+}
+saveRDS(runtimes, file="RunTimes.rds")
+
+# ##### Extracts Run Times
+# library(plyr)
+# library(reshape2)
+# simtable <- read.csv("S0_simtable.csv")
+# runtimes <- readRDS("RunTimes.rds")
+# runtimes <- cbind(simtable[,1:2], runtimes)
+# r.melt <- melt(runtimes, id.vars=1:2, value.name="runtime")
+# smurf <- ddply(r.melt, .(modelname), summarize, mu.t=mean(runtime))
+# smurf$minutes <- smurf$mu.t/60
+# smurf

### 2017-03-02 Sims/SimZero/StanPlot_zero.R
Change DIC calculations.
Add post-processing code (StanPlot_zero.R)

@@ -1,0 +1,185 @@
+########## This misnamed code actually does all the immediate post-model processing of data and
+# returns a list object called 'outtable' for each model fit.  Outtable is saved in the '<modelname>_sum.Rdata' file.
+
+
+########## Load Stan output
+library(mcmcse)
+library(coda)
+library(rstan)
+library(hexbin)
+
+simtable <- read.csv("S0_simtable.csv")   # Simulation information is stored here
+
+# Simplify calls to simtable
+fname       <- simtable$fname
+modeloutput <- as.character(simtable$modeloutput)
+parmcode    <- simtable$parmcode
+n_ints      <- simtable$n_ints
+datacode    <- simtable$datacode
+modelmix    <- simtable$modelmix
+datamodelmatch <- simtable$datamodelmatch
+params <- list(c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","lp__"),                     # Exponential
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","lp__"),             # ExpoMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","alpha","lp__"),             # Gamma
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","alpha","lp__"),     # GammaMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","sigma_det","lp__"),         # LogN
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","sigma_det","lp__"), # LogNMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","shape_k","lp__"),           # Weibull
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","shape_k","lp__"))   # WeibullMix
+# Because it's the intercept-only model
+for(l in 1:length(params)) params[[l]] <- params[[l]][-which(params[[l]] %in% c("ba", "bd", "sigma_a", "sigma_d", "lp__"))]
+
+Reps <- 1
+
+for(Rep in 1:Reps){
+  ESS <- quanttable <- NULL; Geweke <- vector('list', nrow(simtable))  # Storage objects
+  setwd(paste0("Sim_",Rep))
+  load("SimData.Rdata")                            # Load data (dat9) for all simulations in Rep.
[…]

### 2017-03-02 Sims/SimZero/GetTimes.R
Tweaks.

@@ -3,6 +3,7 @@ library(rstan)
 simtable <- read.csv("S0_simtable.csv")
 nReps    <- 1
 runtimes <- data.frame(matrix(NA, nrow=nrow(simtable), ncol=nReps))
+names(runtimes) <- paste0("RunTime_", 1:nReps)
 for(Rep in 1:nReps){
   for(m in 1:nrow(simtable)){
     loadfile <- paste0("Sim_", Rep, "/", simtable[m,"fname"], ".Rdata")

### 2017-03-02 Sims/SimZero/StanPlot_zero.R

@@ -174,7 +174,7 @@ for(Rep in 1:Reps){
     save(outtable, file=paste(fname[i],"_sum.Rdata",sep=""))
     
     
-    rm(list=ls()[-match(c("i","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","quanttable","Geweke","ESS","datacode","modelmix","datamodelmatch"),ls())])
+    rm(list=ls()[-match(c("i","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","quanttable","Geweke","ESS","datacode","modelmix","datamodelmatch","uncounted"),ls())])
     gc(verbose=T)
   }
   

### 2017-03-02 Sims/SimZero/StanPlot_zero.R
Tweak

@@ -106,9 +106,11 @@ for(Rep in 1:Reps){
     Geweke[[i]] <- geweke.diag(gew.ext2)[[1]]
     
     ##### Effective Sample Sizes
-    ESS <- data.frame(parameter = rownames(outtable[[1]]),
+    ESS <- rbind(ESS, 
+                 data.frame(parameter = rownames(outtable[[1]]),
                       n_eff = ess(gew.ext2),  # Using the 'mcmcse' package
                       fname = fname[i])
+                 )
     
     ########## DIC -- there used to be three versions of this.  But I'm only keeping the first.
     # dev1: p(n|betas, gamma, shape)

### 2017-03-02 Sims/SimZero/SimulateData_zero.R
Upscaling to 100 replicates.

@@ -19,7 +19,7 @@ Mix  <- as.logical(substr(Simpars$model,2,2)) # Whether mixture or non-mixture
 ##### Loop to simulated datasets from parsmeters
 # for generating multiple replicates of simulations
 
-for(Rep in 1:16){
+for(Rep in 1:100){
   for(j in 1:nrow(Simpars)){
     ##### Start Generating Data
     # Simulated site abundances

### 2017-03-02 Sims/SimZero/StanPlot_zero.R

@@ -32,7 +32,7 @@ for(l in 1:length(params)) params[[l]] <- params[[l]][-which(params[[l]] %in% c(
 Reps <- 1
 
 for(Rep in 1:Reps){
-  ESS <- quanttable <- NULL; Geweke <- vector('list', nrow(simtable))  # Storage objects
+  ESS <- NULL; Geweke <- vector('list', nrow(simtable))  # Storage objects
   setwd(paste0("Sim_",Rep))
   load("SimData.Rdata")                            # Load data (dat9) for all simulations in Rep.
                                                    # Also includes the total number of birds uncounted (uncounted)
@@ -109,7 +109,8 @@ for(Rep in 1:Reps){
     ESS <- rbind(ESS, 
                  data.frame(parameter = rownames(outtable[[1]]),
                       n_eff = ess(gew.ext2),  # Using the 'mcmcse' package
-                      fname = fname[i])
+                      fname = fname[i],
+                      Rep   = paste0("Rep_", Rep))
                  )
     
     ########## DIC -- there used to be three versions of this.  But I'm only keeping the first.

### 2017-03-02 Sims/SimZero/Run_SimZero_slurm.R
Upscale to 100 reps. Test for adequate sample sizes as we fit models.  (Increase iterations if failure).

@@ -1,9 +1,9 @@
 ########## Code to fit SimZero models on slurm
 set.seed(NULL)
-bundle = 1    # How many models will be fit by a single call to this script
+bundle = 5    # How many models will be fit by a single call to this script
 
 # Code to replicate S0_sumtable according to the number of replicates
-NumReps  <- 16
+NumReps  <- 100
 simtable <- read.csv("S0_simtable.csv")       # Load table of datasets
 simtable <- do.call('rbind', rep(list(simtable), times=NumReps))
 simtable$Rep <- rep(1:NumReps, each=nrow(simtable)/NumReps)
@@ -42,11 +42,26 @@ for(indx in j:(j+bundle-1)){
   if(simtable$parmcode[indx] %in% 7:8) record.list <- c(record.list, "shape_k")
   
   library(rstan)
+  library(mcmcse)
   # Compile/Load Stan model
   m = stan_model(paste0(simtable$modelname[indx],".stan"), auto_write=rstan_options("auto_write" = TRUE))
   
   # For most runs, it was 60K iter.  For a few troublesome models, I boosted it.
-  assign(as.character(simtable$modeloutput[indx]), sampling(m, data=dataset, chains=1, iter=60000, thin=10, pars=record.list, init=list(initlist)))
+  insuff <- TRUE
+  iter   <- 60000
+  fail   <- 0
+  while(insuff){
+    assign(as.character(simtable$modeloutput[indx]), 
+           sampling(m, data=dataset, chains=1, iter=iter*(2^fail), thin=10, pars=record.list, 
+                    init=list(initlist)))
+    postdraws <- as.data.frame(eval(parse(text=as.character(simtable$modeloutput[indx]))))
+    if(min(ess(postdraws)) > 10){
+      insuff=FALSE
+    } else {
+      fail <- fail + 1
+    }
+  }
+  
   save(list=as.character(simtable$modeloutput[indx]), file=paste0("Sim_",Rep,"/",simtable$fname[indx],".Rdata"))
   
   rm(list = ls()[-which(ls() %in% c("indx", "j", "bundle", "simtable"))])

### 2017-03-04 Sims/SimZero/Run_SimZero_slurm.R
Tweak.

@@ -55,7 +55,7 @@ for(indx in j:(j+bundle-1)){
            sampling(m, data=dataset, chains=1, iter=iter*(2^fail), thin=10, pars=record.list, 
                     init=list(initlist)))
     postdraws <- as.data.frame(eval(parse(text=as.character(simtable$modeloutput[indx]))))
-    if(min(ess(postdraws)) > 10){
+    if(min(ess(postdraws)) > 1000){
       insuff=FALSE
     } else {
       fail <- fail + 1

### 2017-03-04 Sims/SimZero/GetTimes.R
Change number of reps.

@@ -1,7 +1,7 @@
 ##### Gets Run Times
 library(rstan)
 simtable <- read.csv("S0_simtable.csv")
-nReps    <- 1
+nReps    <- 100
 runtimes <- data.frame(matrix(NA, nrow=nrow(simtable), ncol=nReps))
 names(runtimes) <- paste0("RunTime_", 1:nReps)
 for(Rep in 1:nReps){

### 2017-03-04 Sims/SimZero/GetTimes.R
Change number of reps.  Collect minimum estimated sample size across parameters.

@@ -1,8 +1,10 @@
 ##### Gets Run Times
 library(rstan)
+library(mcmcse)
 simtable <- read.csv("S0_simtable.csv")
-nReps    <- 1
-runtimes <- data.frame(matrix(NA, nrow=nrow(simtable), ncol=nReps))
+nReps    <- 100
+ess <- runtimes <- data.frame(matrix(NA, nrow=nrow(simtable), ncol=nReps))
+names(ess)      <- paste0("ESS_", 1:nReps)
 names(runtimes) <- paste0("RunTime_", 1:nReps)
 for(Rep in 1:nReps){
   for(m in 1:nrow(simtable)){
@@ -10,8 +12,11 @@ for(Rep in 1:nReps){
     if(file.exists(loadfile)){
       load(loadfile)
       runtimes[m, Rep] <- sum(get_elapsed_time(eval(parse(text=as.character(simtable[m,"modeloutput"])))))
+      postdraws        <- as.data.frame(eval(parse(text=as.character(simtable[m,"modeloutput"]))))
+      ess[m, Rep]      <- min(ess(postdraws))
     } else {
-      runtimes[m,Rep] <- NA}
+      runtimes[m,Rep]  <- NA
+    }
   }
 }
 saveRDS(runtimes, file="RunTimes.rds")

### 2017-03-04 Sims/SimZero/GetTimes.R
Forgotten line.

@@ -20,6 +20,7 @@ for(Rep in 1:nReps){
   }
 }
 saveRDS(runtimes, file="RunTimes.rds")
+saveRDS(ess, file="ESS.rds")
 
 # ##### Extracts Run Times
 # library(plyr)

### 2017-03-04 Draft.tex
Last Adam updates to reviewer's comments and draft until: (i) simulations done, or (ii) Philip and Jarad review.
Discussion of the generalized gamma distribution.

@@
 For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families. 
+Because our main goal with these studies was to demonstrate poor performance under model misspecification, and because the computation time for sampling from some models was large, we simulated only a small number of replicates.
 In all simulation studies, we focused on accuracy in estimation of $\pdet$ which then translated into estimation of abundance.
@@
 
+Just as the exponential TTDD is a special case of the two-parameter gamma and Weibull TTDDs, so all four TTDDs in our analysis are special cases the three-parameter generalized gamma distribution.
+A generalized gamma TTDD encompasses a diversity hazard functions \citep{Cox2007}, eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
+However, maximum likelihood estimation of the generalized gamma has historically suffered from computational difficulties, unsatisifactory asymptotic normality at large sample sizes, and non-unique roots to the likelihood equation \citep{CoorayAnand2008, NoufailyJones2013}.
+It may well be possible to implement our model with a generalized gamma TTDD, but especially when we consider the right-truncation of data from point-count surveys, we think model convergence would not be a trivial problem.
+
 Study design can address some issues associated with non-constant detection.

### 2017-03-04 ReviewerComments.tex
Discussion of the generalized gamma distribution.

@@
 Additionally, its focus is on flexible domains of attraction, whereas the TTDDs in our analysis are all from the Gumbel domain of attraction.
-\adam{After researching the generalized gamma, add comments in the paragraph beginning `If the estimation of abundance is the primary interest...'.}
+
+Here are the comments we have added:
+``Just as the exponential TTDD is a special case of the two-parameter gamma and Weibull TTDDs, so all four TTDDs in our analysis are special cases the three-parameter generalized gamma distribution.
+A generalized gamma TTDD encompasses a diversity hazard functions \citep{Cox2007}, eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
+However, maximum likelihood estimation of the generalized gamma has historically suffered from computational difficulties, unsatisifactory asymptotic normality at large sample sizes, and non-unique roots to the likelihood equation \citep{CoorayAnand2008, NoufailyJones2013}.
+It may well be possible to implement our model with a generalized gamma TTDD, but especially when we consider the right-truncation of data from point-count surveys, we think model convergence would not be a trivial problem."
+\adam{This is an area of active research.  I feel like my tone is not optimistic enough.  Then again, it's not worth micro-wordsmithing.}
 \end{quote}
@@
 We have added this as Figure \ref{M-fig:schematic}.
-\adam{I have not citedthe figure in the text.  I am not convinced I need to.}
+\adam{I have not cited the figure in the text.  I am not convinced I need to.}
 \end{quote}

### 2017-03-05 Sims/SimZero/Run_SimZero_slurm.R
Modify model fitting code to fit specified models.

@@ -1,6 +1,6 @@
 ########## Code to fit SimZero models on slurm
 set.seed(NULL)
-bundle = 5    # How many models will be fit by a single call to this script
+bundle   <- 1    # How many models will be fit by a single call to this script
 
 # Code to replicate S0_sumtable according to the number of replicates
 NumReps  <- 100
@@ -8,15 +8,18 @@ simtable <- read.csv("S0_simtable.csv")       # Load table of datasets
 simtable <- do.call('rbind', rep(list(simtable), times=NumReps))
 simtable$Rep <- rep(1:NumReps, each=nrow(simtable)/NumReps)
 
-# Collect array number from slurm and add one
+# Rows of simtable that I want to (re-)run
+# indices  <- 1:nrow(simtable)   # Run the full simulation, all replicates
+# Repeat models that have effective sample size less than 1000 or NA (didn't fit before)
+ess      <- readRDS("ESS.rds")
+indices  <- which(is.na(ess) | ess<1000)
+
+# Collect array number from slurm and add one.  
+# 'j' identifies which models will be fit based on array number and bundle number
 ss = Sys.getenv("SLURM_ARRAY_TASK_ID") 
 j  = as.numeric(ss) * bundle + 1
 
-## List of models to run, in case I only want to run a subset
-# analysis_dupes <- readRDS("Data/A_Dupes.rds")
-# indices <- analysis_dupes[j:(j+bundle-1)]
-
-for(indx in j:(j+bundle-1)){
+for(indx in indices[j:(j+bundle-1)]){
   Rep <- simtable$Rep[indx]
   load(paste0("Sim_", Rep, "/SimData.Rdata"))  # Load all datasets in Rep.  Object name is 'dat9'
   dataset <- dat9[[simtable$datacode[indx]]]

### 2017-03-06 Sims/SimZero/StanPlot_zero.R
Change Rep Number

@@ -29,7 +29,7 @@ params <- list(c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","lp__"),   
 # Because it's the intercept-only model
 for(l in 1:length(params)) params[[l]] <- params[[l]][-which(params[[l]] %in% c("ba", "bd", "sigma_a", "sigma_d", "lp__"))]
 
-Reps <- 1
+Reps <- 100
 
 for(Rep in 1:Reps){
   ESS <- NULL; Geweke <- vector('list', nrow(simtable))  # Storage objects

### 2017-03-06 Sims/SimZero/StanPlot_zero.R
Tweak.

@@ -177,7 +177,7 @@ for(Rep in 1:Reps){
     save(outtable, file=paste(fname[i],"_sum.Rdata",sep=""))
     
     
-    rm(list=ls()[-match(c("i","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","quanttable","Geweke","ESS","datacode","modelmix","datamodelmatch","uncounted"),ls())])
+    rm(list=ls()[-match(c("i","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","Geweke","ESS","datacode","modelmix","datamodelmatch","uncounted"),ls())])
     gc(verbose=T)
   }
   

### 2017-03-06 Sims/SimZero/StanPlot_zero.R
Tweak.

@@ -177,7 +177,7 @@ for(Rep in 1:Reps){
     save(outtable, file=paste(fname[i],"_sum.Rdata",sep=""))
     
     
-    rm(list=ls()[-match(c("i","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","Geweke","ESS","datacode","modelmix","datamodelmatch","uncounted"),ls())])
+    rm(list=ls()[-match(c("i","Rep","dat9","fname","params","modeloutput","simtable","n_ints","parmcode","Geweke","ESS","datacode","modelmix","datamodelmatch","uncounted"),ls())])
     gc(verbose=T)
   }
   

### 2017-03-06 Sims/SimZero/StanPlot_zero.R
Tweak.

@@ -73,6 +73,7 @@ for(Rep in 1:Reps){
     Pval.param[2] <- sum( counted/(counted+tot.unc) < p_global ) / NumDraws
     # Posterior Pr( E[pdet] < posterior p_global ), where E[...] calculated from simulation parameters
     Pval.param[3] <- sum( Simpars$pdet[datacode[i]] < p_global ) / NumDraws
+    names(Pval.param) <- c("Uncounted", "p_global", "p80")
     
     if(datamodelmatch[i]){
       ### Posterior p-values for all model parameters noted in 'params' up top

### 2017-03-06 Sims/SimZero/Compile_simzero.R
Code to compile summary files from all model fits in all replicates.

@@ -1,0 +1,79 @@
+########## Code to gather results from across all simulations
+library(reshape2)
+library(ggplot2)
+library(plyr)
+
+nReps <- 100
+
+##### Load pre-saved summary tables
+simtable    <- read.csv("S0_simtable.csv")
+fname       <- simtable$fname
+modeloutput <- simtable$modeloutput
+parmcode    <- simtable$parmcode
+n_ints      <- simtable$n_ints
+datacode    <- simtable$datacode
+modelmix    <- simtable$modelmix
+datamodelmatch <- simtable$datamodelmatch
+simdat <- as.character(simtable$simdat)
+params <- list(c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","lp__"),                     # Exponential
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","lp__"),             # ExpoMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","alpha","lp__"),             # Gamma
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","alpha","lp__"),     # GammaMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","sigma_det","lp__"),         # LogN
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","sigma_det","lp__"), # LogNMix
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","shape_k","lp__"),           # Weibull
+               c("intcpt_a","ba","intcpt_d","bd","sigma_a","sigma_d","gamma","shape_k","lp__"))   # WeibullMix
+# Because it's the intercept-only model
+for(l in 1:length(params)) params[[l]] <- params[[l]][-which(params[[l]] %in% c("ba", "bd", "sigma_a", "sigma_d", "lp__"))]
+
+parmests <- count.det <- posteriordist <- dic <- true.unc <- true.pglobal <- post.unc.p <- NULL
+
+for(Rep in 1:nReps){
+  setwd(paste0("Sim_",Rep))
+  
+  # For each of the 14 simulated datasets, collect the true number of uncounted individuals and 
+  # the true global_p (proportion of simulated birds that are detected)
+  load("SimData.Rdata")
[…]

### 2017-03-07 Sims/SimZero/Compile_simzero.R
Code to generate plots and tables from the p50 simulation.  (Will be very similar to that for p80 and p95).

@@ -1,4 +1,6 @@
 ########## Code to gather results from across all simulations
+# Everything gets saved to a single file: Outputs.Rdata
+
 library(reshape2)
 library(ggplot2)
 library(plyr)

### 2017-03-07 Sims/SimZero/Outputs_Code_zero.R

@@ -1,0 +1,436 @@
+library(ggplot2)
+library(reshape2)
+library(plyr)
+last <- function(x) substring(x, nchar(as.character(x)))  # Function to extract last character from a string
+# Lookup tables
+familytab <- data.frame(Letter=c("E","G","L","W"), ModelFamily=c("Exponential", "Gamma", "Lognormal", "Weibull"))
+familytab2 <- data.frame(Letter2=c("E","G","L","W"), DataFamily=c("Exponential", "Gamma", "Lognormal", "Weibull"))
+
+data.handle <- function(df){
+  df$Letter  <- substr(df$fname,1,1)   # Family of model
+  df$Letter2 <- substr(df$simdat,1,1)  # Family of data
+  df <- join(df, familytab, type="left", by="Letter")
+  df <- join(df, familytab2, type="left", by="Letter2")
+  df$familymatch <- with(df, Letter==Letter2) # Is model from the correct family (regardless of mixture)?
+  return(df)
+}
+
+truepdet <- 0.5
+
+simtable <- read.csv("S0_simtable.csv")  # Load simtable
+load("Outputs_p50.Rdata")  # Load model fits
+for(i in 1:length(outputs)) assign(names(outputs)[i], outputs[[i]]) # Convert model fits from list into objects
+
+##### There used to be a section here for posterior predictive p-values
+# See First_simulations folder for code
+
+###### Posterior p-values for global estimates of pdet
+# There's a choice here. 
+# Choice one: use the p-values associated with "p_global" and make comparisons to the data realized in simulation.
+# Choice two: use the p-values associated with (misnamed) "p80".  This seems more correct.
+p_glob <- subset(post.unc.p, parameter=="p80")
+
+# Lots of data handling
+names(p_glob)[1]     <- "pval_pglob"
+p_glob               <- data.handle(p_glob)
+p_glob$datapeaked    <- as.logical(last(p_glob$simdat))                       # Is data peaked
+p_glob$datamixlabel  <- "Nomix"                                               # Label whether data mixture
+p_glob$datamixlabel[p_glob$datamix] <- "Mix"
+p_glob$datapeaklabel <- "Nonpeaked"
+p_glob$datapeaklabel[p_glob$datapeaked] <- "Peaked"
[…]

### 2017-04-03 Draft.tex
Updates to draft and reviewer's comments.

@@
 \usepackage{natbib}
-% \usepackage{color}
+\usepackage{color}
 % \usepackage[dvipsnames]{xcolor}
@@
 
-% \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
+\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
 % \newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
@@
 \newcommand{\Fm}{F_T^{(M)}}
+\newcommand{\nobs}{n^{(obs)}}
 
@@
 Each stand had up to four sites with sufficient geographical distance between sites to reduce or eliminate overlapping territories.
-The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s. 
-A total of 947 Ovenbirds were counted with a maximum of eight in any single survey.  
+This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s.
 % The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
@@
 
-Before considering interval censoring and explanatory variables, we first present the scenario of exact time to detections with no explanatory variables. 
+\adam{I think we start by describing our joint likelihood: p(counts $| N, \pdet$) and p(times $|$ counts,TTDD) and prior.
+We explain that, in isolation, $N$ and $\pdet$ are not identifiable in the first piece, which is why we use the second piece to inform our estimates of $\pdet$.
+Then, in the body of section \ref{sec:exact_time}, we introduce a second paragraph which defines our distribution for $\nobs$ and completes the data likelihood.
+The discussion of identifiability would travel with the distribution of $\nobs$ and be cleaned up a little.
+Perhaps even split the discussion of TTDDs into another subsection.
+}
+
+Before considering interval censoring and explanatory variables, we first present the scenario of exact times to detections with no explanatory variables. 
 We then incorporate interval censoring and follow with inclusion of fixed and random effects for abundance and detection.
@@
 Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
-All three could be reasonable TTDD choices if, for instance, detection rates vary among individuals.
+All three may be reasonable TTDD choices if, for instance, detection rates vary among individuals.
 This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population.
@@
 For the third study, we included fixed and random effects for both abundance and detection and again compared the distribution families. 
-Because our main goal with these studies was to demonstrate poor performance under model misspecification, and because the computation time for sampling from some models was large, we simulated only a small number of replicates.
 In all simulation studies, we focused on accuracy in estimation of $\pdet$ which then translated into estimation of abundance.
@@
 In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
-Detection rates $h(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
+Detection rates $h(t)$ of peaked distributions generally increase over time, while detection rates of nonpeaked distributions generally decrease over time.
 More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked.
@@
 
-\ifdic
-\begin{table}[ht]\centering\small
-\begin{tabular}{l|l|l|l|cccc|cccc||r}
- \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
- \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-  \hline
-  \hline
- \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\
- & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
- & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
-  \cline{3-13}
- & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
-  \cline{3-13}
- & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
- & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
- & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
-  \cline{2-13}
-& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\
- & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
- & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
-  \cline{3-13}
- & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
-  \cline{3-13}
- & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
- & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
- & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
-   \hline
-\end{tabular}
-\caption{Summary of mixture vs. non-mixture model fits.  
-In all cases, the inference model family matches the dataset family.  
-Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
+\begin{figure}\centering
+\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"}
+\caption{
+Comparisons of data and models with and without mixture TTDDs.
+\textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.
+Dashed lines show unbiased estimation.
+\textit{Right}: Coverage of 50\% credible intervals across 100 replicate simulations.
+Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.
+\textit{Rows}: Data simulated from non-mixture (upper) or mixture (lower) TTDDs.
+\textit{Columns}:  Data simulated from nonpeaked (left), exponential (center), or peaked (right) TTDDs.
+In these simulations, each inference model was only fit to datasets from the same TTDD family (e.g.
+lognormal to lognormal).
 }
-\label{tbl:sim1}
-\end{table}
+\label{fig:mixture_fig} 
+\end{figure}
 
-\else
+% FIRST SUBMISSION TABLE
+%\ifdic
+%\begin{table}[ht]\centering\small
+%\begin{tabular}{l|l|l|l|cccc|cccc||r}
+% \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
+% \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+%  \hline
+%  \hline
+% \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\
+% & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
+% & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
+%  \cline{3-13}
+% & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
+%  \cline{3-13}
+% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
+% & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
+% & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
+%  \cline{2-13}
+%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\
+% & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
+% & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
+%  \cline{3-13}
+% & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
+%  \cline{3-13}
+% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
+% & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
+% & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
+%   \hline
+%\end{tabular}
+%\caption{Summary of mixture vs. non-mixture model fits.  
+%In all cases, the inference model family matches the dataset family.  
+%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+%50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+%$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
+%}
+%\label{tbl:sim1}
+%\end{table}
+%
+%\else
+%
+%\begin{table}[ht]\centering\small
+%\begin{tabular}{l|l|l|l|cccc|cccc}
+% \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+% \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+%  \hline
+%  \hline
+% \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 \\
+% & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00  \\ 
+% & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 \\ 
+%  \cline{3-12}
+% & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88  \\ 
+%  \cline{3-12}
+% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 \\ 
+% & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 \\ 
+% & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 \\ 
+%  \cline{2-12}
+%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 \\
+% & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 \\ 
+% & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
+%  \cline{3-12}
+% & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 \\ 
+%  \cline{3-12}
+% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
+% & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 \\ 
+% & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
+%   \hline
+%\end{tabular}
+%\caption{Summary of mixture vs. non-mixture model fits.  
+%In all cases, the inference model family matches the dataset family.  
+%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+%50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+%}
+%\label{tbl:sim1}
+%\end{table}
+%\fi
 
-\begin{table}[ht]\centering\small
-\begin{tabular}{l|l|l|l|cccc|cccc}
- \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
- \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-  \hline
-  \hline
- \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 \\
- & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00  \\ 
- & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 \\ 
-  \cline{3-12}
- & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88  \\ 
-  \cline{3-12}
- & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 \\ 
- & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 \\ 
- & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 \\ 
-  \cline{2-12}
-& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 \\
- & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 \\ 
- & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
-  \cline{3-12}
- & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 \\ 
-  \cline{3-12}
- & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
- & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 \\ 
- & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
-   \hline
-\end{tabular}
-\caption{Summary of mixture vs. non-mixture model fits.  
-In all cases, the inference model family matches the dataset family.  
-Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-}
-\label{tbl:sim1}
-\end{table}
-\fi
+% THIS IS SIMILAR TO THE ABOVE TABLE FOR ALL FOUR PDET SIMULATIONS.
+% WE DEEMED IT TO OPPRESSIVE AND INSTEAD GENERATED THE FIGURES CURRENTLY IN THE PLOT
+%\begin{table}[ht]\centering\footnotesize
+%\setlength{\tabcolsep}{4pt} %default is 6
+%\begin{tabular}{l|l|l|l|cccc|cccc||cccc|cccc}
+% \multicolumn{4}{c|}{ } & \multicolumn{8}{c}{\underline{Non-mixture model}} & \multicolumn{8}{c}{\underline{Mixture model}} \\
+% \multicolumn{4}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
+% \multicolumn{4}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
+%  \hline
+%  \hline
+%\parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.69 & 0.75 & 0.95 & 0.11 & 0.78 & 0.45 & 0.59 & 0.76 & 0.78 & 0.83 & 0.96 & 0.00 & 0.35 & 0.55 & 0.61 \\
+%   &  &  & Lognormal & 0.54 & 0.59 & 0.71 & 0.94 & 0.81 & 0.45 & 0.29 & 0.47 & 0.68 & 0.73 & 0.82 & 0.96 & 0.05 & 0.63 & 0.63 & 0.44 \\ 
+%   &  &  & Weibull & 0.55 & 0.60 & 0.72 & 0.94 & 0.79 & 0.59 & 0.41 & 0.43 & 0.69 & 0.73 & 0.82 & 0.96 & 0.16 & 0.59 & 0.59 & 0.52 \\ 
+%\cline{3-20}
+%   &  &  & Exponential & 0.49 & 0.64 & 0.80 & 0.95 & 0.51 & 0.50 & 0.54 & 0.48 & 0.45 & 0.61 & 0.78 & 0.94 & 0.47 & 0.33 & 0.48 & 0.42 \\ 
+%\cline{3-20}
+%   & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.48 & 0.64 & 0.79 & 0.95 & 0.57 & 0.46 & 0.45 & 0.44 & 0.57 & 0.68 & 0.81 & 0.95 & 0.42 & 0.51 & 0.45 & 0.40 \\ 
+%   &  &  & Lognormal & 0.49 & 0.64 & 0.80 & 0.95 & 0.53 & 0.36 & 0.52 & 0.44 & 0.54 & 0.67 & 0.81 & 0.95 & 0.37 & 0.44 & 0.53 & 0.44 \\ 
+%   &  &  & Weibull & 0.48 & 0.61 & 0.78 & 0.95 & 0.40 & 0.46 & 0.38 & 0.49 & 0.59 & 0.69 & 0.81 & 0.96 & 0.37 & 0.46 & 0.49 & 0.42 \\ 
+%\cline{2-20}
+%\cline{2-20}
+%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.62 & 0.60 & 0.64 & 0.85 & 0.42 & 0.78 & 0.11 & 0.03 & 0.71 & 0.71 & 0.75 & 0.91 & 0.10 & 0.81 & 0.64 & 0.31 \\
+%   &  &  & Lognormal & 0.50 & 0.52 & 0.58 & 0.85 & 0.95 & 0.12 & 0.04 & 0.03 & 0.64 & 0.67 & 0.74 & 0.92 & 0.09 & 0.93 & 0.51 & 0.38 \\ 
+%   &  &  & Weibull & 0.50 & 0.51 & 0.54 & 0.71 & 0.92 & 0.23 & 0.01 & 0.01 & 0.64 & 0.66 & 0.70 & 0.87 & 0.39 & 0.92 & 0.46 & 0.28 \\ 
+%\cline{3-20}
+%   &  &  & Exponential & 0.88 & 0.93 & 0.96 & 0.99 & 0.00 & 0.00 & 0.00 & 0.00 & 0.50 & 0.63 & 0.78 & 0.94 & 0.54 & 0.45 & 0.42 & 0.51 \\ 
+%\cline{3-20}
+% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.24 & 0.26 & 0.28 & 0.37 & 0.00 & 0.00 & 0.00 & 0.00 & 0.52 & 0.58 & 0.72 & 0.94 & 0.76 & 0.51 & 0.40 & 0.41 \\ 
+%   &  &  & Lognormal & 0.18 & 0.20 & 0.22 & 0.27 & 0.00 & 0.00 & 0.00 & 0.00 & 0.49 & 0.61 & 0.78 & 0.94 & 0.66 & 0.58 & 0.46 & 0.43 \\ 
+%   &  &  & Weibull & 0.18 & 0.19 & 0.22 & 0.28 & 0.00 & 0.00 & 0.00 & 0.00 & 0.52 & 0.56 & 0.69 & 0.93 & 0.80 & 0.44 & 0.33 & 0.47 \\ 
+%  \hline
+%\end{tabular}
+%\caption{Summary of mixture vs. non-mixture model fits over 100 replicates.
+%In all cases, the inference model family matches the dataset family.  
+%Column titles denote the true detection probability (e.g. `p80' means $\pdet = 0.80$). 
+%Medians are averaged across all replicates.
+%Coverage is for 50\% credible intervals.
+%For nominal coverage, we expect values between (0.41, 0.59) 95\% of the time.}
+%\label{tbl:sim1}
+%\end{table}
 
@@
 
+\begin{figure}\centering
+\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"}
+\caption{
+Comparisons of TTDDs across families.
+\textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.
+Dashed lines show unbiased estimation.
+\textit{Right}: Coverage of 50\% credible intervals across 100 replicate simulations.
+Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.
+\textit{Rows}: Denote the TTDD family used to simulated data.
+\textit{Columns}:  Data simulated from nonpeaked (left) or exponential/peaked (right) TTDDs.
+In these simulations, we used only mixture datasets and inference models.
+}
+\label{fig:family_fig}
+\end{figure}
+
 Figure \ref{fig:sim2typical} presents a representative example of posterior distributions for $\pdet$ for data and models from the exponential and gamma mixture families. 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Oral Prelim/Posteriors_Families_draft"}
+\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"}
 \caption{\label{fig:sim2typical} 
-Representative example of the posterior distributions for $\pdet$ from exponential (light solid line) and gamma (dark solid line) mixture models fit to data simulated from nonpeaked gamma (shape$<$1), exponential (shape$=$1), and peaked gamma (shape$>$1) mixture TTDDs, with true detection probabilities shown for comparison (dashed vertical line, $\pdet=0.8$).
+Representative examples of posterior distributions for $\pdet$ (left panel) and survey-level log-scale abundance (right panel).
+Data are simulated from nonpeaked gamma (left column), exponential (center), and peaked gamma (right) mixture TTDDs.
+True detection probabilities are shown for comparison (dashed vertical line) and vary by row.
+Inference models are either exponential mixture (light gray) or gamma mixture (dark gray) TTDDs.
 }
@@
 
-\ifdic
-\begin{table}[ht]
-\footnotesize\centering
-\begin{tabular}{l|l|l|ccccc|ccccc}
- \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
- \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-  \hline
-\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\
-& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
-& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
-\cline{2-13}
-& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
-\cline{2-13}
-& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
-& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
-& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
-   \hline
-\end{tabular}
-\vspace{0.5cm}\\
-\begin{tabular}{l|l|l|ccccc|ccccc}
- \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
- \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-  \hline
-\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\
-& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
-& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
-\cline{2-13}
-& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
-\cline{2-13}
-& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
-& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
-& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
-   \hline
-\end{tabular}
-\caption{Summary of mixture inference models of all families fit to mixture datasets.  
-Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
-\label{tbl:sim2}
-\end{table}
+% FIRST SUBMISSION TABLE
+%\ifdic
+%\begin{table}[ht]
+%\footnotesize\centering
+%\begin{tabular}{l|l|l|ccccc|ccccc}
+% \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
+% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+%  \hline
+%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\
+%& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
+%& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
+%\cline{2-13}
+%& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
+%\cline{2-13}
+%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
+%& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
+%& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
+%   \hline
+%\end{tabular}
+%\vspace{0.5cm}\\
+%\begin{tabular}{l|l|l|ccccc|ccccc}
+% \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
+% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
+%  \hline
+%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\
+%& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
+%& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
+%\cline{2-13}
+%& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
+%\cline{2-13}
+%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
+%& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
+%& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
+%   \hline
+%\end{tabular}
+%\caption{Summary of mixture inference models of all families fit to mixture datasets.  
+%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+%50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
+%$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
+%\label{tbl:sim2}
+%\end{table}
+%
+%\else
+%
+%\begin{table}[ht]
+%\footnotesize\centering
+%\begin{tabular}{l|l|l|cccc|cccc}
+% \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\%  & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+%  \hline
+%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 0.76 & 0.41 & 0.69 & 1.00 \\
+%& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 0.85 & 0.72 & 0.44 & 0.69 \\ 
+%& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.79 & 0.49 & 0.69 & 1.00 \\ 
+%\cline{2-11}
+%& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & 0.68 & 0.24 & 0.25 & 0.81 \\ 
+%\cline{2-11}
+%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
+%& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 0.84 & 0.73 & 0.44 & 0.81 \\ 
+%& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 0.66 & 0.15 & 0.19 & 0.75 \\
+%   \hline
+%\end{tabular}
+%\vspace{0.5cm}\\
+%\begin{tabular}{l|l|l|cccc|cccc}
+% \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
+%  \hline
+%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.67 & 0.24 & 0.31 & 1.00 \\
+%& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & 0.77 & 0.48 & 0.44 & 1.00 \\ 
+%& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
+%\cline{2-11}
+%& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & 0.63 & 0.20 & 0.25 & 0.75 \\ 
+%\cline{2-11}
+%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & 0.79 & 0.52 & 0.69 & 0.88 \\ 
+%& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & 0.89 & 0.88 & 0.12 & 0.69 \\ 
+%& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
+%   \hline
+%\end{tabular}
+%\caption{Summary of mixture inference models of all families fit to mixture datasets.  
+%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
+%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+%50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+%\label{tbl:sim2}
+%\end{table}
+%\fi
 
-\else
-
-\begin{table}[ht]
-\footnotesize\centering
-\begin{tabular}{l|l|l|cccc|cccc}
- \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
- \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\%  & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-  \hline
-\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 0.76 & 0.41 & 0.69 & 1.00 \\
-& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 0.85 & 0.72 & 0.44 & 0.69 \\ 
-& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.79 & 0.49 & 0.69 & 1.00 \\ 
-\cline{2-11}
-& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & 0.68 & 0.24 & 0.25 & 0.81 \\ 
-\cline{2-11}
-& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
-& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 0.84 & 0.73 & 0.44 & 0.81 \\ 
-& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 0.66 & 0.15 & 0.19 & 0.75 \\
-   \hline
-\end{tabular}
-\vspace{0.5cm}\\
-\begin{tabular}{l|l|l|cccc|cccc}
- \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
- \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-  \hline
-\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.67 & 0.24 & 0.31 & 1.00 \\
-& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & 0.77 & 0.48 & 0.44 & 1.00 \\ 
-& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
-\cline{2-11}
-& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & 0.63 & 0.20 & 0.25 & 0.75 \\ 
-\cline{2-11}
-& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & 0.79 & 0.52 & 0.69 & 0.88 \\ 
-& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & 0.89 & 0.88 & 0.12 & 0.69 \\ 
-& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
-   \hline
-\end{tabular}
-\caption{Summary of mixture inference models of all families fit to mixture datasets.  
-Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
-\label{tbl:sim2}
-\end{table}
-\fi
+% THIS IS SIMILAR TO THE ABOVE TABLE FOR ALL FOUR PDET SIMULATIONS.
+% WE DEEMED IT TO OPPRESSIVE AND INSTEAD GENERATED THE FIGURES CURRENTLY IN THE PLOT
+%\begin{table}[ht]
+%\footnotesize\centering\setlength{\tabcolsep}{4pt} %default is 6
+%\begin{tabular}{l|l|l|cccc|cccc||cccc|cccc}
+% \multicolumn{3}{c}{ } & \multicolumn{8}{c}{\underline{Exponential mixture model}} & \multicolumn{8}{c}{\underline{Gamma mixture model}} \\
+% \multicolumn{3}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
+% \multicolumn{3}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
+%  \hline
+%  \hline
+%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.84 & 0.83 & 0.86 & 0.94 & 0.00 & 0.04 & 0.25 & 0.38 & 0.63 & 0.63 & 0.65 & 0.81 & 0.23 & 0.87 & 0.16 & 0.10 \\
+%   &  & Lognormal & 0.89 & 0.91 & 0.93 & 0.97 & 0.00 & 0.00 & 0.00 & 0.07 & 0.64 & 0.67 & 0.74 & 0.92 & 0.09 & 0.93 & 0.51 & 0.38 \\ 
+%   &  & Weibull & 0.86 & 0.87 & 0.89 & 0.94 & 0.00 & 0.01 & 0.11 & 0.37 & 0.63 & 0.64 & 0.66 & 0.82 & 0.18 & 0.97 & 0.22 & 0.15 \\ 
+%\cline{2-19}
+%   &  & Exponential & 0.50 & 0.63 & 0.78 & 0.94 & 0.54 & 0.45 & 0.42 & 0.51 & 0.49 & 0.55 & 0.61 & 0.82 & 0.77 & 0.39 & 0.12 & 0.09 \\ 
+%\cline{2-19}
+%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.31 & 0.33 & 0.37 & 0.55 & 0.05 & 0.02 & 0.00 & 0.00 & 0.45 & 0.50 & 0.64 & 0.91 & 0.60 & 0.21 & 0.10 & 0.10 \\ 
+%   &  & Lognormal & 0.26 & 0.30 & 0.34 & 0.47 & 0.00 & 0.00 & 0.00 & 0.00 & 0.49 & 0.61 & 0.78 & 0.94 & 0.66 & 0.58 & 0.46 & 0.43 \\ 
+%   &  & Weibull & 0.30 & 0.32 & 0.38 & 0.57 & 0.05 & 0.00 & 0.00 & 0.00 & 0.43 & 0.47 & 0.57 & 0.84 & 0.51 & 0.13 & 0.02 & 0.00 \\ 
+%   \hline
+%\end{tabular}
+%\vspace{0.5cm}\\
+%\begin{tabular}{l|l|l|cccc|cccc||cccc|cccc}
+% \multicolumn{3}{c}{ } & \multicolumn{8}{c}{\underline{Lognormal mixture model}} & \multicolumn{8}{c}{\underline{Weibull mixture model}} \\
+% \multicolumn{3}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
+% \multicolumn{3}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
+%  \hline
+% \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.71 & 0.71 & 0.75 & 0.91 & 0.10 & 0.81 & 0.64 & 0.31 & 0.63 & 0.63 & 0.67 & 0.87 & 0.54 & 0.94 & 0.34 & 0.22 \\
+%   &  & Lognormal & 0.78 & 0.82 & 0.88 & 0.96 & 0.01 & 0.19 & 0.33 & 0.46 & 0.68 & 0.73 & 0.81 & 0.95 & 0.21 & 0.68 & 0.59 & 0.57 \\ 
+%   &  & Weibull & 0.73 & 0.75 & 0.79 & 0.91 & 0.03 & 0.54 & 0.66 & 0.32 & 0.64 & 0.66 & 0.70 & 0.87 & 0.39 & 0.92 & 0.46 & 0.28 \\ 
+%\cline{2-19}
+%   &  & Exponential & 0.54 & 0.60 & 0.70 & 0.92 & 0.84 & 0.70 & 0.32 & 0.33 & 0.54 & 0.58 & 0.64 & 0.88 & 0.87 & 0.69 & 0.24 & 0.25 \\ 
+%\cline{2-19}
+% & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.52 & 0.58 & 0.72 & 0.94 & 0.76 & 0.51 & 0.40 & 0.41 & 0.55 & 0.62 & 0.77 & 0.97 & 0.77 & 0.59 & 0.46 & 0.26 \\ 
+%   &  & Lognormal & 0.59 & 0.71 & 0.85 & 0.97 & 0.52 & 0.48 & 0.24 & 0.27 & 0.64 & 0.77 & 0.90 & 0.98 & 0.43 & 0.34 & 0.13 & 0.01 \\ 
+%   &  & Weibull & 0.49 & 0.53 & 0.65 & 0.89 & 0.75 & 0.31 & 0.22 & 0.09 & 0.52 & 0.56 & 0.69 & 0.93 & 0.80 & 0.44 & 0.33 & 0.47 \\ 
+%   \hline
+%\end{tabular}
+%\caption{Summary of mixture inference models for all families.  
+%Column titles denote the true detection probability (e.g. `p80' means $\pdet = 0.80$). 
+%Medians are averaged across all replicates.
+%Coverage is for 50\% credible intervals.
+%For nominal coverage, we expect values between (0.41, 0.59) 95\% of the time.}
+%\label{tbl:sim2}
+%\end{table}
 
@@
 
-The results from this simulation are qualitatively similar to that Ovenbird analysis (Figure \ref{ovenposteriors}) and thus we only briefly review the results here and provide the corresponding figures and tables in the Supplementary Material.
+The results from this simulation are qualitatively similar to the Ovenbird analysis (Figure \ref{ovenposteriors}) and thus we only briefly review the results here and provide the corresponding figures and tables in the Supplementary Material.
 Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).

### 2017-04-03 ReviewerComments.tex
Updates to draft and reviewer's comments.

@@
 Our manuscript explores distinct topics in that our model is for removal-only data, whereas their models are for removal-distance data.
-Also, we explicitly investigate non-constant detection distributions, which has not been their focus.
+Also, we specifically investigate non-constant detection distributions, which has not been their focus.
 \end{quote}
@@
 \end{enumerate}
+\begin{quote}
+We address each of these in turn as they appear in the comments below.
+\end{quote}
 
@@
 \begin{quote}
-We have prefaced the results with the clause: ``Based on simulation analyses, ..."
+We have prefaced the statement with the clause: ``Based on simulation analyses, ..."
 \end{quote}
@@
 We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
-\adam{I am unclear what the reviewer means by `providing citations for the cdf and pdfs outlined here for reference'.}
 \end{quote}
@@
 We have added the following sentences to \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
-All three could be reasonable TTDD choices if, for instance, detection rates vary among individuals.
+All three may be reasonable TTDD choices if, for instance, detection rates vary among individuals.
 This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population."
@@
 That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
-The dynamics of the system may be more complicated than we can appreciate.
+If the dynamics of the system are more complicated than we can directly model, the best we can do is find a TTDD that reasonably describes the observed pattern.
 For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
+Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences for tail probability.
 \end{quote}
@@
 ``In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
-Detection rates $h(t)$ of peaked distributions generally decrease over time, while detection rates of nonpeaked distributions generally increase over time.
+Detection rates $h(t)$ of peaked distributions generally increase over time, while detection rates of nonpeaked distributions generally decrease over time.
 More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked."
-We do not believe this distinction requires a separate explanation in the Introduction.
-\jarad{Can we reword to remove peaked and nonpeaked here?}
+
+We do not believe this distinction requires a separate explanation in the Introduction.  It distinguishes between scenarios but is not essential to framing the problem of non-constant detection.
 \end{quote}
@@
 \item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
-\jarad{Computation time.}
 \begin{quote}
-We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
-However, it is informative.
-TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
+We agree that an intercept-only model is simplistic and will not be of much practical use in an applied setting.
+However, the intercept-only simulation is informative, because models that perform poorly in simple applications can hardly be trusted in more complex applications.
 
-On the practical side, while simulation studies are hampered by long computation times, the typical user of this model will only need to apply it to a comparatively smaller number of datasets.
+Meanwhile, a simulation study involving effects on abundance and detection is computationally expensive. \adam{Reference exact duration required}.
+Fortunately, while simulation studies are hampered by long computation times, the typical user of this model will only apply it to a small number of datasets, so computation time will not be prohibitive.
 \end{quote}
+
 \item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
 \begin{quote}
-\adam{Investigate more/different simulations}
+We have extended our simulation study to include a variety of detection probabilities rather than a variety of sample sizes.
+Both affect estimation and are worthy of investigation, but since the performance of unknown-$N$ binomial models generally deteriorates as values of $\pdet$ decrease, we chose to direct our attentions there.
+Sections \ref{M-sec:mixture} and \ref{M-sec:family} have been reworked to detail the updated findings.
 \end{quote}
@@
 \begin{quote}
-We appreciate that these are important decision in the design of surveys, but we have not investigated either of these issues with respect to bias and precision.
-Not having done so, we prefer not to speculate.
+These are important decision in the design of surveys, but they are beyond the scope we have defined for this study.
+Not having explicitly investigated them, we prefer not to speculate.
 %\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
 We have separated out a paragraph in the Discussion posing some design-related questions.
-\adam{Insert a statement about the NRRI one-minute intervals relative to two- or three-interval designs?}
 \end{quote}
@@
 \begin{quote}
-We have retained just the 50\% coverage.
-\adam{Waiting until all new simulations are run.}
+We agree that using both is redundant.
+The two measures are largely the same, and we note where they differ in Sections \ref{M-sec:mixture} and \ref{M-sec:family}.
+We now choose to report only 50\% credible intervals.
+This choice more effectively highlights over/undercoverage in figures \ref{M-fig:mixture_fig} and \ref{M-fig:family_fig}.
+
 \end{quote}
@@
 \begin{quote}
-Clarification: 947 detections at 381 sites (which is \textit{really} a lot of detections).
-We have placed this information in Section \ref{M-sec:data}: ``The data included 65 sites and a total of 381 surveys with site specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s.
-A total of 947 Ovenbirds were counted with a maximum of eight in any single survey."
+Clarification: We have 947 detections at 381 sites (which is \textit{really} a lot of detections).
+We now clearly state in Section \ref{M-sec:data}: ``This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s."
 \end{quote}
@@
 The cited credible intervals were for \textit{uncounted} birds not total abundance.
-We have now changed the figure to $\log_{10}$(Abundance) to avoid confusion.
+We have now changed the figure to $\log_{10}$(total abundance) to avoid confusion.
 The gamma TTDD 95\% credible interval for total abundance is (1007, 2836).
@@
 \begin{quote}
-By the precautionary principle, we believe the burden of proof should rest with the analyst who wishes to assume constant detection rather than with the analyst who wishes to account for non-constant detection.
+This last clause implies that a constant-rate model should be the default.
+We disagree.
+Based on our results, applying the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection.
 This is a central point in our manuscript: the constant-rate assumption should not be the default.
@@
 This manuscript applies to removal-only analyses.
-Still, the concept should carry over to removal-distance analyses.
+Still, the concept can be carried over to removal-distance analyses.
 The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
@@
 \begin{quote}
-We have not been clear enough in stating our goals for the simulations.
-We have added the following line in Section \ref{M-sec:sim}: ``Because our main goal with these studies was to demonstrate poor performance under model misspecification, and because the computation time for sampling from some models was large, we simulated only a small number of replicates."
-\adam{Can we run more simulations?  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wider range of simulated $\pdet$.}
+We answer this comments in two parts.
+First, we agree that the number of simulations was inadequate for diagnosing model accuracy.
+We have therefore increased to 100 the number of simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family}.
+This has been made possible by improvements to Stan \citep{Rstan2016} and by streamlining our code for homogeneous surveys.
+\adam{Add reference to summary of run times vis-a-vis the third simulation study}
+
+Second, although the original number of simulations was inadequate for characterizing model accuracy, it nonetheless highlighted scenarios of gross inaccuracy, such as: (i) omitting a mixture component in modeling a peaked mixture dataset, or (ii) assuming constant detection when the rate varies during the survey period.
 \end{quote}
@@
 Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det).
-Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det)
+Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det).
+(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
+Without variation in ``true" simulated scenarios, it is easy to inadvertendly ``stack the deck" in favor or against a particular combination of settings (Particularly since your ``true" value of p is high: 0.8).
+A wide range of simulated truths is also necessary because to be useful, these methods should be widely applicable in the
+tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
 \begin{quote}
-\adam{We should run more simulations.  Reviewer 1 wants more variation in $n^{(obs)}$.  Reviewer 2 wants a wider range of simulated $\pdet$.}
-\end{quote}
+We have repeated the simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family} at true detection probabilities $\pdet = 0.50, 0.65, 0.80$, and 0.95.
+The findings have significantly added to our results and conclusions.
 
-(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
-\begin{quote}
-It is not a hard boundary.  
-But the boundary is a property of estimating $N$ in Binomial unknown-$N$, unknown-$p$ problems.
-Olkin's calculation results from the relative sample mean and sample variance of large samples.  
+With reference to Olkin's finding that point estimates of $N_s$ are unstable unless $\pdet>0.41$, it is not a hard boundary.  
+It varies by context, but the boundary is a general property of estimating $N$ in binomial unknown-$N$, unknown-$p$ problems.
+Olkin's calculation derives from the relative sample mean and sample variance of repeated large samples with the same values of $N$ and $\pdet$.
+As such, it is a best-case scenario.
 Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
@@
 \end{itemize}
-\adam{Since I've cited the above articles, is there a reason to work them into the manuscript?  I don't feel they add to our story.}
 \end{quote}
-
- Without variation in ``true" simulated scenarios, it is easy to inadvertendly ``stack the deck" in favor or against a particular combination of settings (Particularly since your ``true" value of p is high: 0.8).
-A wide range of simulated truths is also necessary because to be useful, these methods should be widely applicable in the
-tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
-
-\jarad{Although we should look at more simulations. The purpose of the current values
-was to give some credence to the Ovenbird analysis.}
 
@@
 \begin{quote}
-We have posterior estimates of abundance from our model fits.  We have added them to the manuscript.\\
-\adam{Thoughts on how to do this:
-\begin{enumerate}
-\item Generate Tables 1 \& 2 for $\hat{N}$ as well as $\pdet$, but this gets crowded.
-\item Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$, but averaging across replicates is not a trivial issue
-\item Supplement Figure 1 with a row of representative abundance posteriors (no averaging required)
-\item Following the profile likelihood comment, generate a plot of $\hat{N}$ vs. $\pdet$ for a representative sample.
-\end{enumerate}
-}
-\jarad{I think 3.}
+With the expansion of our simulations to include multiple true values for $\pdet$, we have switched from tabular summaries of results to graphical summaries (though the tabular summaries will still be available in the supplemental material).
+We think the most informative way to present results on the abundance scale is to add them to the Figure \ref{M-fig:sim2typical} results on detection probabilities.
+This figure presents examples of entire posterior distributions rather than just median and coverage point estimates.
+The inclusion of abundance estimates in Figure \ref{M-fig:sim2typical} should convey much the same information as would profile likelihoods.
+\adam{Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$?  But averaging across replicates is not a trivial issue.}
 \end{quote}
@@
 \begin{quote}
-\adam{
-\begin{itemize}
-\item The Solymos package is built for BUGS and JAGS.  Most of the tools are easily hand-coded, I think.
-\item To simplify our lives, we could choose to attempt this for only quick-running models on the actual data...
-there's no reason to suspect identifiability issues for the gamma TTDD that would not appear for the Weibull or lognormal.
-\item Technical note: they suggest using narrower priors as the number of clones increases.  Likewise, we could input posterior means as initial values.
-\end{itemize}
-}
-\jarad{I'm not inclined to do this.}
+Thank you for pointing out \citet{Lele2010}.
+We found the article to be interesting and novel, using data cloning to exploit relationships among data likelihood, priors, and identifiability.
+While the approach could certainly be applied to our model in its more complex forms (e.g. with fixed and random effects), to do so in this manuscript would distract from our main objectives.
+We have introduced more discussion of identifiability and flat data likelihoods in \adam{location}, and we have cited \citep{Lele2010} there.
+
+\adam{YOU HAD BETTER DO THIS!}
+
+As an informational note, the \texttt{R} package by Solymos was constructed for BUGS and JAGS, though it would not be difficult to adapt data cloning for Stan.
 \end{quote}
@@
 \begin{quote}
-We appreciate the merit of the reviewer's suggestion and have added comments in the Discussion.
-We focus our comments on a possible generalized gamma TTDD, which is an umbrella distribution that includes exponential, gamma, lognormal, and Weibull distributions as special cases.
+We appreciate the merit of the reviewer's suggestion and have added related comments in the Discussion.
+We focus on a possible generalized gamma TTDD, which is an umbrella distribution that includes exponential, gamma, lognormal, and Weibull distributions as special cases.
 The generalized Pareto analysis in \citet{Beisel2007} requires observations from the tail of the TTDD distribution, which we do not have.
@@
 It may well be possible to implement our model with a generalized gamma TTDD, but especially when we consider the right-truncation of data from point-count surveys, we think model convergence would not be a trivial problem."
-\adam{This is an area of active research.  I feel like my tone is not optimistic enough.  Then again, it's not worth micro-wordsmithing.}
 \end{quote}
@@
 \begin{quote}
-In the Introduction, we now clearly describe a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
-Later in the same paragraph, we state ``We apply the term `mixture model' to any model with a mixture TTDD."
-This is contrasted in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model \dotso, which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
+In the Introduction, we now clearly refer to a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
+We define our use of `mixture model' later in the paragraph:  ``We apply the term `mixture model' to any model with a mixture TTDD."
+This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso], which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
 \end{quote}

### 2017-04-03 Supplement.tex
Updates to draft and reviewer's comments.

@@
 %\end{sidewaystable}
+
+\begin{table}[ht]
+\footnotesize\centering
+\begin{tabular}{l|l|l|l|cccc|cccc}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+ \multicolumn{4}{c|}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+  \hline
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.84 & 0.11 & 0.94 & 0.76 & 0.92 & 0.00 & 0.70 \\
+   &  &  & Lognormal & 0.54 & 0.60 & 0.81 & 1.00 & 0.68 & 0.88 & 0.05 & 0.79 \\ 
+   &  &  & Weibull & 0.55 & 0.61 & 0.79 & 0.99 & 0.69 & 0.84 & 0.16 & 0.87 \\ 
+  \cline{3-12}
+ & & &  Exponential & 0.49 & 0.45 & 0.51 & 0.93 & 0.45 & 0.30 & 0.47 & 0.86 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.48 & 0.46 & 0.57 & 0.90 & 0.57 & 0.72 & 0.42 & 0.85 \\ 
+   &  &  & Lognormal & 0.49 & 0.45 & 0.53 & 0.93 & 0.54 & 0.69 & 0.37 & 0.89 \\ 
+   &  &  & Weibull & 0.48 & 0.48 & 0.40 & 0.91 & 0.59 & 0.74 & 0.37 & 0.80 \\ 
+  \cline{2-12}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.62 & 0.76 & 0.42 & 0.99 & 0.71 & 0.86 & 0.10 & 0.89 \\
+   &  &  & Lognormal & 0.50 & 0.47 & 0.95 & 1.00 & 0.64 & 0.84 & 0.09 & 0.97 \\ 
+   &  &  & Weibull & 0.50 & 0.47 & 0.92 & 1.00 & 0.64 & 0.77 & 0.39 & 0.98 \\ 
+  \cline{3-12}
+ & & &   Exponential & 0.88 & 1.00 & 0.00 & 0.00 & 0.50 & 0.49 & 0.54 & 0.95 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.24 & 0.01 & 0.00 & 0.00 & 0.52 & 0.53 & 0.76 & 0.99 \\ 
+   &  &  & Lognormal & 0.18 & 0.00 & 0.00 & 0.00 & 0.49 & 0.47 & 0.66 & 0.97 \\ 
+   &  &  & Weibull & 0.18 & 0.00 & 0.00 & 0.00 & 0.52 & 0.52 & 0.80 & 1.00 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture vs. non-mixture model fits when the detection probability is $\pdet =0.50$.  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+
+\vspace{1.2cm}
+
+\begin{tabular}{l|l|l|l|cccc|cccc}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+ \multicolumn{4}{c|}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+  \hline
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}}& Gamma & 0.69 & 0.59 & 0.78 & 0.98 & 0.78 & 0.79 & 0.35 & 0.90 \\
+   &  &  & Lognormal & 0.59 & 0.31 & 0.45 & 0.91 & 0.73 & 0.69 & 0.63 & 0.90 \\ 
+   &  &  & Weibull & 0.60 & 0.38 & 0.59 & 0.99 & 0.73 & 0.68 & 0.59 & 0.94 \\
+  \cline{3-12}
+   &  &  & Exponential & 0.64 & 0.47 & 0.50 & 0.85 & 0.61 & 0.29 & 0.33 & 0.82 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}}& Gamma & 0.64 & 0.44 & 0.46 & 0.89 & 0.68 & 0.66 & 0.51 & 0.85 \\ 
+   &  &  & Lognormal & 0.64 & 0.47 & 0.36 & 0.87 & 0.67 & 0.63 & 0.44 & 0.79 \\ 
+   &  &  & Weibull & 0.61 & 0.40 & 0.46 & 0.88 & 0.69 & 0.67 & 0.46 & 0.88 \\ 
+  \cline{2-12}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}}& Gamma & 0.60 & 0.36 & 0.78 & 1.00 & 0.71 & 0.63 & 0.81 & 0.99 \\
+   &  &  & Lognormal & 0.52 & 0.13 & 0.12 & 0.74 & 0.67 & 0.54 & 0.93 & 1.00 \\ 
+   &  &  & Weibull & 0.51 & 0.17 & 0.23 & 0.82 & 0.66 & 0.52 & 0.92 & 1.00 \\ 
+  \cline{3-12}
+   &  &  & Exponential & 0.93 & 1.00 & 0.00 & 0.00 & 0.63 & 0.46 & 0.45 & 0.87 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.26 & 0.00 & 0.00 & 0.00 & 0.58 & 0.34 & 0.51 & 0.95 \\ 
+   &  &  & Lognormal & 0.20 & 0.00 & 0.00 & 0.00 & 0.61 & 0.41 & 0.58 & 0.92 \\ 
+   &  &  & Weibull & 0.19 & 0.00 & 0.00 & 0.00 & 0.56 & 0.34 & 0.44 & 0.95 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture vs. non-mixture model fits when the detection probability is $\pdet =0.65$.  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+
+\end{table}
+
+\begin{table}[ht]\centering\footnotesize
+\begin{tabular}{l|l|l|l|cccc|cccc}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+ \multicolumn{4}{c|}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+  \hline
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.75 & 0.39 & 0.45 & 0.92 & 0.83 & 0.62 & 0.55 & 0.91 \\
+   &  &  & Lognormal & 0.71 & 0.29 & 0.29 & 0.82 & 0.82 & 0.61 & 0.63 & 0.91 \\ 
+   &  &  & Weibull & 0.72 & 0.34 & 0.41 & 0.86 & 0.82 & 0.60 & 0.59 & 0.93 \\ 
+  \cline{3-12}
+    &  &  & Exponential & 0.80 & 0.49 & 0.54 & 0.88 & 0.78 & 0.32 & 0.48 & 0.86 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.79 & 0.44 & 0.45 & 0.87 & 0.81 & 0.59 & 0.45 & 0.90 \\ 
+   &  &  & Lognormal & 0.80 & 0.48 & 0.52 & 0.95 & 0.81 & 0.60 & 0.53 & 0.92 \\ 
+   &  &  & Weibull & 0.78 & 0.37 & 0.38 & 0.84 & 0.81 & 0.57 & 0.49 & 0.92 \\ 
+  \cline{2-12}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.64 & 0.11 & 0.11 & 0.59 & 0.75 & 0.38 & 0.64 & 1.00 \\
+   &  &  & Lognormal & 0.58 & 0.06 & 0.04 & 0.27 & 0.74 & 0.33 & 0.51 & 0.98 \\ 
+   &  &  & Weibull & 0.54 & 0.04 & 0.01 & 0.23 & 0.70 & 0.27 & 0.46 & 1.00 \\ 
+  \cline{3-12}
+   &  &  & Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.78 & 0.45 & 0.42 & 0.86 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.72 & 0.30 & 0.40 & 0.80 \\ 
+   &  &  & Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.78 & 0.45 & 0.46 & 0.87 \\ 
+   &  &  & Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.69 & 0.29 & 0.33 & 0.88 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture vs. non-mixture model fits when the detection probability is $\pdet =0.80$.  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+
+\vspace{1.2cm}
+% Note: both tables are in the same 'table' environment.
+
+\begin{tabular}{l|l|l|l|cccc|cccc}
+ \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
+ \multicolumn{4}{c|}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+  \hline
+ \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.95 & 0.44 & 0.59 & 0.95 & 0.96 & 0.62 & 0.61 & 0.96 \\
+   &  &  & Lognormal & 0.94 & 0.42 & 0.47 & 0.87 & 0.96 & 0.65 & 0.44 & 0.90 \\ 
+   &  &  & Weibull & 0.94 & 0.46 & 0.43 & 0.90 & 0.96 & 0.67 & 0.52 & 0.84 \\ 
+  \cline{3-12}
+   &  &  & Exponential & 0.95 & 0.54 & 0.48 & 0.90 & 0.94 & 0.37 & 0.42 & 0.89 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.95 & 0.43 & 0.44 & 0.89 & 0.95 & 0.50 & 0.40 & 0.88 \\ 
+   &  &  & Lognormal & 0.95 & 0.50 & 0.44 & 0.88 & 0.95 & 0.54 & 0.44 & 0.88 \\ 
+   &  &  & Weibull & 0.95 & 0.53 & 0.49 & 0.91 & 0.96 & 0.65 & 0.42 & 0.85 \\ 
+  \cline{2-12}
+& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.85 & 0.04 & 0.03 & 0.22 & 0.91 & 0.25 & 0.31 & 0.77 \\
+   &  &  & Lognormal & 0.85 & 0.03 & 0.03 & 0.16 & 0.92 & 0.28 & 0.38 & 0.86 \\ 
+   &  &  & Weibull & 0.71 & 0.01 & 0.01 & 0.02 & 0.87 & 0.22 & 0.28 & 0.68 \\ 
+  \cline{3-12}
+   &  &  & Exponential & 0.99 & 1.00 & 0.00 & 0.00 & 0.94 & 0.41 & 0.51 & 0.85 \\ 
+  \cline{3-12}
+ & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.37 & 0.00 & 0.00 & 0.00 & 0.94 & 0.39 & 0.41 & 0.83 \\ 
+   &  &  & Lognormal & 0.27 & 0.00 & 0.00 & 0.00 & 0.94 & 0.40 & 0.43 & 0.83 \\ 
+   &  &  & Weibull & 0.28 & 0.00 & 0.00 & 0.00 & 0.93 & 0.36 & 0.47 & 0.80 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of mixture vs. non-mixture model fits when the detection probability is $\pdet =0.95$.  
+In all cases, the inference model family matches the dataset family.  
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\end{table}
+
+\begin{table}[ht]\centering\footnotesize
+\centering
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\%  & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.84 & 0.99 & 0.00 & 0.05 & 0.71 & 0.86 & 0.10 & 0.89 \\
+   &  & Lognormal & 0.89 & 1.00 & 0.00 & 0.00 & 0.78 & 0.93 & 0.01 & 0.56 \\ 
+   &  & Weibull & 0.86 & 1.00 & 0.00 & 0.02 & 0.73 & 0.89 & 0.03 & 0.82 \\ 
+\cline{2-11}
+   &  & Exponential & 0.50 & 0.49 & 0.54 & 0.95 & 0.54 & 0.57 & 0.84 & 1.00 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.31 & 0.06 & 0.05 & 0.32 & 0.52 & 0.53 & 0.76 & 0.99 \\ 
+   &  & Lognormal & 0.26 & 0.02 & 0.00 & 0.09 & 0.59 & 0.68 & 0.52 & 0.91 \\ 
+   &  & Weibull & 0.30 & 0.05 & 0.05 & 0.28 & 0.49 & 0.46 & 0.75 & 1.00 \\ 
+   \hline
+\end{tabular}
+\vspace{0.5cm}\\
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.63 & 0.80 & 0.23 & 0.99 & 0.63 & 0.74 & 0.54 & 0.98 \\
+   &  & Lognormal & 0.64 & 0.84 & 0.09 & 0.97 & 0.68 & 0.83 & 0.21 & 0.95 \\ 
+   &  & Weibull & 0.63 & 0.81 & 0.18 & 0.99 & 0.64 & 0.77 & 0.39 & 0.98 \\ 
+\cline{2-11}
+   &  & Exponential & 0.49 & 0.45 & 0.77 & 1.00 & 0.54 & 0.56 & 0.87 & 0.99 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.45 & 0.36 & 0.60 & 1.00 & 0.55 & 0.58 & 0.77 & 0.98 \\ 
+   &  & Lognormal & 0.49 & 0.47 & 0.66 & 0.97 & 0.64 & 0.75 & 0.43 & 0.87 \\ 
+   &  & Weibull & 0.43 & 0.30 & 0.51 & 0.97 & 0.52 & 0.52 & 0.80 & 1.00 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of models fits across family of TTDD when true detection probability is $\pdet = 0.50$.
+All data and inference models have mixture components.
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+
+\vspace{0.5cm}
+% Note: both tables are in the same 'table' environment.
+
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\%  & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.83 & 0.96 & 0.04 & 0.21 & 0.71 & 0.63 & 0.81 & 0.99 \\
+   &  & Lognormal & 0.91 & 1.00 & 0.00 & 0.00 & 0.82 & 0.86 & 0.19 & 0.70 \\ 
+   &  & Weibull & 0.87 & 0.98 & 0.01 & 0.09 & 0.75 & 0.72 & 0.54 & 0.97 \\ 
+\cline{2-11}
+   &  & Exponential & 0.63 & 0.46 & 0.45 & 0.87 & 0.60 & 0.38 & 0.70 & 1.00 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.33 & 0.01 & 0.02 & 0.05 & 0.58 & 0.34 & 0.51 & 0.95 \\ 
+   &  & Lognormal & 0.30 & 0.00 & 0.00 & 0.00 & 0.71 & 0.68 & 0.48 & 0.88 \\ 
+   &  & Weibull & 0.32 & 0.00 & 0.00 & 0.00 & 0.53 & 0.25 & 0.31 & 0.89 \\ 
+   \hline
+\end{tabular}
+\vspace{0.5cm}\\
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.63 & 0.43 & 0.87 & 1.00 & 0.63 & 0.45 & 0.94 & 1.00 \\
+   &  & Lognormal & 0.67 & 0.54 & 0.93 & 1.00 & 0.73 & 0.66 & 0.68 & 0.98 \\ 
+   &  & Weibull & 0.64 & 0.46 & 0.97 & 1.00 & 0.66 & 0.52 & 0.92 & 1.00 \\ 
+\cline{2-11}
+   &  & Exponential & 0.55 & 0.24 & 0.39 & 0.95 & 0.58 & 0.36 & 0.69 & 1.00 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.50 & 0.16 & 0.21 & 0.65 & 0.62 & 0.44 & 0.59 & 0.98 \\ 
+   &  & Lognormal & 0.61 & 0.41 & 0.58 & 0.92 & 0.77 & 0.79 & 0.34 & 0.78 \\ 
+   &  & Weibull & 0.47 & 0.11 & 0.13 & 0.48 & 0.56 & 0.34 & 0.44 & 0.95 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of models fits across family of TTDD when true detection probability is $\pdet = 0.65$.
+All data and inference models have mixture components.
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:sim2}
+\end{table}
+
+\begin{table}[ht]
+\footnotesize\centering
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\%  & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.86 & 0.83 & 0.25 & 0.59 & 0.75 & 0.38 & 0.64 & 1.00 \\
+   &  & Lognormal & 0.93 & 1.00 & 0.00 & 0.01 & 0.88 & 0.80 & 0.33 & 0.72 \\ 
+   &  & Weibull & 0.89 & 0.93 & 0.11 & 0.36 & 0.79 & 0.49 & 0.66 & 0.99 \\ 
+\cline{2-11}
+   &  & Exponential & 0.78 & 0.45 & 0.42 & 0.86 & 0.70 & 0.25 & 0.32 & 0.91 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.37 & 0.00 & 0.00 & 0.00 & 0.72 & 0.30 & 0.40 & 0.80 \\ 
+   &  & Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 0.85 & 0.77 & 0.24 & 0.65 \\ 
+   &  & Weibull & 0.38 & 0.00 & 0.00 & 0.00 & 0.65 & 0.16 & 0.22 & 0.54 \\ 
+   \hline
+\end{tabular}
+\vspace{0.5cm}\\
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.16 & 0.85 & 0.67 & 0.23 & 0.34 & 0.99 \\
+   &  & Lognormal & 0.74 & 0.33 & 0.51 & 0.98 & 0.81 & 0.55 & 0.59 & 0.97 \\ 
+   &  & Weibull & 0.66 & 0.17 & 0.22 & 0.87 & 0.70 & 0.27 & 0.46 & 1.00 \\ 
+\cline{2-11}
+   &  & Exponential & 0.61 & 0.11 & 0.12 & 0.52 & 0.64 & 0.21 & 0.24 & 0.85 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.64 & 0.09 & 0.10 & 0.41 & 0.77 & 0.49 & 0.46 & 0.93 \\ 
+   &  & Lognormal & 0.78 & 0.45 & 0.46 & 0.87 & 0.90 & 0.90 & 0.13 & 0.40 \\ 
+   &  & Weibull & 0.57 & 0.04 & 0.02 & 0.19 & 0.69 & 0.29 & 0.33 & 0.88 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of models fits across family of TTDD when true detection probability is $\pdet = 0.80$.
+All data and inference models have mixture components.
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:sim2}
+
+\vspace{0.5cm}
+
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\%  & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.94 & 0.30 & 0.38 & 0.83 & 0.91 & 0.25 & 0.31 & 0.77 \\
+   &  & Lognormal & 0.97 & 0.91 & 0.07 & 0.44 & 0.96 & 0.72 & 0.46 & 0.82 \\ 
+   &  & Weibull & 0.94 & 0.29 & 0.37 & 0.76 & 0.91 & 0.25 & 0.32 & 0.76 \\ 
+\cline{2-11}
+   &  & Exponential & 0.94 & 0.41 & 0.51 & 0.85 & 0.92 & 0.26 & 0.33 & 0.82 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.55 & 0.00 & 0.00 & 0.00 & 0.94 & 0.39 & 0.41 & 0.83 \\ 
+   &  & Lognormal & 0.47 & 0.00 & 0.00 & 0.00 & 0.97 & 0.80 & 0.27 & 0.65 \\ 
+   &  & Weibull & 0.57 & 0.00 & 0.00 & 0.00 & 0.89 & 0.08 & 0.09 & 0.43 \\ 
+   \hline
+\end{tabular}
+\vspace{0.5cm}\\
+\begin{tabular}{l|l|l|cccc|cccc}
+ \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
+ \multicolumn{3}{c}{ } & Med $p$ & Q($p$) & 50\% & 90\% & Med $p$ & Q($p$) & 50\% & 90\% \\ 
+  \hline
+\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.81 & 0.11 & 0.10 & 0.35 & 0.87 & 0.23 & 0.22 & 0.68 \\
+   &  & Lognormal & 0.92 & 0.28 & 0.38 & 0.86 & 0.95 & 0.55 & 0.57 & 0.92 \\ 
+   &  & Weibull & 0.82 & 0.10 & 0.15 & 0.34 & 0.87 & 0.22 & 0.28 & 0.68 \\ 
+\cline{2-11}
+   &  & Exponential & 0.82 & 0.09 & 0.09 & 0.38 & 0.88 & 0.20 & 0.25 & 0.65 \\ 
+\cline{2-11}
+& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}}  & Gamma & 0.91 & 0.09 & 0.10 & 0.42 & 0.97 & 0.79 & 0.26 & 0.63 \\ 
+   &  & Lognormal & 0.94 & 0.40 & 0.43 & 0.83 & 0.98 & 0.98 & 0.01 & 0.12 \\ 
+   &  & Weibull & 0.84 & 0.01 & 0.00 & 0.03 & 0.93 & 0.36 & 0.47 & 0.80 \\ 
+   \hline
+\end{tabular}
+\caption{Summary of models fits across family of TTDD when true detection probability is $\pdet = 0.95$.
+All data and inference models have mixture components.
+Med $p$: average across simulations of the posterior median of $\pdet$.  
+Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
+50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\end{table}
+
 \FloatBarrier

### 2017-04-03 Sims/SimZero/Outputs_Code_zero.R

@@ -2,8 +2,8 @@
 # As input, it needs: S0_simtable.csv, "Outputs_p##.Rdata", "Simpars_p##.Rdata"
 
 # Specify which pdet we're analyzing
-truepdet <- 0.65
-fileID   <- "65"
+truepdet <- 0.5
+fileID   <- "50"
 
 library(ggplot2)
 library(reshape2)
@@ -22,7 +22,7 @@ data.handle <- function(df){
   return(df)
 }
 
-simtable <- read.csv("S0_simtable.csv")     # Load simtable
+simtable <- read.csv("S0_simtable.csv")    # Load simtable
 load(paste0("Outputs", fileID, ".Rdata"))  # Load model fits
 for(i in 1:length(outputs)) assign(names(outputs)[i], outputs[[i]]) # Convert model fits from list into objects
 
@@ -51,7 +51,8 @@ p_glob$MixPeakLabel  <- with(p_glob, paste(datamixlabel, datapeaklabel))      # 
 # Posterior estimates of detection probabilities -- By mix/non-mix and by family
 CD.melt <- melt(subset(count.det, parameter=="p_global"), id.vars=11:ncol(count.det)) # Exclude 'uncounted' estimates
 CD.melt <- CD.melt[,-which(names(CD.melt) %in% c("indx"))]
-pglobs  <- join(true.pglobal, simtable)  # Construct a table of true pdet values with same variables as posterior estimates
+pglobs  <- join(true.pglobal, simtable)  # Construct a table of "true pdet values" with same variables as posterior estimates
+                                         # These pdet values are realizations from each simulated dataset
 pglobs$parameter <- "p_global"
 pglobs$variable  <- "True"
 CD.melt <- rbind(CD.melt, pglobs)
@@ -99,12 +100,8 @@ print(ggplot(data=subset(tempS2, variable %in% c("X2.5.","X97.5."))) +
 
 # There used to be a caterpillar plot of just lognormal data/models to provide an example for the Oral Prelim
 
-print(ggplot(data=subset(tempS2, variable %in% c("X2.5.","X97.5."))) + 
-        geom_line(aes(x=value, y=as.factor(Rep), group=Rep)) + 
-        theme(axis.title.x=element_blank(),
-              axis.text.x=element_blank(),
[…]

### 2017-04-06 Draft.tex
Updates to manuscript.

@@
 
+\renewcommand{\floatpagefraction}{0.8}  % Changes settings for when a figure hogs a page all to itself
+
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
@@
 \newcommand{\Fm}{F_T^{(M)}}
+\newcommand{\dt}{\text{det}}
+\newcommand{\ftdt}{f_{T|\dt}(t|\dt)}
 \newcommand{\nobs}{n^{(obs)}}
@@
 % Johnson2008: all birds present are available for detection
-A typical assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007, LeeMarsden2008}.  
+A standard assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007, LeeMarsden2008}.  
 In particular, animal behaviors such as intermittent singing in birds and frogs or diving in whales \citep{Scott2005, Diefenbach2007, Reidy2011}, differences in behavior across subgroups of animals \citep{Otis1978, Farnsworth2002}, observer impacts on animal behaviors \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}, and variations in observer effort, e.g.
@@
 Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown and may be the primary quantity of interest.  
-We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture}, which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
+We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture} --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
 For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
@@
 The latter is not possible when detection rates are not constant.  
-Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
+Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf (Figure \ref{fig:schematic}), we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.
 Indeed our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine equal-duration intervals \citep{Reidy2011, Amundson2014}.
@@
 \caption{\label{fig:schematic} 
-Illustration of fitting mixture exponential (left) and mixture gamma (right) time-to-detection distributions (TTDDs) to multinomial removal-sampled observations (center) assuming no survey-level heterogeneity.
-The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component (light gray rectangle), which constitute $\gamma$ and $1-\gamma$ proportions of the population, respectively.
+Illustration of fitting mixture exponential (left) and mixture gamma (right) time-to-detection distributions (TTDDs) to interval-censored removal-sampled observations (center).
+The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component of first-interval detections (light gray rectangle), constituting $\gamma$ and $1-\gamma$ proportions of the population, respectively.
 We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving an estimated proportion $1-\pdet$ undetected.
@@
 
-\adam{I think we start by describing our joint likelihood: p(counts $| N, \pdet$) and p(times $|$ counts,TTDD) and prior.
-We explain that, in isolation, $N$ and $\pdet$ are not identifiable in the first piece, which is why we use the second piece to inform our estimates of $\pdet$.
-Then, in the body of section \ref{sec:exact_time}, we introduce a second paragraph which defines our distribution for $\nobs$ and completes the data likelihood.
-The discussion of identifiability would travel with the distribution of $\nobs$ and be cleaned up a little.
-Perhaps even split the discussion of TTDDs into another subsection.
-}
-
 Before considering interval censoring and explanatory variables, we first present the scenario of exact times to detections with no explanatory variables. 
@@
 
-\subsection{Exact time to detection}\label{sec:exact_time}
+\subsection{Distributions for exact times to detection}\label{sec:exact_time}
 
@@
 Assuming detection times for all individuals at a survey are independent, identically distributed according to a common time-to-detection distribution (TTDD), we define $T_{sb}$ as a random variable with cumulative distribution function (cdf) $F_T(t)$ and probability density function (pdf) $f_T(t)$.
-In reality, times to first detection are often truncated due to a finite survey length of $C$, meaning that each individual has a detection probability $\pdet=F_T(C)$.  
-The conditional distribution of observed detection times consequently has pdf $f_{T|det}(t)= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|det}(t) = \int_0^t f_{T|det}(x) dx$, and instantaneous detection rate, or hazard function, $h(t) = f_T(t) / [1-F_T(t)]$.
-%$T_{sb|det} \ind f_{T|det}(t_{sb})$ where $T_{sb|det}$ is observed only for those $n_s^{(obs)}$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$.
+In practice, times to first detection are truncated due to a finite survey length of $C$, meaning each individual has detection probability $\pdet=F_T(C)$.  
+The conditional distribution of observed detection times consequently has pdf $\ftdt= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|\dt}(t|\dt) = \int_0^t f_{T|\dt}(x|\dt) dx$, and instantaneous detection rate, or hazard function, $h(t) = f_T(t) / [1-F_T(t)]$.
+%$T_{sb|\dt} \ind f_{T|\dt}(t_{sb}|\dt)$ where $T_{sb|\dt}$ is observed only for those $\nobs_s$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$.
 
 A common choice for TTDD is an exponential distribution, i.e. $T_{sb}\ind \mbox{Exp}(\varphi)$, which imposes a constant first detection rate, i.e. $h(t) = \varphi$.
-Choosing another TTDD can allow for a systematic non-constant detection regime. 
-For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) organisms acclimate and gradually return to constant detection, a gamma TTDD could be appropriate.
-%In addition to an exponential and gamma TTDD, we also consider Weibull and lognormal as these distributions are often used in survival analysis. 
-Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
-All three may be reasonable TTDD choices if, for instance, detection rates vary among individuals.
-This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population.
+By choosing another TTDD, we can allow for a systematic non-constant detection regime. 
+For example, to model an observer effect where: (i) the observer's arrival suppresses or stimulates detectable cues, but (ii) individuals acclimate and gradually return to constant detection, a gamma TTDD would be appropriate.
+Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to increase or decrease during the survey.
+All three TTDDs may provide reasonable empirical approximations of non-constant detection, though the shapes of the distributions differ, potentially leading to differing inference.
+For instance, when detection rates vary across individuals, the result is a marginal detection rate that decreases over time.
+Whether the marginal rate is best approximated by a gamma, lognormal, Weibull, or some other TTDD depends on just how rates vary across individuals.
 
@@
 The exponential distribution is a special case of both the gamma and Weibull distributions when $\alpha=1$. 
-We employ a log link for $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time.
+We employ a log link to model $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time.
 
-We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $n_{s}^{(obs)} \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
-Using maximum likelihood, we can estimate the parameters in a TTDD from exact times to first detection and thus estimate $\pdet$ and its uncertainty.
-With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
-In these models, point estimates of $N_s$ are unstable unless $\pdet>0.41$ \citep{Olkin1981}.
-One approach to regularizing these estimates is to construct a hierarchical model for the site-specific abundance, e.g. $N_s\ind \Po(\lambda)$ \citep{Raftery1988, Royle2004NMixture}. 
-With this assumption, we can decompose $N_{s}$ into observed and unobserved portions: $n^{(obs)} \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
-Although alternative distributions could be considered, e.g.
+\subsection{TTDDs in an N-mixture model}
+
+A basic N-mixture model describes observed survey-level abundance $\nobs$ with a hierarchy where $\nobs_s \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$ and $N_s \ind Po(\lambda)$.
+We can decompose $N_{s}$ into observed and unobserved portions: $\nobs \ind Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \ind \Po\left(\lambda[1-\pdet]\right)$.
+Although alternative distributions can be considered, e.g.
+negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed, and thus we will use the Poisson assumption here.
+
+The above definitions complete our exact-time homogenous-survey data model, consisting of distributions for counts and observed detection times:
+\begin{align}
+\nobs_s &\ind \Po(\lambda \pdet) \nonumber\\
+p(\textbf{t}_s) &= \prod\limits_{b=1 \dotso\nobs_s} f_{T|\dt}(t_{sb} | \det, \alpha, \varphi) \label{eq:simplemodel}\\
+\pdet &= F_T(C|\alpha,\varphi) \nonumber
+\end{align}
+where $\textbf{t}_s$ is a vector of observed times at survey $s$.
+
+%We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $\nobs_s \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
+%Using maximum likelihood, we can estimate the parameters of a TTDD from the observed times to first detection and thereby estimate $\pdet$ and its uncertainty.
+%With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
+%In these models, point estimates of $N_s$ are unstable unless $\pdet>0.41$ \citep{Olkin1981}.
+%One approach to regularizing these estimates is to construct a hierarchical model for site-specific abundance, e.g. $N_s\ind \Po(\lambda)$ \citep{Raftery1988, Royle2004NMixture}. 
+%Assuming Poisson abundance, we can decompose $N_{s}$ into observed and unobserved portions: $\nobs \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
+%Although alternative distributions could be considered, e.g.
 negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed, and thus we will use the Poisson assumption here.
@@
 Let $C_i$ for $i=1,\dots,I$ indicate the right endpoint of the $i$th interval then $C_I$ is the total survey duration and, letting $C_0=0$, the $i$th interval is $(C_{i-1},C_{i}]$. 
-Let $n_{si}$ be the number of individuals counted during interval $i$ on survey $s$, $n_{s}^{(obs)} = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
-Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(n_{s}^{(obs)}, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and is calculated from the TTDD: $p_{si} = F_{T|det}(C_i) - F_{T|det}(C_{i-1})$.
+Let $n_{si}$ be the number of individuals counted during interval $i$ on survey $s$, $\nobs_s = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(\nobs_s, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and is calculated from the TTDD: $p_{si} = F_{T|\dt}(C_i|\dt) - F_{T|\dt}(C_{i-1}|\dt)$.
 
@@
 \begin{align}
-n_s^{(obs)} &\ind \Po(\lambda_s \pdet_s) \nonumber\\
-\vn_s &\ind \Mult(n_s^{(obs)}, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
+\nobs_s &\ind \Po(\lambda_s \pdet_s) \nonumber\\
+\vn_s &\ind \Mult(\nobs_s, \vp_s); \qquad \vp_s = (p_{s1},\dots,p_{sI}) \nonumber\\
 \pdet_s &= F_T(C_I|\alpha,\varphi_s) \nonumber \\
@@
 
-We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2015}.  
+We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2016}.  
 %Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
@@
 
-We focus inference on the overall detection probability across all surveys: $\pdet = \sum\limits_{s}n_{s}^{(obs)}\big/$ $\sum\limits_{s}N_{s}$.  
+We focus inference on the overall detection probability across all surveys: $\pdet = \sum\limits_{s}\nobs_s \big/$ $\sum\limits_{s}N_{s}$.  
 Within simulation studies, we compared model estimates to true $\pdet$ values by percent coverage and by posterior p-values --- the proportion of MCMC samples having a smaller value than the true value.
@@
 Detection rates $h(t)$ of peaked distributions generally increase over time, while detection rates of nonpeaked distributions generally decrease over time.
-More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked.
+More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a nonpeaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked.
 
@@
 
-To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 16 intercept-only datasets from each of 14 TTDDs: each combination of peaked/non-peaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal, where all exponential models are non-peaked.
-We chose the number of surveys (381) and true parameter values (Table \ref{S-simvals}) to mimic values from the Ovenbird analysis (Section \ref{sec:ovenbirds}).  
-In particular, we set parameters such that (i) the overall expected detection probability was 0.80, (ii) for mixture datasets, $\gamma = 0.65$ (meaning 35\% of individuals were immediately detected), (iii) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (iv) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
+To assess the need for incorporating a mixture component to increase the probability of detection in the initial interval as discussed in Section \ref{s:interval}, we simulated 5600 intercept-only datasets: 100 replicates using 4 values of $\pdet$ (0.50, 0.65, 0.80, and 0.95) from each of 14 TTDDs (each combination of peaked/nonpeaked, mixture/non-mixture, and exponential/gamma/Weibull/lognormal, where exponential models are considered nonpeaked).
+We chose true parameter values (Table \ref{S-simvals}) and the number of surveys (381) to mimic the Ovenbird analysis (Section \ref{sec:ovenbirds}).  
+In particular, we set parameters such that (i) in nonpeaked models, 70\% of \textit{detected} individuals were observed during the first two minutes, and (ii) in peaked models, the detection mode for `hard to detect' individuals occured at 5 minutes.
 
-Each dataset was fit with two models: mixture and non-mixture version of the distribution family, e.g. exponential, used to simulate the data.
-For each dataset-analysis combination, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effective samples for all parameters.
+We fit each dataset with two models: mixture and non-mixture versions of the distribution family, e.g. exponential, used to simulate the data.
+For each dataset-analysis combination, we sampled $>$ 90,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effective samples for all parameters.
 
-We summarized the analysis by calculating the average, across the simulations, of the posterior median of $\pdet$ (Med.
-$p$), the average proportion of the posterior distribution of $\pdet$ larger than the truth (Q($p$)), and average coverage for 50\% and 90\% credible intervals.
-If the analyses are providing a reasonable estimate of $\pdet$, we would expect Med.
-$p$ to be around the true value of 0.8, Q($p$) to be around 0.5 indicating that half of the posterior is above and below the true value of 0.8, and the coverage to be close to their credibility.
-Table \ref{tbl:sim1} provides a summary of these quantities.
-When a mixture model is used to simulate the data (bottom half of the table), there is clearly a benefit to using a mixture model for inference.  
-Using a non-mixture model for inference, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ (Q($p$) $\approx 1$) and the other models underestimating.
-When a non-mixture model is used to simulate the data (top half of the table), there are no clearly discernable differences between the ability of a non-mixture or mixture model to capture $\pdet$. 
-For nonpeaked non-mixture data (top 4 lines), the mixture model is less biased for lognormal and Weibull scenarios (Q($p$)$\approx 0.5$), and CIs are 15-40\% narrower (results not shown).
+For each dataset and inference model combination, we summarize the analysis across simulations by averaging the posterior median and reporting coverage for 90\% credible intervals.
+If analyses are providing reasonable estimates of $\pdet$, we expect the average median to be unbiased and the coverage to be close to 90\%.
+Figure \ref{fig:mixture_fig} provides a summary of these quantities. 
+When a mixture model is used to simulate the data (lower row of plots), there is clearly a benefit to using a mixture model for inference.  
+Using a non-mixture model for inference, the credible interval coverage is near zero for most models with the exponential model overestimating $\pdet$ and the other models underestimating. 
+When a non-mixture model is used to simulate the data (upper row of plots), there are no clearly discernable differences between the ability of a non-mixture or mixture model to capture $\pdet$. 
 These results support the general default use of a mixture model over a non-mixture model.
+
+For nonpeaked datasets, estimates of $\pdet$ from the same TTDD inference model differed by only 1-5\% between $\pdet=0.50$ and $\pdet=0.65$ simulations (Tables \ref{S-tbl:mix50} to \ref{S-tbl:mix95}), and credible intervals had roughly the same widths.
+This suggests that, when data are nonpeaked with true detection probabilities less than 65-80\%, the patterns of detections over time are insufficient for distinguishing between moderate and low values of $\pdet$.
+In these cases, mixture inference models estimated higher detection probabilities than did non-mixture models.
 
@@
 \caption{
-Comparisons of data and models with and without mixture TTDDs.
+Comparisons of mixture and non-mixture TTDDs.
 \textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.
 Dashed lines show unbiased estimation.
-\textit{Right}: Coverage of 50\% credible intervals across 100 replicate simulations.
+\textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.
 Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.
@@
 \textit{Columns}:  Data simulated from nonpeaked (left), exponential (center), or peaked (right) TTDDs.
-In these simulations, each inference model was only fit to datasets from the same TTDD family (e.g.
+Each inference model was fit only to datasets from the same TTDD family (e.g.
 lognormal to lognormal).
@@
 Now we turn to model misspecification of the distribution family. 
-We simulated 16 intercept-only datasets from the 7 different mixture TTDD models using the same settings as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
-%Again, we ran 60,000 iterations which showed no evidence of lack of convergence according to the Geweke diagnostic and reached over 1,000 effect samples for all parameters.
+We simulated 100 replicates of intercept-only datasets from the 7 different mixture TTDD models using the same detection probabilties and parameters as in the previous section, and we fit them with mixture models from each of exponential, gamma, lognormal, and Weibull families.
+
+\begin{figure}\centering
+\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"}
+\caption{\label{fig:sim2typical} 
+Representative examples of posterior distributions for $\pdet$ (left panel) and survey-level log-scale abundance (right panel).
+Data are simulated from nonpeaked gamma (left column), exponential (center), and peaked gamma (right) mixture TTDDs.
+True detection probabilities vary by row and are shown for comparison (dashed vertical line).
+Inference models are either exponential mixture (light gray) or gamma mixture (dark gray) TTDDs.
+}
+\end{figure}
+
+Figure \ref{fig:sim2typical} illustrates representative examples of posterior distributions for $\pdet$ and site-level log(abundance) when data and models were from the exponential and gamma mixture TTDDs.
+Posterior distributions under exponential inference models accurately captured true detection probabilities when the simulation model was exponential, but they overestimated (underestimated) detection probabilities when the simulation model was nonpeaked (peaked).
+In contrast, gamma family posteriors, which have added flexibility from having a shape parameter, accurately captured the truth in most scenarios although with increased uncertainty.
+
+Figure \ref{fig:family_fig} and Tables \ref{S-tbl:fam50} to \ref{S-tbl:fam95} summarize posterior estimates of $\pdet$ and coverage from 90\% credible intervals across all TTDD families.
+The poorest estimation of $\pdet$ occurred for the exponential inferential model when the data simulation model had a peak, because the two parameters (rate and mixing parameter) did not provide enough flexibility to adequately fit a TTDD with both an initial increase and a delayed mode.
+As a result, the exponential model underestimated the actual detection probability.
+In contrast, the exponential model typically overestimated detection probability for nonpeaked simulated data.
+However, exponential model estimates were both less biased and more precise when the data actually derived from an exponential mechanism.
 
@@
 Dashed lines show unbiased estimation.
-\textit{Right}: Coverage of 50\% credible intervals across 100 replicate simulations.
+\textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.
 Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.
-\textit{Rows}: Denote the TTDD family used to simulated data.
-\textit{Columns}:  Data simulated from nonpeaked (left) or exponential/peaked (right) TTDDs.
-In these simulations, we used only mixture datasets and inference models.
+\textit{Rows}:  Data simulated from mixture exponential, gamma, lognormal, or Weibull TTDDs.
+\textit{Columns}: Data simulated from nonpeaked or exponential (left) or peaked (right) TTDDs.
+All data and inference models used mixture TTDDs.
 }
 \label{fig:family_fig}
-\end{figure}
-
-Figure \ref{fig:sim2typical} presents a representative example of posterior distributions for $\pdet$ for data and models from the exponential and gamma mixture families. 
-In this example, the posterior distribution under an exponential inference model accurately captures the true detection probability when the simulation model is an exponential, but overestimates (underestimates) the detection probability when the simulation model has a small (large) gamma shape parameter.
-In contrast, the posterior from a gamma family, which has the additional flexibility of the shape parameter, is able to accurately capture the truth in all scenarios with increased uncertainty.
-
-To summarize these findings across the various simulation and inference models, Table \ref{tbl:sim2} provides simulation-averaged Median estimates of $\pdet$ (Med $p$), proportion of the $\pdet$ posterior larger than the truth ($Q(p)$), and 50\% and 90\% coverage proportions.
-Each table quadrant assesses a model used for inference compared to the 7 different models used for simulation. 
-The poorest estimation of $\pdet$ occurs for the exponential inferential model when the model used for simulation has a peak, because the two parameters (rate and mixing parameter) do not provide enough flexibility for the mixture exponential distribution to adequately fit a TTDD with both an initial increase and a delayed mode.
-In this situation, the exponential model underestimates the actual detection probability.
-In contrast, if the simulated data are non-peaked, the exponential model typically overestimates the actual detection probability. 
-However, exponential model estimates are both less biased and more precise when the data actually do derive from an exponential mechanism.
-
-\begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"}
-\caption{\label{fig:sim2typical} 
-Representative examples of posterior distributions for $\pdet$ (left panel) and survey-level log-scale abundance (right panel).
-Data are simulated from nonpeaked gamma (left column), exponential (center), and peaked gamma (right) mixture TTDDs.
-True detection probabilities are shown for comparison (dashed vertical line) and vary by row.
-Inference models are either exponential mixture (light gray) or gamma mixture (dark gray) TTDDs.
-}
 \end{figure}
@@
 
-When comparing the different three-parameter TTDDs, model misspecification is not as serious an issue because these models can adequately account for the interval-censored times to detection. 
-Nonetheless, it appears the gamma mixture model is better able to account for data from most non-peaked datasets, while gamma and Weibull mixture models have comparable performance across peaked datasets.
-In the gamma quadrant of Table \ref{tbl:sim2}, the average posterior median is near the truth of 0.8, the average proportion of the posterior distribution of $\pdet$ is near 0.5, and the credible intervals have coverage near their credibility.
-In contrast, when the lognormal model is used for inference, it performs worse across nearly all data types.
-However, gamma models required $\sim 10$ times the computational time as did lognormal and Weibull models and almost 25 times that of exponential models.
+When comparing the different three-parameter TTDDs, model misspecification was not as serious an issue because the models could better account for the patterns in time to detection.
+Even so, estimates of $\pdet$ amongst the three models differed by as much as 0.15.
+As in the previous simulation study, estimates of $\pdet$ from nonpeaked datasets changed little as true values of $\pdet$ decreased below 65-80\%.
+For nonpeaked datasets, gamma TTDDs produced larger estimates of $\pdet$ than did Weibull TTDDs, with lognormal TTDDs producing the lowest of all.
+For peaked datasets, the order of Weibull estimates were larger than gamma estimates.
+Our results do not favor the use of one of these TTDDs over the others.
 
@@
 The previous sections studied effects of time to detection assumptions in the context of no explanatory variables.
-We now incorporate fixed and random effects for abundance and detection. 
+We now incorporate fixed and random effects for abundance and detection to ascertain whether models differ in their estimates of effect sizes.
 We simulated data from each of the 7 mixture TTDDs and fit models from exponential, gamma, lognormal, and Weibull mixture models.
+
 We simulated data using the median posterior parameter estimates obtained in the analysis of Ovenbird data in Section \ref{sec:ovenbirds}.
-Because the fitted Ovenbird models did not yield peaked distributions, we simulated peaked datasets by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data in the previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
+As those models differed in their estimates of $\pdet$, so the simulated datasets featured different true values of $\pdet$.
+Because fitted Ovenbird models did not yield peaked distributions, we simulated peaked data by: (i) using the same intercepts, shape parameters, and mixing parameters as for peaked data ($\pdet = 0.80$) in previous simulations, (ii) using median covariate and random effects from the Ovenbird estimates, and (iii) scaling the detection intercept and random effect to achieve true detection probabilities $\approx 0.8$ with a detection mode at 5 minutes.
 See Table \ref{S-tbl:sim3} for actual parameter values.
-Due to the computation time involved in estimating models with these fixed and random effects, we simulated each TTDD only once.
-To obtain reasonable convergence diagnostics and effective sample sizes, these analyses ranged from 250,000-375,000 iterations.
 Because of the difficulty in integrating random effects over all sites, approximate posterior distributions for the study-wide marginal $\pdet$ were obtained by simulating data from each MCMC sample and calculating the proportion of simulated Ovenbirds that were observed.
 
+Computation times for this simulation study were much greater than for the other studies because partitions of the cdf, e.g.
+$\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
+Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
+Due to the computation times involved, we fit each model only once.
+
 The results from this simulation are qualitatively similar to the Ovenbird analysis (Figure \ref{ovenposteriors}) and thus we only briefly review the results here and provide the corresponding figures and tables in the Supplementary Material.
-Patterns in posterior estimates of site-specific detection probabilities, $\pdet_s$, with respect to mixture and family TTDD forms were the same as in the previous simulation studies -- the inclusion of explanatory variables did not make models more robust to violations of mixture- and constant-detection assumptions (Figures \ref{S-pdet_cater_correct} and \ref{S-pdet_cater_family}).
-Posteriors for abundance fixed and random effects were the same regardless of what TTDD was assumed (Section \ref{S-s:sim3} and figures therein).  
-Posteriors for the mixing parameter $\gamma$ and detection fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
+Patterns in posterior estimates of overall detection probabilities with respect to family TTDD forms were largely the same as in the previous simulation studies for appropriate values of $\pdet$ -- the inclusion of explanatory variables did not make models more robust to violations of constant-detection assumptions (Figures \ref{S-pdet_cater_family}).
+Posteriors for abundance-related fixed and random effects were the same regardless of which TTDD was assumed (Section \ref{S-s:sim3} and figures therein).  
+Posteriors for the mixing parameter $\gamma$ and detection-related fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
 %It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.
@@
 
-We formulated a model for interval-censored time to detection data that allows for non-constant detection rates.
+%We formulated a model for interval-censored time to detection data that allows for non-constant detection rates.
+We formulated a model for single-species removal-sampled point-count survey data that allows for non-constant detection rates.
+The model accommodates both interval-censored and exact times to detection.
 Our model adopts a time-to-event approach within a hierarchical N-mixture framework, and it allows times to first detection to be modeled according to flexibly defined TTDD families.
@@
 Because the exponential TTDD is a special case of both gamma and Weibull TTDDs, we can interpret the differences in estimation between models as resulting from the information conveyed by the assumption of constant detection.
-%Phrased differently, in the Bayesian context, the exponential model is equivalent to a gamma or Weibull model where the prior for the shape parameter $\alpha$ is defined: Pr($\alpha=1) = 1$; and the differences between the models results from the strength of this prior.
-Among the TTDDs modeled in our study, the gamma mixture TTDD gave the most accurate estimates across data distributions, though the Weibull mixture TTDD was comparable for peaked datasets.
+%Among the TTDDs modeled in our study, the gamma mixture TTDD gave the most accurate estimates across data distributions, though the Weibull mixture TTDD was comparable for peaked datasets.
 
 We have additionally demonstrated for non-constant models the utility of using a mixture TTDD formulation.
-Inference models with a mixture component are accurate whether the data have a mixture or not, whereas inference models without the mixture are badly biased when the data do feature a mixture.  
-Mixture models even outperform non-mixture models when the data are non-mixture and nonpeaked.
-%For a discussion of mixtures as heterogeneity, see Pledger 2000, dating all the way back to Otis 1978.
+Inference models with a mixture component are accurate under most scenarios whether the data have a mixture or not, whereas inference models without the mixture can be badly biased when the data do feature a mixture.
 
@@
 
-If the estimation of abundance is the primary interest, then the choice of TTDD has large consequences, and we may reasonably ask whether removal sampled point-count surveys are adequate for the purpose.
-The strategy behind removal sampling is to use the pattern of detections over time to estimate the proportion of individuals that would be detected if only the observation period lasted longer.  
-As such, it is entirely based upon extrapolation and correctly estimating the probability in the tail of the TTDD based on recorded observations.
-An assumption of constant detection places constraints on the amount of uncertainty in the extrapolated tail but at the cost of potentially sizable bias.
-When we allow for non-constant detection, bias is less problematic but estimation of the tail probability becomes less certain.
+If the estimation of abundance is the primary interest, then the choice of TTDD has large consequences, and we may reasonably ask when removal sampled point-count surveys are adequate.
+Based on our results, we would be skeptical of abundance estimates that indicate a nonpeaked TTDD with median posterior estimates of $\pdet$ below 0.75 (using a mixture Weibull as a yardstick).
+Reduced model performance in the presence of low detection probabilities is common among abundance models.
+We recommend the discussion in \citet{CoullAgresti1999}, elaborating key points here.
+The essential problem is a flat log-likelihood.
+For single-pass surveys, we cannot rely on repeated counts to improve our estimates.
+Instead, the strategy of removal sampling is to use the observed pattern of detection times to estimate the proportion of individuals that would be detected if only the observation period lasted longer.
+As such, it is entirely reliant upon extrapolation based on an accurate fit of $\ftdt$.
+When true detection probabilities are low, $\ftdt$ for nonpeaked data becomes flat.
+Such a flat observed pattern provides little information and is well approximated by a wide variety of TTDDs, which vary greatly in their tail probabilities.
+An assumption of constant detection limits the shape of $\ftdt$ and thereby constrains uncertainty but at the cost of potentially sizable bias.
 
@@
 Consequently, our application of a TTDD represents an averaging across distance classes, which induces systematic bias in estimates of abundance \citep{EffordDawson2009, Laake2011, Solymos2013}.  
-To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \citeauthor{Borchers2016}.  
+To be consistent with the continuous time-to-event approach, distance can be incorporated into the detection model as an event-level modifier as is done in \cite{Borchers2016}.  
 This approach is distinct from earlier integrations of removal- and distance sampling, where distance has been treated as an interval-/survey-level modifier \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}.

### 2017-04-06 ReviewerComments.tex
Updates to manuscript.

@@
 \begin{quote}
-We have added the following in Section \ref{M-sec:exact_time} to clarify: ``We employ a log link for $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time."
+We have added the following in Section \ref{M-sec:exact_time} to clarify: ``We employ a log link to model $\varphi$, and therefore our model is equivalent to a generalized linear model with a log link on the mean detection time."
 
@@
 \begin{quote}
-\adam{Address this point}
+\adam{Address this point in the abstract}
 \end{quote}
@@
 \begin{quote}
-We have added the following sentences to \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to either increase or decrease during the survey; however, the shapes of these distributions differ, most notably in tail probability.
-All three may be reasonable TTDD choices if, for instance, detection rates vary among individuals.
-This leads to marginal detection rates that decrease over time as individuals with higher detection rates are removed from the population."
+We have added the following sentences to \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to increase or decrease during the survey.
+All three TTDDs may provide reasonable empirical approximations of non-constant detection, though the shapes of the distributions differ, potentially leading to differing inference.
+For instance, when detection rates vary across individuals, the result is a marginal detection rate that decreases over time.
+Whether the marginal rate is best approximated by a gamma, lognormal, or Weibull TTDD depends on just how rates vary across individuals."
 
@@
 We agree that an intercept-only model is simplistic and will not be of much practical use in an applied setting.
-However, the intercept-only simulation is informative, because models that perform poorly in simple applications can hardly be trusted in more complex applications.
+However, the intercept-only simulation is informative all the same, because models that perform poorly in simple applications can hardly be trusted in more complex applications.
 
-Meanwhile, a simulation study involving effects on abundance and detection is computationally expensive. \adam{Reference exact duration required}.
+Meanwhile, as disscussed later, a simulation study involving effects on abundance and detection is computationally expensive.
 Fortunately, while simulation studies are hampered by long computation times, the typical user of this model will only apply it to a small number of datasets, so computation time will not be prohibitive.
@@
 \begin{quote}
-\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
+Average run times for the different simulation studies are now discussed in the second paragraph of Section \ref{M-sec:simfull}.
 \end{quote}
@@
 We agree that using both is redundant.
-The two measures are largely the same, and we note where they differ in Sections \ref{M-sec:mixture} and \ref{M-sec:family}.
-We now choose to report only 50\% credible intervals.
-This choice more effectively highlights over/undercoverage in figures \ref{M-fig:mixture_fig} and \ref{M-fig:family_fig}.
-
+We now choose to report only 90\% credible intervals.
 \end{quote}
@@
 This has been made possible by improvements to Stan \citep{Rstan2016} and by streamlining our code for homogeneous surveys.
-\adam{Add reference to summary of run times vis-a-vis the third simulation study}
+See the second paragraph of Section \ref{M-sec:simfull} for a summary of run times for the models currently.
 
@@
 We define our use of `mixture model' later in the paragraph:  ``We apply the term `mixture model' to any model with a mixture TTDD."
-This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso], which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
+This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso] --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
 \end{quote}
@@
 \adam{
-This potentially requires a lot of rewiring.
-Two options are: (i) highlight each component of Equation (1) as we address it in the text, because we do address them all, or (ii) at the end of Section 3.2, provide a simpler summary of the model without mixtures and without covariates.
-I do think we provide the model pieces one-by-one, but we could be more heavy-handed (in a good way).
+To avoid potentially a lot of rewriting, I think we should add an abbreviated version of Equation (1) at the end of Section 3.2.
+I am struggling with this at present.
 }
-\jarad{But it appears the reviewer wants it up front. 
-I need to go back and reread to determine if this seems feasible. 
-I thought we did start without mixtures and without covariates.}
 \end{quote}
@@
 We have split the cited paragraph and a few others.
-\adam{The only times I feel I have cited many articles en masse is when I've said ``Many authors have done this.'}
 \end{quote}
@@
 We have added this as Figure \ref{M-fig:schematic}.
-\adam{I have not cited the figure in the text.  I am not convinced I need to.}
 \end{quote}
@@
 \begin{quote}
-\adam{
-The main application I've encountered before is \citet{Borchers2013}.
-There are technical differences between that approach and our parameteric survival analysis approach, mainly due to the continuous and interval-censored handling of time.
-For instance, I suspect that for analysis, as with \citet{Farnsworth2002} and its derivatives, the hidden Markov model would require an artificial division of the observation period into equal-intervals intervals.
-One cool feature of that approach is that it directly handles bout-singing through an availability state (which may be stochastically `available' or `not available') in conjunction with a detection model.
-I am not sure that model would serve well for a problem like ours where we are looking for systematic non-constant detection --- i.e., the Markov model would require time-dependent transition matrices.
-Additionally, at least in the case of \citet{Borchers2013}, the model utilizes separate observations of animal behaviors to estimate transitions among hidden states (e.g., moving/singing/etc.).
-In their case, they analyzed radio-tagged whales to estimate transitions for diving states.
+The reviewer's comments are thought-provoking but ultimately lead to a set of analytical strategies that are different than those we investigate in our manuscript.
 
-Those are just some preliminary thoughts.
-I think the reviewer's comment is thought-provoking but not applicable to the manuscript we have crafted.}
+Within the world of survey sampling, one of the better known Markov models is \citet{Borchers2013}.
+They model marine mammal surveys, where the biological states correspond to depth zones of diving, and only mammals in the surface state are detectable.
+One strength of their approach is how easily it handles transitions between behavioral states during the survey.
+For non-constant detection in avian point-count surveys, Markov models appear particularly well-suited to account for bout singing --- i.e., when a bird alternates between periods of calling and silence.
+Our parametric survival analysis only handles bout singing marginally.
+
+However, for other forms of non-constant detection (e.g. observer effects), it is not clear that Markov models offer an advantage.
+A Markov model would feature presumably at least three classes (unavailable, available, and detected), with the chief focus of inference being estimation of the time-dependent transition matrix.
+Assumptions about the form of that matrix would pose similar modeling challenges to those we faced in parametric survival analysis.
+Markov models do enable the incorporation of supplementary behavioral observations --- for example, \citet{Borchers2013} use data from radio-tagged whales to estimate transition matrices --- but we have no such data to supplement our avian counts.
+In one regard, the Markov model is a little more inelegant in that it requires discrete-time observations, whereas the parametric survival analysis in our presentation can utilize exact time-to-detection data.
+
+We reiterate: the comments about Markov models are thought-provoking but are ultimately something different than what we examine in this manuscript.
 \end{quote}

### 2017-04-06 Supplement.tex
Updates to manuscript.

@@
 \usepackage{setspace}
+\usepackage{url}
+\usepackage{xr}
 \usepackage{listings}
@@
 \usepackage{rotating}
+
+\externaldocument[M-]{Draft}
 
@@
   \hline
-  Gamma & Non-mixture & Non-peaked &  & -3.279 & 0.257 \\ 
+  Gamma & Non-mixture & Nonpeaked &  & -3.279 & 0.257 \\ 
   & & Peaked &  & -0.746 & 3.371 \\ 
-  & Mixture & Non-peaked & 0.65 & -2.773 & 0.577 \\ 
+  & Mixture & Nonpeaked & 0.65 & -2.773 & 0.577 \\ 
   & & Peaked & 0.65 & -1.210 & 2.491 \\ 
   \hline
-  Lognormal & Non-mixture & Non-peaked &  & -0.341 & 2.330 \\ 
+  Lognormal & Non-mixture & Nonpeaked &  & -0.341 & 2.330 \\ 
   & & Peaked &  & -1.872 & 0.512 \\ 
-  & Mixture & Non-peaked & 0.65 & -1.462 & 1.674 \\ 
+  & Mixture & Nonpeaked & 0.65 & -1.462 & 1.674 \\ 
   & & Peaked & 0.65 & -1.992 & 0.618 \\ 
   \hline
-  Weibull & Non-mixture & Non-peaked &  & -1.165 & 0.418 \\ 
+  Weibull & Non-mixture & Nonpeaked &  & -1.165 & 0.418 \\ 
   & & Peaked &  & -2.042 & 1.829 \\ 
-  & Mixture & Non-peaked & 0.65 & -2.063 & 0.687 \\ 
+  & Mixture & Nonpeaked & 0.65 & -2.063 & 0.687 \\ 
   & & Peaked & 0.65 & -2.201 & 1.621 \\ 
@@
 & Mixture & &0.124&0.012&0.064&-1.07&-0.068&-0.146&0.03&0.312&0.27 \\
-Gamma & Non-mixture & Non-peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
+Gamma & Non-mixture & Nonpeaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
 & & Peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
-& Mixture & Non-peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
+& Mixture & Nonpeaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
 & & Peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
-Lognormal & Non-mixture & Non-peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
+Lognormal & Non-mixture & Nonpeaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
 & & Peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
-& Mixture & Non-peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
+& Mixture & Nonpeaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
 & & Peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
-Weibull & Non-mixture & Non-peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
+Weibull & Non-mixture & Nonpeaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
 & & Peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
-& Mixture & Non-peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
+& Mixture & Nonpeaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
 & & Peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
@@
 & Mixture & &1.097&-1.97&0.13&0.091&0.259&0.647& \\
-Gamma & Non-mixture & Non-peaked &1.383&-3.941&0.121&0.072&0.39& & 0.316 \\
+Gamma & Non-mixture & Nonpeaked &1.383&-3.941&0.121&0.072&0.39& & 0.316 \\
 & & Peaked &1.134&-0.746&0.121&0.072&0.39& & 3.371 \\
-& Mixture & Non-peaked &1.215&-2.7&0.121&0.079&0.332&0.79& 0.597 \\
+& Mixture & Nonpeaked &1.215&-2.7&0.121&0.079&0.332&0.79& 0.597 \\
 & & Peaked &1.134&-1.21&0.121&0.079&0.332&0.65& 2.491 \\
-Lognormal & Non-mixture & Non-peaked &1.613&-2.444&0.151&0.077&0.353& & 3.287 \\
+Lognormal & Non-mixture & Nonpeaked &1.613&-2.444&0.151&0.077&0.353& & 3.287 \\
 & & Peaked &1.134&-1.872&0.151&0.077&0.353& & 0.512 \\
-& Mixture & Non-peaked &1.282&-2.051&0.14&0.079&0.312&0.706& 1.467 \\
+& Mixture & Nonpeaked &1.282&-2.051&0.14&0.079&0.312&0.706& 1.467 \\
 & & Peaked &1.134&-1.992&0.14&0.079&0.312&0.65& 0.618 \\
-Weibull & Non-mixture & Non-peaked &1.57&-3.137&0.131&0.074&0.383& & 0.387 \\
+Weibull & Non-mixture & Nonpeaked &1.57&-3.137&0.131&0.074&0.383& & 0.387 \\
 & & Peaked &1.134&-2.042&0.131&0.074&0.383& & 1.829 \\
-& Mixture & Non-peaked &1.294&-2.354&0.122&0.078&0.344&0.814& 0.653 \\
+& Mixture & Nonpeaked &1.294&-2.354&0.122&0.078&0.344&0.814& 0.653 \\
 & & Peaked &1.134&-2.201&0.122&0.078&0.344&0.65& 1.621 \\
@@
 \end{tabular}
-\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.
+\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates (Section \ref{M-sec:simfull}).
 Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
@@
 %   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
-%  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
+%  Gamma & Non-mixture & Nonpeaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
 %   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
-%   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
+%   & Mixture & Nonpeaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
 %   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
-%  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
+%  Lognormal & Non-mixture & Nonpeaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
 %   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
-%   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
+%   & Mixture & Nonpeaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
 %   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
-%  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
+%  Weibull & Non-mixture & Nonpeaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
 %   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
-%   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
+%   & Mixture & Nonpeaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
 %   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
@@
 50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:mix50}
 
@@
 50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
-
+\label{tbl:mix65}
 \end{table}
@@
 50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:mix80}
 
@@
 50\% and 90\% coverage is expressed as the proportion of 100 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:mix95}
 \end{table}
@@
 50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:fam50}
 
@@
 50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
-\label{tbl:sim2}
+\label{tbl:fam65}
 \end{table}
@@
 50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
-\label{tbl:sim2}
+\label{tbl:fam80}
 
@@
 50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
+\label{tbl:fam95}
 \end{table}
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-Each column presents one family of simulated dataset.  
-Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-`X' marks the expected marginal probability of detection based on true parameter values.}
+%\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
+%\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
+%Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
+%Each column presents one family of simulated dataset.  
+%Upper plots show non-mixture datasets; lower plots show mixture datasets.  
+%`X' marks the expected marginal probability of detection based on true parameter values.}
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
-\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.
-All data and inference models include mixtures.
+\caption{\label{pdet_cater_family}  Simulation results from Section \ref{M-sec:simfull}.
+Caterpillar plots of posterior 50\% (orange) and 95\% (black) credible intervals for the probability of detection $\pdet$ across all surveys.
+Data and models include fixed and random effects for both abundance and detection processes.
+All data and inference models include mixture components.  
 Each plot presents one simulated dataset.  

### 2017-04-06 Sims/SimZero/Outputs_Code_zero.R

@@ -220,122 +220,122 @@ print(
 
 
 
-########## Generate Summary Tables -- Mix versus Non-Mix
-# Table of coverages for each dataset and inference model combo (averaged over reps) plus average posterior p-value
-cover.p <- ddply(p_glob, .(simdat, ModelFamily, DataFamily, modelmix), summarize, 
-                 C50=sum(abs(pval_pglob-0.5)<0.25)/length(unique(p_glob$Rep)), 
-                 C90=sum(abs(pval_pglob-0.5)<0.45)/length(unique(p_glob$Rep)), avg.postp=mean(pval_pglob))
-
-Tbl.mix.m  <- subset(cover.p, DataFamily==ModelFamily & modelmix)[,c("simdat", "C50", "C90", "avg.postp")]  # Table for mixture inference models
-Tbl.mix.nm <- subset(cover.p, DataFamily==ModelFamily & !modelmix)[,c("simdat", "C50", "C90", "avg.postp")] # Table for non-mixture inference models
-# Mix.mean and NM.mean are average p.values, not average detection probabilities
-# Mix50, Mix90, etc. are coverage proportions
-names(Tbl.mix.m)[2:4] <- c("Mix50", "Mix90", "Mix.avg.postp")
-names(Tbl.mix.nm)[2:4] <- c("NM50", "NM90", "NM.avg.postp")
-Tbl.mix <- join(Tbl.mix.nm, Tbl.mix.m, by=c("simdat"))
-
-# Average median(gamma) across Reps.
-# This table consists only of mixture inference models (but all data models). 
-# Note: the true value of gamma is 5/6 for the p50 simulations
-parmests  <- data.handle(parmests)
-avggamma  <- ddply(subset(parmests, parameter=="gamma" & DataFamily==ModelFamily),
-                  .(simdat), summarize, avg.med.gamma = mean(X50.))
-# Tbl.mix <- join(Tbl.mix, avggamma) # No longer reporting this optional column
-
-### Average posterior median estimate of pdet ###
-count.det <- data.handle(count.det)
-avgpdet <- ddply(subset(count.det, parameter=="p_global" & DataFamily==ModelFamily),
[…]

### 2017-04-06 ReviewerComments.tex
More updates.

@@
 \begin{quote}
-\adam{Address this point in the abstract}
+\adam{I certainly don't intend to address either of these in the Abstract.
+Both are discussed in the text.
+Computation times are not addressed in the Discussion.
+Let me know if you think otherwise.}
 \end{quote}
@@
 
-Meanwhile, as disscussed later, a simulation study involving effects on abundance and detection is computationally expensive.
+Meanwhile, as discussed later, a simulation study involving effects on abundance and detection is computationally expensive.
 Fortunately, while simulation studies are hampered by long computation times, the typical user of this model will only apply it to a small number of datasets, so computation time will not be prohibitive.
@@
 \begin{quote}
-Average run times for the different simulation studies are now discussed in the second paragraph of Section \ref{M-sec:simfull}.
+Average run times for the different simulation studies are now discussed in the third paragraph of Section \ref{M-sec:simfull}:
+
+``Computation times for this simulation study were much greater than for the other studies because partitions of the cdf, e.g.
+$\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
+Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
+Due to the computation times involved, we fit each model only once."
 \end{quote}
@@
 \begin{quote}
-\adam{What is a `discrete-count survey'?  I have not found a clear reference.\\
-I accede the `large sample' point.
-But based on our findings, I'd advocate against modeling abundance/perception from any small dataset, because analysis requires either a strong assumption about detection rates or a strong prior.
-Any comment on this point would be strengthened if we ran small-dataset simulations.\\
-I think the low-availability stipulation is redundant.  If availability is really high, then a removal analysis is not necessary at all.}
-\jarad{But how do you know availability is high?}
+We hope that clarifications above about abundance estimates and uncertainty resolve some of the reviewer's hesitations.
+We have added single-species' to the draft.
+
+\adam{Perhaps more explanation is required here.}
 \end{quote}
@@
 We disagree.
-Based on our results, applying the precautionary principle, the burden of proof should rest with the analyst who wishes to assume constant detection.
+Based on our results and applying the precautionary principle, we believe the burden of proof should rest with the analyst who wishes to assume constant detection.
 This is a central point in our manuscript: the constant-rate assumption should not be the default.
@@
 This manuscript applies to removal-only analyses.
-Still, the concept can be carried over to removal-distance analyses.
+Still, the concept can be carried over to removal-distance analyses, as we describe in a paragraph of the Discussion.
 The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
@@
 \begin{quote}
-\adam{This last recommendation seems feasible, though I think our final paragraph already addesses some of these issues.}
-\jarad{If we already address it, then you should just quote from the manuscript.}
+\adam{I do not deem this necessary.
+We have not explicitly studied sample size.
+The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
 \end{quote}
-\vspace{1.5cm}
 
@@
 The inclusion of abundance estimates in Figure \ref{M-fig:sim2typical} should convey much the same information as would profile likelihoods.
-\adam{Supplement Tables 1 \& 2 with some manner of confidence intervals for $\hat{N} / N$?  But averaging across replicates is not a trivial issue.}
 \end{quote}
@@
 While the approach could certainly be applied to our model in its more complex forms (e.g. with fixed and random effects), to do so in this manuscript would distract from our main objectives.
-We have introduced more discussion of identifiability and flat data likelihoods in \adam{location}, and we have cited \citep{Lele2010} there.
 
-\adam{YOU HAD BETTER DO THIS!}
+We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicity discussion of them in the Discussion.
 
@@
 \begin{quote}
-\adam{
-To avoid potentially a lot of rewriting, I think we should add an abbreviated version of Equation (1) at the end of Section 3.2.
-I am struggling with this at present.
-}
+We have split the first section of the model exposition into two sections: `Distributions for exact times to detection' and `TTDDs in an N-mixture model'.
+We think this makes the flow of arguments clearer.
+We have also introduced a new Equation (1) --- precursor to the old Equation (1) --- that provides the basic data model for exact-time homogeneous survey data.
+This also should make the model progression more lucid.
 \end{quote}

### 2017-04-06 Sims/SimZero/Outputs_Code_zero.R

@@ -449,7 +449,7 @@ grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, 
 
 lp <- ggplot(subset(ov, statistic=="Med.p" & familymatch)) + 
   geom_point(aes(x=as.numeric(pdet), y=value, shape=mt), size=2.25) +
-  geom_segment(aes(x=50 , y=0.5 , xend=95, yend=0.95), linetype=2, size=0.6) +
+  geom_segment(aes(x=50 , y=0.5 , xend=95, yend=0.95), color="gray60", linetype=2, size=0.5) +
   facet_grid(datamix~peaked) +
   scale_shape_manual(values=c(15,16,17,18,0,1,2,5), name="Inference Model") +
   scale_x_continuous(breaks = seq(50, 95, 15)) +
@@ -458,8 +458,8 @@ lp <- ggplot(subset(ov, statistic=="Med.p" & familymatch)) +
   ylab("Estimated Detection Probability") + xlab("Actual Detection Probability")
 
 rp <- ggplot(subset(ov, statistic=="C90" & familymatch)) + 
-  geom_hline(aes(yintercept=0.835), linetype=2, size=1, color="gray50") +
-  geom_hline(aes(yintercept=0.955), linetype=2, size=1, color="gray50") +
+  geom_hline(aes(yintercept=0.835), linetype=2, size=1, color="gray60") +
+  geom_hline(aes(yintercept=0.955), linetype=2, size=1, color="gray60") +
   geom_point(aes(x=as.numeric(pdet), y=value, shape=mt), size=2.25) +
   facet_grid(datamix~peaked) +
   scale_shape_manual(values=c(15,16,17,18,0,1,2,5), name="Inference Model") +
@@ -473,8 +473,8 @@ dev.off()
 
 # Family comparison plots
 lf <- ggplot(subset(ov, statistic=="Med.p" & datamix=="Mixed" & modelmix)) + 
+  geom_segment(aes(x=50 , y=0.5 , xend=95, yend=0.95), color="gray60", linetype=2, size=0.5) +
   geom_point(aes(x=as.numeric(pdet), y=value, shape=ModelFamily), size=2.25) +
-  geom_segment(aes(x=50 , y=0.5 , xend=95, yend=0.95), linetype=2, size=0.6) +
   facet_grid(DataFamily~peaked2) +
   scale_shape_manual(values=c(0,1,2,5), name="Inference Model") +
[…]

### 2017-04-18 ReviewerComments.tex
Add my comments on reviewer response

@@
 We address each of these in turn as they appear in the comments below. 
+\jarad{Answer them here even if it repeats information below.}
 \end{quote}
@@
 Let me know if you think otherwise.}
+\jarad{I think that's fine. In responding to the review, we can state our 
+opinion that the abstract is not the appropriate place and point the reviewer
+to where the content can be found.} 
 \end{quote}
@@
 We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
+\jarad{We probably need to add more, e.g. intended audience. Philip?} 
 \end{quote}
@@
 For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
-Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences for tail probability.
+Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences detection probability.
 \end{quote}
@@
 More formally, we define a peaked TTDD as having a mode greater than zero (or $C_1$ for lognormal) while a non-peaked TTDD has a mode of zero (or less than $C_1$), but we consider exponential TTDDs to be neither peaked nor nonpeaked."
+
+\jarad{We should probably remove TTDD from the introduction.}
 
@@
 \begin{quote}
+\jarad{I think we can thank reviewers here for the push to do more simulation 
+studies that resulted in our realization of a computational savings allowing
+us to do more analyses.}
 We have extended our simulation study to include a variety of detection probabilities rather than a variety of sample sizes.
 Both affect estimation and are worthy of investigation, but since the performance of unknown-$N$ binomial models generally deteriorates as values of $\pdet$ decrease, we chose to direct our attentions there.
+\jarad{Don't we have different sample sizes due to the variation in detection
+probabilities?}
 Sections \ref{M-sec:mixture} and \ref{M-sec:family} have been reworked to detail the updated findings.
@@
 We have separated out a paragraph in the Discussion posing some design-related questions.
+\jarad{Be more specific.}
 \end{quote}
@@
 Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
+\jarad{Too many significant digits.}
 Due to the computation times involved, we fit each model only once."
@@
 We now choose to report only 90\% credible intervals.
+\jarad{Are 50\% included in supplement?}
 \end{quote}
@@
 We now clearly state in Section \ref{M-sec:data}: ``This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s."
+
+We believe this uncertainty accurately reflects the uncertainty in estimation
+of detection probability and abundance. 
+The apparent certainty in previously published studies relies heavily on the 
+assumption of constant detection rate.
 \end{quote}
@@
 The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
+\jarad{This is ongoing work for us.}
 \end{quote}
@@
 The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
+\jarad{I'll have to re-read, but overall providing a summary for the reader
+seems like a good idea.}
 \end{quote}
@@
 Second, although the original number of simulations was inadequate for characterizing model accuracy, it nonetheless highlighted scenarios of gross inaccuracy, such as: (i) omitting a mixture component in modeling a peaked mixture dataset, or (ii) assuming constant detection when the rate varies during the survey period.
+\jarad{Is the second point necessary?}
 \end{quote}
@@
 The findings have significantly added to our results and conclusions.
+\jarad{Should we summarize the results here?}
 
@@
 \end{itemize}
+\jarad{I don't think this parameter is necessary.}
 \end{quote}
@@
 
-We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicity discussion of them in the Discussion.
+We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicit discussion of them in the Discussion.
 
 As an informational note, the \texttt{R} package by Solymos was constructed for BUGS and JAGS, though it would not be difficult to adapt data cloning for Stan.
+\jarad{Remove paragraph.}
 \end{quote}
@@
 This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso] --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
+\jarad{We might not need the ``which..''.}
 \end{quote}
@@
 We have added this as Figure \ref{M-fig:schematic}.
+\jarad{Thank the reviewer for the suggestion.}
 \end{quote}

### 2017-04-21 Draft.tex
Revisions of Reviewers' Comments.  Small changes to Draft.

@@
 $\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
-Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
+Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2.0, 3.0, 3.3, and 5.3 minutes for the intercept-only models.
 Due to the computation times involved, we fit each model only once.

### 2017-04-21 ReviewerComments.tex
Revisions of Reviewers' Comments.  Small changes to Draft.

--- 2017-04-06 ReviewerComments.tex
@@
 \newcommand{\ind}{\stackrel{ind}{\sim}}
+\newcommand{\dt}{\text{det}}
+\newcommand{\ftdt}{f_{T|\dt}(t|\dt)}
 \newcommand{\Fm}{F_T^{(M)}}
@@
 \begin{document}
+
+\begin{quote}
+We thank the reviewers for their comments.
+We believe the manuscript is clearly a better document in consequence.
+In particular, the reviewers encouraged us to generate more simulations --- in terms of both the number of replicates and the variety in true detection probabilities.
+In pursuing these recommendations, we improved our simulation code to be more efficient by orders of magnitude for the covariate-free simulations.
+Our results now more fully describe model performance across a range of true detection probabilities, and the patterns observed are more compelling on account of the greater number of replicates.
+\end{quote}
+
+\newpage
+
 \itshape
@@
 \item What types of data are relevant to each examined distribution (e.g., gamma, Weibull)?
+
+\begin{quote}\label{ans:one}
+Thank you for raising this point.
+Unlike the exponential and gamma TTDDs, the lognormal and Weibull do not have an immediate biological mechanism for interpreting hazard rates.
+To address this issue, we have added to Section \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to increase or decrease during the survey.
+All three TTDDs may provide reasonable empirical approximations of non-constant detection, though the shapes of the distributions differ, potentially leading to differing inference.
+For instance, when detection rates vary across individuals, the result is a marginal detection rate that decreases over time.
+Whether the marginal rate is best approximated by a gamma, lognormal, or Weibull TTDD depends on just how rates vary across individuals."
+
+We definitely appreciate the importance of providing biological justification for statistical assumptions.
+It is always a good foundation.
+That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
+If the dynamics of the system are more complicated than we can directly model, the best we can do is find a TTDD that reasonably describes the observed pattern.
+For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
+Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences for detection probability.
+
+If an analyst does have a clear detection rate mechanism in mind, they can always fit our model using the appropriate hazard function and the relationship between it and its corresponding TTDD: $h(t) = f_T(t) / [1-F_T(t)]$.
+\end{quote}
+
 \item What sample sizes (i.e., number of detections) are needed to obtain reliable estimates with CIs that are not so wide as to be meaningless?
+
+\begin{quote}
+We fear the original version of Figure \ref{M-ovenposteriors} was confusing and led the reviewer to believe that CIs were larger than they in fact are.
+It showed posterior credible intervals for \textit{uncounted} abundance, rather than the more intuitive \textit{total} abundance.
+We have fixed the figure.
+We hope the reviewer will agree that CIs are not always so wide as to be meaningless.
+
+One reviewer recommended we simulate across multiple sample sizes; another reviewer recommended that we simulate across a variety of true detection probabilities.
+Both are useful suggestions.
+However, we decided to prioritize the simulation across detection probabilities, since papers on N-mixture models more often emphasize this side of model performance.
+\end{quote}
+
 \item What are approximate computation times for models of different complexity (e.g., days? weeks?).
+\begin{quote}
+Thanks to the encouragement from the reviewers to run more simulations, we revisited our code and improved efficiency by orders of magnitude for fitting interval-only models.
+So, the first two simulation studies now use 100 replicates rather than then original 16.
+We cannot realize such savings for models with heterogeneous detection across surveys.
+We now cite runtime statistics in Section \ref{M-sec:simfull}: ``Average times for exponential, lognormal, Weibull, and gamma [heterogeneous] model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.''
+\end{quote}
+
 \end{enumerate}
-\begin{quote}
-We address each of these in turn as they appear in the comments below.
-\end{quote}
 
@@
 \begin{quote}
-\adam{I certainly don't intend to address either of these in the Abstract.
-Both are discussed in the text.
-Computation times are not addressed in the Discussion.
-Let me know if you think otherwise.}
+
+As noted above, we think imprecision appeared more exaggerated in our first submission than it in fact was, and we have improved computation times.
+We address both issues in the text but feel that neither quite belongs in the Abstract.
+
+We address computation times in Section \ref{M-sec:simfull}.
+We address imprecision multiple places in the simulation results, Ovenbird results, and Discussion.
 \end{quote}
@@
 \begin{quote}
-We have added the following sentences to \ref{M-sec:exact_time}: ``Like the gamma TTDD, Weibull and lognormal TTDDs offer the flexibility of a two-parameter form and allow rates to increase or decrease during the survey.
-All three TTDDs may provide reasonable empirical approximations of non-constant detection, though the shapes of the distributions differ, potentially leading to differing inference.
-For instance, when detection rates vary across individuals, the result is a marginal detection rate that decreases over time.
-Whether the marginal rate is best approximated by a gamma, lognormal, or Weibull TTDD depends on just how rates vary across individuals."
-
-We appreciate the importance of providing biological justification for statistical assumptions.
-It is always a good foundation.
-That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
-If the dynamics of the system are more complicated than we can directly model, the best we can do is find a TTDD that reasonably describes the observed pattern.
-For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
-Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences for tail probability.
+See the answer to Item 1 on Page~\pageref{ans:one} above for our full response.
 \end{quote}
@@
 \begin{quote}
-We have given the definitions of peaked/nonpeaked their own paragraph in the Simulation Studies methodology (Section \ref{M-sec:sim}):
+We have separated the definitions of peaked/nonpeaked into own paragraph in the Simulation Studies methodology (Section \ref{M-sec:sim}):
 ``In the following analyses we distinguish two categories of purely continuous TTDDs: peaked and nonpeaked.  
@@
 \begin{quote}
-We have extended our simulation study to include a variety of detection probabilities rather than a variety of sample sizes.
+We thank the reviewers for pushing us to expand our simulation studies.
+It led us to improvements in our simulation code, more compelling results (through increased replication), and characterization of model performance across more scenarios.
+One reviewer asked us to simulated across a variety of detection probabilities, while the other asked us to simulation across sample sizes.
 Both affect estimation and are worthy of investigation, but since the performance of unknown-$N$ binomial models generally deteriorates as values of $\pdet$ decrease, we chose to direct our attentions there.
@@
 %\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
-We have separated out a paragraph in the Discussion posing some design-related questions.
+We have separated out a paragraph in the Discussion posing some design-related issues: (i) longer surveys, (ii) introduction of settling down periods, and (iii) collection of complete detection histories (not just first detections).
 \end{quote}
@@
 $\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
-Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
+Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2.0, 3.0, 3.3, and 5.3 minutes for the intercept-only models.
 Due to the computation times involved, we fit each model only once."
@@
 We now choose to report only 90\% credible intervals.
+We still report 50\% intervals in the supplement.
 \end{quote}
@@
 \begin{quote}
-Clarification: We have 947 detections at 381 sites (which is \textit{really} a lot of detections).
+Clarification: We actually have 947 detections at 381 sites (which is \textit{really} a lot).
 We now clearly state in Section \ref{M-sec:data}: ``This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s."
+We address CI widths in the next response.
 \end{quote}
@@
 We hope that clarifications above about abundance estimates and uncertainty resolve some of the reviewer's hesitations.
-We have added single-species' to the draft.
+We have added `single-species' to the manuscript.
+\end{quote}
 
-\adam{Perhaps more explanation is required here.}
-\end{quote}
 where non-constant detection through time is suspected based on the behavioral ecology of the species.
+
 \begin{quote}
@@
 The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
+This is ongoing work for us.
 \end{quote}
@@
 \begin{quote}
-We answer this comments in two parts.
-First, we agree that the number of simulations was inadequate for diagnosing model accuracy.
-We have therefore increased to 100 the number of simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family}.
+We agree that the number of simulations was inadequate for diagnosing model accuracy.
+We have increased to 100 the number of simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family}.
 This has been made possible by improvements to Stan \citep{Rstan2016} and by streamlining our code for homogeneous surveys.
-See the second paragraph of Section \ref{M-sec:simfull} for a summary of run times for the models currently.
-
-Second, although the original number of simulations was inadequate for characterizing model accuracy, it nonetheless highlighted scenarios of gross inaccuracy, such as: (i) omitting a mixture component in modeling a peaked mixture dataset, or (ii) assuming constant detection when the rate varies during the survey period.
 \end{quote}
@@
 The findings have significantly added to our results and conclusions.
+The extra simulations demonstrate that, for nonpeaked datasets, the model returns similar detection estimates whether the true value of $\pdet$ is 0.50 or 0.65 or even somewhat 0.80.
+For exponential and peaked datasets, the model is more able to return unbiased estimates.
+Other major findings of our original manuscript -- the inadequacy of non-mixture or constant-rate models in many cases -- remain true across wide ranges of true values for $\pdet$.
 
-With reference to Olkin's finding that point estimates of $N_s$ are unstable unless $\pdet>0.41$, it is not a hard boundary.  
-It varies by context, but the boundary is a general property of estimating $N$ in binomial unknown-$N$, unknown-$p$ problems.
-Olkin's calculation derives from the relative sample mean and sample variance of repeated large samples with the same values of $N$ and $\pdet$.
-As such, it is a best-case scenario.
-Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
-\begin{itemize}
-\item \citet{Veech2016}: advise caution when $p<0.5$ in the context of individual-level detection effects (no time-to-detection data). 
-\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
-\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
-\end{itemize}
+%With reference to Olkin's finding that point estimates of $N_s$ are unstable unless $\pdet>0.41$, it is not a hard boundary.  
+%It varies by context, but the boundary is a general property of estimating $N$ in binomial unknown-$N$, unknown-$p$ problems.
+%Olkin's calculation derives from the relative sample mean and sample variance of repeated large samples with the same values of $N$ and $\pdet$.
+%As such, it is a best-case scenario.
+%Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
+%\begin{itemize}
+%\item \citet{Veech2016}: advise caution when $p<0.5$ in the context of individual-level detection effects (no time-to-detection data). 
+%\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
+%\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
+%\end{itemize}
 \end{quote}
@@
 
-We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicity discussion of them in the Discussion.
+We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicit discussion of them in the Discussion:
 
-As an informational note, the \texttt{R} package by Solymos was constructed for BUGS and JAGS, though it would not be difficult to adapt data cloning for Stan.
+``Reduced model performance in the presence of low detection probabilities is common among abundance models.
+We recommend the discussion in \citet{CoullAgresti1999}, elaborating key points here.
+The essential problem is a flat log-likelihood.
+For single-pass surveys, we cannot rely on repeated counts to improve our estimates.
+Instead, the strategy of removal sampling is to use the observed pattern of detection times to estimate the proportion of individuals that would be detected if only the observation period lasted longer.
+As such, it is entirely reliant upon extrapolation based on an accurate fit of $\ftdt$.
+When true detection probabilities are low, $\ftdt$ for nonpeaked data becomes flat.
+Such a flat observed pattern provides little information and is well approximated by a wide variety of TTDDs, which vary greatly in their tail probabilities.
+An assumption of constant detection limits the shape of $\ftdt$ and thereby constrains uncertainty but at the cost of potentially sizable bias.''
 \end{quote}
@@
 \begin{quote}
+Thank you for the suggestion.
 We have added this as Figure \ref{M-fig:schematic}.
+We agree that the figure is quite effective and should be a useful guide to readers of the paper.
 \end{quote}

### 2017-05-08 Draft.tex
Add reference to Figure 1 in Section 3.3

@@
 Let $n_{si}$ be the number of individuals counted during interval $i$ on survey $s$, $\nobs_s = \sum_{i=1}^I n_{si}$, and $\vn_{s}=(n_{s1},\dots,n_{sI})$.
-Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(\nobs_s, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and is calculated from the TTDD: $p_{si} = F_{T|\dt}(C_i|\dt) - F_{T|\dt}(C_{i-1}|\dt)$.
+Assuming independence amongst individuals and sites, we have $\vn_{s} \ind \Mult \left(\nobs_s, \vp_{s}\right)$, where $\vp_{s}=(p_{s1},\dots,p_{sI})$ and $p_{si} = F_{T|\dt}(C_i|\dt) - F_{T|\dt}(C_{i-1}|\dt) = \int_{C_{i-1}}^{C_i} f_{T|\dt}(t) dt$,
+see Figure \ref{fig:schematic}.
 

### 2017-05-08 Draft.tex
Add \jarad comments

@@
 \usepackage{color}
-% \usepackage[dvipsnames]{xcolor}
+\usepackage[dvipsnames]{xcolor}
 \usepackage[margin=1cm]{caption}
@@
 \newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-% \newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
+\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
 
@@
 However, this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}. 
-We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
-We apply the term `mixture model' to any model with a mixture TTDD.
+We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates
+and apply the term \emph{mixture TTDD} when the TTDD has a discrete and continuous component.
 
@@
 We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture} --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
+\jarad{I think we should remove the ---.}
 For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).
@@
 We specify a mixture TTDD with mixing parameter $\gamma\in[0,1]$, a point-mass during the first observation interval, and a continuous-time detection distribution $\Fm(t)$.
+\jarad{Why is the superscript M?}
 The mixture TTDD cdf is defined: $F_T(t) = (1-\gamma) + \gamma \Fm(t)$ for $t>0$.
@@
 We centered and standardized all continuous covariates prior to fitting models.
-We ran chains 250,000-375,000 iterations; Geweke diagnostics showed no indication of lack of fit, and effecive sample sizes were over 1000 for all parameters.
+We ran chains 250,000-375,000 iterations; Geweke diagnostics showed no indication of lack of fit, and effective sample sizes were over 1000 for all parameters.
 

### 2017-05-09 Draft.tex
Modifications to manuscript.

--- 2017-05-08 Draft.tex
@@
 \newcommand{\ind}{\stackrel{ind}{\sim}}
-\newcommand{\Fm}{F_T^{(M)}}
+\newcommand{\Fm}{F_T^{(C)}}
 \newcommand{\dt}{\text{det}}
@@
 However, this is not yet the standard \citep{Solymos2013, Amundson2014, Reidy2016}. 
-We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates.
-We apply the term `mixture model' to any model with a mixture TTDD.
+We consider the choice of whether to include a mixture component in conjunction with TTDDs with non-constant rates
+and apply the term \emph{mixture TTDD} when the TTDD has a discrete and continuous component.
 
 Unlike most survival analyses, the number of individuals $N$ present at a survey is unknown and may be the primary quantity of interest.  
-We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture} --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
+We embed the TTDD in a hierarchical framework for multinomial counts called an N-mixture model \citep{Wyatt2002, Royle2004NMixture}, which is an entirely different use of `mixture' from the mixture models in the previous paragraph.
 For our purposes, the N-mixture framework provides three clear benefits: 1) it handles counts within a flexible multinomial data framework \citep{RoyleDorazio2006} which accords with the interval-censored data collection that is customary in point-count surveys \citep{Ralph1995}, 2) the hierarchical structure readily lends itself to including abundance- and detection-related covariates and random effects \citep{Dorazio2005, Etterson2009, Amundson2014}, and 3) for a Bayesian analysis, we can sample the posterior joint distribution of N-mixture parameters straight-forwardly using Markov chain Monte Carlo (MCMC).

### 2017-05-11 ReviewerComments.tex
Make Reviewers' Comments consistent with current manuscript.

@@
 In the Introduction, we now clearly refer to a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
-We define our use of `mixture model' later in the paragraph:  ``We apply the term `mixture model' to any model with a mixture TTDD."
-This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso] --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
-\jarad{We might not need the ``which..''.}
+We define our use of `mixture model' later in the paragraph:  ``[We] apply the term \emph{mixture TTDD} when the TTDD has a discrete and continuous component."
+This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model \mbox{[\dotso]}, which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
 \end{quote}

### 2017-05-18 Draft.tex
Fix remaining two comments to reviewers

@@
 Just as the exponential TTDD is a special case of the two-parameter gamma and Weibull TTDDs, so all four TTDDs in our analysis are special cases the three-parameter generalized gamma distribution.
-A generalized gamma TTDD encompasses a diversity hazard functions \citep{Cox2007}, eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
+A generalized gamma TTDD encompasses a diversity of hazard functions \citep{Cox2007}, 
+eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
 However, maximum likelihood estimation of the generalized gamma has historically suffered from computational difficulties, unsatisifactory asymptotic normality at large sample sizes, and non-unique roots to the likelihood equation \citep{CoorayAnand2008, NoufailyJones2013}.

### 2017-05-18 ReviewerComments.tex
Fix remaining two comments to reviewers

@@
 \begin{quote}
-We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
-\jarad{We probably need to add more, e.g. intended audience. Philip?} 
+Since we and the AE feel this comparison is useful, 
+we have retained the comparison. 
+But we have tried to make the equivalence more explicit: ``We analyze times to first 
+detection as time-to-event data, as is done in parametric survival analysis, ..."
 \end{quote}
@@
 \begin{quote}
-\adam{I do not deem this necessary.
-We have not explicitly studied sample size.
-The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
-\jarad{I'll have to re-read, but overall providing a summary for the reader
-seems like a good idea.}
+We have attempted to address these comments in the discussion:
+``If the estimation of abundance is the primary interest, 
+then the choice of TTDD has large consequences ... we would be skeptical of 
+abundance estimates that indicate a nonpeaked TTDD with median posterior 
+estimates of $\pdet$ below 0.75 ...
+An assumption of constant detection limits the shape of $\ftdt$ and 
+thereby constrains uncertainty but at the cost of potentially sizable bias...
+all found constant-detection models to be flawed, 
+producing underestimates of abundance and too-narrow error estimates; 
+these resulted in inadequate coverage and also overstatement of effect 
+significance.''
+
+We have also included computation times in Section \ref{M-sec:simfull}: 
+``Average times for exponential, lognormal, Weibull, and gamma model fits were 
+2.6, 5.1, 5.3, and 25+ hours [when mixed effects were included in both abundance 
+and detection].'' Since these types of surveys are typically conducted yearly, 
+we do not feel even a one day runtime is arduous.
 \end{quote}

### 2017-05-18 ReviewerComments.tex
Fix typo

@@
 \begin{quote}
-We have attempted to address these comments in the discussion:
+We have addressed these comments in the discussion:
 ``If the estimation of abundance is the primary interest, 

### 2017-07-13 Draft.tex
Remove comments

@@
 \usepackage[margin=1cm]{caption}
-%\usepackage{tabularx, booktabs}
-%\usepackage{float}
 \usepackage{adjustbox}
@@
 
-\newif\ifdic   % Type \ifdic at the beginning of any optional section... use \fi to denote the end of the optional section
-  % See: http://tex.stackexchange.com/questions/33576/conditional-typesetting-build
-%\dictrue         % comment out if I want to hide DIC and model fit paragraphs
-
 \renewcommand{\floatpagefraction}{0.8}  % Changes settings for when a figure hogs a page all to itself
-
-\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
-\newcommand{\jarad}[1]{{\color{Orange} JARAD: #1}}
 
@@
 
-%\tableofcontents
-
 \begin{abstract}
@@
 The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption can often lead to biased estimates.
-%Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.  
-%Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  
 We consider a class of N-mixture models that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
-%We consider a class of N-mixture models that allow for detection heterogeniety through a mixture component, flexible two-parameter time-to-detection distributions (TTDDs), and allows for fixed and random effects for both abundance and detection.
-%Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution (TTDD) embedded within a hierarchical N-mixture framework.
 Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
@@
 We demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs but that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
-%We find that modeling a TTDD by using a mixture component for increased probability of detection in the initial observation period is necessary when data have a chance of arising from distributions of this nature.
-%In addition, the flexibility offered by two-parameter distributions, e.g. gamma, can outperform exponential based models when the truth is exponential or non-exponential.
-%Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection and demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs and that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
-% We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  
-% Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.
-% Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.
 
@@
 Removal sampling, where individuals are solely counted on their first capture, provides one established methodology for estimating detection probabilities \citep{Farnsworth2002}.
-% cost-effective in that it requires neither multiple observers nor multiple site visits.
-%Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
-%Assuming that individual capture rates remain constant over all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time .
-%\citet{Farnsworth2002} adapted removal sampling to point-count surveys, suggesting that if: (i) observers record only the first time-to-detection for each animal, and (ii) times to first detection are interval-censored, then data will be similar in for to traditional removal sampled data.
-%In the analysis of such data, the observation period is subdivided into unit intervals of equal duration, each of which is equivalent to a trapping session.
-%Farnsworth's removal assumptions:\\
-%(i) Closed population during survey\\
-%(ii) no double-counting\\
-%(iii) easy-to-detect group is entirely counted during first interval\\
-%(iv) TTDD for hard-to-detect group is correct after the end of the observation period\\
-%(v) if using limited radius counts, they are accurate
-% Johnson2008: all birds present are available for detection
 A standard assumption in removal sampling is a constant detection rate throughout the observation period, but this assumption is often unjustified \citep{Alldredge2007, LeeMarsden2008}.  
@@
 
-%Application of this methodology generally assumes that organisms at a survey site are detected at a constant rate \citep{Farnsworth2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}; however, this assumption is often not justified in the field.
-%Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.
-
 Framing a model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  
@@
 
-% \adam{Based on further lit review, I think I need to insert a paragraph here about what time-varying modeling has been done.  
-% In the fisheries world, models of time-varying catchability have been implemented -- see Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples.  
-% Also, Alldredge 2007 implements a complete-history of detection model (technically mark-recapture, not removal) that estimates interval-specific detection probabilities (complete with covariates and detection heterogeneity).
-% However, it cannot be applied in a removal context, because removal data contains only first times to detection, and the Alldredge model requires subsequent observations -- removal models compensate by specifying a TTDD, thus enabling us to extrapolate beyond the observation period.}
-
 Section \ref{sec:data} provides a description of the interval-censored time-to-detection avian point count data under consideration.
 Section \ref{sec:model} introduces an N-mixture model with a generically defined TTDD for estimating abundance from removal-sampled point-count surveys.
-%In addition, we discuss reasonable default priors and MCMC analysis used to estimate parameters in these models.
 Section \ref{sec:sim} provides three simulation studies to assess the impact of TTDD choice on estimated detection probability. 
@@
 Our analysis is motivated by avian point-count surveys in Chippewa National Forest from 2008-2013 as part of the Minnesota Forest Breeding Bird Project (MNFB) \citep{Hanowski1995}.  
-%However, the model we describe is appropriate for point-count surveys of any species, so long as time-to-first-detection data are available.  
 For our analysis, we focus on Ovenbird counts selected from one habitat type: sawtimber red pine stands with no recent logging activity.  
@@
 This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s.
-% The counts of Ovenbirds by interval, across all sites, were 596, 69, 62, 68, 43, 33, 35, 27, and 14.
 
@@
 The conditional distribution of observed detection times consequently has pdf $\ftdt= f_T(t)/F_T(C)$ for $0<t<C$, cdf $F_{T|\dt}(t|\dt) = \int_0^t f_{T|\dt}(x|\dt) dx$, and instantaneous detection rate, or hazard function, $h(t) = f_T(t) / [1-F_T(t)]$.
-%$T_{sb|\dt} \ind f_{T|\dt}(t_{sb}|\dt)$ where $T_{sb|\dt}$ is observed only for those $\nobs_s$ individuals whose observed time to first detection is less than $C$, i.e. $t_{sb}<C$.
 
@@
 where $\textbf{t}_s$ is a vector of observed times at survey $s$.
-
-%We model the number of individuals at survey $s$ for which $t_{sb}$ is actually observed as $\nobs_s \ind \mbox{Binomial}\left(N_{s}, \pdet\right)$. 
-%Using maximum likelihood, we can estimate the parameters of a TTDD from the observed times to first detection and thereby estimate $\pdet$ and its uncertainty.
-%With an estimate and uncertainty for $\pdet$, we are in a scenario of a binomial model with unknown $N_s$ and ``known'' $p$ and thus can estimate $N_s$.
-%In these models, point estimates of $N_s$ are unstable unless $\pdet>0.41$ \citep{Olkin1981}.
-%One approach to regularizing these estimates is to construct a hierarchical model for site-specific abundance, e.g. $N_s\ind \Po(\lambda)$ \citep{Raftery1988, Royle2004NMixture}. 
-%Assuming Poisson abundance, we can decompose $N_{s}$ into observed and unobserved portions: $\nobs \sim Po(\lambda \pdet)$ and, independently, $n_{s}^{(unobs)} \sim \Po\left(\lambda[1-\pdet]\right)$.
-%Although alternative distributions could be considered, e.g.
-negative binomial, our experience with Ovenbird point counts suggests that, after accounting for appropriate explanatory variables, the resulting abundances are likely underdispersed rather than overdispersed, and thus we will use the Poisson assumption here.
 
@@
 We fit the models by MCMC sampling using the Bayesian statistical software Stan, implemented via the R package \texttt{rstan} version 2.8.0 \citep{Rstan2016}.  
-%Stan model code and \texttt{rstan} code to run the models can be found in the Supplementary Materials.  
 We discarded half of the iterations as warmup and thinned by 10.  
@@
 For most models, we accepted Stan defaults for initial values; however, gamma and Weibull models sometimes failed to run unless care was taken in the specification of initial values.
-
-%Thus, the link functions in our detection model are:
-%\begin{alignat}{3}
-%&\text{Exponential:\;} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}])\\
-%&\text{Gamma:} &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \log(\alpha)\\
-%&\text{Weibull:}  &&\log(\varphi_{s}) &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
-%&\text{Lognormal:} &&-\mu_{s} &&= \text{Intercept}^D + \vX_{s}^D\vbeta^D + \vZ_{s}^D\vxi^D = -\log(E[T_{sq}]) + \sigma_{det}^2/2
-%\end{alignat}
-
-%\subsection{Abundance Model}
-
-% We distinguish between non-constant detection and detection heterogeneity, which can be modeled in similar ways but refer to different mechanisms.  
-% We reserve `detection heterogeneity' for differences across population subgroups.  
-% It is often modeled using a mixture of detection distributions.  
-% \citet{Farnsworth2002} introduce a finite-mixture for detection probabilities distinguishing between easy-to-detect individuals, which are always detected in first observation interval, and hard-to-detect individuals, which are detected at a constant rate.
-%  Heterogeneity at the individual level can be modeled as random effects in the calculation of detection probabilities \citep{DorazioRoyle2003, Mantyniemi2005}.  
-% We use `non-constant detection' when detection rates for an organism change with time.  
-% Models constructed to address detection heterogeneity may provide reasonable estimates for non-constant detection scenarios and vice versa \citep{Mantyniemi2005}.
-
-%\adam{Individual effects on detection might suffer from identifiability issues. This was the subject of a dust-up between Pledger and Dorazio/Royle from 2003-2005.}
-
-\ifdic\
-\subsection{Model diagnostics}
-
-To assess goodness of fit, we relied on posterior predictive checks \citep{Gelman1996}.  
-%Posterior predictive checks assess whether a fitted model can effectively replicate datasets that look like the original data in key aspects, which are measured by check statistics.  
-In particular, we are concerned with the goodness of fit for the TTDD distribution, so we defined a check statistic for each observation interval:
-\[D(n_i) = \sum\limits_{s} n_{si} \big/ \sum\limits_{si} n_{si}\]
-which is the proportion of the total count that occurs during interval $i$ marginally across all surveys.  
-The posterior predictive p-value associated with each $D(n_i)$ is the proportion of replicate check statistics that have a larger value than the check statistic for the actual data.  
-Posterior predictive p-values near to 0 or 1 are potentially indicators of a misspecified model.
-We simulated replicate datasets and calculated the check statistic at each iteration of our MCMC sampling.
-
-We focus inference on the overall detection probability across all surveys: $\pdet = \sum\limits_{s}\nobs_s \big/$ $\sum\limits_{s}N_{s}$.  
-Within simulation studies, we compared model estimates to true $\pdet$ values by percent coverage and by posterior p-values --- the proportion of MCMC samples having a smaller value than the true value.
-
-We compared different models by means of their goodness of fit and deviance information criterion (DIC) \citep{Spiegelhalter2002}.  
-For the full model with random effects (Ovenbirds and Sim3), which is a missing data model, we used formulation DIC$_7$ from \citet{Celeux2006} for its ease of computation, but we acknowledge that it suffers some difficulties \citep{Celeux2006, Li2014}.
-\fi
 
@@
 
-% FIRST SUBMISSION TABLE
-%\ifdic
-%\begin{table}[ht]\centering\small
-%\begin{tabular}{l|l|l|l|cccc|cccc||r}
-% \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c||}{\underline{Mixture model}} & \\
-% \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-%  \hline
-%  \hline
-% \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 & 0.24 \\
-% & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00 & -0.67 \\ 
-% & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 & 0.37 \\ 
-%  \cline{3-13}
-% & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88 & 1.48 \\ 
-%  \cline{3-13}
-% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 & 0.96 \\ 
-% & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 & 0.97 \\ 
-% & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 & 1.28 \\ 
-%  \cline{2-13}
-%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 & 0.51 \\
-% & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 & 0.33 \\ 
-% & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 & 2.40 \\ 
-%  \cline{3-13}
-% & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 & 122.81 \\ 
-%  \cline{3-13}
-% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 & 31.84 \\ 
-% & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 & 63.65 \\ 
-% & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 & 28.30 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Summary of mixture vs. non-mixture model fits.  
-%In all cases, the inference model family matches the dataset family.  
-%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-%50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-%$\Delta$ DIC: mean difference in DIC between the incorrect-mixture model minus the true model.
-%}
-%\label{tbl:sim1}
-%\end{table}
-%
-%\else
-%
-%\begin{table}[ht]\centering\small
-%\begin{tabular}{l|l|l|l|cccc|cccc}
-% \multicolumn{4}{c|}{ } & \multicolumn{4}{c|}{\underline{Non-mixture model}} & \multicolumn{4}{c}{\underline{Mixture model}} \\
-% \multicolumn{4}{c|}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-%  \hline
-%  \hline
-% \parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.76 & 0.41 & 0.75 & 0.88 & 0.84 & 0.66 & 0.56 & 0.94 \\
-% & & &   Lognormal & 0.66 & 0.17 & 0.31 & 0.62 & 0.78 & 0.48 & 0.62 & 1.00  \\ 
-% & & &   Weibull & 0.69 & 0.25 & 0.38 & 0.75 & 0.79 & 0.51 & 0.75 & 1.00 \\ 
-%  \cline{3-12}
-% & & &   Exponential & 0.80 & 0.54 & 0.44 & 0.94 & 0.79 & 0.38 & 0.31 & 0.88  \\ 
-%  \cline{3-12}
-% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} &  Gamma & 0.80 & 0.55 & 0.50 & 1.00 & 0.82 & 0.70 & 0.38 & 0.88 \\ 
-% & & &   Lognormal & 0.78 & 0.31 & 0.50 & 0.94 & 0.80 & 0.48 & 0.62 & 1.00 \\ 
-% & & &   Weibull & 0.79 & 0.49 & 0.56 & 0.94 & 0.82 & 0.66 & 0.44 & 0.88 \\ 
-%  \cline{2-12}
-%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.17 & 0.12 & 0.81 & 0.76 & 0.41 & 0.69 & 1.00 \\
-% & & &   Lognormal & 0.56 & 0.04 & 0.00 & 0.31 & 0.72 & 0.30 & 0.44 & 0.88 \\ 
-% & & &   Weibull & 0.51 & 0.02 & 0.00 & 0.06 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
-%  \cline{3-12}
-% & & &   Exponential & 0.96 & 1.00 & 0.00 & 0.00 & 0.77 & 0.37 & 0.38 & 0.94 \\ 
-%  \cline{3-12}
-% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.28 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
-% & & &   Lognormal & 0.22 & 0.00 & 0.00 & 0.00 & 0.76 & 0.36 & 0.38 & 0.94 \\ 
-% & & &   Weibull & 0.22 & 0.00 & 0.00 & 0.00 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Summary of mixture vs. non-mixture model fits.  
-%In all cases, the inference model family matches the dataset family.  
-%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-%50\% and 90\% coverage is expressed as the proportion of 16 simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-%}
-%\label{tbl:sim1}
-%\end{table}
-%\fi
-
-% THIS IS SIMILAR TO THE ABOVE TABLE FOR ALL FOUR PDET SIMULATIONS.
-% WE DEEMED IT TO OPPRESSIVE AND INSTEAD GENERATED THE FIGURES CURRENTLY IN THE PLOT
-%\begin{table}[ht]\centering\footnotesize
-%\setlength{\tabcolsep}{4pt} %default is 6
-%\begin{tabular}{l|l|l|l|cccc|cccc||cccc|cccc}
-% \multicolumn{4}{c|}{ } & \multicolumn{8}{c}{\underline{Non-mixture model}} & \multicolumn{8}{c}{\underline{Mixture model}} \\
-% \multicolumn{4}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
-% \multicolumn{4}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
-%  \hline
-%  \hline
-%\parbox[t]{2mm}{\multirow{14}{*}{\rotatebox[origin=c]{90}{TTDD used to simulate data}}} & \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Non-mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.67 & 0.69 & 0.75 & 0.95 & 0.11 & 0.78 & 0.45 & 0.59 & 0.76 & 0.78 & 0.83 & 0.96 & 0.00 & 0.35 & 0.55 & 0.61 \\
-%   &  &  & Lognormal & 0.54 & 0.59 & 0.71 & 0.94 & 0.81 & 0.45 & 0.29 & 0.47 & 0.68 & 0.73 & 0.82 & 0.96 & 0.05 & 0.63 & 0.63 & 0.44 \\ 
-%   &  &  & Weibull & 0.55 & 0.60 & 0.72 & 0.94 & 0.79 & 0.59 & 0.41 & 0.43 & 0.69 & 0.73 & 0.82 & 0.96 & 0.16 & 0.59 & 0.59 & 0.52 \\ 
-%\cline{3-20}
-%   &  &  & Exponential & 0.49 & 0.64 & 0.80 & 0.95 & 0.51 & 0.50 & 0.54 & 0.48 & 0.45 & 0.61 & 0.78 & 0.94 & 0.47 & 0.33 & 0.48 & 0.42 \\ 
-%\cline{3-20}
-%   & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.48 & 0.64 & 0.79 & 0.95 & 0.57 & 0.46 & 0.45 & 0.44 & 0.57 & 0.68 & 0.81 & 0.95 & 0.42 & 0.51 & 0.45 & 0.40 \\ 
-%   &  &  & Lognormal & 0.49 & 0.64 & 0.80 & 0.95 & 0.53 & 0.36 & 0.52 & 0.44 & 0.54 & 0.67 & 0.81 & 0.95 & 0.37 & 0.44 & 0.53 & 0.44 \\ 
-%   &  &  & Weibull & 0.48 & 0.61 & 0.78 & 0.95 & 0.40 & 0.46 & 0.38 & 0.49 & 0.59 & 0.69 & 0.81 & 0.96 & 0.37 & 0.46 & 0.49 & 0.42 \\ 
-%\cline{2-20}
-%\cline{2-20}
-%& \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.62 & 0.60 & 0.64 & 0.85 & 0.42 & 0.78 & 0.11 & 0.03 & 0.71 & 0.71 & 0.75 & 0.91 & 0.10 & 0.81 & 0.64 & 0.31 \\
-%   &  &  & Lognormal & 0.50 & 0.52 & 0.58 & 0.85 & 0.95 & 0.12 & 0.04 & 0.03 & 0.64 & 0.67 & 0.74 & 0.92 & 0.09 & 0.93 & 0.51 & 0.38 \\ 
-%   &  &  & Weibull & 0.50 & 0.51 & 0.54 & 0.71 & 0.92 & 0.23 & 0.01 & 0.01 & 0.64 & 0.66 & 0.70 & 0.87 & 0.39 & 0.92 & 0.46 & 0.28 \\ 
-%\cline{3-20}
-%   &  &  & Exponential & 0.88 & 0.93 & 0.96 & 0.99 & 0.00 & 0.00 & 0.00 & 0.00 & 0.50 & 0.63 & 0.78 & 0.94 & 0.54 & 0.45 & 0.42 & 0.51 \\ 
-%\cline{3-20}
-% & & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.24 & 0.26 & 0.28 & 0.37 & 0.00 & 0.00 & 0.00 & 0.00 & 0.52 & 0.58 & 0.72 & 0.94 & 0.76 & 0.51 & 0.40 & 0.41 \\ 
-%   &  &  & Lognormal & 0.18 & 0.20 & 0.22 & 0.27 & 0.00 & 0.00 & 0.00 & 0.00 & 0.49 & 0.61 & 0.78 & 0.94 & 0.66 & 0.58 & 0.46 & 0.43 \\ 
-%   &  &  & Weibull & 0.18 & 0.19 & 0.22 & 0.28 & 0.00 & 0.00 & 0.00 & 0.00 & 0.52 & 0.56 & 0.69 & 0.93 & 0.80 & 0.44 & 0.33 & 0.47 \\ 
-%  \hline
-%\end{tabular}
-%\caption{Summary of mixture vs. non-mixture model fits over 100 replicates.
-%In all cases, the inference model family matches the dataset family.  
-%Column titles denote the true detection probability (e.g. `p80' means $\pdet = 0.80$). 
-%Medians are averaged across all replicates.
-%Coverage is for 50\% credible intervals.
-%For nominal coverage, we expect values between (0.41, 0.59) 95\% of the time.}
-%\label{tbl:sim1}
-%\end{table}
-
-% Median posterior abundance from mixture models was unbiased for non-mixture data and positively biased by only 4-15\% for mixture data.  
-% Median posterior abundance from non-mixture models applied to non-mixture data were unbiased (exponential and peaked) or biased by 5-24\% (nonpeaked); however, for mixture data, the non-mixture models were biased by 20-60\% (nonpeaked), -17\% (exponential), and 183-265\% (peaked).
-% For the time-varying distributions, non-mixture models not only had greater bias, but 95\% credible intervals were generally 10-40\% wider (470\% wider for peaked mixture data).
-% For exponential datasets, the non-mixture model credible intervals were 28\% narrower for non-mixture data and an unrealistic 95\% narrower for mixture data.
-
-% Possible other comments
-%(1) For nonpeaked data, the true inference model is biased low\\
-%(2) Mixture inference models generate higher estimates than non-mixture models\\
-%(3) True models are generally good\\
-%(4) Estimates of non-mixture peaked data are much more precise\\
-%(5) Note that plots of TTDD at the median are indistinguishable from the truth except in the blatant fail scenario
-
 \subsection{Constant vs. non-constant detection mixture TTDDs}\label{sec:family}
@@
 
-% FIRST SUBMISSION TABLE
-%\ifdic
-%\begin{table}[ht]
-%\footnotesize\centering
-%\begin{tabular}{l|l|l|ccccc|ccccc}
-% \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Exponential mixture model}} & \multicolumn{5}{c}{\underline{Gamma mixture model}} \\
-% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-%  \hline
-%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 1.54 & 0.76 & 0.41 & 0.69 & 1.00 & --- \\
-%& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 2.08 & 0.85 & 0.72 & 0.44 & 0.69 & 0.93 \\ 
-%& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.67 & 0.79 & 0.49 & 0.69 & 1.00 & -0.08 \\ 
-%\cline{2-13}
-%& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & --- & 0.68 & 0.24 & 0.25 & 0.81 & -0.97 \\ 
-%\cline{2-13}
-%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 6.63 & 0.74 & 0.33 & 0.31 & 0.88 & --- \\ 
-%& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 18.54 & 0.84 & 0.73 & 0.44 & 0.81 & 0.71 \\ 
-%& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 1.91 & 0.66 & 0.15 & 0.19 & 0.75 & -0.11 \\
-%   \hline
-%\end{tabular}
-%\vspace{0.5cm}\\
-%\begin{tabular}{l|l|l|ccccc|ccccc}
-% \multicolumn{3}{c}{ } & \multicolumn{5}{c}{\underline{Lognormal mixture model}} & \multicolumn{5}{c}{\underline{Weibull mixture model}} \\
-% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC & Med. $p$ & Q($p$) & 50\% & 90\% & $\Delta$ DIC \\ 
-%  \hline
-%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.51 & 0.67 & 0.24 & 0.31 & 1.00 & 0.03 \\
-%& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & --- & 0.77 & 0.48 & 0.44 & 1.00 & 0.36 \\ 
-%& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.38 & 0.71 & 0.32 & 0.44 & 1.00 & --- \\ 
-%\cline{2-13}
-%& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & -0.19 & 0.63 & 0.20 & 0.25 & 0.75 & -0.62 \\ 
-%\cline{2-13}
-%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & -0.72 & 0.79 & 0.52 & 0.69 & 0.88 & 0.27 \\ 
-%& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & --- & 0.89 & 0.88 & 0.12 & 0.69 & 1.52 \\ 
-%& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & -0.50 & 0.70 & 0.29 & 0.56 & 0.94 & --- \\ 
-%   \hline
-%\end{tabular}
-%\caption{Summary of mixture inference models of all families fit to mixture datasets.  
-%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-%50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.  
-%$\Delta$ DIC: mean difference in DIC between the incorrect-family model minus the true model.}
-%\label{tbl:sim2}
-%\end{table}
-%
-%\else
-%
-%\begin{table}[ht]
-%\footnotesize\centering
-%\begin{tabular}{l|l|l|cccc|cccc}
-% \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Exponential mixture model}} & \multicolumn{4}{c}{\underline{Gamma mixture model}} \\
-% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\%  & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-%  \hline
-%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.88 & 0.91 & 0.06 & 0.69 & 0.76 & 0.41 & 0.69 & 1.00 \\
-%& &   Lognormal & 0.92 & 0.99 & 0.00 & 0.12 & 0.85 & 0.72 & 0.44 & 0.69 \\ 
-%& &   Weibull & 0.87 & 0.83 & 0.31 & 0.38 & 0.79 & 0.49 & 0.69 & 1.00 \\ 
-%\cline{2-11}
-%& &   Exponential & 0.77 & 0.37 & 0.38 & 0.94 & 0.68 & 0.24 & 0.25 & 0.81 \\ 
-%\cline{2-11}
-%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.36 & 0.00 & 0.00 & 0.00 & 0.74 & 0.33 & 0.31 & 0.88 \\ 
-%& &   Lognormal & 0.34 & 0.00 & 0.00 & 0.00 & 0.84 & 0.73 & 0.44 & 0.81 \\ 
-%& &   Weibull & 0.39 & 0.00 & 0.00 & 0.00 & 0.66 & 0.15 & 0.19 & 0.75 \\
-%   \hline
-%\end{tabular}
-%\vspace{0.5cm}\\
-%\begin{tabular}{l|l|l|cccc|cccc}
-% \multicolumn{3}{c}{ } & \multicolumn{4}{c}{\underline{Lognormal mixture model}} & \multicolumn{4}{c}{\underline{Weibull mixture model}} \\
-% \multicolumn{3}{c}{ } & Med. $p$ & Q($p$) & 50\% & 90\% & Med. $p$ & Q($p$) & 50\% & 90\% \\ 
-%  \hline
-%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.65 & 0.16 & 0.19 & 0.81 & 0.67 & 0.24 & 0.31 & 1.00 \\
-%& &   Lognormal & 0.72 & 0.30 & 0.44 & 0.88 & 0.77 & 0.48 & 0.44 & 1.00 \\ 
-%& &   Weibull & 0.68 & 0.23 & 0.31 & 0.88 & 0.71 & 0.32 & 0.44 & 1.00 \\ 
-%\cline{2-11}
-%& &   Exponential & 0.61 & 0.12 & 0.12 & 0.44 & 0.63 & 0.20 & 0.25 & 0.75 \\ 
-%\cline{2-11}
-%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.65 & 0.13 & 0.06 & 0.56 & 0.79 & 0.52 & 0.69 & 0.88 \\ 
-%& &   Lognormal & 0.76 & 0.36 & 0.38 & 0.94 & 0.89 & 0.88 & 0.12 & 0.69 \\ 
-%& &   Weibull & 0.58 & 0.03 & 0.00 & 0.12 & 0.70 & 0.29 & 0.56 & 0.94 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Summary of mixture inference models of all families fit to mixture datasets.  
-%Med $p$: average across simulations of the posterior median of $\pdet$ (true value = 0.80).  
-%Q($p$): average proportion of the posterior distribution of $\pdet$ that is larger than the true value.  
-%50\% and 90\% coverage is expressed as the proportion of simulations for which the true value of $\pdet$ lies within the appropriate credible interval.}
-%\label{tbl:sim2}
-%\end{table}
-%\fi
-
-% THIS IS SIMILAR TO THE ABOVE TABLE FOR ALL FOUR PDET SIMULATIONS.
-% WE DEEMED IT TO OPPRESSIVE AND INSTEAD GENERATED THE FIGURES CURRENTLY IN THE PLOT
-%\begin{table}[ht]
-%\footnotesize\centering\setlength{\tabcolsep}{4pt} %default is 6
-%\begin{tabular}{l|l|l|cccc|cccc||cccc|cccc}
-% \multicolumn{3}{c}{ } & \multicolumn{8}{c}{\underline{Exponential mixture model}} & \multicolumn{8}{c}{\underline{Gamma mixture model}} \\
-% \multicolumn{3}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
-% \multicolumn{3}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
-%  \hline
-%  \hline
-%\parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.84 & 0.83 & 0.86 & 0.94 & 0.00 & 0.04 & 0.25 & 0.38 & 0.63 & 0.63 & 0.65 & 0.81 & 0.23 & 0.87 & 0.16 & 0.10 \\
-%   &  & Lognormal & 0.89 & 0.91 & 0.93 & 0.97 & 0.00 & 0.00 & 0.00 & 0.07 & 0.64 & 0.67 & 0.74 & 0.92 & 0.09 & 0.93 & 0.51 & 0.38 \\ 
-%   &  & Weibull & 0.86 & 0.87 & 0.89 & 0.94 & 0.00 & 0.01 & 0.11 & 0.37 & 0.63 & 0.64 & 0.66 & 0.82 & 0.18 & 0.97 & 0.22 & 0.15 \\ 
-%\cline{2-19}
-%   &  & Exponential & 0.50 & 0.63 & 0.78 & 0.94 & 0.54 & 0.45 & 0.42 & 0.51 & 0.49 & 0.55 & 0.61 & 0.82 & 0.77 & 0.39 & 0.12 & 0.09 \\ 
-%\cline{2-19}
-%& \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.31 & 0.33 & 0.37 & 0.55 & 0.05 & 0.02 & 0.00 & 0.00 & 0.45 & 0.50 & 0.64 & 0.91 & 0.60 & 0.21 & 0.10 & 0.10 \\ 
-%   &  & Lognormal & 0.26 & 0.30 & 0.34 & 0.47 & 0.00 & 0.00 & 0.00 & 0.00 & 0.49 & 0.61 & 0.78 & 0.94 & 0.66 & 0.58 & 0.46 & 0.43 \\ 
-%   &  & Weibull & 0.30 & 0.32 & 0.38 & 0.57 & 0.05 & 0.00 & 0.00 & 0.00 & 0.43 & 0.47 & 0.57 & 0.84 & 0.51 & 0.13 & 0.02 & 0.00 \\ 
-%   \hline
-%\end{tabular}
-%\vspace{0.5cm}\\
-%\begin{tabular}{l|l|l|cccc|cccc||cccc|cccc}
-% \multicolumn{3}{c}{ } & \multicolumn{8}{c}{\underline{Lognormal mixture model}} & \multicolumn{8}{c}{\underline{Weibull mixture model}} \\
-% \multicolumn{3}{c|}{ } & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}} & \multicolumn{4}{c}{\underline{Posterior median}} & \multicolumn{4}{c}{\underline{Coverage}}\\
-% \multicolumn{3}{c|}{ } & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 & p50 & p65 & p80 & p95 \\
-%  \hline
-% \parbox[t]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{Data Mixture}}} & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Nonpk.}}} & Gamma & 0.71 & 0.71 & 0.75 & 0.91 & 0.10 & 0.81 & 0.64 & 0.31 & 0.63 & 0.63 & 0.67 & 0.87 & 0.54 & 0.94 & 0.34 & 0.22 \\
-%   &  & Lognormal & 0.78 & 0.82 & 0.88 & 0.96 & 0.01 & 0.19 & 0.33 & 0.46 & 0.68 & 0.73 & 0.81 & 0.95 & 0.21 & 0.68 & 0.59 & 0.57 \\ 
-%   &  & Weibull & 0.73 & 0.75 & 0.79 & 0.91 & 0.03 & 0.54 & 0.66 & 0.32 & 0.64 & 0.66 & 0.70 & 0.87 & 0.39 & 0.92 & 0.46 & 0.28 \\ 
-%\cline{2-19}
-%   &  & Exponential & 0.54 & 0.60 & 0.70 & 0.92 & 0.84 & 0.70 & 0.32 & 0.33 & 0.54 & 0.58 & 0.64 & 0.88 & 0.87 & 0.69 & 0.24 & 0.25 \\ 
-%\cline{2-19}
-% & \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{Peaked}}} & Gamma & 0.52 & 0.58 & 0.72 & 0.94 & 0.76 & 0.51 & 0.40 & 0.41 & 0.55 & 0.62 & 0.77 & 0.97 & 0.77 & 0.59 & 0.46 & 0.26 \\ 
-%   &  & Lognormal & 0.59 & 0.71 & 0.85 & 0.97 & 0.52 & 0.48 & 0.24 & 0.27 & 0.64 & 0.77 & 0.90 & 0.98 & 0.43 & 0.34 & 0.13 & 0.01 \\ 
-%   &  & Weibull & 0.49 & 0.53 & 0.65 & 0.89 & 0.75 & 0.31 & 0.22 & 0.09 & 0.52 & 0.56 & 0.69 & 0.93 & 0.80 & 0.44 & 0.33 & 0.47 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Summary of mixture inference models for all families.  
-%Column titles denote the true detection probability (e.g. `p80' means $\pdet = 0.80$). 
-%Medians are averaged across all replicates.
-%Coverage is for 50\% credible intervals.
-%For nominal coverage, we expect values between (0.41, 0.59) 95\% of the time.}
-%\label{tbl:sim2}
-%\end{table}
-
 When comparing the different three-parameter TTDDs, model misspecification was not as serious an issue because the models could better account for the patterns in time to detection.
@@
 Our results do not favor the use of one of these TTDDs over the others.
-
-%These results support the use of a non-constant rate mixture TTDD model unless you are absolutely sure your data arose from an exponential mixture TTDD.
-
-% Gamma models were more accurate than others for nonpeaked mixture datasets.  
-% Gamma and Weibull models seemed equally accurate for peaked mixture datasets.  
-% Also, 95\% credible intervals for gamma models were 5-15\% narrower than for Weibull models and 15-25\% narrower than for lognormal models.
-% 
-% Median posterior abundance for exponential models were positiviely biased only 4\% for exponential data, while time-varying TTDD estimates were biased from 19-34\% with average 95\% credible intervals being 3.5 to 5.5 times wider.
-% For nonpeaked mixture data, posterior median abundance was negatively biased 8-13\% for exponential models, was unbiased for gamma models and had positive bias of 5-20\% for Weibull and 13-23\% for lognormal.
-% So, in terms of raw abundance, the magnitude of bias for nonpeaked data from exponential models was commensurate (but in the opposite direction) with magnitudes from time-varying TTDD models; however, credible intervals from exponential models were only 10-30\% as wide as from other models, thus explaining the poor coverage in Table \ref{tbl:sim2}.
-% For peaked mixture data, exponential model median posterior abundance was biased 126\%, while time-varying TTDD estimates were only 3-23\% biased; exponential model credible intervals were from 2-12 times wider than from time-varying TTDD models.
 
@@
 Posteriors for the mixing parameter $\gamma$ and detection-related fixed and random effects were the same across gamma, lognormal, and Weibull mixture models but were narrower and location-shifted for the exponential mixture model.
-%It should be remembered that detection covariate effect estimates are conditional on the estimated mixing parameter $\gamma$.
-
-\ifdic
-\subsection{Model Fits}\label{sec:fits}
-
-In Sim1, DIC and posterior predictive checks could not easily differentiate a mixture from a non-mixture model for inference when a nonpeaked simulation model was used, even though non-mixture models displayed bias and poor coverage.
-Both tools clearly preferred mixture models when the data arose from a mixture exponential distribution or any of the peaked mixture distributions. 
-In these cases, posterior predictive p-values demonstrated that, relative to the observed data, the estimated TTDD placed too many counts during these intervals and too few in the first and later intervals.
-
-In Sim2, DIC did show a clear difference for exponential mixture fits of gamma and lognormal peaked datasets, and posterior predictive checks revealed lack-of-fit in the same scenarios, yielding p-values of 1.00 in the second observation interval.
-However, in all other cases, neither DIC nor posterior predictive checks evidenced preference for any model, even when bias and coverage across simulations suggested one was inferior.
-
-For Sim3, DIC was effective in the same scenarios where it was effective in Sim1 and Sim2.  
-Additionally, for exponential mixture data, DIC preferred the exponential mixture TTDD to time-varying mixture TTDDs.  
-Unfortunately, DIC could also be misleading.  
-DIC favored the exponential mixture to time-varying mixtures (by 4-8) when the data were in fact lognormal nonpeaked data.  
-Yet the posterior 95\% credible interval for $\pdet$ from the exponential mixture model did not contain the true value of $\pdet$, while the true value was within a central 60\% credible interval for all three other mixture models.
-In conjunction with the ineffectiveness of DIC for nonpeaked data in Sim2, this last finding calls into question the utility of DIC for nonpeaked datasets.
-
-DIC performance in Sim3 may differ from Sims 1 \& 2 not only because of the addition of fixed and random effects but also because DIC was calculated according to a different method.
-\fi
 
@@
 
-% Effects: * - means signif
-% Observer variation - Farnsworth2002, LinkSauer98, LinkSauer97, Dief07 cites Sauer94 & Dief03, Simons et al03
-% First-year observer - LinkSauer98
-% Time of day - Farnsworth2002, Soly13*, Amundson
-% Season (jdate) - Farnsworth2002, Soly13*, Amundson, Dief07
-% Quadratic effects - Soly13
-% Tree cover (density) - Soly13
-% Year (on detection!) - Reidy11, see also Norvell03
-% See Table 1 of Johnson2008, see McShea and Rappole, lit review from Warren2013, Rosenstock02
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{lccccccccc}
-%  \hline
-%Minutes & 0-2 & 2-3 & 3-4 & 4-5 & 5-6 & 6-7 & 7-8 & 8-9 & 9-10 \\ 
-%Count & 596 & 69 & 62 & 68 & 43 & 33 & 35 & 27 & 14 \\ 
-%\hline
-%\multicolumn{10}{l}{Posterior predictive check $p$-values by model:}\\
-%\hline
-%Expo & 0.000 & 1.000 & 1.000 & 0.347 & 0.590 & 0.372 & 0.007 & 0.006 & 0.190 \\ 
-%  Gamma & 0.582 & 0.820 & 0.394 & 0.006 & 0.362 & 0.637 & 0.241 & 0.519 & 0.977 \\ 
-%  LogNormal & 0.571 & 0.943 & 0.471 & 0.005 & 0.322 & 0.554 & 0.165 & 0.407 & 0.974 \\ 
-%  Weibull & 0.582 & 0.862 & 0.401 & 0.005 & 0.359 & 0.616 & 0.221 & 0.509 & 0.982 \\ 
-%\hline
-%  Exponential Mix & 0.533 & 0.804 & 0.577 & 0.021 & 0.501 & 0.668 & 0.190 & 0.338 & 0.918 \\ 
-%  Gamma Mix & 0.593 & 0.720 & 0.431 & 0.015 & 0.439 & 0.670 & 0.246 & 0.476 & 0.947 \\ 
-%  LogNormal Mix & 0.576 & 0.806 & 0.508 & 0.016 & 0.421 & 0.627 & 0.197 & 0.413 & 0.951 \\ 
-%  Weibull Mix & 0.596 & 0.734 & 0.411 & 0.013 & 0.421 & 0.663 & 0.252 & 0.492 & 0.958 \\ 
-%   \hline
-%\end{tabular}
-%\caption{\label{tbl:ovencounts} Ovenbird counts and posterior predictive p-values -- Pr$\left(D(n_i)^{(rep)} > D(n_i)^{(obs)}\right)$ -- by observation interval for each model across all surveys.
-P-values near zero or one may indicate poor model fit.}
-%\end{table}
-
 Figure \ref{ovenposteriors} presents posterior medians and credible intervals for model parameters, overall detection probability $\pdet$, and the logarithm of total Ovenbird abundance.  
 Estimates for the shape parameter $\alpha$ from the gamma and Weibull models are consistent with the data arising from an exponential distribution, although the uncertainty on this parameter remains relatively large.
-%Based on shape parameter estimates and patterns in the estimation of $\gamma$, $\beta^D$, $\pdet$, and total abundance across models, Ovenbird data most likely arose from an exponential or nonpeaked mixture TTDD.
 
@@
 
-\ifdic
-DIC calculations preferred the exponential mixture TTDD to the lognormal, Weibull, and gamma mixtures (differences of 6.95, 10.7, and 11.5, respectively).  
-These differences closely mirror those from the exponential mixture dataset in Sim3, but they also mirror those from the nonpeaked lognormal mixture dataset, where the exponential mixture 95\% credible interval for abundance did not contain the true value, and its posterior median was negatively biased by 18\% of the actual abundance.
-\fi
-
 In spite of the similarity of effect parameter estimates, the posterior distributions for detection probability and abundance differ greatly between the exponential and non-exponential models.  
 It is clear that the assumption of constant detection leads to much higher and more precise estimates of detection than would be obtained if we are unwilling to make that assumption.
-%The Ovenbird analysis highlights the ambiguity that emerged from the simulation studies: the results were entirely consistent with data generated from a finite-mixture involving a constant-rate $F_{T1}$ TTDD, but they were also consistent with finite-mixtures involving a time-varying TTDD.
 
@@
 
-%We formulated a model for interval-censored time to detection data that allows for non-constant detection rates.
 We formulated a model for single-species removal-sampled point-count survey data that allows for non-constant detection rates.
@@
 Because the exponential TTDD is a special case of both gamma and Weibull TTDDs, we can interpret the differences in estimation between models as resulting from the information conveyed by the assumption of constant detection.
-%Among the TTDDs modeled in our study, the gamma mixture TTDD gave the most accurate estimates across data distributions, though the Weibull mixture TTDD was comparable for peaked datasets.
 
@@
 These findings may well not hold if the same covariate is modeled in both abundance and detection models \citep{Kery2008}.
-
-\ifdic
-In a limited set of situations, posterior predictive checks and DIC can identify when non-mixture or exponential TTDDs inadequately describe the marginal pattern of detections over time.  
-But in many situations, neither tool points a clear way to select the correct family of TTDD.  
-\fi
 
@@
 
-%With the increasing use of microphone arrays for automated detection, it is tempting to think TTDD models may be valuable for microphone-collected data.  
-%This should be the case when variations in detection result solely from organismal behaviors, such as bout singing, or if we wish to model individual random effects in the detection model.  
-%However, some explanations for non-constant detection are explicitly related to the observation process --- either effects of the observer on animal behavior or variations in observer effort caused by survey methodology.
-%The assumption of constant detection seems more plausible in studies where observers are absent, though detection heterogeneity may still exist across subgroups of the population.
-
 Versions of time-varying models have been described for trap-based removal sampling and continuous-time capture-recapture.  
@@
 \citet{WangLoneragan1996}, \citet{HwangChao2002}, and \citet{Mantyniemi2005} all found constant-detection models to be flawed, producing underestimates of abundance and too-narrow error estimates; these resulted in inadequate coverage and also overstatement of effect significance.
-%Schnute: allowed time-varying detection which collapsed to a mixture model
-%Wang and Lonergan: \pi_s ~ beta for each interval at a site... not particularly applicable in an avian framework (with arbitrarily partitioned), except perhaps for bout singing?
-%- didn't have group heterogeneity
-%- constant models are overly precise (factor of 3)
-%Mantyniemi: individual-level variation in \pi_s ~ beta
-%- large number of trapping sessions required to overwhelm prior on variance (8-12)
-%- could also be viewed in terms of percent of total caught
-%- constant detection model badly biased
-%- constant detection model made indexing unreliable, since pairwise differences could be magnified
-%Laplanche: model selection with site-level abundance variation, site-level catchability variation, and individual-level catchability variation.
-%- only the abundance proved significant
-%Hwang and Chao: time-varying capture-recapture
-%- use a nonpeaked hazard function
-%- again, the constant model underestimates abundance, with coverage between 0-6\% for 95\% credible intervals (even allowing for covariate-based individual effects and behavioral responses).
 
@@
 This is an area of ongoing exploration.
-%Our adoption of a time-heterogeneous continuous time-to-detection approach in our model mirrors continuous time work in mark-and-recapture \citep{HwangChao2002, FarcomeniScacciatelli2013, Borchers2014}.
-
-%Examples of the interval-censored approach.
-%\citep{Farnsworth2002, Royle2004Generalized, Farnsworth2005, Alldredge2007, EffordDawson2009, Etterson2009, Reidy2011, Solymos2013, Amundson2014, Reidy2016}.
-
-%In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}...
-citations too numerous to enumerate.
-%In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
-%Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
-%The probability of detection is thus formulated as availability times perceptibility: $\pdet = P_aP_b$.  
-%However, our analysis makes possible a different strategy that is more consistent with a time-to-event conceptualization of the problem.  
-%Rather than define availability and perceptibility as independent detection events over the course of the observation period, we can define them at the event level.  
-%In terms of the model we have presented in this paper, this is equivalent to redefining the time-to-detection hazard function, $h_{sq}(t)$ --- which is now a function of the distance to individual $q$ --- as the product of the time-to-availability hazard function $h_{s}^A(t)$ and the perceptibility function $g(d_q)$.
-%So, $h_{sq}(t) = g(d_q) h_{s}^A(t)$.  
-%From standard survival analysis results, $p_{sq}^{(det)} = 1 - \exp\left(-g(d_q)\int\limits_0^{C_I} h_{s}^A(u)du\right) = 1 - \left[1 - F_T^A(C_I|\varphi_s, \boldsymbol{\theta}) \right]^{g(d_q)}$.
-%In short, the time-to-event conceptualization frames perceptibility as a rate-of-availability multiplier rather than as a probability-of-availability multiplier.  
-%In a context where rates of availability are allowed to vary from individual to individual, this makes an intuitive sense --- a bird that sings twenty times during the observation period should have its overall detection probability discounted much less than a bird at the same distance which only sings once.
-
-% Skeptics of removal sampling: Johnson08, EffordDawson09, Reidy11
 
@@
 We believe that more informative data collection, such as complete time-to-detection histories and microphone arrays, offer more effective tools for time-to-event modeling going forward.
-
-% \adam{Topics I have omitted from the Discussion (which reflects my own judgment, but you may disagree):
-% \begin{enumerate}
-% \item Selection of priors (especially for intercepts; also, the possibility of a spike-and-slab for $\alpha$)
-% % Note: now that I've seen van Dishoeck and Mantyniemi, perhaps I should worry about informativeness in priors.}
-% \item Values for $\pdet$ used in simulations
-% \end{enumerate}
-% }
 

### 2017-07-13 Supplement.tex
Remove comments

@@
 \end{sidewaystable}
-%\begin{sidewaystable}[ht]
-%\centering
-%\begin{tabular}{lllrrrrrrrrrrrrrrrr}
-%  \hline
-%Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
-%  \hline
-%Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
-%   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
-%  Gamma & Non-mixture & Nonpeaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
-%   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
-%   & Mixture & Nonpeaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
-%   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
-%  Lognormal & Non-mixture & Nonpeaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
-%   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
-%   & Mixture & Nonpeaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
-%   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
-%  Weibull & Non-mixture & Nonpeaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
-%   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
-%   & Mixture & Nonpeaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
-%   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
-%   \hline
-%\end{tabular}
-%\begin{tabular}
-%\end{tabular}
-%\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.
-Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
-%\end{sidewaystable}
 
@@
 \begin{figure}[h!]\centering
-%\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
-%\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
-%Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
-%Each column presents one family of simulated dataset.  
-%Upper plots show non-mixture datasets; lower plots show mixture datasets.  
-%`X' marks the expected marginal probability of detection based on true parameter values.}
 \includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
@@
 
-%\small
-%\section{Stan Models}\label{s:models}
-%\subsection{Exponential Model}
-%\lstinputlisting[breaklines]{./OVEN/Expo_model.stan}
-%\subsection{Exponential Mixture Model}
-%\lstinputlisting[breaklines]{./OVEN/ExpoMix.stan}
-%\subsection{Gamma Model}
-%\lstinputlisting[breaklines]{./OVEN/Gamma_model.stan}
-%\subsection{Gamma Mixture Model}
-%\lstinputlisting[breaklines]{./OVEN/GammaMix.stan}
-%\subsection{Lognormal Model}
-%\lstinputlisting[breaklines]{./OVEN/LogNormal_model.stan}
-%\subsection{Lognormal Mixture Model}
-%\lstinputlisting[breaklines]{./OVEN/LogNormalMix.stan}
-%\subsection{Weibull Model}
-%\lstinputlisting[breaklines]{./OVEN/Weibull_model.stan}
-%\subsection{Weibull Mixture Model}
-%\lstinputlisting[breaklines]{./OVEN/WeibullMix.stan}
-
 \end{document}

### 2017-07-14 Draft.tex
Comment out unnecessary packages
Corrected 'Overbird'. Replaced .pdf's with .jpg's in Draft.tex Still need to update Supplement.tex

@@
 \usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsfonts}
-\usepackage{parskip}
+% \usepackage{amssymb}
+% \usepackage{amsfonts}
+% \usepackage{parskip}
 \usepackage{natbib}
-\usepackage{color}
-\usepackage[dvipsnames]{xcolor}
+% \usepackage{color}
+% \usepackage[dvipsnames]{xcolor}
 \usepackage[margin=1cm]{caption}
 \usepackage{adjustbox}
-\usepackage[toc,page]{appendix}
-\usepackage{enumitem}
-\usepackage{multirow}
-\usepackage{setspace}
-\usepackage{url}
-\usepackage{xr}
-\usepackage{lineno}
+% \usepackage[toc,page]{appendix}
+% \usepackage{enumitem}
+% \usepackage{multirow} 
+\usepackage{setspace} % for \doublespacing
+% \usepackage{url}
+\usepackage{xr} % for cross referencing to supplement
+% \usepackage{lineno}
 
@@
 In addition, models with a mixture component can outperform non-mixture models even when the truth is non-mixture.  
-Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
+Finally, we analyze an Ovenbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
 We demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs but that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"}
+\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"} % Source file: Oral Prelim/TTDD_Demo.R
 \caption{\label{fig:schematic} 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"}
+\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
 \caption{
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"}
+\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"} % Source file: Sims/SimZero/RepresentativePosteriors.R
 \caption{\label{fig:sim2typical} 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"}
+\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
 \caption{
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors}
+\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors} % Source file: OVEN/Compile.R
 \caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.

### 2017-07-15 Supplement.tex
Commented out unused packages. Fixes #11.
Mostly: converting pdf images to jpg images.

@@
 \usepackage[margin=1.0in]{geometry}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsfonts}
-\usepackage{parskip}
-\usepackage[round]{natbib}
-\usepackage{caption}
-\usepackage{tabularx, booktabs}
-\usepackage{float, placeins}
-\usepackage{adjustbox}
-\usepackage{multirow}
-\usepackage{setspace}
-\usepackage{url}
-\usepackage{xr}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{rotating}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage{amsfonts}
+%\usepackage{parskip}
+%\usepackage[round]{natbib}
+%\usepackage{caption}
+%\usepackage{tabularx, booktabs}
+%\usepackage{float}
+\usepackage{placeins}     % Enables \FloatBarrier, which enforces a break between table/figures
+%\usepackage{adjustbox}
+\usepackage{multirow}     % Allows tables with vertically merged cells
+%\usepackage{setspace}
+%\usepackage{url}
+\usepackage{xr}    % Links to external document
+%\usepackage{listings}
+%\usepackage{verbatim}
+%\usepackage{graphicx}
+\usepackage{rotating}     % For landscape-oriented table
 
@@
 \newcommand{\pdet}{p^{(det)}}
+% Plug-in caption for figures created using \sfig
 \newcommand{\capt}[1]{Caterpillar plots of posterior parameter estimates from all models fit to the simulated #1 dataset (from simulations involving covariates and random effects).
@@
 Parameters designated with an 'A' are abundance parameters; those designated with a 'D' are detection parameters.}
+% Include graphics for posterior estimates of dataset parameters
 \newcommand{\sfig}[2]{\begin{figure}
-  \includegraphics[width=1.0\textwidth, page=#2]{Sims/SimFull/Posteriors_all_simdat.pdf}
+  \includegraphics[width=1.0\textwidth]{Sims/SimFull/Posteriors_all_simdat#2.jpg} % Source file: Sims/SimFull/Outputs_Code_sf.R
   \caption{\capt{#1}}
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
+\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family} % Source file: Sims/SimFull/Outputs_Code_sf.R
 \caption{\label{pdet_cater_family}  Simulation results from Section \ref{M-sec:simfull}.

### 2017-07-15 Sims/SimZero/Outputs_Code_zero.R

@@ -467,7 +467,8 @@ rp <- ggplot(subset(ov, statistic=="C90" & familymatch)) +
   scale_y_continuous(breaks = seq(0, 1, 0.5), limits=c(-0.05, 1.05)) +
   theme_bw() + ylab("Coverage") + xlab("Actual Detection Probability")
 
-pdf("mixture_fig.pdf", width=10, height=4.7)
+#pdf("mixture_fig.pdf", width=10, height=4.7)
+jpeg("mixture_fig.jpg", width=10, height=4.7, units="in", res=300)
 grid_arrange_shared_legend(lp, rp, position="bottom")
 dev.off()
 
@@ -491,7 +492,8 @@ rf <- ggplot(subset(ov, statistic=="C90" & datamix=="Mixed" & modelmix)) +
   scale_y_continuous(breaks = seq(0, 1, 0.5), limits=c(-0.05, 1.05)) +
   theme_bw() + ylab("Coverage") + xlab("Actual Detection Probability")
 
-pdf("family_fig.pdf", width=10, height=4.7)
+#pdf("family_fig.pdf", width=10, height=4.7)
+jpeg("family_fig.jpg", width=10, height=4.7, units="in", res=300)
 grid_arrange_shared_legend(lf, rf, position="bottom")
 dev.off()
  

### 2017-07-26 Draft.tex
Remove commented packages
Use graphicspath
Add directories for graphicspath

@@
 \usepackage{amsmath}
-% \usepackage{amssymb}
-% \usepackage{amsfonts}
-% \usepackage{parskip}
 \usepackage{natbib}
-% \usepackage{color}
-% \usepackage[dvipsnames]{xcolor}
 \usepackage[margin=1cm]{caption}
 \usepackage{adjustbox}
-% \usepackage[toc,page]{appendix}
-% \usepackage{enumitem}
-% \usepackage{multirow} 
 \usepackage{setspace} % for \doublespacing
-% \usepackage{url}
 \usepackage{xr} % for cross referencing to supplement
-% \usepackage{lineno}
+
+\graphicspath{{"Oral Prelim/"}{Sims/SimZero/}{OVEN/oven_sum/}}
 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"} % Source file: Oral Prelim/TTDD_Demo.R
+\includegraphics[width=\textwidth]{AMRO_JABES} % Source file: Oral Prelim/TTDD_Demo.R
 \caption{\label{fig:schematic} 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
+\includegraphics[width=\textwidth]{mixture_fig} % Source file: Sims/SimZero/Outputs_Code_zero.R
 \caption{
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"} % Source file: Sims/SimZero/RepresentativePosteriors.R
+\includegraphics[width=\textwidth]{Posteriors_Families_draft} % Source file: Sims/SimZero/RepresentativePosteriors.R
 \caption{\label{fig:sim2typical} 
@@
 \begin{figure}\centering
-\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
+\includegraphics[width=\textwidth]{family_fig} % Source file: Sims/SimZero/Outputs_Code_zero.R
 \caption{
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors} % Source file: OVEN/Compile.R
+\includegraphics[width=0.95\textwidth]{OVEN_Posteriors} % Source file: OVEN/Compile.R
 \caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.

### 2017-07-31 Supplement.tex
Remove commented packages
Use graphicspath

@@
 \usepackage[margin=1.0in]{geometry}
-%\usepackage{amsmath}
-%\usepackage{amssymb}
-%\usepackage{amsfonts}
-%\usepackage{parskip}
-%\usepackage[round]{natbib}
-%\usepackage{caption}
-%\usepackage{tabularx, booktabs}
-%\usepackage{float}
 \usepackage{placeins}     % Enables \FloatBarrier, which enforces a break between table/figures
-%\usepackage{adjustbox}
 \usepackage{multirow}     % Allows tables with vertically merged cells
-%\usepackage{setspace}
-%\usepackage{url}
 \usepackage{xr}    % Links to external document
-%\usepackage{listings}
-%\usepackage{verbatim}
-%\usepackage{graphicx}
 \usepackage{rotating}     % For landscape-oriented table
+
+\graphicspath{{Sims/SimFull/}}
 
@@
 \newcommand{\pdet}{p^{(det)}}
-% Plug-in caption for figures created using \sfig
 \newcommand{\capt}[1]{Caterpillar plots of posterior parameter estimates from all models fit to the simulated #1 dataset (from simulations involving covariates and random effects).
@@
 Parameters designated with an 'A' are abundance parameters; those designated with a 'D' are detection parameters.}
-% Include graphics for posterior estimates of dataset parameters
 \newcommand{\sfig}[2]{\begin{figure}
-  \includegraphics[width=1.0\textwidth]{Sims/SimFull/Posteriors_all_simdat#2.jpg} % Source file: Sims/SimFull/Outputs_Code_sf.R
+  \includegraphics[width=1.0\textwidth]{Posteriors_all_simdat#2.jpg} % Source file: Sims/SimFull/Outputs_Code_sf.R
   \caption{\capt{#1}}
@@
 \vspace{1.2cm}
-% Note: both tables are in the same 'table' environment.
 
@@
 \begin{figure}[h!]\centering
-\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family} % Source file: Sims/SimFull/Outputs_Code_sf.R
+\includegraphics[width=0.98\textwidth]{pdet_cater_family} % Source file: Sims/SimFull/Outputs_Code_sf.R
 \caption{\label{pdet_cater_family}  Simulation results from Section \ref{M-sec:simfull}.
@@
 
-\subsection{Posteriors parameter estimates from simulations including covariates and random effects}\label{s:sim3}
+\subsection{Posterior parameter estimates from simulations including covariates and random effects}\label{s:sim3}
 \sfig{non-mixture exponential}{1}

### 2017-08-15 OralPrelim/TTDD_Demo.R
Fixing Figure 1.
Renaming "Oral Prelim" directory to "OralPrelim"

@@ -1,0 +1,294 @@
+########## Generate figures of mixture expo and mixture gamma distributions to illustrate
+# how these models fit the data
+
+setwd("C:/Users/Adam/Documents/Adam/IAState/Stat 599JN/TimeToDetect/OralPrelim")
+#load("../OVEN/Data9.Rdata")
+load("AMRO.Rdata") # There's a sharper contrast when we use AMRO
+
+# Read in and restructure data
+counts    <- colSums(dat$y)
+cntbyint  <- counts
+counts[1] <- counts[1]/2
+names(counts) <- paste0(c(0,2:9)," - ", 2:10)
+
+### MLEs for TTDDs
+
+# Exponential MLE -- outputs estimate as (phi, gamma)
+phi8.llik <- function(parameters, cnt, tau=c(0,2:10)){
+  ints  <- length(cnt)
+  phi   <- parameters[1]
+  gamma <- parameters[2]
+  llk <- cnt[1] * log(1-gamma + gamma * (pexp(tau[2],phi) - pexp(tau[1],phi)))
+  for (j in 2:ints) llk <- llk + cnt[j]*log(gamma * (pexp(tau[j+1],phi) - pexp(tau[j],phi)))
+  llk <- llk - sum(cnt)*log(1-gamma + gamma*pexp(max(tau),phi))
+  return(unname(llk))
+}
+efit <- optim(c(0.13, 0.6), fn=phi8.llik, gr=NULL, cnt=cntbyint, control=list(fnscale=-1))$par
+p.e <- efit[2]*pexp(10,efit[1]) + (1-efit[2])
+
+# Gamma MLE -- outputs estimate as (alpha, phi, gamma)
+# Ignore the warnings
+gamma.llik <- function(parameters, cnt, tau=c(0,2:10)){
+  ints  <- length(cnt)
+  alpha <- parameters[1]
+  phi   <- parameters[2]
+  gamma <- parameters[3]
+  llik <- cnt[1] * log(1-gamma + gamma * (pgamma(tau[2],alpha,phi) - pgamma(tau[1],alpha,phi)))
+  for(j in 2:ints) llik <- llik + 
+      cnt[j] * log(gamma * (pgamma(tau[j+1],alpha,phi) - pgamma(tau[j],alpha,phi)))
[…]
